### 2015-03-25 paper1.tex
waka

\documentclass[11pt]{article}

\usepackage{graphicx}
\usepackage{amssymb,amsmath}
\usepackage{textcomp}

%\usepackage{subcaption}
%\usepackage{float}
\usepackage{setspace}
\usepackage{fullpage}
\usepackage[font=scriptsize]{caption}
%\usepackage{fullpage}
%\setcounter{secnumdepth}{1}
\begin{document}

\title{Pattern adaptation and normalization reweighting}
\author{Kari Y. Lam, Zachary M. Westrick, Others TBA}
\maketitle

\begin{abstract}
I'm MC Zack and I've got something to say. Hey. Hey. Hey. Hey. Hey
\end{abstract}

\section{Introduction}

\begin{figure}
\begin{center}
  \includegraphics[scale=0.45]{simulated.png}
  \caption{\label{fig:figure1} This is what a figure looks like}
  \end{center}
\end{figure}

\section{Methods}
Look an equation array
\begin{equation}
\begin{array}{l}
R_i^{\mathrm{input}}(\theta) = c g_i^{\mathrm{input}}\exp(-(\theta - \theta_i)^2 / (2\sigma^2_{\mathrm{input}}))
\\
R_i^{\mathrm{output}}(\theta) = g_i^{\mathrm{output}}\sum_{j=0}^N R_j^{\mathrm{input}}(\theta) \exp(-(\theta_i - \theta_j)^2/(2\sigma^2_{\mathrm{output}})) .
\end{array}
\end{equation}
\bibliographystyle{plain}
\bibliography{paper}

\end{document}

### 2015-03-25 paper1.tex
wrote description of simulated data generation

@@
 
-\title{Pattern adaptation and normalization reweighting}
+\title{GRN inference in multiple species}
 \author{Kari Y. Lam, Zachary M. Westrick, Others TBA}
@@
 \section{Methods}
+\subsection{GRN}
+are good
+\subsection{fused GRN}
+yeees
+\subsection{Simulated data}
+Generation of simulated data begins with the production of random orthology mappings.
+We produce a one-to-one orthology by pairing random genes until a specified fraction have been assigned orthologs.
+This process is carried out seperately for TFs and non-TF genes, so that TFs and non-TF genes are never assigned to be orthologous.
+We then produce a pair of random networks ($B^1$ and $B^2$) as follows.
+For each unfilled entry in $B^1$ or $B^2$, we enumerate the set $C$ consisting of the entry along with every entry in either matrix to which it is fused.
+With probability equal to the sparsity rate we assign every entry in $C$ to be 0, otherwise we sample a value $v \sim \mathcal{N}(0,1)$ and independently assign each entry in $C$ to $v + \mathcal{N}(0, \sigma_f^2)$.
+$\sigma_f$ is a parameter that controls the distribution of differences in the values of fused coefficients, so that the nonzero coefficients of $B^1, B^2$ are distributed as $\mathcal{N}(0, 1 + \sigma_f^2)$.
+
+Given a network $B$, we generate $N$ samples of gene expressions at each of two timepoints.
+The condition by gene expression matrix for timepoint one, $Y_{T1}$, is sampled randomly from a multivariate Gaussian distribution with identity covariance matrix.
+$X_{T1}$ is the TF expression sub-matrix of $Y_{T1}$, and consists of columns of $Y_{T1}$ that correspond to TFs.
+The gene expression matrix at timepoint two, $Y_{T2}$ is sampled as $Y_{T2} = Y_{T1} + BX_{T1}$.
+This process is carried out separately for each network.
+
+Following generation of simulated data, we may introduce error into the orthology mapping.
+This can take the form of discarding a specified fraction of true orthologies (governed by a false-negative rate), or by introducing random false orthologies (governed by a false-positive rate).
+For convenience, the false-positive rate is specified in units of the number of true orthologs, and not the number of possible orthologs.
+\subsection{Bacterial data}
+is good
 Look an equation array

### 2015-03-30 paper1.tex
more
hi

@@
 \documentclass[11pt]{article}
-
 \usepackage{graphicx}
+\usepackage{mathtools}
 \usepackage{amssymb,amsmath}
 \usepackage{textcomp}
+\DeclareMathOperator*{\argmin}{arg\,min}
 
@@
 \subsection{GRN}
-are good
-\subsection{fused GRN}
-yeees
+We model gene regulatory networks using ordinary differential equations where the change in expression of some gene $x_{i}$ is described as a linear combination of regulators:
+\begin{equation}
+\frac{\mathrm d}{\mathrm d t} = -\alpha_{i}x_{i} + \sum \beta_{i,j}x_{j}
+\end{equation}
+where $\alpha$ is the degradation rage and $\beta$ is a set of parameters to be estimated.
+
+\noindent We approach the learning of the $\beta$ matrix using regression. We solve
+\begin{equation}
+\argmin_\beta\vert \vert X\beta - Y_2 \vert \vert ^2
+\end{equation}
+where the matrix $X \in R^{n \times k}$ is the expression matrix with $n$ conditions and $k$ transcription factors.
+$X$ is a submatrix of $Y \in R^{n \times m}$, the gene expression matrix with $m$ genes.
+Typically, $n < k < m$, so this problem is ill-posed without regularization.
+We use ridge regression and solve
+\begin{equation}
+\argmin_\beta\vert \vert X\beta - Y_2 \vert \vert ^2 + \lambda_R \vert \vert \beta \vert \vert ^2
+\end{equation} 
+with $\lambda_R$ as the weight of the regularization term.
+
+\subsection{Fused GRN}
+Related species are expected to be governed by similar but not necessarily identical gene regulatory networks.
+Pooling data from multiple species drowns out species-specific mechanisms, but learning GRNs separately limits the availability of training data.
+We approach the problem of cross-species GRN inference by learning networks simultaneously with an additional constraint in the form of an L2 penalty on differences between interactions between orthologous TF, gene pairs.
+The assumption is that orthologous genes play analogous roles in their respective GRNs, and therefore, the interactions between orthologous TF, gene pairs should have similar weights.
+\\
+
+We solve
+\begin{equation}
+\argmin_\beta \vert \vert X^1 \beta^1 - Y_2 ^1 \vert \vert ^2 + \vert \vert X^2 \beta^2 - Y_2 ^2 \vert \vert ^2 + \lambda_R \vert \vert \beta^1 + \beta^2 \vert \vert ^2 + \lambda_S \sum\limits_{i^1, i^2, j^1, j^2 \in C} (\beta^1_{i^1,j^1} - \beta^2_{i^2,j^2})^2
+\end{equation}
+where $C$ is the set of orthologous TF, gene pairs.\newline
+
+If $\lambda_S = 0$, this is equivalent to solving the two regression problems separately, while $\lambda_S = \infty$ corresponds to requiring equality between constrained entries of $B^1, B^2$ where in the case of 1 to 1 correspondence weights, this is equivalent to pooling the data for the two species and solving one regression problem.
+\\
+
+We use an L2 norm because it can be solved in closed form and because the structure of differences between corresponding weights may not be sparse, but should simply be made small.
+
 \subsection{Simulated data}
@@
 We produce a one-to-one orthology by pairing random genes until a specified fraction have been assigned orthologs.
-This process is carried out seperately for TFs and non-TF genes, so that TFs and non-TF genes are never assigned to be orthologous.
+This process is carried out separately for TFs and non-TF genes, so that TFs and non-TF genes are never assigned to be orthologous.
 We then produce a pair of random networks ($B^1$ and $B^2$) as follows.
@@
 $X_{T1}$ is the TF expression sub-matrix of $Y_{T1}$, and consists of columns of $Y_{T1}$ that correspond to TFs.
-The gene expression matrix at timepoint two, $Y_{T2}$ is sampled as $Y_{T2} = Y_{T1} + BX_{T1}$.
+Treating the decay rate as 0, the gene expression matrix at timepoint two, $Y_{T2}$ is sampled as $Y_{T2} = Y_{T1} + BX_{T1}$.
 This process is carried out separately for each network.
@@
 For convenience, the false-positive rate is specified in units of the number of true orthologs, and not the number of possible orthologs.
+
+For the purposes of evaluating simulated network recovery, we define a gold standard network as the support of the beta matrices.
+Priors used in network inference are interactions from the gold standard.
+The list of priors can be be manipulated to include false positives and false negatives as with the generation of orthologs.
+
 \subsection{Bacterial data}

### 2015-03-30 paper1.tex

@@
+<<<<<<< HEAD
+\documentclass[11pt]{article}
+
+\usepackage{graphicx}
+\usepackage{amssymb,amsmath}
+\usepackage{textcomp}
+
+%\usepackage{subcaption}
+%\usepackage{float}
+\usepackage{setspace}
+\usepackage{fullpage}
+\usepackage[font=scriptsize]{caption}
+%\usepackage{fullpage}
+%\setcounter{secnumdepth}{1}
+\begin{document}
+
+\title{GRN inference in multiple species}
+\author{Kari Y. Lam, Zachary M. Westrick, Others TBA}
+\maketitle
+
+\begin{abstract}
+I'm MC Zack and I've got something to say. Hey. Hey. Hey. Hey. Hey
+\end{abstract}
+
+\section{Introduction}
+
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{simulated.png}
+  \caption{\label{fig:figure1} This is what a figure looks like}
+  \end{center}
+\end{figure}
+
+\section{Methods}
+\subsection{GRN}
+are good
+\subsection{fused GRN}
+yeees
+\subsection{Simulated data}
+Generation of simulated data begins with the production of random orthology mappings.
+We produce a one-to-one orthology by pairing random genes until a specified fraction have been assigned orthologs.
+This process is carried out seperately for TFs and non-TF genes, so that TFs and non-TF genes are never assigned to be orthologous.
+We then produce a pair of random networks ($B^1$ and $B^2$) as follows.
+For each unfilled entry in $B^1$ or $B^2$, we enumerate the set $C$ consisting of the entry along with every entry in either matrix to which it is fused.
+With probability equal to the sparsity rate we assign every entry in $C$ to be 0, otherwise we sample a value $v \sim \mathcal{N}(0,1)$ and independently assign each entry in $C$ to $v + \mathcal{N}(0, \sigma_f^2)$.
+$\sigma_f$ is a parameter that controls the distribution of differences in the values of fused coefficients, so that the nonzero coefficients of $B^1, B^2$ are distributed as $\mathcal{N}(0, 1 + \sigma_f^2)$.
+
+Given a network $B$, we generate $N$ samples of gene expressions at each of two timepoints.
+The condition by gene expression matrix for timepoint one, $Y_{T1}$, is sampled randomly from a multivariate Gaussian distribution with identity covariance matrix.
+$X_{T1}$ is the TF expression sub-matrix of $Y_{T1}$, and consists of columns of $Y_{T1}$ that correspond to TFs.
+The gene expression matrix at timepoint two, $Y_{T2}$ is sampled as $Y_{T2} = Y_{T1} + BX_{T1}$.
+This process is carried out separately for each network.
+
+Following generation of simulated data, we may introduce error into the orthology mapping.
+This can take the form of discarding a specified fraction of true orthologies (governed by a false-negative rate), or by introducing random false orthologies (governed by a false-positive rate).
+For convenience, the false-positive rate is specified in units of the number of true orthologs, and not the number of possible orthologs.
+Priors are defined by the support of $B$.
+We can introduce false-positive or false-negative errors, as in the orthology mapping, in order to study the effect of faulty prior information.
+\subsection{Bacterial data}
+is good
+Look an equation array
+\begin{equation}
+\begin{array}{l}
+R_i^{\mathrm{input}}(\theta) = c g_i^{\mathrm{input}}\exp(-(\theta - \theta_i)^2 / (2\sigma^2_{\mathrm{input}}))
+\\
+R_i^{\mathrm{output}}(\theta) = g_i^{\mathrm{output}}\sum_{j=0}^N R_j^{\mathrm{input}}(\theta) \exp(-(\theta_i - \theta_j)^2/(2\sigma^2_{\mathrm{output}})) .
+\end{array}
+\end{equation}
+\bibliographystyle{plain}
+\bibliography{paper}
+
+\end{document}
+
+=======
 \documentclass[11pt]{article}
@@
 \end{document}
+
+>>>>>>> 34faf0e6aea8fc5a3af9cb49bfcff0b2e3eaa7ed

### 2015-03-31 paper1.tex
waka

--- 2015-03-25 paper1.tex
@@
 \documentclass[11pt]{article}
-
 \usepackage{graphicx}
+%\usepackage{mathtools}
 \usepackage{amssymb,amsmath}
 \usepackage{textcomp}
+\DeclareMathOperator*{\argmin}{arg\,min}
 
@@
 \subsection{GRN}
-are good
-\subsection{fused GRN}
-yeees
+We begin with an existing framework for estimating the interaction weights in a gene regulatory network (GRN).
+In the Inferelator algorithm, linear differential equations are used to model changes in gene expression.
+The rate at which the expression of gene $i$, $x_i$, changes is governed by degradation of existing transcripts with rate $\alpha$ plus a linear combination of transcription factor (TF) expressions.
+\begin{equation}
+\frac{\mathrm d}{\mathrm d t} x_i = -\alpha_{i}x_{i} + \sum \beta_{i,j}x_{j}
+\end{equation}
+$\beta$, the matrix defining the influence of each TF on each gene, is what we hope to estimate.
+The decay rate $\alpha$ is treated as fixed for all genes, and is set assuming a time-constant of 10 minutes [CITATIONS], as in [CITATIONS].
+Let $x_i(t)$ be the expression of gene $i$ at time $t$.
+Given time-series data on the expression of gene $i$ at timepoints $t_k$ and $t_{k+1}$, we can approximate the rate of change of $x_i$ as $x_i'(t_k)=\frac{x_i(t_{k+1})-x_i(t_k)}{t_{k+1}-t_k}$.
+This gives us, for each gene $i$ and time $t_{k}$ an equation
+
+\begin{equation}
+\frac{x_i(t_k)-x_i(t_1)}{t_2-t_1} + \alpha_{i}x_{i}(t_k)= \sum \beta_{i,j}x_{j}(t_k)
+\end{equation}
+\noindent We can summarize these equations in matrix form as
+\begin{equation}
+Y = \beta X
+\end{equation}
+\noindent We approach the learning of the $\beta$ matrix using regression.
+Because there are typically far fewer conditions than possible regressors (TFs), we introduce a ridge regularization constraint with weight $\lambda_R$ and solve
+\begin{equation}
+\argmin_\beta\vert \vert X\beta - Y_2 \vert \vert ^2 + \lambda_R \vert \vert \beta \vert \vert ^2
+\end{equation}
+This is essentially the same basic formulation as was used in the Inferelator algorithm, which we extend to the case of simultaneously inferring the GRNs of multiple species.
+
+\subsection{Fused GRN}
+Related species are expected to be governed by similar but not necessarily identical gene regulatory networks.
+We approach the problem of cross-species GRN inference by introducing constraints into the above regression formulation to penalize differences between interaction weights that are expected to be similar to one another based on gene orthology.
+We can then solve the penalized regression problems simultaneously, in order to obtain a GRN for each species.
+Consider the case of organisms $A$ and $B$, governed by GRNs $\beta^A$ and $\beta^B$ (the following approach applies equally well to more than two species).
+If TF $g^A$ in organism $A$ and TF $h^B$ in organism $B$ are orthologs, and gene $k^A$ and $l^B$ are orthologs, then we expect that the $g^A \rightarrow k^A$ interaction weight should be similar to the $h^B \rightarrow l^B$ interaction weight, and we introduce a fusion constraint between these analogous interactions.
+In terms of the above regression formulation, we expect that $\beta^A_{g,h} \approx \beta^B_{k,l}$, and include a penalty term $\lambda_S(\beta^A_{g,h} - \beta^B_{k,l})^2$ to the quantity being minimized in order to ensure similarity.
+The parameter $\lambda_S$ controls how highly to penalize differences between fused coefficients.
+This gives us the final equation to be minimized
+\begin{equation}
+\argmin_{(\beta^A, \beta^B)} \displaystyle\sum_{S \in (A, B)} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\beta^{S_1}_{g,h} \approx \beta^{S_2}_{k,l}} \lambda_S(\beta^{S_1}_{g,h} - \beta^{S_2}_{k,l})^2
+\end{equation}
+Where the second sum is over pairs of interactions with fusion constraints.
+Note that each term in the quantity being minimized is an L2 norm of a linear combination of gene expressions, so the entire problem can be solved through linear regression with a suitable design matrix [DETAILS SHOULD GO SOMEWHERE ELSE, OR NOWHERE].
+Because pairs of genes can be solved independently as long as they are not linked by fusion constraints, or a chain of fusion constraints, the problem of cross-species GRN inference decomposes into a large number of manageable regression problems as long as the orthology mappings are sparse.
+
+[DIGRESSION?]
+
+When there is a one-to-one orthology between the species being considered (ie different cell-lines of the same organism), the choice of $\lambda_S$ allows one to interpolate between fitting each network independently ($\lambda_S=0$) and pooling data together as if it came from one source ($\lambda_S=\inf$).
+Part of the appeal of the approach, however, is that it allows pooling of data even when there is incomplete orthology.
+By introducing constraints on the similarity of individual interactions, rather than on the networks as a whole [CITE IBM?], we can pool some information across species even when an arbitrarily small fraction of genes have orthologs.
+This is particularly useful when dealing with a large number of species; pairwise orthologies may be nearly complete even when the number of genes present in every organism is small.
+\subsubsection{Equivalent prior}
+I want to talk about the fusion constraints as a Gaussian prior, and talk about the adjustment stuff
+
 \subsection{Simulated data}

### 2015-04-03 paper1.tex
added alternative penalty section

@@
 \title{GRN inference in multiple species}
-\author{Kari Y. Lam, Zachary M. Westrick, Others TBA}
+\author{Kari Y. Lam, Zachary M. Westrick, Frank Underwood}
 \maketitle
@@
 \begin{abstract}
-I'm MC Zack and I've got something to say. Hey. Hey. Hey. Hey. Hey
+HELLOOOOOOOOOOOOOOOOOOOOOOOO
 \end{abstract}
@@
 \end{equation}
-$\beta$, the matrix defining the influence of each TF on each gene, is what we hope to estimate.
-The decay rate $\alpha$ is treated as fixed for all genes, and is set assuming a time-constant of 10 minutes [CITATIONS], as in [CITATIONS].
+We are interested in learning $\beta$, the matrix defining the influence of each TF on each gene.
+We fix the decay rate $\alpha$ for all genes, and set it assuming a time-constant of 10 minutes [CITATIONS], as in [CITATIONS].
 Let $x_i(t)$ be the expression of gene $i$ at time $t$.
@@
 \begin{equation}
-\frac{x_i(t_k)-x_i(t_1)}{t_2-t_1} + \alpha_{i}x_{i}(t_k)= \sum \beta_{i,j}x_{j}(t_k)
+\frac{x_i(t_{k+1})-x_i(t_k)}{t_{k+1}-t_k} + \alpha_{i}x_{i}(t_k)= \sum \beta_{i,j}x_{j}(t_k)
 \end{equation}
@@
 \subsection{Fused GRN}
-Related species are expected to be governed by similar but not necessarily identical gene regulatory networks.
-We approach the problem of cross-species GRN inference by introducing constraints into the above regression formulation to penalize differences between interaction weights that are expected to be similar to one another based on gene orthology.
+We expect related species to be governed by similar but not necessarily identical gene regulatory networks.
+We approach the problem of cross-species GRN inference by introducing constraints into the above regression formulation to penalize differences between interaction weights that are expected to be similar based on gene orthology.
 We can then solve the penalized regression problems simultaneously, in order to obtain a GRN for each species.
-Consider the case of organisms $A$ and $B$, governed by GRNs $\beta^A$ and $\beta^B$ (the following approach applies equally well to more than two species).
+Consider the case of organisms $A$ and $B$, governed by GRNs $\beta^A$ and $\beta^B$ (the following approach applies equally well to more than two species but for simplicity we continue with the case of two species).
 If TF $g^A$ in organism $A$ and TF $h^B$ in organism $B$ are orthologs, and gene $k^A$ and $l^B$ are orthologs, then we expect that the $g^A \rightarrow k^A$ interaction weight should be similar to the $h^B \rightarrow l^B$ interaction weight, and we introduce a fusion constraint between these analogous interactions.
-In terms of the above regression formulation, we expect that $\beta^A_{g,h} \approx \beta^B_{k,l}$, and include a penalty term $\lambda_S(\beta^A_{g,h} - \beta^B_{k,l})^2$ to the quantity being minimized in order to ensure similarity.
+In terms of the above regression formulation, we expect that $\beta^A_{g,k} \approx \beta^B_{h,l}$, and include a penalty term $\lambda_S(\beta^A_{g,k} - \beta^B_{h,l})^2$ to the quantity being minimized in order to ensure similarity.
 The parameter $\lambda_S$ controls how highly to penalize differences between fused coefficients.
@@
 \begin{equation}
-\argmin_{(\beta^A, \beta^B)} \displaystyle\sum_{S \in (A, B)} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\beta^{S_1}_{g,h} \approx \beta^{S_2}_{k,l}} \lambda_S(\beta^{S_1}_{g,h} - \beta^{S_2}_{k,l})^2
+\argmin_{(\beta^A, \beta^B)} \displaystyle\sum_{S \in (A, B)} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\beta^{S_1}_{g,k} \approx \beta^{S_2}_{h,l}} \lambda_S(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})^2
 \end{equation}
@@
 Note that each term in the quantity being minimized is an L2 norm of a linear combination of gene expressions, so the entire problem can be solved through linear regression with a suitable design matrix [DETAILS SHOULD GO SOMEWHERE ELSE, OR NOWHERE].
-Because pairs of genes can be solved independently as long as they are not linked by fusion constraints, or a chain of fusion constraints, the problem of cross-species GRN inference decomposes into a large number of manageable regression problems as long as the orthology mappings are sparse.
+The $\beta$ vectors corresponding to a pair of genes can be solved independently as long as the pair is not linked by fusion constraints, or a chain of fusion constraints; thus the problem of cross-species GRN inference decomposes into a large number of manageable regression problems as long as the orthology mappings are sparse.
 
@@
 I want to talk about the fusion constraints as a Gaussian prior, and talk about the adjustment stuff
+
+\subsection{Alternative penalty functions}
+Sometimes gene orthology is not a good predictor of functional similarity.
+To simultaneously learn when a gene's function has diverged from that of its ortholog, and improve model selection, we consider alternative penalized least squares approaches.
+Our fusion penalty pulls beta values for analogous interactions towards each other; since we use an L2 penalty, interaction weights which differ from each other the most are excessively penalized.
+SCAD [Fan Li 2001] and MCP [Zhang] are two alternative penalties which diminish this bias.
+Fan and Li point out that the bias can be removed by choosing a constant penalty beyond some threshold.
+This is analogous to 'unfusing' certain interactions, and easing up on the demand that certain beta values be close in value when doing so strongly opposes minimizing the mean squared error component of the penalty function.
+SCAD and MCP both use an L1 penalty near the origin, setting small estimated coefficients to zero and choosing a sparse solution.
+Because we do not expect most fused beta values to be identical, but rather similar, we amend the alternative fusion penalty to be L2 near the origin, resulting in:
+\begin{equation}
+p_{\lambda,\gamma}(\theta) = \left\{
+    \begin{array}{lr}
+    \lambda\theta^2-{ \theta^2 \over 2\gamma} & \text{if } \theta \leq \gamma\lambda\\
+    {1 \over 2}\gamma\lambda^2 & \text{if } \theta > \gamma\lambda
+    \end{array}
+    \right. 
+\end{equation}
+It is more intuitive to think about the derivative of the penalty:
+\begin{equation}
+p'_{\lambda,\gamma}(\theta) = \left\{
+    \begin{array}{lr}
+    2\lambda\theta - {\theta \over \gamma} & \text{if } \theta \leq \gamma\lambda\\
+    0 & \text{if } \theta > \gamma\lambda
+    \end{array}
+    \right.
+\end{equation}
+
+The MCP-like penalty begins by applying the same rate of penalization as the L2, but continuously relaxes the penalization until the rate drops to 0 and the penalty saturates. 
+As in [Fan Li], we use local quadratic approximation and solve by iteratively computing ridge regression.
 
@@
 \subsection{Bacterial data}
+We
+
 is good

### 2015-04-06 paper1.tex
alt penalty

@@
 \begin{abstract}
-HELLOOOOOOOOOOOOOOOOOOOOOOOO
+sup sup sup
 \end{abstract}
@@
 Sometimes gene orthology is not a good predictor of functional similarity.
-To simultaneously learn when a gene's function has diverged from that of its ortholog, and improve model selection, we consider alternative penalized least squares approaches.
+To learn when a gene's function has diverged from that of its ortholog and to improve model selection, we consider alternative penalized least squares approaches.
 Our fusion penalty pulls beta values for analogous interactions towards each other; since we use an L2 penalty, interaction weights which differ from each other the most are excessively penalized.
-SCAD [Fan Li 2001] and MCP [Zhang] are two alternative penalties which diminish this bias.
 Fan and Li point out that the bias can be removed by choosing a constant penalty beyond some threshold.
-This is analogous to 'unfusing' certain interactions, and easing up on the demand that certain beta values be close in value when doing so strongly opposes minimizing the mean squared error component of the penalty function.
-SCAD and MCP both use an L1 penalty near the origin, setting small estimated coefficients to zero and choosing a sparse solution.
-Because we do not expect most fused beta values to be identical, but rather similar, we amend the alternative fusion penalty to be L2 near the origin, resulting in:
+This is analogous to 'unfusing' certain interactions, and easing up on the demand that certain beta values be close in value when doing so is in strong opposition to minimizing the mean squared error component of the penalty function.
+SCAD [Fan Li 2001] and MCP [Zhang] are two alternative penalties which have this feature.
+We make some modifications and test both.
+Both use an L1 penalty near the origin, setting small estimated coefficients to zero and choosing a sparse solution.
+Because we do not expect most fused beta values to be identical, but rather similar, we amend the alternative fusion penalty to be L2 near the origin, resulting in modified MCP:
 \begin{equation}
@@
 The MCP-like penalty begins by applying the same rate of penalization as the L2, but continuously relaxes the penalization until the rate drops to 0 and the penalty saturates.
+
+The rationale behind SCAD is similar: both penalties apply the same rate of penalization as the lasso and reduce that rate to 0 as the absolute value of $\theta$ grows.
+What differs is the transition between the lasso and the flat penalty.
+In SCAD, both the derivative of the penalty as well as where the transition occurs depend on the same parameter, $\lambda$.
+In our version, we decouple these, and as with MCP, we alter the penalty so that it is L2-like near the origin to get:
+
+\begin{equation}
+p_{\lambda,\gamma}(\theta) = \left\{
+    \begin{array}{lr}
+    \lambda\theta^2 & \text{if } \theta \leq {a \over 2}\\
+    {a\lambda\theta - 0.5(\theta^2 + \lambda^2) \over a - 1} & \text{if } \lambda < \theta \leq {a \over 2}\\
+    {\lambda^2(a^2-1) \over 2(a-1)} & \text{if } \theta > a
+    \end{array}
+    \right. 
+\end{equation}
+
+\begin{equation}
+p'_{\lambda,\gamma}(\theta) = \left\{
+    \begin{array}{lr}
+    2\lambda\theta & \text{if } \theta \leq \lambda\\
+    \text{max}(\lambda_S(a-\theta),0) & \text{if } \theta > {a \over 2}
+    \end{array}
+    \right.
+\end{equation}
+
 As in [Fan Li], we use local quadratic approximation and solve by iteratively computing ridge regression.

### 2015-04-06 paper1.tex
blegh

--- 2015-03-31 paper1.tex
@@
 Where the second sum is over pairs of interactions with fusion constraints.
-Note that each term in the quantity being minimized is an L2 norm of a linear combination of gene expressions, so the entire problem can be solved through linear regression with a suitable design matrix [DETAILS SHOULD GO SOMEWHERE ELSE, OR NOWHERE].
+Note that each term in the quantity being minimized is an L2 norm of a linear combination of gene expressions, so the entire problem can be solved through linear regression with a suitable design matrix[DETAILS SHOULD GO SOMEWHERE or nowhere?].
 Because pairs of genes can be solved independently as long as they are not linked by fusion constraints, or a chain of fusion constraints, the problem of cross-species GRN inference decomposes into a large number of manageable regression problems as long as the orthology mappings are sparse.
@@
 \end{document}
-
->>>>>>> 34faf0e6aea8fc5a3af9cb49bfcff0b2e3eaa7ed

### 2015-04-06 paper1.tex

--- 2015-04-03 paper1.tex
@@
 \subsection{Fused GRN}
-We expect related species to be governed by similar but not necessarily identical gene regulatory networks.
+Related species should be governed by similar but not necessarily identical gene regulatory networks.
 We approach the problem of cross-species GRN inference by introducing constraints into the above regression formulation to penalize differences between interaction weights that are expected to be similar based on gene orthology.
@@
 \end{equation}
+
 Where the second sum is over pairs of interactions with fusion constraints.
 Note that each term in the quantity being minimized is an L2 norm of a linear combination of gene expressions, so the entire problem can be solved through linear regression with a suitable design matrix [DETAILS SHOULD GO SOMEWHERE ELSE, OR NOWHERE].
-The $\beta$ vectors corresponding to a pair of genes can be solved independently as long as the pair is not linked by fusion constraints, or a chain of fusion constraints; thus the problem of cross-species GRN inference decomposes into a large number of manageable regression problems as long as the orthology mappings are sparse.
+The regulators of a pair of genes can be solved independently as long as the pair is not linked by fusion constraints, or a chain of fusion constraints; thus the problem of cross-species GRN inference decomposes into a large number of manageable regression problems as long as the orthology mappings are sparse.
 
@@
 \subsection{Alternative penalty functions}
-Sometimes gene orthology is not a good predictor of functional similarity.
-To simultaneously learn when a gene's function has diverged from that of its ortholog, and improve model selection, we consider alternative penalized least squares approaches.
-Our fusion penalty pulls beta values for analogous interactions towards each other; since we use an L2 penalty, interaction weights which differ from each other the most are excessively penalized.
-SCAD [Fan Li 2001] and MCP [Zhang] are two alternative penalties which diminish this bias.
-Fan and Li point out that the bias can be removed by choosing a constant penalty beyond some threshold.
-This is analogous to 'unfusing' certain interactions, and easing up on the demand that certain beta values be close in value when doing so strongly opposes minimizing the mean squared error component of the penalty function.
-SCAD and MCP both use an L1 penalty near the origin, setting small estimated coefficients to zero and choosing a sparse solution.
-Because we do not expect most fused beta values to be identical, but rather similar, we amend the alternative fusion penalty to be L2 near the origin, resulting in:
+Fusion constraints are intended to penalize dissimilarity between interactions thought to be evolutionarily analogous.
+Because fusion constraints are L2, interaction weights which differ from each other by a large amount are excessively penalized, which effectively ensures that fused interactions are assigned similar weights.
+This may be inappropriate for interactions which are identified based on orthology as being analogous, but which are no longer similar due to evolutionary changes.
+We propose that a saturating penalty may be useful for dealing with uncertainty about which interactions are conserved.
+
+With a saturating fusion penalty, fused interactions which appear to be very different based on expression data are allowed to unfuse from one another.
+A closely related problem has been studied in the context of LASSO regularization, where it was shown by Fan and Li that using a saturating penalty retains many of LASSO's desireable properties while removing its bias towards 0.
+They further showed that, although the resulting loss-function is nonconvex, good results can be obtained with a local quadratic approximation of gradient descent.
+Several saturating penalties, such as SCAD [Fan Li 2001] and MCP [Zhang], have been discussed in the context of sparse regression.
+We introduce a modified form of MCP to the problem of penalizing differences between fused coefficients.
+The principle difference between the penalty we adopt and SCAD/MCP is that both of these penalties are L1 like at the origin, producing sparse solutions.
+Because we are penalizing differences in interaction weights, rather than the weights themselves, there's no particular reason to prefer sparsity.
+
+We use a penalty on the difference between fused coefficients $\theta$ which is L2 like at the origin, begins to level out at $\theta = \frac{a}{2}$, and saturates at $\theta = a$.
+Written in terms of its derivative, the penalty $p'_{\lambda, a}$
+
+\begin{equation}
+p'_{\lambda,a}(\theta) = \left\{
+    \begin{array}{lr}
+    2\lambda\theta & \text{if } \theta \leq \lambda\\
+    \text{max}(\lambda_S(a-\theta),0) & \text{if } \theta > {a \over 2}
+    \end{array}
+    \right.
+\end{equation}
+
+We also implement MCP [IS THIS MODIFIED TO BE L2 like at the origin?]
+
 \begin{equation}
@@
 \end{equation}
-It is more intuitive to think about the derivative of the penalty:
+
+With derivative
+
 \begin{equation}
@@
 
-The MCP-like penalty begins by applying the same rate of penalization as the L2, but continuously relaxes the penalization until the rate drops to 0 and the penalty saturates. 
-As in [Fan Li], we use local quadratic approximation and solve by iteratively computing ridge regression.
+As in [Fan Li], we use solve using iterative local quadratic approximation. Specifically,
+
+\begin{equation}
+\begin{array}{l}
+\beta^S(t)\text{ is the network on iteration } t
+\\
+\text{for each fused  } B^{S_1}_{g,k} \approx B^{S_2}_{h,l}  \text{ we define } 
+\\
+\theta(0)=0
+\\
+\theta(t) = \vert B^{S_1}_{g,k} - B^{S_2}_{h,l} \vert
+\\
+\text{and introduce a fusion constraint } \lambda = \frac{p'(\theta(t))}{2\theta(t)} 
+\\
+\beta^S(t+1) \text{ is obtained by fitting the ridge-fused model with fusion constraints given by the above }\lambda s
+\end{array}
+\end{equation}
 
@@
 \subsection{Bacterial data}
-We
+TBA
 
-is good
-Look an equation array
-\begin{equation}
-\begin{array}{l}
-R_i^{\mathrm{input}}(\theta) = c g_i^{\mathrm{input}}\exp(-(\theta - \theta_i)^2 / (2\sigma^2_{\mathrm{input}}))
-\\
-R_i^{\mathrm{output}}(\theta) = g_i^{\mathrm{output}}\sum_{j=0}^N R_j^{\mathrm{input}}(\theta) \exp(-(\theta_i - \theta_j)^2/(2\sigma^2_{\mathrm{output}})) .
-\end{array}
-\end{equation}
 \bibliographystyle{plain}

### 2015-04-27 paper1.tex
hi

@@
 
-We also implement MCP [IS THIS MODIFIED TO BE L2 like at the origin?]
+We also implement MCP:
 
@@
 
-As in [Fan Li], we use solve using iterative local quadratic approximation. Specifically,
+As in [Fan Li], we use solve using iterative local quadratic approximation.
+
+Specifically, $\beta^S(t)$ is the network on iteration $t$.
 
@@
 \begin{array}{l}
-\beta^S(t)\text{ is the network on iteration } t
-\\
-\text{for each fused  } B^{S_1}_{g,k} \approx B^{S_2}_{h,l}  \text{ we define } 
+\text{for each fused  } B^{S_1}_{g,k} \approx B^{S_2}_{h,l}  \text{ we define: } 
 \\
@@
 \\
-\beta^S(t+1) \text{ is obtained by fitting the ridge-fused model with fusion constraints given by the above }\lambda s
 \end{array}
 \end{equation}
+
+$\beta^S(t+1)$ is obtained by fitting the ridge-fused model with fusion constraints given by the above $\lambda_S$.
+This is nice because all our penalties can be treated as L2 and therefore retain the properties of ridge regression.
 
@@
 
+\subsection{Optimization and model selection}
+
 \subsection{Bacterial data}
 TBA
+
+\section{Results}
+
+\subsection{Fusing similar network has the same effect as increasing the number of experimental conditions}
+When the sample size of gene expression increases, the performance
+
+\subsection{SCAD successfully identifies and unfuses 'neofunctionalized' genes}
 

### 2015-04-28 paper1.tex
wrote up some results

@@
 \section{Results}
+We tested the ability of fused regression to improve network inference and to learn conservation of gene function in both real and simulated data.
+We assessed performance using area under the precision recall curve on a leave-out set.
+We used cross-validation to define leave-out sets.
 
-\subsection{Fusing similar network has the same effect as increasing the number of experimental conditions}
-When the sample size of gene expression increases, the performance
+\subsection{Fusing similar networks has the same effect as increasing the number of experimental conditions}
+We assessed whether using a fusion penalty on similar networks improved our ability to learn the underlying networks.
+We used sets of two networks with 20 TFs and 200 genes each, where each set had varying proportions of orthologous TFs and genes.
+As we increased network similarity, we saw greater improvements in performance from fused regression, and more accurate network recovery with higher fusion penalty weight.
+We show that these improvements are comparable to adding more data points when learning networks individually; by taking advantage of what is known about similarity between genes and pathways, we are able to move closer towards understanding the gene regulatory network without collecting more data.
+
+\subsection{Fused regression improves performance on both the constrained and non-constrained parts of the network}
+We wanted to know if using fused regression on networks with only conserved subgraphs affected the recovery of the non constrained parts of the networks.
+We used sets of two networks with 20 TFs and 200 genes each, where each set had 0\%, 25\%, 50\%, or 75\% of the TFs and genes in orthology groups.
+We varied lamS and computed AUPR using cross-validation on the portion of the network that had fusion constraints, the portion without any orthology information, and on the whole network.
+On the fused networks, performance improved as the fusion penalty weight increased; interestingly, performance gains are seen even the portion of the network where no orthology information is known.
+By constraining part of an underdetermined system, we get gains in both parts.
 
 \subsection{SCAD successfully identifies and unfuses 'neofunctionalized' genes}
+We recognize that orthology prediction is not necessarily a perfect proxy for functional conservation.
+Therefore, it is important that our method be able to learn what to "fuse" and what to "unfuse." We generated synthetic fused networks and introduce error in the fusion constraints by adding false positives and negatives to the orthology information given to the solver.
+We then examined those interactions which are truly analogous (interactions between correctly fused gene/TF pairs) and interactions which are labeled analogous but represent edges between gene/TF pairs where one or both nodes in the pair is incorrectly fused.
+Using our method, these latter interactions are successfully identified and the force pulling their weights towards each other is reduced; the former interactions are left alone.
+
+\subsection{SCAD improves network inference}
+We wanted to determine whether SCAD is able to improve learning of GRNs.
+That is, beyond correctly identifying functional conservation and recovering the performance lost by fusing neofunctionalized genes, is our method able to produce more accurate networks than those learned without fusion?
 

### 2015-04-29 paper1.tex
hello

--- 2015-04-27 paper1.tex
@@
 \section{Results}
+We tested the ability of fused regression to improve network inference performance, as well as the ability of adaptive-fusion to learn the conservation of gene interactions.
+We measured network inference performance with the area under the precision recall curve (AUPR).
 
-\subsection{Fusing similar network has the same effect as increasing the number of experimental conditions}
-When the sample size of gene expression increases, the performance
+\subsection{Network similarity and fusion performance}
+We assessed how varying the similarity of simulated networks affected network inference performance with fusion.
+The main factors governing similarity between our generated networks were the extent (and accuracy) of the orthology mapping, and the magnitude of variability between analogous interactions.
+We conducted a series of simulations to assess the effect of varying orthology coverage on networks in which analogous interactions varied little from one another.
+In order to determine the effect of the size of the orthology mapping on network inference performance, we simulated a series of 20 TFs by 200 genes networks at several values of percent orthology coverage.
+
+As expected, increasing the weight of the fusion constraint $\lambda_S$ improved network recovery.
+Networks with higher orthology coverage saw a greater benefit from fusion, although the improvement was not limited to analogous interactions.
+One way of assessing the effect of fusion on network inference performance is to compute an ``exchange-rate'' between conditions in each network.
+Suppose you are interested in determining a regulatory network for species A, a model organism which is experimentally inconvenient.
+A related species, B, is easier to collect data for, but of less intrinsic interest.
+Our simulation results suggest that adding conditions from species B to a fused network inference problem has an effect on network A recovery that is similar to adding conditions from A at a reduced rate.
+
+We also assessed the effect of varying the standard deviation of the distribution of differences between analogous interactions on network performance. FILL IN. MENTION PRIOR INTERPRETATION
+
+\subsection{Fused regression improves performance on both the constrained and non-constrained parts of the network}
+We wanted to know if using fused regression on networks with only conserved subgraphs affected the recovery of the non constrained parts of the networks.
+We used sets of two networks with 20 TFs and 200 genes each, where each set had 0\%, 25\%, 50\%, or 75\% of the TFs and genes in orthology groups.
+We varied lamS and computed AUPR using cross-validation on the portion of the network that had fusion constraints, the portion without any orthology information, and on the whole network.
+On the fused networks, performance improved as the fusion penalty weight increased; interestingly, performance gains are seen even the portion of the network where no orthology information is known.
+By constraining part of an underdetermined system, we obtain gains in both parts.
 
 \subsection{SCAD successfully identifies and unfuses 'neofunctionalized' genes}
+We recognize that orthology prediction is not necessarily a perfect proxy for functional conservation.
+We implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions.
+Pairs of interactions which are very dissimilar even after fusion, which sit in the flat portion of this penalty function, are effectively ``unfused''.
+We can interpret this ``unfusing'' as evidence for neofunctionalization of one or more of the genes involved in this pair of analogous interactions.
+
+We performanced a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved when orthology information is faulty.
+We generated synthetic fused networks and introduce error in the fusion constraints by adding false positives and negatives to the orthology information given to the solver.
+Because we know which entries in the orthology mapping are ``fake'' (not reflected in the generation of the networks), we can label fusion constraints that involve one or more ``fake'' mappings as being ``fake''.
+We verified that adaptive-fusion unfused mostly ``fake'' interactions, while leaving truly analogous interactions fused.
+
+\subsection{SCAD improves network inference}
+We wanted to determine whether SCAD is able to improve learning of GRNs.
+That is, beyond correctly identifying functional conservation and recovering the performance lost by fusing neofunctionalized genes, is our method able to produce more accurate networks than those learned without fusion?
 

### 2015-04-29 paper1.tex
added figs

@@
 
+\subsection{Model selection}
+
 \subsection{Simulated data}
@@
 
-\subsection{Optimization and model selection}
+\subsection{Optimization}
 
@@
 
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{l2fusionquick.png}
+  \caption{\label{fig:figure1} The x-axis represents lamS and the y-axis represents the proportion of nodes which are orthologous.
+As the networks increase in similarity, performance improves and benefits from a large fusion penalty weight}
+  \end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{increase_data2.png}
+  \caption{\label{fig:figure1} We fuse two networks, using lamS=0 and nonzero lamS, and hold the number of conditions in network 1 constant while varying the number of data points for network 2.
+We plot the performance of network 2.}
+  \end{center}
+\end{figure}
+
 \subsection{Network similarity and fusion performance}
@@
 
-As expected, increasing the weight of the fusion constraint $\lambda_S$ improved network recovery.
+As expected, increasing the weight of the fusion constraint $\lambda_S$ improved network recovery (Figure 2).
 Networks with higher orthology coverage saw a greater benefit from fusion, although the improvement was not limited to analogous interactions.
@@
 A related species, B, is easier to collect data for, but of less intrinsic interest.
-Our simulation results suggest that adding conditions from species B to a fused network inference problem has an effect on network A recovery that is similar to adding conditions from A at a reduced rate.
+Our simulation results suggest that adding conditions from species B to a fused network inference problem has an effect on network A recovery that is similar to adding conditions from A at a reduced rate (Figure 3).
 
@@
 
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{con_noncon.png}
+  \caption{\label{fig:figure4} We looked at four different sets of networks, where each set had different amounts of orthologous genes.
+We plotted the performance on the whole network, the constrained, and non-constrained portions of the network as we increased the fusion penalty weight.}
+  \end{center}
+\end{figure}
+
 \subsection{Fused regression improves performance on both the constrained and non-constrained parts of the network}
-We wanted to know if using fused regression on networks with only conserved subgraphs affected the recovery of the non constrained parts of the networks.
+We wanted to know if using fused regression on networks with only conserved subgraphs affected the recovery of the non constrained parts of the networks (Figure 4).
 We used sets of two networks with 20 TFs and 200 genes each, where each set had 0\%, 25\%, 50\%, or 75\% of the TFs and genes in orthology groups.
@@
 By constraining part of an underdetermined system, we obtain gains in both parts.
+
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{plot_betas_scad_l2.png}
+  \caption{\label{fig:figure1} This is what a figure looks like}
+  \end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{plot_betas_scad.png}
+  \caption{\label{fig:figure1} This is what a figure looks like}
+  \end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{scadf.png}
+  \caption{\label{fig:figure1} This is what a figure looks like}
+  \end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{scadt.png}
+  \caption{\label{fig:figure1} This is what a figure looks like}
+  \end{center}
+\end{figure}
 
@@
 
-We performanced a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved when orthology information is faulty.
+We performanced a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved when orthology information is faulty (Figure 5-8).
 We generated synthetic fused networks and introduce error in the fusion constraints by adding false positives and negatives to the orthology information given to the solver.
@@
 
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{test_scad_opt_params3.png}
+  \caption{\label{fig:figure1} This is what a figure looks like}
+  \end{center}
+\end{figure}
+
 \subsection{SCAD improves network inference}
@@
 That is, beyond correctly identifying functional conservation and recovering the performance lost by fusing neofunctionalized genes, is our method able to produce more accurate networks than those learned without fusion?
+(Figure 9)
 

### 2015-04-30 paper1.tex
added figs

@@
 We varied lamS and computed AUPR using cross-validation on the portion of the network that had fusion constraints, the portion without any orthology information, and on the whole network.
-On the fused networks, performance improved as the fusion penalty weight increased; interestingly, performance gains are seen even the portion of the network where no orthology information is known.
-By constraining part of an underdetermined system, we obtain gains in both parts.
+On the fused networks, performance improved as the fusion penalty weight increased; interestingly, performance gains were seen even the portion of the network where no orthology information is known.
+By constraining part of an underdetermined system, we obtained gains in both parts.
 
@@
 Pairs of interactions which are very dissimilar even after fusion, which sit in the flat portion of this penalty function, are effectively ``unfused''.
-We can interpret this ``unfusing'' as evidence for neofunctionalization of one or more of the genes involved in this pair of analogous interactions.
+We can interpret this ``unfusing'' as evidence for neofunctionalization of one or more of the genes involved in this pair of interactions.
 
-We performanced a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved when orthology information is faulty (Figure 5-8).
-We generated synthetic fused networks and introduce error in the fusion constraints by adding false positives and negatives to the orthology information given to the solver.
+We performed a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved when orthology information is faulty, to mimic the effects of predicting orthologs which are not functionally similar (Figure 5-8).
+We generated synthetic fused networks and introduced error in the fusion constraints by adding false positives and negatives to the orthology information given to the solver.
 Because we know which entries in the orthology mapping are ``fake'' (not reflected in the generation of the networks), we can label fusion constraints that involve one or more ``fake'' mappings as being ``fake''.
@@
 That is, beyond correctly identifying functional conservation and recovering the performance lost by fusing neofunctionalized genes, is our method able to produce more accurate networks than those learned without fusion?
-(Figure 9)
+(Figure 9) We again generated synthetic fused networks and added false positive and negative orthologs, then used fused L2 and adaptive fusion approaches to solve the underlying network.
 

### 2015-04-30 paper1.tex
hi

@@
 
-\subsection{SCAD successfully identifies and unfuses 'neofunctionalized' genes}
+\subsection{Adaptive fusion successfully identifies and unfuses 'neofunctionalized' genes}
 We recognize that orthology prediction is not necessarily a perfect proxy for functional conservation.
@@
 
-We performed a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved when orthology information is faulty, to mimic the effects of predicting orthologs which are not functionally similar (Figure 5-8).
+We performed a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved when orthology information is faulty, to mimic orthologs which are not functionally analogous  (Figure 5-8).
 We generated synthetic fused networks and introduced error in the fusion constraints by adding false positives and negatives to the orthology information given to the solver.
-Because we know which entries in the orthology mapping are ``fake'' (not reflected in the generation of the networks), we can label fusion constraints that involve one or more ``fake'' mappings as being ``fake''.
+Because we knew which entries in the orthology mapping were ``fake'' (not reflected in the generation of the networks), we could correctly label fusion constraints that involved one or more ``fake'' mappings.
 We verified that adaptive-fusion unfused mostly ``fake'' interactions, while leaving truly analogous interactions fused.
@@
 
-\subsection{SCAD improves network inference}
-We wanted to determine whether SCAD is able to improve learning of GRNs.
+\subsection{Adaptive fusion improves network inference}
+We wanted to determine whether adaptive fusion is able to improve learning of GRNs.
 That is, beyond correctly identifying functional conservation and recovering the performance lost by fusing neofunctionalized genes, is our method able to produce more accurate networks than those learned without fusion?
 (Figure 9) We again generated synthetic fused networks and added false positive and negative orthologs, then used fused L2 and adaptive fusion approaches to solve the underlying network.
+Fused L2 incorporated flawed orthology information, resulting in degradation of performance as we increased orthology error.
+Performance increased using adaptive fusion, pointing to its utility in network inference in addition to identifying neofunctionalization.
 

### 2015-04-30 paper1.tex
hi
more figures

--- 2015-04-30 paper1.tex
@@
 
+\subsection{Beta scaling}
+[INSERT JUSTIFICATION]
+\\
+In previous work, betas were rescaled as to form a matrix of confidence scores $S$ as follows
+\begin{equation}
+S_{i,j} = \frac{\sigma^2_{\text{full model for }y_j}}{\sigma^2_{\text{full model for }y_j \text{ without predictor }i}}
+\end{equation}
+Computing residuals with respect to the data alone would disregard information gained through fusion. Instead, we used an approximation
+\begin{equation}
+S_{i,j} = \frac{\sigma^2_{\text{full model for }y_j}}{\sigma^2_{\text{full model for }y_j} + B_{i,j}^2 \times var(TF_j)}
+\end{equation}
+When the residual is uncorrelated with the prediction, then removing regressor $i$ with weight $B_{i,j}^2$ from the model of $j$ will increase the residual by $B_{i,j}^2$ times the variance of regressor $i$.
+In the case of regularized regression, the regressor and residual need not be uncorrelated, so the approximation will not hold exactly.
+This scaling is similar to rescaling according to variance explained relative to an augmented design matrix that includes fusion constraints.
+We use an approximation because it is simpler to compute.
+The intuition is that a large $B_{i,j}$ may be necessary because TF $i$ varies little across the available data, and these large weights are not suggestive of a true regulatory interaction.
+
 \subsection{Optimization}
@@
 
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{waka7unnormed.png}
+  \caption{\label{fig:figure1} precision recall curves of constrained interactions for several values of lamS, computed across folds containing $\frac{1}{15}th$ of the available B. Subtilis data}
+  \end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{woko7.png}
+  \caption{\label{fig:figure1} normalized precision recall curves of constrained interactions for several values of lamS, computed across folds containing $\frac{1}{15}th$ of the available B.
+Subtilis data.
+Normalization involved dividing the precision each each recall level, for each curve, and for each cross-validation fold by the corresponding $\lambda_S=0$ precision.}
+  \end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{wat7.png}
+  \caption{\label{fig:figure1} As previous figure, but for all interactions.}
+  \end{center}
+\end{figure}
+
 \bibliographystyle{plain}

### 2015-04-30 paper1.tex

@@
 \section{Introduction}
-
-\begin{figure}
-\begin{center}
-  \includegraphics[scale=0.45]{simulated.png}
-  \caption{\label{fig:figure1} This is what a figure looks like}
-  \end{center}
-\end{figure}
 

### 2015-05-04 paper1.tex
hi
abstract

@@
 \title{GRN inference in multiple species}
-\author{Kari Y. Lam, Zachary M. Westrick, Frank Underwood}
+\author{Kari Y. Lam, Zachary M. Westrick, Christian Mueller, Lionel Christiaen, Rich Bonneau}
 \maketitle
@@
 \begin{abstract}
-sup sup sup
+Methods for learning gene regulatory networks are essential to our understanding of cell behavior and response to external stimuli.
+We are interested in improving the accuracy of our networks in order to do meaningful comparisons between networks.
+We present two approaches using fused regression, where we exploit available evolutionary information to constrain an undetermined system, and show that both result in more accurate networks.
+We show that fused L2 is most useful in the case where the networks are very similar (eg different cell types or closely related species), but where data is limited.
+We can leverage our knowledge about the relationship between genes to promote our understanding of the respective networks.
+When data is more readily available, adaptive fusion is useful for identifying conservation between networks.
+
 \end{abstract}
@@
 Where the second sum is over pairs of interactions with fusion constraints.
-Note that each term in the quantity being minimized is an L2 norm of a linear combination of gene expressions, so the entire problem can be solved through linear regression with a suitable design matrix [DETAILS SHOULD GO SOMEWHERE ELSE, OR NOWHERE].
+Note that each term in the quantity being minimized is an L2 norm of a linear combination of gene expressions, so the entire problem can be solved through linear regression with a suitable design matrix.
 The regulators of a pair of genes can be solved independently as long as the pair is not linked by fusion constraints, or a chain of fusion constraints; thus the problem of cross-species GRN inference decomposes into a large number of manageable regression problems as long as the orthology mappings are sparse.
@@
 
+For each fused $B^{S_1}_{g,k} \approx B^{S_2}_{h,l}$ we define:
+
 \begin{equation} 
-\begin{array}{l}
-\text{for each fused  } B^{S_1}_{g,k} \approx B^{S_2}_{h,l}  \text{ we define: } 
-\\
 \theta(0)=0
-\\
+\end{equation}
+\begin{equation}
 \theta(t) = \vert B^{S_1}_{g,k} - B^{S_2}_{h,l} \vert
-\\
-\text{and introduce a fusion constraint } \lambda = \frac{p'(\theta(t))}{2\theta(t)} 
-\\
-\end{array}
 \end{equation}
+
+and introduce a fusion constraint $\lambda = \frac{p'(\theta(t))}{2\theta(t)} $
 
@@
 \subsection{Beta scaling}
-[INSERT JUSTIFICATION]
 \\

### 2015-05-04 paper1.tex
started writing real data results

@@
 Given time-series data on the expression of gene $i$ at timepoints $t_k$ and $t_{k+1}$, we can approximate the rate of change of $x_i$ as $x_i'(t_k)=\frac{x_i(t_{k+1})-x_i(t_k)}{t_{k+1}-t_k}$.
+We can treat steady-state data as having a derivative of zero.
 This gives us, for each gene $i$ and time $t_{k}$ an equation
@@
 \subsection{Beta scaling}
-\\
 In previous work, betas were rescaled as to form a matrix of confidence scores $S$ as follows
@@
 \subsection{Optimization}
-
-\subsection{Bacterial data}
-TBA
+We use cross-validation to select optimal regularization parameters $\lambda_R$ and $\lambda_S$.
+In order to determine the relevant scale of the fusion weight $\lambda_S$, we can compute the correlation between pairs of coefficients linked by fusion constraints.
+The maximal value of $\lambda_S$ should be just large enough so that this correlation is near $1$.
+More details on this later.
 
@@
 
+\section{Bacterial data}
+We used gene-expression data from B.
+Subtilis and B.
+Anthracis in order to assess performance gains of fused regression on real data.
+Our B Subtilis data set consists of 266 time-series and steady-state observations of 4891 genes during development.
+Our B.
+Anthracis dataset consists of 72 time-series and steady-state observations of 5536 genes comprising data from developmental and iron-starvation conditions.
+There were 247 known Transcription Factors (TFs) in the B.
+Subtilis dataset, and 248 TFs in the B.
+Anthracis dataset.
+
+We obtained 1870 one-to-one orthologs from Inparanoid, which produced 177650 fusion-constraints between gene interactions within the two species.
+Although this number seems very large, it represents only $14.7\%$ of the regulatory interaction matrix in B.
+Subtilis and $12.9\%$ in B.
+Anthracis.
+The use of such dissimilar species is a test of whether fusion can improve network-inference even when the overlap in interactions is very small.
+
+In order to evaluate network inference performance, and for use as priors, we used a gold standard of 2896 known B.
+Subtilis interactions with corresponding activation and repression sign.
+Of these 2896 priors, 968 had corresponding interactions in B.
+Anthracis.
+Based on our simulation results, we can expect the greatest gains in network-inference performance from fusion when the species of interest has a small number of available conditions, but data is abundant in a related species.
+However, in order to evaluate performance objectively a gold-standard of known interactions is necessary.
+As a result, we can only evaluate network recovery for B.
+Subtilis, and B.
+Subtilis also has the majority of our conditions.
+In order to simulate the data-poor regime, we subsampled our B.
+Subtilis data.
+Specifically, we divided our B.
+Subtilis data into $k$ folds, and then for each fold fit a network to the B.
+Subtilis data from that fold alone fused to the entire 72 B.
+Anthracis conditions.
+We can then compute performance metrics (ie AUPR) for each fold, and use their variability across folds to test for significance of any improvements relating to the use of fusion constraints.
+
+\subsection{Bacterial performance}
+\subsubsection{Performance with priors}
+We want to show that performance increases with introduction of priors - this is easy/circular but probably important.
+
+\subsubsection{Precision recall}
+Introducing fusion constraints leads to a small but consistent improvement in AUPR.
+
+\begin{center}
+  \begin{tabular}{ l | l | l }
+    \hline
+    $\lambda_S$ & $k$ & $AUPR$ \\ \hline
+    0 & 15 & 0.04 \\
+    0 & 15 & 0.06 \\
+    \hline
+  \end{tabular}
+\end{center}
+
 \begin{figure}

### 2015-05-04 paper1.tex
hi

@@
   \includegraphics[scale=0.45]{l2fusionquick.png}
-  \caption{\label{fig:figure1} The x-axis represents lamS and the y-axis represents the proportion of nodes which are orthologous.
+  \caption{\label{fig:figure1} We generate a series of networks with 20 TFs by 200 genes, with 75\% sparsity, 0.1 measurement noise, and 0.1 fusion noise.
+We hold lamP and lamR constant at 1.0 and 2.0 respectively.
+The x-axis represents lamS and the y-axis represents the proportion of nodes which are orthologous.
 As the networks increase in similarity, performance improves and benefits from a large fusion penalty weight}
@@
   \includegraphics[scale=0.45]{increase_data2.png}
-  \caption{\label{fig:figure1} We fuse two networks, using lamS=0 and nonzero lamS, and hold the number of conditions in network 1 constant while varying the number of data points for network 2.
+  \caption{\label{fig:figure1} We fuse two networks, both with 25 TFs by 50 genes, using lamS=0 and nonzero lamS, and hold the number of conditions in network 1 constant while varying the number of data points for network 2.
 We plot the performance of network 2.}
@@
   \includegraphics[scale=0.45]{con_noncon.png}
-  \caption{\label{fig:figure4} We looked at four different sets of networks, where each set had different amounts of orthologous genes.
+  \caption{\label{fig:figure4} We looked at four different sets of networks, each with 20 TFs by 200 genes, where each set had different amounts of orthologous genes.
 We plotted the performance on the whole network, the constrained, and non-constrained portions of the network as we increased the fusion penalty weight.}
@@
   \includegraphics[scale=0.45]{plot_betas_scad_l2.png}
-  \caption{\label{fig:figure1} This is what a figure looks like}
+  \caption{\label{fig:figure1} We used two networks, each with 20 TFs by 200 genes, 50\% of which are orthologous.
+We include as many false positive orthologs as true positives, and color the points solved using true orthology information green and the points solved using false orthology information red.
+We randomly select an interaction from each network, and color this point blue.
+We solve the networks using fused L2.}
   \end{center}
@@
   \includegraphics[scale=0.45]{plot_betas_scad.png}
-  \caption{\label{fig:figure1} This is what a figure looks like}
+  \caption{\label{fig:figure1} We used two networks, each with 20 TFs by 200 genes, 50\% of which are orthologous.
+We include as many false positive orthologs as true positives, and color the points solved using true orthology information green and the points solved using false orthology information red.
+We randomly select an interaction from each network, and color this point blue.
+We solve the networks using adaptive fusion.}
   \end{center}
@@
   \includegraphics[scale=0.45]{scadf.png}
-  \caption{\label{fig:figure1} This is what a figure looks like}
+  \caption{\label{fig:figure1} Here we use two 10 TF by 200 gene networks, 50\% of which are orthologous.
+We include as many false positive orthologs as we have true positives, and apply adaptive fusion to the beta values with lamS=3.
+We plot the distribution of lamS learned for false orthologs.}
   \end{center}
@@
   \includegraphics[scale=0.45]{scadt.png}
-  \caption{\label{fig:figure1} This is what a figure looks like}
+  \caption{\label{fig:figure1} Here we use two 10 TF by 200 gene networks, 50\% of which are orthologous.
+We include as many false positive orthologs as we have true positives, and apply adaptive fusion to the beta values with lamS=3.
+We plot the distribution of lamS learned for true orthologs.}
   \end{center}
@@
   \includegraphics[scale=0.45]{test_scad_opt_params3.png}
-  \caption{\label{fig:figure1} This is what a figure looks like}
+  \caption{\label{fig:figure1} We use three sets of 20 TF by 200 gene networks, 50\% of which are orthologs, and solve with orthology information containing 0:1, 0.5:1, and 1:1 false positive to true positive.
+We plot AUPR under fused L2 and adaptive fusion.}
   \end{center}

### 2015-05-13 paper1.tex
intro

@@
 
-\title{GRN inference in multiple species}
-\author{Kari Y. Lam, Zachary M. Westrick, Christian Mueller, Lionel Christiaen, Rich Bonneau}
+\title{Fused regression for network inference in multiple species}
+\author{Kari Y. Lam, Zachary M. Westrick, Christian Mueller, Lionel Christiaen, Richard Bonneau}
 \maketitle
@@
 \begin{abstract}
-Methods for learning gene regulatory networks are essential to our understanding of cell behavior and response to external stimuli.
-We are interested in improving the accuracy of our networks in order to do meaningful comparisons between networks.
-We present two approaches using fused regression, where we exploit available evolutionary information to constrain an undetermined system, and show that both result in more accurate networks.
-We show that fused L2 is most useful in the case where the networks are very similar (eg different cell types or closely related species), but where data is limited.
-We can leverage our knowledge about the relationship between genes to promote our understanding of the respective networks.
-When data is more readily available, adaptive fusion is useful for identifying conservation between networks.
-
+Learning gene regulatory networks is essential to our understanding of cell behavior and response to external stimuli.
+Methods for network inference have been developed and used for a wide variety of species; however, these approaches consider the problem as independent in each species even though there is known to be conservation of gene regulation even in distantly related species.
+We introduce a method that allows data to be pooled across species in a principled way based on gene orthology and demonstrate that it improves network inference even when the orthology is incomplete.
+We further refine this method with an algorithm that attempts to learn the true conserved subnetwork from a larger set of potentially conserved interactions.
 \end{abstract}
@@
 \section{Introduction}
+As the volume and variety of genome scale data have exploded, the goal of accurately modeling gene regulatory networks has become more attainable.
+Already, there is enough data in some model organisms for the development of high quality networks; however, this is still not true of many other model systems.
+Luckily, the problem of inferring gene regulatory networks is not an independent problem.
+There has been a great deal of work showing that functional conservation exists in gene regulatory networks even across large evolutionary distance.
+A high quality network in one species should be a useful source when learning the network of a related species.
+But most existing inference algorithms do not incorporate this information.
+
+We employ a linear model of gene expression in which the rate of expression is assumed to be proportional to a weighted sum of transcription factor expression.
+These weights represent the regulatory interactions we wish to learn.
+We can fit these weights simultaneously in multiple species of interest using time series and steady state expression data.
+This is similar to many existing network inference methods and produces identical answers to fitting each species separately.
+We introduce additional constraints on the problem that penalize differences between regulatory weights which are expected to be similar.
+The choice of which interactions to fuse together follows from the observation that orthologous genes are likely to have similar regulatory interactions.
+
+A few approaches to multi-source network inference have been proposed; however, current approaches either require one-to-one orthology [aviv regev paper], which is useful for analyzing cell lines but not for most cross species comparisons, or restrict analysis to the subset of genes with orthologs in every network [ibm paper].
+Our approach allows simultaneous inference of the whole gene regulatory networks of species that may have partial orthology overlap.
+We demonstrate performance gains in networks learned from simulated data and real data from B.
+subtilis and B.
+anthracis, where a large fraction of genes have no orthologs.
+Though this method was developed for multi-species network inference, it is also useful in the easier problem of learning networks in multiple cell lines.
+
+The largest gains in performance are obtained when there is very little data in a species of interest but a large amount of data in a related species.
+But the information flow is two ways and our simulation results suggest that network inference improves in both species.
+This feature of simultaneously fitting networks is particularly useful when there is an adequate amount of data available but spread over too many species.
+
+There exist several tools for identifying orthologous genes using sequence similarity.
+Although sequence similarity is a useful predictor of functional similarity, we recognize that some genes have taken on different functions and therefore may have new regulatory interactions.
+When this occurs, the benefits in overall performance from our method come at the cost of impaired prediction of regulators for these genes.
+The constraints we impose favor weight configurations in which fused interaction weights are similar to one another.
+Our method minimizes a cost function that strives  to simultaneously fit expression data in each species and produce networks that are consistent with these evolutionary constraints.
+When a gene's regulators have changed through evolution, there may be no way of achieving both these goals.
+In this case, the network inference procedure would be improved by relaxing or ignoring these evolutionary constraints.
+However, we don't know a priori which constraints to enforce and which to relax.
+
+We approach this problem by introducing a saturating penalty function based on work in statistics to develop unbiased regularization penalties for regression.
+The resulting cost function is non-convex and difficult to optimize and fares less well in the underdetermined case; however, we can approximate its solution and obtain deeper insights into functional similarity than is available through orthology or the comparison of separately learned networks.
+When the quantity of data is sufficient to produce high quality networks, our adaptive fusion approach improves performance of network inference; this is achieved in part through the identification of presumptive orthologs with different functions.
+In simulated data, when we introduced faulty orthology information, our method not only correctly identified which evolutionary restraints to enforce and which to relax, but also outperformed the control case where no incorrect orthologs were used.
+
+We present two methods to multi-species network inference using fused regression.
+One approach, our fused L2 method, uses an L2 fusion penalty on differences in regulatory weights with evolutionary constraints.
+Additionally, we developed a method, which we call adaptive fusion, using a non-convex saturating fusion penalty for both network inference and network comparison.
+We test the ability of fused L2 and adaptive fusion to improve network recovery on both synthetic data and Bacillus subtilis and Bacillus anthracis.
+We explore the circumstances under which each approach is optimal, and evaluate the robustness of adaptive fusion to incorrect orthology, simulating the biologically relevant case of neofunctionalization.
 

### 2015-05-14 paper1.tex
hi

@@
 \title{Fused regression for network inference in multiple species}
-\author{Kari Y. Lam, Zachary M. Westrick, Christian Mueller, Lionel Christiaen, Richard Bonneau}
+\author{Kari Y. Lam, Zachary M. Westrick, Christian L. M\"{u}ller, Lionel Christiaen, Richard Bonneau}
 \maketitle
@@
 Already, there is enough data in some model organisms for the development of high quality networks; however, this is still not true of many other model systems.
-Luckily, the problem of inferring gene regulatory networks is not an independent problem.
+Luckily, the problems of inferring gene regulatory networks in multiple species are not independent, due to the presence of conserved structure.
 There has been a great deal of work showing that functional conservation exists in gene regulatory networks even across large evolutionary distance.
-A high quality network in one species should be a useful source when learning the network of a related species.
+Abundant data in one species should be a useful source when learning the network of a related species.
 But most existing inference algorithms do not incorporate this information.
@@
 This is similar to many existing network inference methods and produces identical answers to fitting each species separately.
-We introduce additional constraints on the problem that penalize differences between regulatory weights which are expected to be similar.
+To take advantage of cross-species information, we introduce additional constraints on the problem that penalize differences between regulatory weights which are expected to be similar.
 The choice of which interactions to fuse together follows from the observation that orthologous genes are likely to have similar regulatory interactions.
 
-A few approaches to multi-source network inference have been proposed; however, current approaches either require one-to-one orthology [aviv regev paper], which is useful for analyzing cell lines but not for most cross species comparisons, or restrict analysis to the subset of genes with orthologs in every network [ibm paper].
+A few approaches to multi-source network inference have been proposed; however, current approaches either require that the sources have identical sets of genes [aviv regev paper], which is useful for analyzing cell lines but not for most cross species comparisons, or restrict analysis to the subset of genes with orthologs in every species [ibm paper], which limits the applicability of the model.
 Our approach allows simultaneous inference of the whole gene regulatory networks of species that may have partial orthology overlap.
-We demonstrate performance gains in networks learned from simulated data and real data from B.
-subtilis and B.
-anthracis, where a large fraction of genes have no orthologs.
+We demonstrate performance gains in networks learned from simulated data and real data from Bacillus subtilis and Bacillus anthracis, where a large fraction of genes have no orthologs.
 Though this method was developed for multi-species network inference, it is also useful in the easier problem of learning networks in multiple cell lines.
@@
 \section{Methods}
-\subsection{GRN}
+\subsection{Gene regulatory network}
+We represent gene regulatory networks as a matrix of weights, where rows index transcription factors and columns index genes.
+The weight in a given position represents the regulatory network on that TF on that gene; positive weights represent activation, negative weights represent repression, and 0 weights represent the absence of an interaction.
 We begin with an existing framework for estimating the interaction weights in a gene regulatory network (GRN).
 In the Inferelator algorithm, linear differential equations are used to model changes in gene expression.
-The rate at which the expression of gene $i$, $x_i$, changes is governed by degradation of existing transcripts with rate $\alpha$ plus a linear combination of transcription factor (TF) expressions.
+The rate at which the observed mRNA expression of gene $i$, $x_i$, changes is governed by degradation of existing transcripts with rate $\alpha$ plus a linear combination of transcription factor (TF) expressions.
 \begin{equation}
@@
 
-\subsection{Fused GRN}
-Related species should be governed by similar but not necessarily identical gene regulatory networks.
+\subsection{Fused gene regulatory networks}
+We hypothesize that related species should be governed by similar but not necessarily identical gene regulatory networks.
 We approach the problem of cross-species GRN inference by introducing constraints into the above regression formulation to penalize differences between interaction weights that are expected to be similar based on gene orthology.
@@
 In terms of the above regression formulation, we expect that $\beta^A_{g,k} \approx \beta^B_{h,l}$, and include a penalty term $\lambda_S(\beta^A_{g,k} - \beta^B_{h,l})^2$ to the quantity being minimized in order to ensure similarity.
-The parameter $\lambda_S$ controls how highly to penalize differences between fused coefficients.
+The parameter $\lambda_S$ controls how highly to penalize differences between fused coefficients and controls the tradeoff between fitting the expression data and producing a set of networks that conform to evolutionary constraints.
 This gives us the final equation to be minimized
@@
 
-[DIGRESSION?]
-
 When there is a one-to-one orthology between the species being considered (ie different cell-lines of the same organism), the choice of $\lambda_S$ allows one to interpolate between fitting each network independently ($\lambda_S=0$) and pooling data together as if it came from one source ($\lambda_S=\inf$).
@@
 This is particularly useful when dealing with a large number of species; pairwise orthologies may be nearly complete even when the number of genes present in every organism is small.
+
+\subsection{Solving}
+The problem of solving a multi-valued regression problem can be converted into the problem of solving a single valued regression problem by vectorization.
+Specifically, if $Y \in R^{N \times M}$ then we form $A =$ diagonally concatenate $X$, $M$ times (with the remaining entries 0),
+$y =$ vertically concatenate the columns of $Y$
+
+$$\underset{w}{min} ||Aw - y||_2^{~2}$$
+
+We can then define $W_{i,j} = \beta_{j \times N + i}$ which is the desired form.
+This is equivalent to the original problem due to the block structure of matrix multiplication
+
+Given $X_1... X_S$ matrices, $X_i \in R^{N_i \times K_i}$ we define $N = \sum_{i=0}^S N_i$, $K = \sum_{i=0}^SK_i$ and $X \in R^{N \times K}$ by diagonally concatenating $X_1...X_S$.
+In this new matrix, the first $K_1$ features correspond to features of $X_1$, the next $K_2$ features the features of $X_2$ etc.
+Similarly, the first $N_1$ samples correspond to samples of $X_1$, the next $N_2$ to $X_2$ etc.
+Solving the regression problem associated with each source is equivalent to solving the problem $\underset{W}{argmin}||XW-Y||$ where $Y$ is defined in the same way as $X$.
+The weight matrix $W$ has a block diagonal structure:
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[scale=.5,trim=0mm 0mm 0mm 0mm,clip]{direct_solve.eps}
+\end{figure}
+
+The fused multi-source regression problem can solved by constructing a multi-source design matrix $X$ and data matrix $Y$, and then vectorizing.
+The regression problems associated with two columns $g,h$ of $Y$, where $g$ came from $Y_G$ and $h$ came from $Y_H$, are independent, unless there is a constraint chain of length $L$ with the form (where $x$ is arbitrary)
+
+%$$(G, j_0, g, J_1, j_1, m_1) \rightarrow (J_1, j_1, m_1, J_2, j_2, m_2) \rightarrow (J_2,j_2,m_2,J_3,j_3,_m3)...(J_P, j_P, m_P, H, j_{P+1}, h)$$
+$$(G, j_0, g, J_1, j_1, m_1) \rightarrow (J_1, x, m_1, J_2, j_2, m_2)$$
+
+We use depth-first search to identify linked columns of $Y$, and then solve the fused regression problem directly by vectorizing those columns.
+In order to incorporate the fusion penalty, we append to the resulting design matrix one row \emph{per connected constraint}, which contains $\lambda_S$ in the column associated with $G,i,j,$ and $-\lambda_S$ to the column associated with $J, i',j'$ for constraint $(G,i,j,J,i',j')$.
+Call this resulting matrix $X_P$.
+The vector $Y$ is augmented by appending an appropriate number of zeros to produce $Y_P$.
+As a result, each entry in the output of $X_P b$ in the columns associated with each constraint contain $b_{i,j} - b_{i',j,i}$.
+These entries each contribute  $||b_{i,j} - b_{i',\{ji\}}||^{~2}$ to the squared error $||X_Pb - Y_P||_2^{~2}$
+
+In the context of regression for GRN inference with multiple cell sources, each gene $j$ from source $1$ has constraints $\forall_{i \in \{TFs\}} (1, i, j, 2, i’, j’)$ where $i',j'$ are the same gene in a different cell line.
+As a result, each gene has only one constraint and $L=2$.
+%In general, however, solving the fused regression problem directly, even after factoring constraints, involves at least one inversion of an $L\sum_{i=0}^S K_i \times L \sum_{i=0}^S K_i$ matrix, taking $O((L\sum_{i=0}^S K_i)^3)$.
+In general, however, solving the fused regression problem directly, even after factoring constraints, involves at least one inversion of an $p\sum_{i=0}^S K_i \times p \sum_{i=0}^S K_i$ matrix where $p$ is the size of a column subset such that there is a constraint chain $L$ linking an element of each column to an element in every other column in that subset, taking $O((p\sum_{i=0}^S K_i)^3)$.
+Although the sparse structure of $X_P$ makes this somewhat more tractable, it is still impractical for large $L$.
+
 \subsubsection{Equivalent prior}
-I want to talk about the fusion constraints as a Gaussian prior, and talk about the adjustment stuff
+Our L2 fusion penalty is equivalent to assuming a Gaussian prior with variance proportional to $\frac{1}{\lambda_S}$ on differences in parameters with fusion constraints.
+Combined with the regularization constraints, this forms a multivariate Gaussian prior on weights in both networks.
+When using our method on systems with incomplete orthology mapping, this results in different prior probability distributions for those parameters with fusion constraints and those without.
+We adjust for this by solving for a constant to multiply fused interactions by to equalize the volumes of the prior distributions.
+The intent of this adjustment is to ensure that constrained interactions are not on average more highly penalized, which may tend to drive their weight towards zero, causing them to be excluded from the network.
 

### 2015-05-14 paper1.tex
added a bit more on performance

@@
 
-\subsection{Model selection}
-
 \subsection{Simulated data}
@@
 The x-axis represents lamS and the y-axis represents the proportion of nodes which are orthologous.
-As the networks increase in similarity, performance improves and benefits from a large fusion penalty weight}
+As the networks increase in similarity, performance improves and benefits from a large fusion penalty weight.
+NOTE: This figure shows AUPR.
+$R^2$ behaves as expected with respect to $\lambda_S$ (networks with large differences between fused coefficients have best $R^2$ at intermediate $\lambda_S$).
+This doesn't really seem to be the case with AUPR.
+Maybe these early figures should be replaced with $R^2$}
   \end{center}
@@
 Introducing fusion constraints leads to a small but consistent improvement in AUPR.
+Inspection of precision recall curves shows that this improvement is a result of increased precision at low values of recall.
+Essentially, precision increases for the high confidence part of the network but not for the low confidence part of the network.
+As a result, the improvements in performance may be more useful than would be suggested by AUPR alone.
+Interactions suggested by genome wide analysis need to be experimentally verified, and the ranking of interactions produced by a network inference algorithms is a useful guide to what order these experiments should be carried out in.
+The high confidence part of the network represents interactions that are likely candidates for experimental validation, so improvement in recovery of these interactions is particularly useful in guiding research.
 
@@
     \hline
-    $\lambda_S$ & $k$ & $AUPR$ \\ \hline
-    0 & 15 & 0.04 \\
-    0 & 15 & 0.06 \\
+    $\lambda_S$ & $\%$ data & $AUPR constrained$ \\ \hline
+    0 & 0.066 & 0.0390 \\
+    5 & 0.066 & 0.0415 \\
     \hline

### 2015-05-14 paper1.tex
hi

--- 2015-05-14 paper1.tex
@@
 \subsection{Model selection}
+We can define a leave-out set of priors and choose parameters based on cross-validated AUPR on the leave-out set.
 
@@
 We used gene-expression data from B.
-Subtilis and B.
-Anthracis in order to assess performance gains of fused regression on real data.
-Our B Subtilis data set consists of 266 time-series and steady-state observations of 4891 genes during development.
+subtilis and B.
+anthracis in order to assess performance gains of fused regression on real data.
+Our B subtilis data set consists of 266 time-series and steady-state observations of 4891 genes during development.
 Our B.
-Anthracis dataset consists of 72 time-series and steady-state observations of 5536 genes comprising data from developmental and iron-starvation conditions.
+anthracis dataset consists of 72 time-series and steady-state observations of 5536 genes comprising data from developmental and iron-starvation conditions.
 There were 247 known Transcription Factors (TFs) in the B.
-Subtilis dataset, and 248 TFs in the B.
-Anthracis dataset.
+subtilis dataset, and 248 TFs in the B.
+anthracis dataset.
 
@@
 Although this number seems very large, it represents only $14.7\%$ of the regulatory interaction matrix in B.
-Subtilis and $12.9\%$ in B.
-Anthracis.
+subtilis and $12.9\%$ in B.
+anthracis.
 The use of such dissimilar species is a test of whether fusion can improve network-inference even when the overlap in interactions is very small.
@@
 In order to evaluate network inference performance, and for use as priors, we used a gold standard of 2896 known B.
-Subtilis interactions with corresponding activation and repression sign.
+subtilis interactions with corresponding activation and repression sign.
 Of these 2896 priors, 968 had corresponding interactions in B.
-Anthracis.
+anthracis.
 Based on our simulation results, we can expect the greatest gains in network-inference performance from fusion when the species of interest has a small number of available conditions, but data is abundant in a related species.
@@
 As a result, we can only evaluate network recovery for B.
-Subtilis, and B.
-Subtilis also has the majority of our conditions.
+subtilis, and B.
+subtilis also has the majority of our conditions.
 In order to simulate the data-poor regime, we subsampled our B.
-Subtilis data.
+subtilis data.
 Specifically, we divided our B.
-Subtilis data into $k$ folds, and then for each fold fit a network to the B.
-Subtilis data from that fold alone fused to the entire 72 B.
-Anthracis conditions.
+subtilis data into $k$ folds, and then for each fold fit a network to the B.
+subtilis data from that fold alone fused to the entire 72 B.
+anthracis conditions.
 We can then compute performance metrics (ie AUPR) for each fold, and use their variability across folds to test for significance of any improvements relating to the use of fusion constraints.
@@
 \subsubsection{Performance with priors}
-We want to show that performance increases with introduction of priors - this is easy/circular but probably important.
+We want to show that performance increases with introduction of priors - this is easy/circular but important to show that our method can effectively incorporate priors
+
+\subsubsection{Adaptive fusion}
+We want to show that we can successfully identify neofunctionalization; since we do not have a gold standard for this we will introduce false positive orthologs into the orthology information and evaluate the ability of our method to identify true from false orthologs.
 

### 2015-05-21 paper1.tex

@@
 Learning gene regulatory networks is essential to our understanding of cell behavior and response to external stimuli.
-Methods for network inference have been developed and used for a wide variety of species; however, these approaches consider the problem as independent in each species even though there is known to be conservation of gene regulation even in distantly related species.
+Methods for network inference from large datasets have been developed and used for a wide variety of species; however, these approaches consider the problem as independent in each species even though there is known to be conservation of gene regulation even in distantly related species.
 We introduce a method that allows data to be pooled across species in a principled way based on gene orthology and demonstrate that it improves network inference even when the orthology is incomplete.
@@
 We represent gene regulatory networks as a matrix of weights, where rows index transcription factors and columns index genes.
-The weight in a given position represents the regulatory network on that TF on that gene; positive weights represent activation, negative weights represent repression, and 0 weights represent the absence of an interaction.
+The weight in a given position represents the regulatory weight of that TF on that gene; positive weights represent activation, negative weights represent repression, and 0 weights represent the absence of an interaction.
 We begin with an existing framework for estimating the interaction weights in a gene regulatory network (GRN).
@@
 \begin{equation}
-Y = \beta X
+Y = X \beta 
 \end{equation}
@@
 
-\subsection{Solving}
+\subsection{Solving using augmented matrices}
 The problem of solving a multi-valued regression problem can be converted into the problem of solving a single valued regression problem by vectorization.
@@
 The maximal value of $\lambda_S$ should be just large enough so that this correlation is near $1$.
-More details on this later.
 

### 2015-08-27 paper1.tex
changed intro

@@
 As the volume and variety of genome scale data have exploded, the goal of accurately modeling gene regulatory networks has become more attainable.
-Already, there is enough data in some model organisms for the development of high quality networks; however, this is still not true of many other model systems.
-Luckily, the problems of inferring gene regulatory networks in multiple species are not independent, due to the presence of conserved structure.
-There has been a great deal of work showing that functional conservation exists in gene regulatory networks even across large evolutionary distance.
-Abundant data in one species should be a useful source when learning the network of a related species.
-But most existing inference algorithms do not incorporate this information.
+Large-scale data collection efforts have contributed to the development of high quality networks, but there remain many biological processes and organisms for which there does not exist enough data.
+Furthermore, as new technologies are developed and some old ones are replaced, such as RNAseq and microarray, it becomes important to be able to combine different data sets, lest we lose valuable information from previous studies.
+The problems of inferring networks in related species, processes, and using similar technologies likely are not independent ones, and benefit from the infusion of additional data.
 
-We employ a linear model of gene expression in which the rate of expression is assumed to be proportional to a weighted sum of transcription factor expression.
+We present two methods for network inference which use fused regression, a type of penalized least squares regression, in which the difference of regulatory weights which are thought to be similar are penalized.
+Our fused L2 method uses a ridge fusion penalty on these differences, and is useful where the relationships between nodes (similarity of genes between data sets) is reliable.
+Our adaptive fusion uses a non-concave saturating fusion penalty to simultaneously learn multiple networks and conservation of these networks.
+
+To learn gene regulatory networks, we employ a linear model of gene expression in which the rate of expression is assumed to be proportional to a weighted sum of transcription factor expression.
 These weights represent the regulatory interactions we wish to learn.
-We can fit these weights simultaneously in multiple species of interest using time series and steady state expression data.
-This is similar to many existing network inference methods and produces identical answers to fitting each species separately.
-To take advantage of cross-species information, we introduce additional constraints on the problem that penalize differences between regulatory weights which are expected to be similar.
-The choice of which interactions to fuse together follows from the observation that orthologous genes are likely to have similar regulatory interactions.
-
-A few approaches to multi-source network inference have been proposed; however, current approaches either require that the sources have identical sets of genes [aviv regev paper], which is useful for analyzing cell lines but not for most cross species comparisons, or restrict analysis to the subset of genes with orthologs in every species [ibm paper], which limits the applicability of the model.
-Our approach allows simultaneous inference of the whole gene regulatory networks of species that may have partial orthology overlap.
-We demonstrate performance gains in networks learned from simulated data and real data from Bacillus subtilis and Bacillus anthracis, where a large fraction of genes have no orthologs.
-Though this method was developed for multi-species network inference, it is also useful in the easier problem of learning networks in multiple cell lines.
-
-The largest gains in performance are obtained when there is very little data in a species of interest but a large amount of data in a related species.
-But the information flow is two ways and our simulation results suggest that network inference improves in both species.
-This feature of simultaneously fitting networks is particularly useful when there is an adequate amount of data available but spread over too many species.
+We can fit these weights simultaneously in multiple data sources of interest using time series and steady state expression data.
+This is similar to many existing network inference methods and produces identical answers to fitting each network separately.
+To take advantage of cross-dataset information, we introduce additional constraints on the problem that penalize differences between regulatory weights which are expected to be similar.
+Genes are not always regulated by the same transcription factors under different conditions.
+For example, a TF which promotes expression of a gene under a given process may repress that same gene at another point.
+Using data from these conditions to learn one network will prevent the identification of these contradictory regulatory interactions; it is therefore important to be able to combine data in a principled manner.
+We can use knowledge about the conservation of regulation to while leveraging th eadditional data to learn networks describing disparate processes.
+In the case of multiple species, there has been a great deal of work showing that functional conservation exists in gene regulatory networks even across large evolutionary distance.
+We make the assumption that if a transcription factor in one species is orthologous to a TF in a related species, and a gene in the first species is orthologous to a gene in the second, then the interaction between the TF and gene in the first species should be similar to the interaction between the orthologous TF, gene pair in the second species.
 
@@
 Although sequence similarity is a useful predictor of functional similarity, we recognize that some genes have taken on different functions and therefore may have new regulatory interactions.
-When this occurs, the benefits in overall performance from our method come at the cost of impaired prediction of regulators for these genes.
+When this occurs, the benefits in overall performance from our L2 fusion approach come at the cost of impaired prediction of regulators for these genes.
 The constraints we impose favor weight configurations in which fused interaction weights are similar to one another.
-Our method minimizes a cost function that strives  to simultaneously fit expression data in each species and produce networks that are consistent with these evolutionary constraints.
+Our fused L2 method minimizes a cost function that strives to simultaneously fit expression data in each species and produce networks that are consistent with these evolutionary constraints.
 When a gene's regulators have changed through evolution, there may be no way of achieving both these goals.
@@
 
-We present two methods to multi-species network inference using fused regression.
-One approach, our fused L2 method, uses an L2 fusion penalty on differences in regulatory weights with evolutionary constraints.
-Additionally, we developed a method, which we call adaptive fusion, using a non-convex saturating fusion penalty for both network inference and network comparison.
 We test the ability of fused L2 and adaptive fusion to improve network recovery on both synthetic data and Bacillus subtilis and Bacillus anthracis.

### 2015-08-30 paper1.tex
hi

@@
 Furthermore, as new technologies are developed and some old ones are replaced, such as RNAseq and microarray, it becomes important to be able to combine different data sets, lest we lose valuable information from previous studies.
-The problems of inferring networks in related species, processes, and using similar technologies likely are not independent ones, and benefit from the infusion of additional data.
+The problems of inferring networks in related species, processes, and using different technologies likely are not independent ones, and benefit from the infusion of additional data.
+
+The extent to which datasets are related is an important consideration when deciding to combine them.
+Abundant data in one species should be a useful source when learning the network of a related species.
+But often, how much and what of orthologous networks' structure is conserved, is not known a priori.
+We would like, then to be able to learn the conserved subnetwork in order to make better predictions using multi-source network inference.
 
@@
 Using data from these conditions to learn one network will prevent the identification of these contradictory regulatory interactions; it is therefore important to be able to combine data in a principled manner.
-We can use knowledge about the conservation of regulation to while leveraging th eadditional data to learn networks describing disparate processes.
+We can use knowledge about the conservation of regulation while leveraging the additional data to learn networks describing disparate processes.
 In the case of multiple species, there has been a great deal of work showing that functional conservation exists in gene regulatory networks even across large evolutionary distance.
 We make the assumption that if a transcription factor in one species is orthologous to a TF in a related species, and a gene in the first species is orthologous to a gene in the second, then the interaction between the TF and gene in the first species should be similar to the interaction between the orthologous TF, gene pair in the second species.
+By utilizing information about conservation of gene function across data set, species, and biological process, we enable the usage of less directly relevant data and improve network inference.
 
@@
 When a gene's regulators have changed through evolution, there may be no way of achieving both these goals.
-In this case, the network inference procedure would be improved by relaxing or ignoring these evolutionary constraints.
+Similarly, if a gene has different regulators during different processes, the expression data may conflict with the fusion constraints.
+In these cases, the network inference procedure would be improved by relaxing or ignoring these evolutionary constraints.
 However, we don't know a priori which constraints to enforce and which to relax.
@@
 If TF $g^A$ in organism $A$ and TF $h^B$ in organism $B$ are orthologs, and gene $k^A$ and $l^B$ are orthologs, then we expect that the $g^A \rightarrow k^A$ interaction weight should be similar to the $h^B \rightarrow l^B$ interaction weight, and we introduce a fusion constraint between these analogous interactions.
-In terms of the above regression formulation, we expect that $\beta^A_{g,k} \approx \beta^B_{h,l}$, and include a penalty term $\lambda_S(\beta^A_{g,k} - \beta^B_{h,l})^2$ to the quantity being minimized in order to ensure similarity.
+In terms of the above regression formulation, we expect that $\beta^A_{g,k} \approx \beta^B_{h,l}$, and include a penalty term $p_{\lambda_S}(\beta^A_{g,k} - \beta^B_{h,l})$ to the quantity being minimized in order to ensure similarity.
 The parameter $\lambda_S$ controls how highly to penalize differences between fused coefficients and controls the tradeoff between fitting the expression data and producing a set of networks that conform to evolutionary constraints.
@@
 \begin{equation}
-\argmin_{(\beta^A, \beta^B)} \displaystyle\sum_{S \in (A, B)} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\beta^{S_1}_{g,k} \approx \beta^{S_2}_{h,l}} \lambda_S(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})^2
+\argmin_{(\beta^A, \beta^B)} \displaystyle\sum_{S \in (A, B)} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\beta^{S_1}_{g,k} \approx \beta^{S_2}_{h,l}} p_{\lambda_S}(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
 \end{equation}
 
-Where the second sum is over pairs of interactions with fusion constraints.
-Note that each term in the quantity being minimized is an L2 norm of a linear combination of gene expressions, so the entire problem can be solved through linear regression with a suitable design matrix.
+where the second sum is over pairs of interactions with fusion constraints.
+
+When the penalty $p_{\lambda_S}$ is the L2 norm, the entire problem can be solved through linear regression with a suitable design matrix.
 The regulators of a pair of genes can be solved independently as long as the pair is not linked by fusion constraints, or a chain of fusion constraints; thus the problem of cross-species GRN inference decomposes into a large number of manageable regression problems as long as the orthology mappings are sparse.
@@
 
-\subsection{Solving using augmented matrices}
+\subsection{Solving fused L2 problems using augmented matrices}
 The problem of solving a multi-valued regression problem can be converted into the problem of solving a single valued regression problem by vectorization.
@@
 
-We can then define $W_{i,j} = \beta_{j \times N + i}$ which is the desired form.
+We can then define $W_{i,j} = \beta_{(j-1) \times N + i, (j-1) \times M + j}$ which is the desired form.
 This is equivalent to the original problem due to the block structure of matrix multiplication
@@
 The fused multi-source regression problem can solved by constructing a multi-source design matrix $X$ and data matrix $Y$, and then vectorizing.
-The regression problems associated with two columns $g,h$ of $Y$, where $g$ came from $Y_G$ and $h$ came from $Y_H$, are independent, unless there is a constraint chain of length $L$ with the form (where $x$ is arbitrary)
+The regression problems associated with two columns $g,h$ of $Y$, where $g$ came from $Y_1$ and $h$ came from $Y_2$, are independent, unless there is a constraint chain of length $L$ with the form (where $x$ is arbitrary)
 
 %$$(G, j_0, g, J_1, j_1, m_1) \rightarrow (J_1, j_1, m_1, J_2, j_2, m_2) \rightarrow (J_2,j_2,m_2,J_3,j_3,_m3)...(J_P, j_P, m_P, H, j_{P+1}, h)$$
-$$(G, j_0, g, J_1, j_1, m_1) \rightarrow (J_1, x, m_1, J_2, j_2, m_2)$$
+$$(Y_1, j_1, g, Y_2, j_2, h) \rightarrow (Y_2, j_2, h, Y_3, j_3, k)$$
 
+where the $j$s refer to transcription factors.
+In subproblem 1, $j_1$ regulates gene $g$; $j_1$ is orthologous to $j_2$, a transcription factor of subproblem 2, which regulates $h$, a gene which is orthologous to gene $g$.
+There exists a chain because ($j_1$, $g$) is a TF, gene pair which is fused with ($j_2$, $h$), which is fused to ($j_3$, $k$).
 We use depth-first search to identify linked columns of $Y$, and then solve the fused regression problem directly by vectorizing those columns.
-In order to incorporate the fusion penalty, we append to the resulting design matrix one row \emph{per connected constraint}, which contains $\lambda_S$ in the column associated with $G,i,j,$ and $-\lambda_S$ to the column associated with $J, i',j'$ for constraint $(G,i,j,J,i',j')$.
+In order to incorporate the fusion penalty, we append to the resulting design matrix one row \emph{per connected constraint}, which contains $\lambda_S$ in the column associated with $Y_g,i,j,$ and $-\lambda_S$ to the column associated with $Y_h, i',j'$ for constraint $(Y_g,i,j,Y_h,i',j')$.
 Call this resulting matrix $X_P$.
 The vector $Y$ is augmented by appending an appropriate number of zeros to produce $Y_P$.
-As a result, each entry in the output of $X_P b$ in the columns associated with each constraint contain $b_{i,j} - b_{i',j,i}$.
-These entries each contribute  $||b_{i,j} - b_{i',\{ji\}}||^{~2}$ to the squared error $||X_Pb - Y_P||_2^{~2}$
+As a result, each entry in the output of $X_P b$ in the columns associated with each constraint contain $b_{i,j} - b_{i',j'}$.
+These entries each contribute  $||b_{i,j} - b_{i',j'}||^{~2}$ to the squared error $||X_Pb - Y_P||_2^{~2}$
 
 In the context of regression for GRN inference with multiple cell sources, each gene $j$ from source $1$ has constraints $\forall_{i \in \{TFs\}} (1, i, j, 2, i’, j’)$ where $i',j'$ are the same gene in a different cell line.
-As a result, each gene has only one constraint and $L=2$.
+As a result, each gene has only one constraint and the chain has length $L=2$.
 %In general, however, solving the fused regression problem directly, even after factoring constraints, involves at least one inversion of an $L\sum_{i=0}^S K_i \times L \sum_{i=0}^S K_i$ matrix, taking $O((L\sum_{i=0}^S K_i)^3)$.
@@
 
-\subsection{Alternative penalty functions}
+\subsection{Adaptive fusion}
 Fusion constraints are intended to penalize dissimilarity between interactions thought to be evolutionarily analogous.
@@
 \subsection{Optimization}
-We use cross-validation to select optimal regularization parameters $\lambda_R$ and $\lambda_S$.
-In order to determine the relevant scale of the fusion weight $\lambda_S$, we can compute the correlation between pairs of coefficients linked by fusion constraints.
-The maximal value of $\lambda_S$ should be just large enough so that this correlation is near $1$.
+We use glmnet to compute the regularization path for the
+ ridge penalty.
+Given a range of $\lambda_R$ values, we fit models for each $\lambda_R$.
+We use cross validation to select the optimal parameter, the $\lambda_R$ value which minimizes the error of prediction on the leave out set.
 

### 2015-08-30 paper1.tex
ahh
hi

@@
 
-There exist several tools for identifying orthologous genes using sequence similarity.
+There exist several tools for predicting orthology between genes using sequence similarity.
 Although sequence similarity is a useful predictor of functional similarity, we recognize that some genes have taken on different functions and therefore may have new regulatory interactions.
@@
 We approach this problem by introducing a saturating penalty function based on work in statistics to develop unbiased regularization penalties for regression.
-The resulting cost function is non-convex and difficult to optimize and fares less well in the underdetermined case; however, we can approximate its solution and obtain deeper insights into functional similarity than is available through orthology or the comparison of separately learned networks.
-When the quantity of data is sufficient to produce high quality networks, our adaptive fusion approach improves performance of network inference; this is achieved in part through the identification of presumptive orthologs with different functions.
-In simulated data, when we introduced faulty orthology information, our method not only correctly identified which evolutionary restraints to enforce and which to relax, but also outperformed the control case where no incorrect orthologs were used.
+The main difference between the L2 fusion approach and this adaptive fusion approach occurs when the difference between presumed analogous interactions is large despite the fusion penalty: we assume that large differences represent evidence that functions have diverged and represent this, in adaptive fusion, with a relaxation of the fusion penalty.
+The resulting cost function is non-convex and difficult to optimize; however, we can approximate its solution and obtain deeper insights into functional similarity than is available through orthology or the comparison of separately learned networks.
 

### 2015-08-31 paper1.tex

@@
 \usepackage{graphicx}
+\usepackage[pdf]{pstricks}
 %\usepackage{mathtools}
@@
 As the volume and variety of genome scale data have exploded, the goal of accurately modeling gene regulatory networks has become more attainable.
-Large-scale data collection efforts have contributed to the development of high quality networks, but there remain many biological processes and organisms for which there does not exist enough data.
-Furthermore, as new technologies are developed and some old ones are replaced, such as RNAseq and microarray, it becomes important to be able to combine different data sets, lest we lose valuable information from previous studies.
-The problems of inferring networks in related species, processes, and using different technologies likely are not independent ones, and benefit from the infusion of additional data.
+Large-scale data collection efforts have contributed to the development of high quality networks, but there remain many biological processes and organisms for which there does not exist adequate data.
+Furthermore, as new technologies are developed and some old ones are replaced, such as RNAseq and microarray, it becomes important to be able to combine data from multiple platforms, lest we lose valuable information from previous studies.
+The problem of inferring related -- but not necessarily identical -- structure from related -- but not identical -- data is ubiquitous in biology.
+Multi-source network inference has applications for learning networks in multiple related species, for learning networks associated with distinct processes within the same species, and for learning networks based on heterogenous data sources.
 
-The extent to which datasets are related is an important consideration when deciding to combine them.
-Abundant data in one species should be a useful source when learning the network of a related species.
-But often, how much and what of orthologous networks' structure is conserved, is not known a priori.
-We would like, then to be able to learn the conserved subnetwork in order to make better predictions using multi-source network inference.
+Most algorithms for network inference can not use data from multiple sources.
+This is particularly problematic when studying a lesser used model organism, or a biological process that is not widely studied.
+Approaches that do utilize data from multiple sources have largely relied on either learning individual networks and combining the solutions, which is unviable when one or more of the data sources is very limited, or on treating the data as if it came from a single source, which is limited to sources with a high degree of overlap (ie species with many genes in common).
+Moreover, combining the data into a single source assumes that the underlying networks are identical.
+In reality, the similarity of networks in related species is unknown.
+A method is needed that can flexibly combine data sources based on their similarity.
 
-We present two methods for network inference which use fused regression, a type of penalized least squares regression, in which the difference of regulatory weights which are thought to be similar are penalized.
-Our fused L2 method uses a ridge fusion penalty on these differences, and is useful where the relationships between nodes (similarity of genes between data sets) is reliable.
-Our adaptive fusion uses a non-concave saturating fusion penalty to simultaneously learn multiple networks and conservation of these networks.
+We present two methods for network inference based on linear estimates of gene expression dynamics, extending existing linear dynamical-systems methods for network inference.
+The core of both methods is the observation that biological information about the relatedness of two networks can be used to guess which network coefficients should be similar to one another in a multi-source network inference problem (ie orthologous TFs should regulate orthologous genes), and that these constraints can be efficiently represented as penalties in a least-squares regression problem.
+Our first method -- fused ridge -- method uses an L2 penalty on the differences between a priori similar interactions, and is useful where the relationships between networks (similarity of genes between data sets) is reliable.
+In the case where both networks contain an identical set of genes and TFs, this approach can be thought of as parametrically interpolating between treating the data sources separately and combining them together.
+However, our method allows useful pooling of data even when .
+Our second method -- adaptive fusion -- uses a non-concave saturating fusion penalty to simultaneously learn the constrained networks and to learn which constraints should be relaxed (ie which parts of the network are genuinely different).
 
-To learn gene regulatory networks, we employ a linear model of gene expression in which the rate of expression is assumed to be proportional to a weighted sum of transcription factor expression.
-These weights represent the regulatory interactions we wish to learn.
-We can fit these weights simultaneously in multiple data sources of interest using time series and steady state expression data.
-This is similar to many existing network inference methods and produces identical answers to fitting each network separately.
-To take advantage of cross-dataset information, we introduce additional constraints on the problem that penalize differences between regulatory weights which are expected to be similar.
-Genes are not always regulated by the same transcription factors under different conditions.
-For example, a TF which promotes expression of a gene under a given process may repress that same gene at another point.
-Using data from these conditions to learn one network will prevent the identification of these contradictory regulatory interactions; it is therefore important to be able to combine data in a principled manner.
-We can use knowledge about the conservation of regulation while leveraging the additional data to learn networks describing disparate processes.
 In the case of multiple species, there has been a great deal of work showing that functional conservation exists in gene regulatory networks even across large evolutionary distance.
-We make the assumption that if a transcription factor in one species is orthologous to a TF in a related species, and a gene in the first species is orthologous to a gene in the second, then the interaction between the TF and gene in the first species should be similar to the interaction between the orthologous TF, gene pair in the second species.
-By utilizing information about conservation of gene function across data set, species, and biological process, we enable the usage of less directly relevant data and improve network inference.
+We make the assumption that if a pair of TFs are orthologous to one another, then their regulatory effects on a pair of orthologous genes should be similar.
+This orthology information forms the basis of a large number of constraints on the network inference problems, which allow data from one organism to improve network inference on another organism.
+This very general framework for multi-species network inference can be extended to large numbers of distinct organisms, each contributing data with only partially overlapping sets of genes.
 
-There exist several tools for predicting orthology between genes using sequence similarity.
+Orthology -- the measure of gene similarity we use -- is based on readily identifiable sequence similarity.
 Although sequence similarity is a useful predictor of functional similarity, we recognize that some genes have taken on different functions and therefore may have new regulatory interactions.
-When this occurs, the benefits in overall performance from our L2 fusion approach come at the cost of impaired prediction of regulators for these genes.
-The constraints we impose favor weight configurations in which fused interaction weights are similar to one another.
+These cases of neofunctionalization can be difficult to verify by simply comparing independently inferred networks; because network inference is typically underconstrained, there could be other plausible networks in which the coefficients being compared are similar.
+This is in contrast with our multi-source method: because the constraints we impose favor weight configurations in which fused interaction weights are similar to one another, a solution in which a priori similar regulatory interactions are found to be different is strong evidence that they really are different.
 Our fused L2 method minimizes a cost function that strives to simultaneously fit expression data in each species and produce networks that are consistent with these evolutionary constraints.
@@
     \centering
-    \includegraphics[scale=.5,trim=0mm 0mm 0mm 0mm,clip]{direct_solve.eps}
+    \includegraphics[scale=.5,trim=0mm 0mm 0mm 0mm,clip]{direct_solve.png}
 \end{figure}

### 2015-11-07 paper1.tex
hi

@@
 Learning gene regulatory networks is essential to our understanding of cell behavior and response to external stimuli.
-Methods for network inference from large datasets have been developed and used for a wide variety of species; however, these approaches consider the problem as independent in each species even though there is known to be conservation of gene regulation even in distantly related species.
+Methods for network inference from large datasets have been developed and used for a wide variety of species; however, most approaches consider the problem as independent in each species even though there is known to be conservation of gene regulation even in distantly related species.
 We introduce a method that allows data to be pooled across species in a principled way based on gene orthology and demonstrate that it improves network inference even when the orthology is incomplete.
 We further refine this method with an algorithm that attempts to learn the true conserved subnetwork from a larger set of potentially conserved interactions.
+We demonstrate the utility of our method not only in cross species network inference, but also in merging data sets and incorporating binding similarity predictions.
 \end{abstract}
@@
 \section{Introduction}
-As the volume and variety of genome scale data have exploded, the goal of accurately modeling gene regulatory networks has become more attainable.
+As the volume and variety of genome scale data have exploded, the goal of accurately modeling gene regulatory networks has become more attainable \cite{bonneau_predictive_2007, ciofani_validated_2012, carro_transcriptional_2010}.
 Large-scale data collection efforts have contributed to the development of high quality networks, but there remain many biological processes and organisms for which there does not exist adequate data.
@@
 
-Most algorithms for network inference can not use data from multiple sources.
-This is particularly problematic when studying a lesser used model organism, or a biological process that is not widely studied.
-Approaches that do utilize data from multiple sources have largely relied on either learning individual networks and combining the solutions, which is unviable when one or more of the data sources is very limited, or on treating the data as if it came from a single source, which is limited to sources with a high degree of overlap (ie species with many genes in common).
-Moreover, combining the data into a single source assumes that the underlying networks are identical.
-In reality, the similarity of networks in related species is unknown.
-A method is needed that can flexibly combine data sources based on their similarity.
-
 We present two methods for network inference based on linear estimates of gene expression dynamics, extending existing linear dynamical-systems methods for network inference.
-The core of both methods is the observation that biological information about the relatedness of two networks can be used to guess which network coefficients should be similar to one another in a multi-source network inference problem (ie orthologous TFs should regulate orthologous genes), and that these constraints can be efficiently represented as penalties in a least-squares regression problem.
+The core of both methods is the observation that biological information about the relatedness of two networks can be used to approximate which network coefficients should be similar to one another in a multi-source network inference problem (ie orthologous TFs should regulate orthologous genes), and that these constraints can be efficiently represented as penalties in a least-squares regression problem.
 Our first method -- fused ridge -- method uses an L2 penalty on the differences between a priori similar interactions, and is useful where the relationships between networks (similarity of genes between data sets) is reliable.
 In the case where both networks contain an identical set of genes and TFs, this approach can be thought of as parametrically interpolating between treating the data sources separately and combining them together.
-However, our method allows useful pooling of data even when .
+However, our method allows useful pooling of data even when the overlap between genes is incomplete.
 Our second method -- adaptive fusion -- uses a non-concave saturating fusion penalty to simultaneously learn the constrained networks and to learn which constraints should be relaxed (ie which parts of the network are genuinely different).
@@
 This very general framework for multi-species network inference can be extended to large numbers of distinct organisms, each contributing data with only partially overlapping sets of genes.
+We use our knowledge about the relationship between nodes across networks in order to constrain the network inference problem; this approach has many applications, from improving network inference of multiple species to combining data sets to leveraging knowledge about similarity in binding sites.
 
-Orthology -- the measure of gene similarity we use -- is based on readily identifiable sequence similarity.
-Although sequence similarity is a useful predictor of functional similarity, we recognize that some genes have taken on different functions and therefore may have new regulatory interactions.
-These cases of neofunctionalization can be difficult to verify by simply comparing independently inferred networks; because network inference is typically underconstrained, there could be other plausible networks in which the coefficients being compared are similar.
-This is in contrast with our multi-source method: because the constraints we impose favor weight configurations in which fused interaction weights are similar to one another, a solution in which a priori similar regulatory interactions are found to be different is strong evidence that they really are different.
-Our fused L2 method minimizes a cost function that strives to simultaneously fit expression data in each species and produce networks that are consistent with these evolutionary constraints.
-When a gene's regulators have changed through evolution, there may be no way of achieving both these goals.
-Similarly, if a gene has different regulators during different processes, the expression data may conflict with the fusion constraints.
-In these cases, the network inference procedure would be improved by relaxing or ignoring these evolutionary constraints.
+Existing multi species approaches often use orthology as a proxy for functional conservation \cite{penfold_inferring_2015, joshi_multi-species_2015, kashima_simultaneous_2009, zhang2010nearly}, or attempt to learn functional similarity via expression data \cite{gholami_cross-species_2010}.
+Orthology -- the measure of gene similarity we use -- can be approximated using readily identifiable sequence similarity.
+Sequence similarity is often a useful predictor of functional similarity \cite{wilson_assessing_2000}.
+Our fused L2 method minimizes a cost function that strives to simultaneously fit expression data in each species and produce networks that are consistent with evolutionary constraints created using orthology.
+However, we recognize that some genes may have taken on different functions and therefore may have new regulatory interactions.
+Gene duplications and neofunctionalization in multi-species network inference \cite{eisen_phylogenomics:_1998}, and changes in chromatin configuration in multi-cell line network inference \cite{li_role_2007}, are cases where enforcing fusion of constraints does not produce networks which reflect true differences in regulation between networks.
+These cases can be difficult to learn by simply comparing independently inferred networks; because network inference is typically underconstrainted \cite{marbach_revealing_2010-1}, there could be other plausible networks in which the coefficients being compared are similar.
+Our adaptive fusion method allows us to learn differences in networks predicted to be similar based on orthology or gene matching: because the constraints we impose favor weight configurations in which fused interaction weights are similar to one another, a solution in which a priori similar regulatory interactions are found to be different is strong evidence that they really are different.
+When differences in regulatory interactions arise where they are predicted to be similar, the network inference procedure would be improved by relaxing or ignoring these evolutionary constraints.
 However, we don't know a priori which constraints to enforce and which to relax.
 
-We approach this problem by introducing a saturating penalty function based on work in statistics to develop unbiased regularization penalties for regression.
-The main difference between the L2 fusion approach and this adaptive fusion approach occurs when the difference between presumed analogous interactions is large despite the fusion penalty: we assume that large differences represent evidence that functions have diverged and represent this, in adaptive fusion, with a relaxation of the fusion penalty.
+We approach this problem by introducing a saturating penalty function based on work in statistics to develop unbiased regularization penalties for regression \cite{zhang2010nearly, fan2001variable}.
+The main difference between the L2 fusion approach and this adaptive fusion approach occurs when the difference between presumed analogous interactions is large despite the fusion penalty: we assume that large differences represent evidence that gene functions have diverged and represent this, in adaptive fusion, with a relaxation of the fusion penalty.
 The resulting cost function is non-convex and difficult to optimize; however, we can approximate its solution and obtain deeper insights into functional similarity than is available through orthology or the comparison of separately learned networks.
 
-We test the ability of fused L2 and adaptive fusion to improve network recovery on both synthetic data and Bacillus subtilis and Bacillus anthracis.
+We test the ability of fused L2 and adaptive fusion to improve network recovery on both synthetic data and Bacillus subtilis and Bacillus anthracis, showing the applicability of our method in combining different datasets, and leveraging similarity across organisms as well as within a network in order to improve network inference.
 We explore the circumstances under which each approach is optimal, and evaluate the robustness of adaptive fusion to incorrect orthology, simulating the biologically relevant case of neofunctionalization.
@@
 \subsection{Gene regulatory network}
-We represent gene regulatory networks as a matrix of weights, where rows index transcription factors and columns index genes.
+We represent gene regulatory networks as a matrix of weights, where rows index transcription factors (TFs) and columns index genes.
 The weight in a given position represents the regulatory weight of that TF on that gene; positive weights represent activation, negative weights represent repression, and 0 weights represent the absence of an interaction.
 We begin with an existing framework for estimating the interaction weights in a gene regulatory network (GRN).
-In the Inferelator algorithm, linear differential equations are used to model changes in gene expression.
-The rate at which the observed mRNA expression of gene $i$, $x_i$, changes is governed by degradation of existing transcripts with rate $\alpha$ plus a linear combination of transcription factor (TF) expressions.
+In the Inferelator algorithm, linear differential equations are used to model changes in gene expression \cite{bonneau_inferelator:_2006-1} The rate at which the observed mRNA expression of gene $i$, $x_i$, changes is governed by degradation of existing transcripts with rate $\alpha$ plus a linear combination of transcription factor (TF) expressions.
 \begin{equation}
@@
 We are interested in learning $\beta$, the matrix defining the influence of each TF on each gene.
-We fix the decay rate $\alpha$ for all genes, and set it assuming a time-constant of 10 minutes [CITATIONS], as in [CITATIONS].
+We fix the decay rate $\alpha$ for all genes, and set it assuming a time-constant of 10 minutes \cite{hambraeus_genome-wide_2003, selinger_global_2003}, as in \cite{greenfield_robust_2013}.
 Let $x_i(t)$ be the expression of gene $i$ at time $t$.
@@
 \end{equation}
-This is essentially the same basic formulation as was used in the Inferelator algorithm, which we extend to the case of simultaneously inferring the GRNs of multiple species.
+This is similar to the formulation used in the Inferelator algorithm, which we extend to the case of simultaneously inferring the GRNs of multiple species.
 
@@
 Part of the appeal of the approach, however, is that it allows pooling of data even when there is incomplete orthology.
-By introducing constraints on the similarity of individual interactions, rather than on the networks as a whole [CITE IBM?], we can pool some information across species even when an arbitrarily small fraction of genes have orthologs.
+By introducing constraints on the similarity of individual interactions, rather than on the networks as a whole \cite{liu2011temporal}, we can pool some information across species even when an arbitrarily small fraction of genes have orthologs.
 This is particularly useful when dealing with a large number of species; pairwise orthologies may be nearly complete even when the number of genes present in every organism is small.
@@
 Specifically, if $Y \in R^{N \times M}$ then we form $A =$ diagonally concatenate $X$, $M$ times (with the remaining entries 0),
-$y =$ vertically concatenate the columns of $Y$
+$y =$ vertically concatenate the columns of $Y$. We solve
 
-$$\underset{w}{min} ||Aw - y||_2^{~2}$$
+$$\underset{W}{min} ||AW - y||_2^{~2}$$
 
-We can then define $W_{i,j} = \beta_{(j-1) \times N + i, (j-1) \times M + j}$ which is the desired form.
+%We can then define $W_{i,j} = \beta_{(j-1) \times N + i, (j-1) \times M + j}$ which is the desired form.
 This is equivalent to the original problem due to the block structure of matrix multiplication
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[scale=.5,trim=0mm 0mm 0mm 0mm,clip]{direct_solve.png}
-\end{figure}
+%\begin{figure}[h!]
+%    \centering
+%    \includegraphics[scale=.5,trim=0mm 0mm 0mm 0mm,clip]{direct_solve.png}
+%\end{figure}
 
@@
 The vector $Y$ is augmented by appending an appropriate number of zeros to produce $Y_P$.
-As a result, each entry in the output of $X_P b$ in the columns associated with each constraint contain $b_{i,j} - b_{i',j'}$.
-These entries each contribute  $||b_{i,j} - b_{i',j'}||^{~2}$ to the squared error $||X_Pb - Y_P||_2^{~2}$
+As a result, each entry in the output of $X_P b$ in the columns associated with each constraint contain $\lambda_S(b_{i,j} - b_{i',j'})$.
+These entries each contribute  $\lambda_S||b_{i,j} - b_{i',j'}||^{2}$ to the squared error $||X_Pb - Y_P||_2^{~2}$
 
@@
 With a saturating fusion penalty, fused interactions which appear to be very different based on expression data are allowed to unfuse from one another.
-A closely related problem has been studied in the context of LASSO regularization, where it was shown by Fan and Li that using a saturating penalty retains many of LASSO's desireable properties while removing its bias towards 0.
+A closely related problem has been studied in the context of LASSO regularization, where it was shown by Fan and Li that using a saturating penalty retains many of LASSO's desireable properties while removing its bias towards 0 \cite{fan2001variable}.
 They further showed that, although the resulting loss-function is nonconvex, good results can be obtained with a local quadratic approximation of gradient descent.
-Several saturating penalties, such as SCAD [Fan Li 2001] and MCP [Zhang], have been discussed in the context of sparse regression.
+Several saturating penalties, such as SCAD \cite{fan2001variable} and MCP \cite{zhang2010nearly}, have been discussed in the context of sparse regression.
 We introduce a modified form of MCP to the problem of penalizing differences between fused coefficients.
@@
 
-As in [Fan Li], we use solve using iterative local quadratic approximation.
+As in \cite{fan2001variable}, we use solve using iterative local quadratic approximation.
 
@@
 
-\subsection{Model selection}
-We can define a leave-out set of priors and choose parameters based on cross-validated AUPR on the leave-out set.
+%\subsection{Model selection}
+%We can define a leave-out set of priors and choose parameters based on cross-validated AUPR on the leave-out set.
 
@@
 Following generation of simulated data, we may introduce error into the orthology mapping.
-This can take the form of discarding a specified fraction of true orthologies (governed by a false-negative rate), or by introducing random false orthologies (governed by a false-positive rate).
+This can take the form of discarding a specified fraction of true orthologies (governed by a false-negative rate), by introducing random false orthologies (governed by a false-positive rate), or by adding Gaussian noise so that fused interactions are not identical.
 For convenience, the false-positive rate is specified in units of the number of true orthologs, and not the number of possible orthologs.
@@
 
-\begin{figure}
-\begin{center}
-  \includegraphics[scale=0.45]{l2fusionquick.png}
-  \caption{\label{fig:figure1} We generate a series of networks with 20 TFs by 200 genes, with 75\% sparsity, 0.1 measurement noise, and 0.1 fusion noise.
-We hold lamP and lamR constant at 1.0 and 2.0 respectively.
-The x-axis represents lamS and the y-axis represents the proportion of nodes which are orthologous.
-As the networks increase in similarity, performance improves and benefits from a large fusion penalty weight.
-NOTE: This figure shows AUPR.
-$R^2$ behaves as expected with respect to $\lambda_S$ (networks with large differences between fused coefficients have best $R^2$ at intermediate $\lambda_S$).
-This doesn't really seem to be the case with AUPR.
-Maybe these early figures should be replaced with $R^2$}
-  \end{center}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-  \includegraphics[scale=0.45]{increase_data2.png}
-  \caption{\label{fig:figure1} We fuse two networks, both with 25 TFs by 50 genes, using lamS=0 and nonzero lamS, and hold the number of conditions in network 1 constant while varying the number of data points for network 2.
-We plot the performance of network 2.}
-  \end{center}
-\end{figure}
-
 \subsection{Network similarity and fusion performance}
-We assessed how varying the similarity of simulated networks affected network inference performance with fusion.
-The main factors governing similarity between our generated networks were the extent (and accuracy) of the orthology mapping, and the magnitude of variability between analogous interactions.
-We conducted a series of simulations to assess the effect of varying orthology coverage on networks in which analogous interactions varied little from one another.
+Using related data for network inference allows us to borrow statistical power from other sources, effectively increasing the sample size and boosting the sensitivity and specificity of learned interactions.
+The decision of what other sources to draw from is an important one; at one extreme is combining unrelated data, and at the other is resampling from the same experiments.
+We simulated this spectrum by creating pairs of networks and varying the similarity between networks.
+The main factors governing similarity between our generated networks is the extent (and accuracy) of the orthology mapping, and the magnitude of variability between analogous interactions.
+We conducted a series of simulations to assess the effect of increasing orthology coverage on networks.
 In order to determine the effect of the size of the orthology mapping on network inference performance, we simulated a series of 20 TFs by 200 genes networks at several values of percent orthology coverage.
 
-As expected, increasing the weight of the fusion constraint $\lambda_S$ improved network recovery (Figure 2).
-Networks with higher orthology coverage saw a greater benefit from fusion, although the improvement was not limited to analogous interactions.
+As expected, when the conserved subgraph is very similar, increasing the weight of the fusion constraint $\lambda_S$ improved network recovery (Figure 2).
+Networks with higher orthology coverage saw a greater benefit from fusion, although the improvement was not limited to analogous interactions (Figure 4).
+When the conservation is weak, and even the portions of the network which are conserved have deviated from each other significantly, performance may not continue to improve with by increasing the fusion weight; we simulate this by adding noise to the conserved subgraph (Figure 3).
+We show that even for networks where the conserved subgraph is weakly conserved, there exists a 'sweet spot' where fusion regression improves network recovery.
+
 One way of assessing the effect of fusion on network inference performance is to compute an ``exchange-rate'' between conditions in each network.
@@
 A related species, B, is easier to collect data for, but of less intrinsic interest.
-Our simulation results suggest that adding conditions from species B to a fused network inference problem has an effect on network A recovery that is similar to adding conditions from A at a reduced rate (Figure 3).
-
-We also assessed the effect of varying the standard deviation of the distribution of differences between analogous interactions on network performance. FILL IN. MENTION PRIOR INTERPRETATION
-
-\begin{figure}
-\begin{center}
-  \includegraphics[scale=0.45]{con_noncon.png}
-  \caption{\label{fig:figure4} We looked at four different sets of networks, each with 20 TFs by 200 genes, where each set had different amounts of orthologous genes.
-We plotted the performance on the whole network, the constrained, and non-constrained portions of the network as we increased the fusion penalty weight.}
-  \end{center}
-\end{figure}
+Our simulation results suggest that adding conditions from species B to a fused network inference problem has an effect on network A recovery that is similar to adding conditions from A at a reduced rate.
 
 \subsection{Fused regression improves performance on both the constrained and non-constrained parts of the network}
-We wanted to know if using fused regression on networks with only conserved subgraphs affected the recovery of the non constrained parts of the networks (Figure 4).
-We used sets of two networks with 20 TFs and 200 genes each, where each set had 0\%, 25\%, 50\%, or 75\% of the TFs and genes in orthology groups.
-We varied lamS and computed AUPR using cross-validation on the portion of the network that had fusion constraints, the portion without any orthology information, and on the whole network.
-On the fused networks, performance improved as the fusion penalty weight increased; interestingly, performance gains were seen even the portion of the network where no orthology information is known.
-By constraining part of an underdetermined system, we obtained gains in both parts.
-
-\begin{figure}
-\begin{center}
-  \includegraphics[scale=0.45]{plot_betas_scad_l2.png}
-  \caption{\label{fig:figure1} We used two networks, each with 20 TFs by 200 genes, 50\% of which are orthologous.
-We include as many false positive orthologs as true positives, and color the points solved using true orthology information green and the points solved using false orthology information red.
-We randomly select an interaction from each network, and color this point blue.
-We solve the networks using fused L2.}
-  \end{center}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-  \includegraphics[scale=0.45]{plot_betas_scad.png}
-  \caption{\label{fig:figure1} We used two networks, each with 20 TFs by 200 genes, 50\% of which are orthologous.
-We include as many false positive orthologs as true positives, and color the points solved using true orthology information green and the points solved using false orthology information red.
-We randomly select an interaction from each network, and color this point blue.
-We solve the networks using adaptive fusion.}
-  \end{center}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-  \includegraphics[scale=0.45]{scadf.png}
-  \caption{\label{fig:figure1} Here we use two 10 TF by 200 gene networks, 50\% of which are orthologous.
-We include as many false positive orthologs as we have true positives, and apply adaptive fusion to the beta values with lamS=3.
-We plot the distribution of lamS learned for false orthologs.}
-  \end{center}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-  \includegraphics[scale=0.45]{scadt.png}
-  \caption{\label{fig:figure1} Here we use two 10 TF by 200 gene networks, 50\% of which are orthologous.
-We include as many false positive orthologs as we have true positives, and apply adaptive fusion to the beta values with lamS=3.
-We plot the distribution of lamS learned for true orthologs.}
-  \end{center}
-\end{figure}
+Our method learns networks simultaneously, using predictions about conservation of TF-gene interactions to share information between sources.
+Our approach is useful for learning networks from similar sources such as different cell-lines from the same species, where there exists a one-to-one mapping of genes, as well as datasets where the networks are expected to be more different than similar, where there exists limited orthology mapping.
+For the latter case, we are interested in knowing if performance gains from fused regression are limited to those interactions which have fusion constraints.
+To test this, we used sets of 20 TF by 200 gene synthetic networks, with varying proportions of orthology information about TFs and genes.
+We varied lamS and computed AUPR on the portion of the network that had fusion constraints, the portion where TF-gene interactions from one network could be mapped to putative analogous interactions in the second network based off of orthology; the unconstrained subnetwork; and the whole network.
+As expected, performance improved as the fusion penalty weight increased.
+Interestingly, performance gains were seen even in the portion of the network that is unconstrained by fusion.
+By constraining part of an underdetermined system, we obtain gains in even the unconstrained part.
 
 \subsection{Adaptive fusion successfully identifies and unfuses 'neofunctionalized' genes}
-We recognize that orthology prediction is not necessarily a perfect proxy for functional conservation.
-We implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions.
+We recognize that orthology prediction is not a perfect proxy for functional conservation.
+We implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (Figures 5-6).
 Pairs of interactions which are very dissimilar even after fusion, which sit in the flat portion of this penalty function, are effectively ``unfused''.
-We can interpret this ``unfusing'' as evidence for neofunctionalization of one or more of the genes involved in this pair of interactions.
+Because our network inference problem is underdetermined, meaning that multiple networks would fit our data equally well, the ``unfusing'' of constraints is convincing evidence for the rejection of a network which uses orthology prediction to predict similarity in interaction weights.
+''Unfusing'' suggests neofunctionalization of one or more of the genes involved in this pair of interactions.
 
-We performed a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved when orthology information is faulty, to mimic orthologs which are not functionally analogous  (Figure 5-8).
+We performed a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved when orthology information is faulty, to mimic orthologs which are not functionally analogous  (Figure 7-8).
 We generated synthetic fused networks and introduced error in the fusion constraints by adding false positives and negatives to the orthology information given to the solver.
@@
 We verified that adaptive-fusion unfused mostly ``fake'' interactions, while leaving truly analogous interactions fused.
+We compared the recovery of interaction weights which were accurately fused, and recovery of interaction weights which were inaccurately fused due to incorrect orthology information.
+Because fused L2 heavily penalizes large differences between weights which are predicted to be similar (Figures 8-9), the error for inaccurately fused interactions is higher than solving without fusion.
+When we use our adaptive fusion approach, we preserve the gain in performance on accurately fused interactions, while reducing the error for weights of interactions involving "fake" orthologies.
 
-\begin{figure}
-\begin{center}
-  \includegraphics[scale=0.45]{test_scad_opt_params3.png}
-  \caption{\label{fig:figure1} We use three sets of 20 TF by 200 gene networks, 50\% of which are orthologs, and solve with orthology information containing 0:1, 0.5:1, and 1:1 false positive to true positive.
-We plot AUPR under fused L2 and adaptive fusion.}
-  \end{center}
-\end{figure}
-
-\subsection{Adaptive fusion improves network inference}
-We wanted to determine whether adaptive fusion is able to improve learning of GRNs.
-That is, beyond correctly identifying functional conservation and recovering the performance lost by fusing neofunctionalized genes, is our method able to produce more accurate networks than those learned without fusion?
-(Figure 9) We again generated synthetic fused networks and added false positive and negative orthologs, then used fused L2 and adaptive fusion approaches to solve the underlying network.
-Fused L2 incorporated flawed orthology information, resulting in degradation of performance as we increased orthology error.
-Performance increased using adaptive fusion, pointing to its utility in network inference in addition to identifying neofunctionalization.
-
-\section{Bacterial data}
+\subsection{Cross-species network inference using bacterial data}
 We used gene-expression data from B.
@@
 anthracis in order to assess performance gains of fused regression on real data.
-Our B subtilis data set consists of 266 time-series and steady-state observations of 4891 genes during development.
+Our B subtilis data set consists of 360 time-series and steady-state observations of 4891 genes during development.
 Our B.
@@
 
-We obtained 1870 one-to-one orthologs from Inparanoid, which produced 177650 fusion-constraints between gene interactions within the two species.
+We obtained 1870 one-to-one orthologs from Inparanoid \cite{ostlund_inparanoid_2010}, which produced 177650 fusion-constraints between gene interactions within the two species.
 Although this number seems very large, it represents only $14.7\%$ of the regulatory interaction matrix in B.
@@
 
-\subsection{Bacterial performance}
-\subsubsection{Performance with priors}
-We want to show that performance increases with introduction of priors - this is easy/circular but important to show that our method can effectively incorporate priors
-
-\subsubsection{Adaptive fusion}
-We want to show that we can successfully identify neofunctionalization; since we do not have a gold standard for this we will introduce false positive orthologs into the orthology information and evaluate the ability of our method to identify true from false orthologs.
-
-\subsubsection{Precision recall}
-Introducing fusion constraints leads to a small but consistent improvement in AUPR.
+When we applied fused L2 to our subsampled B.
+subtilis and B.
+anthracis datasets, we did not see a significant improvement over solving the networks separately.
 Inspection of precision recall curves shows that this improvement is a result of increased precision at low values of recall.
@@
 
-\begin{center}
-  \begin{tabular}{ l | l | l }
-    \hline
-    $\lambda_S$ & $\%$ data & $AUPR constrained$ \\ \hline
-    0 & 0.066 & 0.0390 \\
-    5 & 0.066 & 0.0415 \\
-    \hline
-  \end{tabular}
-\end{center}
+\subsection{Within-species fusion using similarity of promoter region}
+Information about the similarity of TF-gene interactions can also come from knowledge about the promoter region; this may be a better predictor of target similarity than orthology.
+In bacteria, genes within the same operon are under the control of the same promoter.
+We predict, therefore, that genes within the same operon will be regulated similarly by the same transcription factors.
+We apply fusion regression by creating fusion constraints between a given transcription factor and genes within the same operon, and show a boost in B.
+subtilis network recovery using within-species fusion.
 
-\begin{figure}
-\begin{center}
-  \includegraphics[scale=0.45]{waka7unnormed.png}
-  \caption{\label{fig:figure1} precision recall curves of constrained interactions for several values of lamS, computed across folds containing $\frac{1}{15}th$ of the available B. Subtilis data}
-  \end{center}
-\end{figure}
+\subsection{Integrating datasets using fusion regression}
+Although there are many large-scale collaborations which attempt to make protocols as uniform as possible for comparability between datasets generated by different labs, there still exists technical and biologial variability between many experiments attempting to capture the same or similar experimental condition.
+With the advent of new technology, such as RNAseq and single-cell RNAseq, microarray is no longer the dominant assay for genome-wide expression, but it is important to be able to combine datasets generated using different technologies, lest we lose valuable information.
+Currently, the most widely used approach to combining datasets for network inference is to learn networks from disparate datasets separately, then rank combine the networks as in Marbach et al \cite{marbach_revealing_2010}.
+We compare this approach to using fusion regression.
+We include, along with our B.
+subtilis dataset, a previously published dataset containing 269 samples covering 104 conditions \cite{nicolas2012condition}.
+We compare performance when learning the networks separately, learning the networks separately followed by rank combining, and learning the networks simultaneously using fusion regression, and show improvement on network inference using our fused L2 approach.
 
-\begin{figure}
-\begin{center}
-  \includegraphics[scale=0.45]{woko7.png}
-  \caption{\label{fig:figure1} normalized precision recall curves of constrained interactions for several values of lamS, computed across folds containing $\frac{1}{15}th$ of the available B.
-Subtilis data.
-Normalization involved dividing the precision each each recall level, for each curve, and for each cross-validation fold by the corresponding $\lambda_S=0$ precision.}
-  \end{center}
-\end{figure}
+\subsection{Transcription factor activity estimation integrates into fusion regression approach}
+We test a combination of our fused regression approach with a method for estimating transcription factor activities (TFA).
+Rather than modeling gene expression using transcription factor mRNA abundance, we attempt to fit gene expression as a function of transcription factor activity.
+We estimate TFA based on known regulatory interactions using network component analysis \cite{liao2003network}.
+We randomly divide the prior known interactions in half, and use half to learn TFA.
+The rest we reserve as a gold standard for validation.
+When we incorporate TFA and repeat our experiment combining B.
+subtilis datasets, we see a marked improvement in performance.
 
-\begin{figure}
-\begin{center}
-  \includegraphics[scale=0.45]{wat7.png}
-  \caption{\label{fig:figure1} As previous figure, but for all interactions.}
-  \end{center}
-\end{figure}
-
+%\nocite{*}
 \bibliographystyle{plain}
-\bibliography{paper}
+\bibliography{paper1.bib}
 

### 2015-11-16 paper1.tex
hi

@@
 
-\subsubsection{Equivalent prior}
-Our L2 fusion penalty is equivalent to assuming a Gaussian prior with variance proportional to $\frac{1}{\lambda_S}$ on differences in parameters with fusion constraints.
+\subsection{Iterative solver}
+
+When the number of genes linked by fusion constraints is relatively small, the augmented design matrix approach we have described has the advantage of being efficiently solvable with a closed form solution.
+However, in the general case the augmented system of equations may be too large to solve efficiently with standard methods.
+To address this limitation, we developed an iterative solver that uses coordinate-wise descent to solve for solutions corresponding to a sequence of values of fusion penalty weights; since our fused L2 method uses a convex and differentiable penalty function, this approach converges to a global minimizer.
+Although less efficient than the augmented design matrix approach we developed, the iterative solver has the advantage of computing a solution path for $\lambda_S$.
+
+On each iteration $t$ the iterative solver computes
+
+\begin{equation}
+\text{argmin}_{\beta^S(t) : S_i, S_j \in \text{species}} \displaystyle\sum_{S_i, S_j \in \text{species}} ||X^{S_i}\beta^{S_i}(t) - Y^{S_i}||^2 + \lambda_R||\beta^{S_i}(t)||^2 + \displaystyle \sum_{\beta_{g,k}^{S_i} \approx \beta_{h,l}^{S_j}} \lambda_S(\beta^{S_i}_{g,k}(t) - \beta_{h,l}^{S_j}(t-1)^2)
+\end{equation}
+Note that this is almost identical to equation 5, but now the network $\beta$ is a function of the iteration number $t$.
+On each step, the we compute $\beta$s that minimize a penalized cost function where the fusion penalty encourages similarity of a pair of parameters, using the previous iteration's solution.
+This process is iterated until the estimated $\betas$s converge.
+Because each iteration reduces the error between $\beta(t)$ and $\beta(t-1)$, and because $\beta(t) = \beta(t-1)$ is the globally optimal solution, this process must eventually converge to the same network as (cite 5).
+Although we have not produced bounds on the convergence rate, in practice a small number of iterations are necessary.
+
+\subsection{Fusion and regularization path}
+The iterative solver requires an initial guess $\beta^{S_i}(0)$ for each species $S_i$.
+When the initial guess is close to the final solution, a small number of iterations are necessary.
+Because a small change in $\lambda_S$ will tend to produce a small change in the network, solving for the $\lambda_S+\epsilon$ network given $\beta$s initialized to the $\lambda_S$ network requires a small number of iterations.
+As a consequence, the entire solution path corresponding to a large number of values of $\lambda_S= \lambda_S_1, \lambda_S_2, ..., \lambda_S_{max}$ can be solved efficiently by using the solution of each network as the initial guess for the next.
+
+We use Tibshirani's cyclical coordinate descent algorithms from the 'glmnet' package \cite{friedman_regularization_2010} to compute a ridge regularization path.
+
+\subsection{Optimization}
+Because jointly optimizing $\lambda_R$ and $\lambda_S$ is computationally prohibitive, we opted to optimize the parameters separately.
+First, we use glmnet to compute the regularization path for the ridge penalty, and given a range of $\lambda_R$ values, we fit models for each $\lambda_R$.
+We use cross validation to select the optimal parameter, the $\lambda_R$ value which minimizes the error of prediction on the leave out set.
+Following selection of $\lambda_R$, we search for optimal $\lambda_S$ again using cross validation.
+Note that both parameters are chosen without reference to the gold standard, which is used in a separate evaluation of network quality.
+
+%\subsection{Equivalent prior}
+%Our L2 fusion penalty is equivalent to assuming a Gaussian prior with variance proportional to $\frac{1}{\lambda_S}$ on differences in parameters with fusion constraints.
 Combined with the regularization constraints, this forms a multivariate Gaussian prior on weights in both networks.
@@
 
+Our adaptive penalty function introduces, in addition to regularization and fusion penalty weights $\lambda_R$ and $\lambda_S$, an unknown parameter $a$.
+We could employ grid search using cross-validation to searach for the best parameters, but this can be computationally expensive.
+Moreover, we are primarily interested in using this saturating penalty as a way of testing the hypothesis that conservation in GRNs can be predicted based off of known similarities between genes.
+Therefore, we implement a way of picking $a$ with a user-selected percentile of differences in interaction weights, which represents the user's hypothesis for the proportion of evolutionary constraints which should be relaxed.
+A gold standard can also be used for selecting $a$ by comparing differences in weights for truly analogous interactions with those of randomly paired interactions, and selecting the difference in interaction weights at which no further fusion penalty is incurred.
+
 %\subsection{Model selection}
@@
 
-\subsection{Optimization}
-We use glmnet to compute the regularization path for the
- ridge penalty.
-Given a range of $\lambda_R$ values, we fit models for each $\lambda_R$.
-We use cross validation to select the optimal parameter, the $\lambda_R$ value which minimizes the error of prediction on the leave out set.
-
 \section{Results}
@@
 
-\subsection{Network similarity and fusion performance}
+\subsection{Using fused regression to learn related networks}
 Using related data for network inference allows us to borrow statistical power from other sources, effectively increasing the sample size and boosting the sensitivity and specificity of learned interactions.
+We used synthetic data to approximate two related biological processes and show that simultaneously learning the networks describing them using fused L2 regression to share information about related interactions results in an improvement in our ability to recover the true networks.
+Adding conditions from a related dataset has an effect on network inference which is akin to adding experimental data, without doing additional experiments.
+
 The decision of what other sources to draw from is an important one; at one extreme is combining unrelated data, and at the other is resampling from the same experiments.
 We simulated this spectrum by creating pairs of networks and varying the similarity between networks.
-The main factors governing similarity between our generated networks is the extent (and accuracy) of the orthology mapping, and the magnitude of variability between analogous interactions.
+The main factors governing similarity between our generated networks is the extent (and accuracy) of the orthology mapping, and the variability between analogous, or conserved, interactions.
 We conducted a series of simulations to assess the effect of increasing orthology coverage on networks.
@@
 As expected, when the conserved subgraph is very similar, increasing the weight of the fusion constraint $\lambda_S$ improved network recovery (Figure 2).
-Networks with higher orthology coverage saw a greater benefit from fusion, although the improvement was not limited to analogous interactions (Figure 4).
-When the conservation is weak, and even the portions of the network which are conserved have deviated from each other significantly, performance may not continue to improve with by increasing the fusion weight; we simulate this by adding noise to the conserved subgraph (Figure 3).
+Networks with higher orthology coverage saw a greater benefit from fusion.
+
+When the conservation is weak, and even the portions of the network which are conserved have deviated from each other significantly, performance may not continue to improve by increasing the fusion weight; we simulate this by adding noise to the conserved subgraph (Figure 3).
 We show that even for networks where the conserved subgraph is weakly conserved, there exists a 'sweet spot' where fusion regression improves network recovery.
-
-One way of assessing the effect of fusion on network inference performance is to compute an ``exchange-rate'' between conditions in each network.
-Suppose you are interested in determining a regulatory network for species A, a model organism which is experimentally inconvenient.
-A related species, B, is easier to collect data for, but of less intrinsic interest.
-Our simulation results suggest that adding conditions from species B to a fused network inference problem has an effect on network A recovery that is similar to adding conditions from A at a reduced rate.
 
@@
 Our method learns networks simultaneously, using predictions about conservation of TF-gene interactions to share information between sources.
-Our approach is useful for learning networks from similar sources such as different cell-lines from the same species, where there exists a one-to-one mapping of genes, as well as datasets where the networks are expected to be more different than similar, where there exists limited orthology mapping.
+Our approach is useful for learning networks from similar sources such as different cell-lines from the same species, where there exists a one-to-one mapping of genes, as well as datasets where the networks are expected to be more different than similar, and there exists limited orthology mapping.
 For the latter case, we are interested in knowing if performance gains from fused regression are limited to those interactions which have fusion constraints.
 To test this, we used sets of 20 TF by 200 gene synthetic networks, with varying proportions of orthology information about TFs and genes.
-We varied lamS and computed AUPR on the portion of the network that had fusion constraints, the portion where TF-gene interactions from one network could be mapped to putative analogous interactions in the second network based off of orthology; the unconstrained subnetwork; and the whole network.
+We varied lamS and computed AUPR on the portion of the network that had fusion constraints, the portion where TF-gene interactions from one network could be mapped to putative analogous interactions in the second network based off of orthology; the unconstrained subnetwork; and the whole network (Figure 5).
 As expected, performance improved as the fusion penalty weight increased.
@@
 We implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (Figures 5-6).
-Pairs of interactions which are very dissimilar even after fusion, which sit in the flat portion of this penalty function, are effectively ``unfused''.
-Because our network inference problem is underdetermined, meaning that multiple networks would fit our data equally well, the ``unfusing'' of constraints is convincing evidence for the rejection of a network which uses orthology prediction to predict similarity in interaction weights.
+Pairs of interactions which are very dissimilar even after fusion, which sit in the flat portion of this penalty function, are effectively ``unfused,'' and no further penalty is incurred as differences in interaction weights grow.
+Because our network inference problem is underdetermined, meaning that multiple networks would fit our data equally well, the ``unfusing'' of constraints is convincing evidence for the rejection of a network which uses orthology to predict similarity in interaction weights.
 ''Unfusing'' suggests neofunctionalization of one or more of the genes involved in this pair of interactions.
@@
 We include, along with our B.
-subtilis dataset, a previously published dataset containing 269 samples covering 104 conditions \cite{nicolas2012condition}.
+subtilis dataset, a previously published dataset containing 269 samples covering 104 conditions, obtained using a different microarray and different strain of B.
+subtilis \cite{nicolas2012condition}.
 We compare performance when learning the networks separately, learning the networks separately followed by rank combining, and learning the networks simultaneously using fusion regression, and show improvement on network inference using our fused L2 approach.
+Importantly, our approach simultaneously learns multiple networks, rather than learning networks separately than combining into one.
 
@@
 When we incorporate TFA and repeat our experiment combining B.
-subtilis datasets, we see a marked improvement in performance.
+subtilis datasets, we see a marked improvement in performance from using TFA.
+Our gains from using fused regression are preserved and even increased with the use of TFA.
+
+\section{Discussion}
+The challenge of learning genome-wide transcriptional regulatory networks became possible with the development of assays such as microarray and RNAseq, and the integration of other data types including ChIP SEQ, have improved the quality of GRNs.
+
+We present a method to simultaneously learn related networks by
+
+We present a method for network inference which allows the principled combination of relevant and related data.
+
+As technologies probing gene expression as well mapping chromatin features mature, the potential to create genome-wide maps of gene interactions becomes more realistic.
+The ability to combine diverse data types, becomes
 

### 2015-12-05 paper1.tex
hi

@@
 
-\title{Fused regression for network inference in multiple species}
+\title{Fused regression for multi-source network inference}
 \author{Kari Y. Lam, Zachary M. Westrick, Christian L. M\"{u}ller, Lionel Christiaen, Richard Bonneau}
@@
 \begin{abstract}
-Learning gene regulatory networks is essential to our understanding of cell behavior and response to external stimuli.
-Methods for network inference from large datasets have been developed and used for a wide variety of species; however, most approaches consider the problem as independent in each species even though there is known to be conservation of gene regulation even in distantly related species.
-We introduce a method that allows data to be pooled across species in a principled way based on gene orthology and demonstrate that it improves network inference even when the orthology is incomplete.
+Understanding gene regulatory networks is critical to learning how cells regulate their behavior and respond to external stimuli.
+To this end, methods for global network inference have been developed and applied to a variety of species.
+However, most approaches consider the problem of network inference independently in each species, despite evidence that gene regulation is conserved even in distantly related species.
+We introduce a method for multi-source network inference that allows simultaneous estimation of gene regulatory networks in multiple species or biological processes through the incorporation of priors based on known gene relationships such as orthology.
+This approach improves network inference performance even when the orthology mapping in incomplete.
 We further refine this method with an algorithm that attempts to learn the true conserved subnetwork from a larger set of potentially conserved interactions.
-We demonstrate the utility of our method not only in cross species network inference, but also in merging data sets and incorporating binding similarity predictions.
+We demonstrate the utility of our method not only in cross species network inference, but also in merging data collected on different experimental platform and in incorporating binding similarity predictions.
 \end{abstract}
@@
 Large-scale data collection efforts have contributed to the development of high quality networks, but there remain many biological processes and organisms for which there does not exist adequate data.
-Furthermore, as new technologies are developed and some old ones are replaced, such as RNAseq and microarray, it becomes important to be able to combine data from multiple platforms, lest we lose valuable information from previous studies.
+Furthermore, as new technologies are developed and some old ones are replaced, such as RNAseq and microarray, it becomes important to be able to combine data from multiple platforms, lest we lose valuable information from existing studies.
 The problem of inferring related -- but not necessarily identical -- structure from related -- but not identical -- data is ubiquitous in biology.
-Multi-source network inference has applications for learning networks in multiple related species, for learning networks associated with distinct processes within the same species, and for learning networks based on heterogenous data sources.
+Multi-source network inference has applications for learning multiple networks in multiple related species, for learning networks associated with distinct processes within the same species, and for learning networks based on heterogenous data sources.
 
-We present two methods for network inference based on linear estimates of gene expression dynamics, extending existing linear dynamical-systems methods for network inference.
-The core of both methods is the observation that biological information about the relatedness of two networks can be used to approximate which network coefficients should be similar to one another in a multi-source network inference problem (ie orthologous TFs should regulate orthologous genes), and that these constraints can be efficiently represented as penalties in a least-squares regression problem.
-Our first method -- fused ridge -- method uses an L2 penalty on the differences between a priori similar interactions, and is useful where the relationships between networks (similarity of genes between data sets) is reliable.
+Moreover, as it becomes increasingly possible to learn genome-wide regulatory networks, we can begin to compare and to test whether there is conservation of networks across species and biological processes.
+Our use of model organisms to study biological processes and diseases relevant to humans relies on the assumption of conservation; yet this has not been tested at the genome scale.
+
+We present two methods for network inference based on linear estimates of gene expression dynamics, extending existing linear dynamical-systems methods for network inference \cite{bonneau_predictive_2007, arrieta-ortiz_experimentally_2015}.
+The core of both methods is the observation that biological information about the relatedness of two networks can be used to constrain which network coefficients should be similar to one another in a multi-source network inference problem (ie orthologous TFs should regulate orthologous genes), and that these constraints can be efficiently represented as penalties in a least-squares regression problem.
+Our first method -- fused ridge -- uses an L2 penalty on the differences between a priori similar interactions, and is useful where the relationships between networks (similarity of genes between data sets) is reliable.
 In the case where both networks contain an identical set of genes and TFs, this approach can be thought of as parametrically interpolating between treating the data sources separately and combining them together.
@@
 Our second method -- adaptive fusion -- uses a non-concave saturating fusion penalty to simultaneously learn the constrained networks and to learn which constraints should be relaxed (ie which parts of the network are genuinely different).
+With this approach, we seek to test the hypothesis of conservation of interactions across networks.
 
-In the case of multiple species, there has been a great deal of work showing that functional conservation exists in gene regulatory networks even across large evolutionary distance.
-We make the assumption that if a pair of TFs are orthologous to one another, then their regulatory effects on a pair of orthologous genes should be similar.
-This orthology information forms the basis of a large number of constraints on the network inference problems, which allow data from one organism to improve network inference on another organism.
-This very general framework for multi-species network inference can be extended to large numbers of distinct organisms, each contributing data with only partially overlapping sets of genes.
-We use our knowledge about the relationship between nodes across networks in order to constrain the network inference problem; this approach has many applications, from improving network inference of multiple species to combining data sets to leveraging knowledge about similarity in binding sites.
+In the case of multiple species, there has been a great deal of work showing that functional conservation exists in gene regulatory networks even across large evolutionary distance \cite{satou2006gene, hinman2009evolution,tanay2005conservation,erwin2009evolution}.
+In our fused L2 approach, we make the assumption that if a pair of TFs is orthologous to one another, then their regulatory effects on a pair of orthologous genes should be similar.
+These orthology relationships form the basis of a set of constraints which favor -- but do not require -- networks in which orthologous transctiption factors regulate orthologous genes.
+As a result, data in one species can improve network inference performance in another species  (and vice versa).
+This general framework for multi-species network inference can be extended to an arbitrary number of distinct organisms, each contributing data with only partially overlapping sets of genes.
+This is an advantage over existing approaches to multi-species network inference, which infer only a sub-network for which orthologs exist in every species \cite{joshi_multi-species_2015}.
+This approach of introducting constraints on the similarity between specific regulatory interactions can be extended beyond the case of multi-species network-inference from orthology; any biological prior on similarity can be used in place of orthology.
+For example, we can introduce constraints that favor genes in the same operon or having similar binding sites, having similar regulators.
 
@@
 Sequence similarity is often a useful predictor of functional similarity \cite{wilson_assessing_2000}.
-Our fused L2 method minimizes a cost function that strives to simultaneously fit expression data in each species and produce networks that are consistent with evolutionary constraints created using orthology.
+In multi-species network inference, our fused L2 approach minimizes a cost function that strives to simultaneously fit expression data in each species and produce networks that are consistent with evolutionary constraints created using orthology.
 However, we recognize that some genes may have taken on different functions and therefore may have new regulatory interactions.
@@
 \subsection{Gene regulatory network}
-We represent gene regulatory networks as a matrix of weights, where rows index transcription factors (TFs) and columns index genes.
-The weight in a given position represents the regulatory weight of that TF on that gene; positive weights represent activation, negative weights represent repression, and 0 weights represent the absence of an interaction.
-We begin with an existing framework for estimating the interaction weights in a gene regulatory network (GRN).
-In the Inferelator algorithm, linear differential equations are used to model changes in gene expression \cite{bonneau_inferelator:_2006-1} The rate at which the observed mRNA expression of gene $i$, $x_i$, changes is governed by degradation of existing transcripts with rate $\alpha$ plus a linear combination of transcription factor (TF) expressions.
+We model each gene's expression as a linear combination of transcription factors (TFs), and we seek to identify the TFs which regulate each gene, and the strength of the TF-gene interactions.
+In the Inferelator algorithm, linear differential equations are used to model changes in gene expression \cite{bonneau_inferelator:_2006-1} Our primary data for learning gene regulatory networks is expression data, time-series and steady state experiments.
+We use time-series data to approximate the rate of change of expression, and we treat steady-state experiments as conditions that have reached equilibrium.
+The rate at which $x_{i}$, the observed mRNA expression of gene $i$, changes, is governed by degradation of existing transcripts with rate $\alpha$ plus a linear combination of transcription factor (TF) expressions.
 \begin{equation}
@@
 \end{equation}
-We are interested in learning $\beta$, the matrix defining the influence of each TF on each gene.
+where $\beta_{i,j}$ represents the weight of TF $j$ on gene $i$, and $\alpha$ is the decay rate of gene $i$.
 We fix the decay rate $\alpha$ for all genes, and set it assuming a time-constant of 10 minutes \cite{hambraeus_genome-wide_2003, selinger_global_2003}, as in \cite{greenfield_robust_2013}.
@@
 Given time-series data on the expression of gene $i$ at timepoints $t_k$ and $t_{k+1}$, we can approximate the rate of change of $x_i$ as $x_i'(t_k)=\frac{x_i(t_{k+1})-x_i(t_k)}{t_{k+1}-t_k}$.
-We can treat steady-state data as having a derivative of zero.
+We treat steady-state data as having a derivative of zero.
 This gives us, for each gene $i$ and time $t_{k}$ an equation
@@
 \end{equation}
+
+We are interested in learning $\beta$, the matrix representation of the gene regulatory network, where the weight in a given position represents the regulatory weight of a TF on a gene.
+Positive weights represent activation, negative weights represent repression, and 0 weights represent the absence of an interaction.
+
 \noindent We approach the learning of the $\beta$ matrix using regression.
@@
 We hypothesize that related species should be governed by similar but not necessarily identical gene regulatory networks.
-We approach the problem of cross-species GRN inference by introducing constraints into the above regression formulation to penalize differences between interaction weights that are expected to be similar based on gene orthology.
+We approach the problem of cross-species GRN inference by introducing constraints into the above regression formulation to penalize differences between interaction weights that are expected to be similar based on, say, gene orthology.
 We can then solve the penalized regression problems simultaneously, in order to obtain a GRN for each species.
@@
 \begin{equation}
-\argmin_{(\beta^A, \beta^B)} \displaystyle\sum_{S \in (A, B)} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\beta^{S_1}_{g,k} \approx \beta^{S_2}_{h,l}} p_{\lambda_S}(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
+%\argmin_{(\beta^A, \beta^B)} \displaystyle\sum_{S \in (A, B)} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\beta^{S_1}_{g,k} \approx \beta^{S_2}_{h,l}} p_{\lambda_S}(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
+\argmin_{(\beta^A, \beta^B)} \displaystyle\sum_{S \in (A, B)} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\substack{(g,h) \in orth,\\
+ (h,l) \in orth}}x p_{\lambda_S}(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
 \end{equation}
@@
 
-When the penalty $p_{\lambda_S}$ is the L2 norm, the entire problem can be solved through linear regression with a suitable design matrix.
+\subsection{Fused L2 penalty for multi-source network inference}
+When the penalty $p_{\lambda_S}$ is the L2 norm, as in our fused L2 approach:
+\begin{equation}
+\argmin_{(\beta^A, \beta^B)} \displaystyle\sum_{S \in (A, B)} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \lambda_S \sum_{\substack{(g,h) \in orth,\\
+ (h,l) \in orth}} \vert \vert (\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l}) \vert \vert ^2
+\end{equation}
+
+the entire problem can be solved through linear regression with a suitable design matrix.
 The regulators of a pair of genes can be solved independently as long as the pair is not linked by fusion constraints, or a chain of fusion constraints; thus the problem of cross-species GRN inference decomposes into a large number of manageable regression problems as long as the orthology mappings are sparse.
 
-When there is a one-to-one orthology between the species being considered (ie different cell-lines of the same organism), the choice of $\lambda_S$ allows one to interpolate between fitting each network independently ($\lambda_S=0$) and pooling data together as if it came from one source ($\lambda_S=\inf$).
+In order to better understand this, consider the case where there is a one-to-one orthology between the species being considered (ie different cell-lines of the same organism).
+The choice of $\lambda_S$ allows one to interpolate between fitting each network independently ($\lambda_S=0$) and pooling data together as if it came from one source ($\lambda_S=\inf$).
 Part of the appeal of the approach, however, is that it allows pooling of data even when there is incomplete orthology.
@@
 \subsection{Solving fused L2 problems using augmented matrices}
-The problem of solving a multi-valued regression problem can be converted into the problem of solving a single valued regression problem by vectorization.
+We use the fact that ridge regression can be solved by ordinary least squares regression on an augmented dataset, by appending a scaled identity matrix to the design matrix, and zeros to the response matrix.
+The problem of solving a multi-valued regression problem can also be converted into the problem of solving a single valued regression problem by vectorization.
+
 Specifically, if $Y \in R^{N \times M}$ then we form $A =$ diagonally concatenate $X$, $M$ times (with the remaining entries 0),
@@
 %We can then define $W_{i,j} = \beta_{(j-1) \times N + i, (j-1) \times M + j}$ which is the desired form.
-This is equivalent to the original problem due to the block structure of matrix multiplication
+This is equivalent to the original problem due to the block structure of matrix multiplication.
+In an ordinary ridge regression problem, this is unnecessary, as the columns of the weight matrix can be solved independently.
+However, in fused regression, information is shared between columns, and therefore they must be solved for simultaneously.
+If two entries are in the same fusion constraint, or if there exists a constraint chain containing entries from two columns, those columns must be solved together.
+Two entries may be fused if say, two TFs are orthologous, and two genes are orthologous--then, the interaction between the first TF, gene pair should be similar to the second TF, gene pair.
+A constraint chain can exist if there are complex orthology mappings, say, if there are multiple similar sequence signals, and three or more entries are fused.
 
-Given $X_1... X_S$ matrices, $X_i \in R^{N_i \times K_i}$ we define $N = \sum_{i=0}^S N_i$, $K = \sum_{i=0}^SK_i$ and $X \in R^{N \times K}$ by diagonally concatenating $X_1...X_S$.
-In this new matrix, the first $K_1$ features correspond to features of $X_1$, the next $K_2$ features the features of $X_2$ etc.
-Similarly, the first $N_1$ samples correspond to samples of $X_1$, the next $N_2$ to $X_2$ etc.
-Solving the regression problem associated with each source is equivalent to solving the problem $\underset{W}{argmin}||XW-Y||$ where $Y$ is defined in the same way as $X$.
-The weight matrix $W$ has a block diagonal structure:
+%Given $X_1... X_S$ matrices, $X_i \in R^{N_i \times K_i}$ we define $N = \sum_{i=0}^S N_i$, $K = \sum_{i=0}^SK_i$ and $X \in R^{N \times K}$ by diagonally concatenating $X_1...X_S$.
+%In this new matrix, the first $K_1$ features correspond to features of $X_1$, the next $K_2$ features the features of $X_2$ etc.
+%Similarly, the first $N_1$ samples correspond to samples of $X_1$, the next $N_2$ to $X_2$ etc.
+%Solving the regression problem associated with each source is equivalent to solving the problem $\underset{W}{argmin}||XW-Y||$ where $Y$ is defined in the same way as $X$.
+%The weight matrix $W$ has a block diagonal structure:
 
@@
 The fused multi-source regression problem can solved by constructing a multi-source design matrix $X$ and data matrix $Y$, and then vectorizing.
-The regression problems associated with two columns $g,h$ of $Y$, where $g$ came from $Y_1$ and $h$ came from $Y_2$, are independent, unless there is a constraint chain of length $L$ with the form (where $x$ is arbitrary)
+Suppose we have a constraint with the form
+
+%The regression problems associated with two columns $A,A'$ of $Y$, where $A$ came from $Y_1$ and $A'$ came from $Y_2$, are independent, unless there is a constraint chain of length $L$ with the form
 
 %$$(G, j_0, g, J_1, j_1, m_1) \rightarrow (J_1, j_1, m_1, J_2, j_2, m_2) \rightarrow (J_2,j_2,m_2,J_3,j_3,_m3)...(J_P, j_P, m_P, H, j_{P+1}, h)$$
-$$(Y_1, j_1, g, Y_2, j_2, h) \rightarrow (Y_2, j_2, h, Y_3, j_3, k)$$
+%$$(Y_1, j_1, g, Y_2, j_2, h) \rightarrow (Y_2, j_2, h, Y_3, j_3, k)$$
+$((\beta_1, A, B), (\beta_1, A, C), (\beta_2, A', B'))$
 
-where the $j$s refer to transcription factors.
+Here, we have two datasets, which we refer to as 1 and 2.
+We have information about the similarity of TF A in network 1 to TF A' in network 2, and about the similarity of genes B and C in network 1 to gene B' in network 2.
+We expect, therefore, that if an interaction exists between A and B or A and C in the first network, that these interactions be similar to the interaction between A' and B' in the second network, and we penalize solutions which do not have this feature.
+
+To find all such linked problems, we use depth-first search to identify linked columns of the TF expression matrix $Y$, then solve the fused regression problem directly by vectorizing corresponding columns.
+To incorporate the fusion penalty, we append to the resulting block diagonal design matrix one row per connected constraint, which contains $\lambda_S$ in the column associated with the TF, gene pair (A, B) from dataset 1, and $-\lambda_S$ to the column associated with TF, gene pair (A', B') from dataset 2, as well as another one for (A, C), (A', B').
+The vector $Y$ is augmented by appending an appropriate number of zeros.
+This results in the addition of the penalties, $\lambda_S||b_{A,B} - b_{A',B'}||^{2}$ and $\lambda_S||b_{A,C} - b_{A',B'}||^{2}$ to the squared error $||AW-y||^2$.
+
+In most cases, we have found the direct solve using augmented matrices to be adequate.
+However, there are certainly use cases of fused regression where solving the fused regression problem directly, even after factoring constraints, involves at least one inversion of an $p\sum_{i=0}^S K_i \times p \sum_{i=0}^S K_i$ matrix where $K_i$ is the number of features in each network and $p$ is the size of a column subset such that there is a constraint chain linking an element of each column to an element in every other column in that subset, taking $O((p\sum_{i=0}^S K_i)^3)$.
+Although the sparse structure of $X_P$ makes this somewhat more tractable, it is still impractical for large chains.
+
+%where the $j$s refer to transcription factors.
 In subproblem 1, $j_1$ regulates gene $g$; $j_1$ is orthologous to $j_2$, a transcription factor of subproblem 2, which regulates $h$, a gene which is orthologous to gene $g$.
 There exists a chain because ($j_1$, $g$) is a TF, gene pair which is fused with ($j_2$, $h$), which is fused to ($j_3$, $k$).
-We use depth-first search to identify linked columns of $Y$, and then solve the fused regression problem directly by vectorizing those columns.
-In order to incorporate the fusion penalty, we append to the resulting design matrix one row \emph{per connected constraint}, which contains $\lambda_S$ in the column associated with $Y_g,i,j,$ and $-\lambda_S$ to the column associated with $Y_h, i',j'$ for constraint $(Y_g,i,j,Y_h,i',j')$.
+%We use depth-first search to identify linked columns of $Y$, and then solve the fused regression problem directly by vectorizing those columns.
+%In order to incorporate the fusion penalty, we append to the resulting design matrix one row \emph{per connected constraint}, which contains $\lambda_S$ in the column associated with $Y_g,i,j,$ and $-\lambda_S$ to the column associated with $Y_h, i',j'$ for constraint $(Y_g,i,j,Y_h,i',j')$.
 Call this resulting matrix $X_P$.
-The vector $Y$ is augmented by appending an appropriate number of zeros to produce $Y_P$.
+%The vector $Y$ is augmented by appending an appropriate number of zeros to produce $Y_P$.
 As a result, each entry in the output of $X_P b$ in the columns associated with each constraint contain $\lambda_S(b_{i,j} - b_{i',j'})$.
-These entries each contribute  $\lambda_S||b_{i,j} - b_{i',j'}||^{2}$ to the squared error $||X_Pb - Y_P||_2^{~2}$
+%These entries each contribute  $\lambda_S||b_{i,j} - b_{i',j'}||^{2}$ to the squared error $||X_Pb - Y_P||_2^{~2}$
 
-In the context of regression for GRN inference with multiple cell sources, each gene $j$ from source $1$ has constraints $\forall_{i \in \{TFs\}} (1, i, j, 2, i’, j’)$ where $i',j'$ are the same gene in a different cell line.
-As a result, each gene has only one constraint and the chain has length $L=2$.
+%In the context of regression for GRN inference with multiple cell sources, each gene $j$ from source $1$ has constraints $\forall_{i \in \{TFs\}} (1, i, j, 2, i’, j’)$ where $i',j'$ are the same gene in a different cell line.
+%As a result, each gene has only one constraint and the chain has length $L=2$.
 %In general, however, solving the fused regression problem directly, even after factoring constraints, involves at least one inversion of an $L\sum_{i=0}^S K_i \times L \sum_{i=0}^S K_i$ matrix, taking $O((L\sum_{i=0}^S K_i)^3)$.
-In general, however, solving the fused regression problem directly, even after factoring constraints, involves at least one inversion of an $p\sum_{i=0}^S K_i \times p \sum_{i=0}^S K_i$ matrix where $p$ is the size of a column subset such that there is a constraint chain $L$ linking an element of each column to an element in every other column in that subset, taking $O((p\sum_{i=0}^S K_i)^3)$.
-Although the sparse structure of $X_P$ makes this somewhat more tractable, it is still impractical for large $L$.
+%In general, however, solving the fused regression problem directly, even after factoring constraints, involves at least one inversion of an $p\sum_{i=0}^S K_i \times p \sum_{i=0}^S K_i$ matrix where $p$ is the size of a column subset such that there is a constraint chain $L$ linking an element of each column to an element in every other column in that subset, taking $O((p\sum_{i=0}^S K_i)^3)$.
+%Although the sparse structure of $X_P$ makes this somewhat more tractable, it is still impractical for large $L$.
 
@@
 
-When the number of genes linked by fusion constraints is relatively small, the augmented design matrix approach we have described has the advantage of being efficiently solvable with a closed form solution.
-However, in the general case the augmented system of equations may be too large to solve efficiently with standard methods.
+When the number of genes in each fusion constraints is relatively small, the augmented design matrix approach we have described has the advantage of being efficiently solvable with a closed form solution.
+However, in some cases the augmented system of equations may be too large to solve efficiently with standard methods.
 To address this limitation, we developed an iterative solver that uses coordinate-wise descent to solve for solutions corresponding to a sequence of values of fusion penalty weights; since our fused L2 method uses a convex and differentiable penalty function, this approach converges to a global minimizer.
@@
 \begin{equation}
-\text{argmin}_{\beta^S(t) : S_i, S_j \in \text{species}} \displaystyle\sum_{S_i, S_j \in \text{species}} ||X^{S_i}\beta^{S_i}(t) - Y^{S_i}||^2 + \lambda_R||\beta^{S_i}(t)||^2 + \displaystyle \sum_{\beta_{g,k}^{S_i} \approx \beta_{h,l}^{S_j}} \lambda_S(\beta^{S_i}_{g,k}(t) - \beta_{h,l}^{S_j}(t-1)^2)
+\text{argmin}_{\beta^{S_i(t)}, \beta^{S_j(t)}} \displaystyle\sum_{S_i, S_j \in \text{species}} ||X^{S_i}\beta^{S_i}(t) - Y^{S_i}||^2 + \lambda_R||\beta^{S_i}(t)||^2 + \displaystyle \sum_{\beta_{g,k}^{S_i} \approx \beta_{h,l}^{S_j}} \lambda_S(\beta^{S_i}_{g,k}(t) - \beta_{h,l}^{S_j}(t-1)^2)
 \end{equation}
@@
 Although we have not produced bounds on the convergence rate, in practice a small number of iterations are necessary.
+
+The ierative solver is used for multi-source regression problems with complex orthology/similarity mappings, and also for solving the regularization path in order to pick regularization and fusion penalty weights.
 
@@
 We use Tibshirani's cyclical coordinate descent algorithms from the 'glmnet' package \cite{friedman_regularization_2010} to compute a ridge regularization path.
-
-\subsection{Optimization}
-Because jointly optimizing $\lambda_R$ and $\lambda_S$ is computationally prohibitive, we opted to optimize the parameters separately.
+As jointly optimizing $\lambda_R$ and $\lambda_S$ is computationally prohibitive, we opted to optimize the parameters separately.
 First, we use glmnet to compute the regularization path for the ridge penalty, and given a range of $\lambda_R$ values, we fit models for each $\lambda_R$.
@@
 \subsection{Adaptive fusion}
-Fusion constraints are intended to penalize dissimilarity between interactions thought to be evolutionarily analogous.
+Fusion constraints are intended to penalize dissimilarity between interactions thought to be analogous based on some a priori knowledge.
+For example, orthology can be used to predict which interactions will be similar across species.
 Because fusion constraints are L2, interaction weights which differ from each other by a large amount are excessively penalized, which effectively ensures that fused interactions are assigned similar weights.
@@
 The principle difference between the penalty we adopt and SCAD/MCP is that both of these penalties are L1 like at the origin, producing sparse solutions.
-Because we are penalizing differences in interaction weights, rather than the weights themselves, there's no particular reason to prefer sparsity.
+Some network inference approaches use L1 penalties to produce sparse networks, on the basis that biological networks are thought to be sparse.
+However, since we are penalizing differences in interaction weights, rather than the weights themselves, there's no reason to assume that most differences will be exactly zero.
 
@@
 
-We also implement MCP:
+%We also implement MCP:
 
-\begin{equation}
-p_{\lambda,\gamma}(\theta) = \left\{
-    \begin{array}{lr}
-    \lambda\theta^2-{ \theta^2 \over 2\gamma} & \text{if } \theta \leq \gamma\lambda\\
-    {1 \over 2}\gamma\lambda^2 & \text{if } \theta > \gamma\lambda
-    \end{array}
-    \right. 
-\end{equation}
+%\begin{equation}
+%p_{\lambda,\gamma}(\theta) = \left\{
+%    \begin{array}{lr}
+%    \lambda\theta^2-{ \theta^2 \over 2\gamma} & \text{if } \theta \leq \gamma\lambda\\
+%    {1 \over 2}\gamma\lambda^2 & \text{if } \theta > \gamma\lambda
+%    \end{array}
+%    \right. 
+%\end{equation}
 
-With derivative
+%With derivative
 
-\begin{equation}
-p'_{\lambda,\gamma}(\theta) = \left\{
-    \begin{array}{lr}
-    2\lambda\theta - {\theta \over \gamma} & \text{if } \theta \leq \gamma\lambda\\
-    0 & \text{if } \theta > \gamma\lambda
-    \end{array}
-    \right.
-\end{equation}
+%\begin{equation}
+%p'_{\lambda,\gamma}(\theta) = \left\{
+%    \begin{array}{lr}
+%    2\lambda\theta - {\theta \over \gamma} & \text{if } \theta \leq \gamma\lambda\\
+%    0 & \text{if } \theta > \gamma\lambda
+%    \end{array}
+%    \right.
+%\end{equation}
 
@@
 Our adaptive penalty function introduces, in addition to regularization and fusion penalty weights $\lambda_R$ and $\lambda_S$, an unknown parameter $a$.
-We could employ grid search using cross-validation to searach for the best parameters, but this can be computationally expensive.
+We could employ grid search using cross-validation to searach for the best parameters, but for many data sets, this can be computationally expensive.
 Moreover, we are primarily interested in using this saturating penalty as a way of testing the hypothesis that conservation in GRNs can be predicted based off of known similarities between genes.
@@
 \section{Discussion}
-The challenge of learning genome-wide transcriptional regulatory networks became possible with the development of assays such as microarray and RNAseq, and the integration of other data types including ChIP SEQ, have improved the quality of GRNs.
+We present two methods for the joint learning of gene regulatory networks.
+One approach, fused L2, takes advantage of known similarities in genes across networks, to add constraints to an underdetermined system.
+We can use evolutionary information to predict functional similarity or chromatin features or accessibility to identify binding site similarity.
+The ability to incorporate multiple data sets describing related processes, as well as multiple data types, in a principled manner, helps us take advantage of the breadth of experimentation in biology to better approximate GRNs.
 
-We present a method to simultaneously learn related networks by
+Our adaptive fusion approach allows the testing of the hypothesis that networks consistent with fusion constraints are appropriate.
+Like the fused L2 approach, this method attempts to solve for networks which minimize a penalty function fitting the expression data and the fusion constraints.
+In the case of cross-species network inference, evolutionary links may be useful to guide the sharing of data.
+However, orthology may not always accurately predict functional similarity, and a fusion constraint encouraging orthologous gene, TF pairs to share similar interactions across species may not be supported by expression data.
+Then, when the fusion penalty is at odds with the penalty fitting expression data, there exists a tension between satisfying evolutionary constraints while accurately fitting experimental data.
+Our saturating penalty relaxes the fusion penalty, giving evidence that the fusion constraint is not predictive of gene interaction.
+We may use the analogy of a rubber band, which can become stretched when penalty functions within our objective function oppose each other; the adaptive fusion approach allows the snapping of this once the opposition is great enough.
+For the multiple species case, this represents orthologs which do not share similar interactions.
+When jointly learning networks describing processes in different cell lines, this may identify interesting context-specific behavior.
+Genes may be fused together on the basis of similar binding sites or chromatin features, and the relaxing of the fusion penalty indicates divergence of gene function.
+Adaptive fusion, therefore, is a tool for network inference as well as a method for testing network conservation.
 
-We present a method for network inference which allows the principled combination of relevant and related data.
-
-As technologies probing gene expression as well mapping chromatin features mature, the potential to create genome-wide maps of gene interactions becomes more realistic.
-The ability to combine diverse data types, becomes
+One future aim is the ability to pick the $a$ parameter controlling the saturation point of the adaptive fusion penalty automatically from the expression data; this parameter is biologically interesting and indicates the difference in interaction weights which is great enough that the fusion constraints should be relaxed.
+An exhaustive search through different combinations of putative fusion constraints is prohibitive, but perhaps there is another way which avoids this.
 

### 2015-12-06 paper1.tex
hi

@@
 Therefore, we implement a way of picking $a$ with a user-selected percentile of differences in interaction weights, which represents the user's hypothesis for the proportion of evolutionary constraints which should be relaxed.
-A gold standard can also be used for selecting $a$ by comparing differences in weights for truly analogous interactions with those of randomly paired interactions, and selecting the difference in interaction weights at which no further fusion penalty is incurred.
 
@@
 \subsection{Simulated data}
+We generate simulated data to evaluate the ability of our fused L2 approach to learn the true network and to show that sharing information between similar but not identical data sources results in more accurate network recovery.
+Because the purpose of adaptive fusion is to learn the structure of conservation while also learning networks, using simulated data allows us to test the method relative to a known gold standard for network conservation.
+Our simulations are based on the model $Y=X\beta+\sigma \epsilon$ where $\epsilion \sim N(0,1)$. Each data set consists of a training set and test set.
+
 Generation of simulated data begins with the production of random orthology mappings.
@@
 For the purposes of evaluating simulated network recovery, we define a gold standard network as the support of the beta matrices.
-Priors used in network inference are interactions from the gold standard.
-The list of priors can be be manipulated to include false positives and false negatives as with the generation of orthologs.
+Priors used in network inference are interactions present in the gold standard.
+The list of priors can be be manipulated to include false positives and false negatives as with ortholog mapping.
 

### 2015-12-07 paper1.tex
zack edits

@@
 However, we recognize that some genes may have taken on different functions and therefore may have new regulatory interactions.
-Gene duplications and neofunctionalization in multi-species network inference \cite{eisen_phylogenomics:_1998}, and changes in chromatin configuration in multi-cell line network inference \cite{li_role_2007}, are cases where enforcing fusion of constraints does not produce networks which reflect true differences in regulation between networks.
-These cases can be difficult to learn by simply comparing independently inferred networks; because network inference is typically underconstrainted \cite{marbach_revealing_2010-1}, there could be other plausible networks in which the coefficients being compared are similar.
+Gene duplications and neofunctionalization in multi-species network inference \cite{eisen_phylogenomics:_1998}, and changes in chromatin configuration in multi-cell line network inference \cite{li_role_2007}, are cases where assumptions about functional conservation may not hold.
+
+Understanding which interactions are functionally dissimilar is of direct biological interest, but existing approaches to network inference are unable to effectively rule out conservation.
+Observing a large difference in the weights of regulatory interactions obtained though independent inference of multiple networks is the best evidence against conservation available with existing network inference techniques.
+This evidence is very weak, however; because network inference is typically underconstrainted \cite{marbach_revealing_2010-1}, there could be a different set of networks for which conservation does hold, and which fit the data almost as well.
 Our adaptive fusion method allows us to learn differences in networks predicted to be similar based on orthology or gene matching: because the constraints we impose favor weight configurations in which fused interaction weights are similar to one another, a solution in which a priori similar regulatory interactions are found to be different is strong evidence that they really are different.
-When differences in regulatory interactions arise where they are predicted to be similar, the network inference procedure would be improved by relaxing or ignoring these evolutionary constraints.
-However, we don't know a priori which constraints to enforce and which to relax.
 
@@
 \section{Methods}
+Although our approach is generalizable to a wide variety of multi-source network inference problems, we begin with the concrete example of network inference in two related species.
+Our approach to multi-species network inference is based on the hypothesis that gene regulation in related species is governed by similar but not necessarily identical gene regulatory networks, due to conservation of function through evolution.
+We represent conservation of network function by introducing constraints into the objective function for network inference that penalize differences between the weights of regulatory interactions believed to be a priori similar.
+These constraints favor the generation of similar networks for related species, and in the generally under-constrained regime of network inference can improve the accuracy of network recovery.
+We then go on to introduce a method to test the assumption of conserved network structure, and to relax the associated constraints on pairs of interactions for which data does not support conservation.
+
 \subsection{Gene regulatory network}
-We model each gene's expression as a linear combination of transcription factors (TFs), and we seek to identify the TFs which regulate each gene, and the strength of the TF-gene interactions.
-In the Inferelator algorithm, linear differential equations are used to model changes in gene expression \cite{bonneau_inferelator:_2006-1} Our primary data for learning gene regulatory networks is expression data, time-series and steady state experiments.
-We use time-series data to approximate the rate of change of expression, and we treat steady-state experiments as conditions that have reached equilibrium.
+We model the transcription rate of each gene as a weighted sum of transcription factor (TF) expression, and seek to identify the identities and regulatory weights of these TFs.
+This formulation matches that of the existing Inferelator algorithm, which models gene expression with linear differential equations \cite{bonneau_inferelator:_2006-1}.
+Our primary data for learning gene regulatory networks is expression data, consisting of time-series and steady state experiments.
+We use time-series data to approximate the rate of change of expression, and treat steady-state experiments as having reached equilibrium (constant expression).
 The rate at which $x_{i}$, the observed mRNA expression of gene $i$, changes, is governed by degradation of existing transcripts with rate $\alpha$ plus a linear combination of transcription factor (TF) expressions.
@@
 \end{equation}
+for time series
+\begin{equation}
+0 = \sum \beta_{i,j}x_{j}(t_k)
+\end{equation}
+for steady state
 \noindent We can summarize these equations in matrix form as
@@
 
-\noindent We approach the learning of the $\beta$ matrix using regression.
+\noindent The matrix $\beta$ can be solved using linear regression.
 Because there are typically far fewer conditions than possible regressors (TFs), we introduce a ridge regularization constraint with weight $\lambda_R$ and solve
@@
 \end{equation}
-This is similar to the formulation used in the Inferelator algorithm, which we extend to the case of simultaneously inferring the GRNs of multiple species.
+This is similar to the formulation used in the Inferelator algorithm, which we extend to the case of simultaneously inferring multiple networks.
 
 \subsection{Fused gene regulatory networks}
-We hypothesize that related species should be governed by similar but not necessarily identical gene regulatory networks.
-We approach the problem of cross-species GRN inference by introducing constraints into the above regression formulation to penalize differences between interaction weights that are expected to be similar based on, say, gene orthology.
+
+Information about the conserved structure of gene regulation is introduced through the incorporation of constraints into the above regression formulation.
+These constraints penalize differences between interaction weights in the networks of multiple species that are expected to be similar based on prior biological knowledge.
 We can then solve the penalized regression problems simultaneously, in order to obtain a GRN for each species.
@@
 If TF $g^A$ in organism $A$ and TF $h^B$ in organism $B$ are orthologs, and gene $k^A$ and $l^B$ are orthologs, then we expect that the $g^A \rightarrow k^A$ interaction weight should be similar to the $h^B \rightarrow l^B$ interaction weight, and we introduce a fusion constraint between these analogous interactions.
-In terms of the above regression formulation, we expect that $\beta^A_{g,k} \approx \beta^B_{h,l}$, and include a penalty term $p_{\lambda_S}(\beta^A_{g,k} - \beta^B_{h,l})$ to the quantity being minimized in order to ensure similarity.
-The parameter $\lambda_S$ controls how highly to penalize differences between fused coefficients and controls the tradeoff between fitting the expression data and producing a set of networks that conform to evolutionary constraints.
+In terms of the above regression formulation, we expect that $\beta^A_{g,k} \approx \beta^B_{h,l}$, and include a penalty term $p_{\lambda_S}(\beta^A_{g,k} - \beta^B_{h,l})$ in the quantity being minimized in order to encourage similarity.
+The function $p_{\lambda_S}(x)$ controls the shape of the relationship between weight dissimilarity and penalty, while parameter $\lambda_S$ controls the overall scaling of the penalization of differences between fused coefficients.
+$\lambda_S$ controls the tradeoff between fitting the expression data alone and producing a set of networks that conform to evolutionary prior knowledge.
 This gives us the final equation to be minimized
@@
 %\argmin_{(\beta^A, \beta^B)} \displaystyle\sum_{S \in (A, B)} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\beta^{S_1}_{g,k} \approx \beta^{S_2}_{h,l}} p_{\lambda_S}(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
-\argmin_{(\beta^A, \beta^B)} \displaystyle\sum_{S \in (A, B)} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\substack{(g,h) \in orth,\\
+\argmin_{\beta^S} \displaystyle\sum_{S \in \text{species}} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\substack{(g,h) \in orth,\\
  (h,l) \in orth}}x p_{\lambda_S}(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
@@
 where the second sum is over pairs of interactions with fusion constraints.
+In the fused L2 algorithm we develop, the penalty function is equal to the L2 norm of the difference in regulatory weight of fused coefficients.
+That is, $p_{\lambda_S}(x)=x^2$.
+The overall objective function for network inference is then:
 
-\subsection{Fused L2 penalty for multi-source network inference}
-When the penalty $p_{\lambda_S}$ is the L2 norm, as in our fused L2 approach:
 \begin{equation}
-\argmin_{(\beta^A, \beta^B)} \displaystyle\sum_{S \in (A, B)} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \lambda_S \sum_{\substack{(g,h) \in orth,\\
+\argmin_{\beta^S} \displaystyle\sum_{S \in \text{species}} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \lambda_S \sum_{\substack{(g,h) \in orth,\\
  (h,l) \in orth}} \vert \vert (\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l}) \vert \vert ^2
@@
 
-the entire problem can be solved through linear regression with a suitable design matrix.
-The regulators of a pair of genes can be solved independently as long as the pair is not linked by fusion constraints, or a chain of fusion constraints; thus the problem of cross-species GRN inference decomposes into a large number of manageable regression problems as long as the orthology mappings are sparse.
+Because every component of the objective function is an L2 norm, the entire problem is convex, and can be solved through linear regression with a suitable design matrix.
+%The regulators of a pair of genes can be solved independently as long as the pair is not linked by fusion constraints, or a chain of fusion constraints; thus the problem of cross-species GRN inference decomposes into a large number of manageable regression problems as long as the orthology mappings are sparse.
 
-In order to better understand this, consider the case where there is a one-to-one orthology between the species being considered (ie different cell-lines of the same organism).
+In order to better understand the behavior of this objective function, consider the case where there is a one-to-one orthology between the species being considered (ie different cell-lines of the same organism).
 The choice of $\lambda_S$ allows one to interpolate between fitting each network independently ($\lambda_S=0$) and pooling data together as if it came from one source ($\lambda_S=\inf$).
@@
 \subsection{Solving fused L2 problems using augmented matrices}
-We use the fact that ridge regression can be solved by ordinary least squares regression on an augmented dataset, by appending a scaled identity matrix to the design matrix, and zeros to the response matrix.
-The problem of solving a multi-valued regression problem can also be converted into the problem of solving a single valued regression problem by vectorization.
+We begin with the problem of constructing a design matrix to solve a fused L2 regression problem with a single response variable, because problems with multiple response variables and with multiple data sources can be converted into this form through vectorization.
+We then go on to show that, although the vectorized solution involves solving an impractically large system of equations, under typical biological conditions the structure of constraints allow the problem to be broken up into many smaller subproblems.
 
-Specifically, if $Y \in R^{N \times M}$ then we form $A =$ diagonally concatenate $X$, $M$ times (with the remaining entries 0),
-$y =$ vertically concatenate the columns of $Y$. We solve
+We use the fact that ridge constraints can be incorporated into a least-squares regression problem by appending a scaled identity matrix to the design matrix, and a corresponding number of zeros to the response vector.
+Similarly, a fusion constraint $\lambda_S (\beta_{i} - \beta{j})^2$ can be incorporated into a least-squares regression problem by appending a row containing $\sqrt{lambda_S}$ in the $i$th position, $-\sqrt{lambda_S}$ in the $j$th position, and $0$s elsewhere to the design matrix, and zero to the response vector.
+
+In order to convert an optimization over multiple response variables (indexed by $i$) and multiple sources (indexed by $j$) with design matrices $X^{S_j}$ and response variables $y_i^{S_j}$ into an optimization with a single source and response variable, we vectorize as follows.
+Let $G$ be a set containing $(i,j)$ for each of the data sources and response variables currently being optimized.
+We construct the matrix $A$ by diagonally concatenating each of $X^{S_j}$ (with the remaining entries 0) and vertically concatenating each of $y_j^{S_j}$.
+We then solve
+%Specifically, if $Y \in R^{N \times M}$ then we form $A =$ diagonally concatenate $X$, $M$ times (with the remaining entries 0),
+%$y =$ vertically concatenate the columns of $Y$. We solve
 
@@
 This is equivalent to the original problem due to the block structure of matrix multiplication.
-In an ordinary ridge regression problem, this is unnecessary, as the columns of the weight matrix can be solved independently.
-However, in fused regression, information is shared between columns, and therefore they must be solved for simultaneously.
-If two entries are in the same fusion constraint, or if there exists a constraint chain containing entries from two columns, those columns must be solved together.
-Two entries may be fused if say, two TFs are orthologous, and two genes are orthologous--then, the interaction between the first TF, gene pair should be similar to the second TF, gene pair.
-A constraint chain can exist if there are complex orthology mappings, say, if there are multiple similar sequence signals, and three or more entries are fused.
+In an ordinary regression problem each response variable can be solved independently as a large number of small regression problems.
+However, in fused regression, we append additional rows to the design matrix $A$ that link entries of $W$ associated with different response variables.
+As a result, these linked response variables must be solved simultaneously through vectorization.
+Two response variables are linked by a fusion constraint if any of the regulatory weights affecting those genes are linked by a fusion constraint.
+Two response variables must be solved simultaneously if there is any chain of linked response variables connecting them.
+However, every other response variable can be solved separately.
+In biological terms, the regulators of two genes (whether in the same species, or different species) must be solved together if there is a fusion constraint linking those genes' regulators, or if there is a chain of such constraints.
+This is important because vectorization produces a design matrix with a number of rows and columns equal to the sum of the number of rows and columns of the design matrices associated with each of the response variables (genes) being solved for.
+If the networks for a large number of Genes are solved simultaneously, the system of equations can quickly become intractible to solve.
 
-%Given $X_1... X_S$ matrices, $X_i \in R^{N_i \times K_i}$ we define $N = \sum_{i=0}^S N_i$, $K = \sum_{i=0}^SK_i$ and $X \in R^{N \times K}$ by diagonally concatenating $X_1...X_S$.
-%In this new matrix, the first $K_1$ features correspond to features of $X_1$, the next $K_2$ features the features of $X_2$ etc.
-%Similarly, the first $N_1$ samples correspond to samples of $X_1$, the next $N_2$ to $X_2$ etc.
-%Solving the regression problem associated with each source is equivalent to solving the problem $\underset{W}{argmin}||XW-Y||$ where $Y$ is defined in the same way as $X$.
-%The weight matrix $W$ has a block diagonal structure:
-
-%\begin{figure}[h!]
-%    \centering
-%    \includegraphics[scale=.5,trim=0mm 0mm 0mm 0mm,clip]{direct_solve.png}
-%\end{figure}
-
-The fused multi-source regression problem can solved by constructing a multi-source design matrix $X$ and data matrix $Y$, and then vectorizing.
-Suppose we have a constraint with the form
-
-%The regression problems associated with two columns $A,A'$ of $Y$, where $A$ came from $Y_1$ and $A'$ came from $Y_2$, are independent, unless there is a constraint chain of length $L$ with the form
-
-%$$(G, j_0, g, J_1, j_1, m_1) \rightarrow (J_1, j_1, m_1, J_2, j_2, m_2) \rightarrow (J_2,j_2,m_2,J_3,j_3,_m3)...(J_P, j_P, m_P, H, j_{P+1}, h)$$
-%$$(Y_1, j_1, g, Y_2, j_2, h) \rightarrow (Y_2, j_2, h, Y_3, j_3, k)$$
-$((\beta_1, A, B), (\beta_1, A, C), (\beta_2, A', B'))$
-
-Here, we have two datasets, which we refer to as 1 and 2.
-We have information about the similarity of TF A in network 1 to TF A' in network 2, and about the similarity of genes B and C in network 1 to gene B' in network 2.
-We expect, therefore, that if an interaction exists between A and B or A and C in the first network, that these interactions be similar to the interaction between A' and B' in the second network, and we penalize solutions which do not have this feature.
-
-To find all such linked problems, we use depth-first search to identify linked columns of the TF expression matrix $Y$, then solve the fused regression problem directly by vectorizing corresponding columns.
-To incorporate the fusion penalty, we append to the resulting block diagonal design matrix one row per connected constraint, which contains $\lambda_S$ in the column associated with the TF, gene pair (A, B) from dataset 1, and $-\lambda_S$ to the column associated with TF, gene pair (A', B') from dataset 2, as well as another one for (A, C), (A', B').
-The vector $Y$ is augmented by appending an appropriate number of zeros.
-This results in the addition of the penalties, $\lambda_S||b_{A,B} - b_{A',B'}||^{2}$ and $\lambda_S||b_{A,C} - b_{A',B'}||^{2}$ to the squared error $||AW-y||^2$.
-
+In order to avoid this difficulty, we use depth-first search to identify linked columns of each TF expression matrix $Y^{S_j}$, then forming design and response matrices through vectorization.
+We can then incorporate fusion constraints as in the case of single-source single response-variable fused regression.
 In most cases, we have found the direct solve using augmented matrices to be adequate.
-However, there are certainly use cases of fused regression where solving the fused regression problem directly, even after factoring constraints, involves at least one inversion of an $p\sum_{i=0}^S K_i \times p \sum_{i=0}^S K_i$ matrix where $K_i$ is the number of features in each network and $p$ is the size of a column subset such that there is a constraint chain linking an element of each column to an element in every other column in that subset, taking $O((p\sum_{i=0}^S K_i)^3)$.
-Although the sparse structure of $X_P$ makes this somewhat more tractable, it is still impractical for large chains.
+This is because the structure of orthology tends to be sparse, so only a small number of genes must be solved at once.
+However, in the general case of $c_i$ response variables from each of $i \in \text{species}$ data sources the design matrix to be solved is of size $\displaystyle \sum_{c_i} rows(X^{S_i}) \times \displaystyle \sum_{c_i} cols(X^{S_i})$.
+Although the sparse structure of the augmented design matrix makes this somewhat more tractable, it is still impractical for large constraint chains.
 
@@
 On each step, the we compute $\beta$s that minimize a penalized cost function where the fusion penalty encourages similarity of a pair of parameters, using the previous iteration's solution.
-This process is iterated until the estimated $\betas$s converge.
+This process is iterated until the estimated $\beta$s converge.
 Because each iteration reduces the error between $\beta(t)$ and $\beta(t-1)$, and because $\beta(t) = \beta(t-1)$ is the globally optimal solution, this process must eventually converge to the same network as (cite 5).
@@
 Moreover, we are primarily interested in using this saturating penalty as a way of testing the hypothesis that conservation in GRNs can be predicted based off of known similarities between genes.
-Therefore, we implement a way of picking $a$ with a user-selected percentile of differences in interaction weights, which represents the user's hypothesis for the proportion of evolutionary constraints which should be relaxed.
+Therefore, we implement a way of selecting $a$ to be equal to a certain percentile of the distribution of differences in fused interaction weights from the networks fit without fusion.
+This choice represents the user's hypothesis for the fraction of both networks that have unrelated function, and should therefore be unfused.
 
@@
 This scaling is similar to rescaling according to variance explained relative to an augmented design matrix that includes fusion constraints.
-We use an approximation because it is simpler to compute.
-The intuition is that a large $B_{i,j}$ may be necessary because TF $i$ varies little across the available data, and these large weights are not suggestive of a true regulatory interaction.
+The intuition for scaling that a large $B_{i,j}$ may be inferred because of overfitting with a TF $i$ varies little across the available data, or because the regulatory weight is strong.
+Only in the latter case is a large weight suggestive of a true regulatory interaction.
 
@@
 \section{Discussion}
-We present two methods for the joint learning of gene regulatory networks.
-One approach, fused L2, takes advantage of known similarities in genes across networks, to add constraints to an underdetermined system.
-We can use evolutionary information to predict functional similarity or chromatin features or accessibility to identify binding site similarity.
-The ability to incorporate multiple data sets describing related processes, as well as multiple data types, in a principled manner, helps us take advantage of the breadth of experimentation in biology to better approximate GRNs.
 
-Our adaptive fusion approach allows the testing of the hypothesis that networks consistent with fusion constraints are appropriate.
+We present a general framework for the simultaneous inference of gene regulatory networks in multiple sources based on the introduction of constraints that penalize differences in regulatory weights thought to be similar based on prior biological knowledge.
+In order to solve the fused network inference problem, we develop two methods for the joint learning of gene regulatory networks, based on the use of different fusion penalties.
+The first method, fused L2, penalizes the squared difference of the regulatory weights of fused interactions.
+These constraints improve the performance of the underdetermined network inference problem, by allowing information to be shared across sources.
+We apply this algorithm to the problem of network inference in two distantly related biological organisms -- B.
+Subtilis and B.
+Anthracis -- and show that network recovery is improved through the introduction of fusion constraints between pairs of orthologous genes.
+We further demonstrate the viability of fused L2 as a method for combining data from multiple experimental platforms, where fusion is between identical pairs of genes, and as a method for the incorporation of information about operon structure, where fusion reflects the assumption that genes in the same operons have similar regulators.
+Because the algorithm we developed can accomodate constraints between arbitrary pairs of regulatory interactions, any biological prior representing information about expected regulatory similarity can be represented, even if the prior provides no information about the magnitude or direction of regulation.
+The ability to incorporate multiple data sets describing related processes, as well as multiple data types, in a principled manner, helps us take advantage of the breadth of experimentation in biology to better learn the structure of gene regulation.
+This approach is particularly interesting in light of the diversity of model organisms used in modern biology.
+Different model systems provide different advantages and disadvantages for experimental design, but without a principled mechanism for combining data from multiple sources, it is difficult to fully leverage data obtained from even a slightly different model system.
+
+Although it is important to take advantage of the similarities of related organisms for generating improved models of gene regulation, it is also critically important to understand how systems differ from one another.
+Our cross-species network inference method is premised on the assumption that orthologous genes have similar regulators.
+This is a common assumption when comparing model systems, which is difficult to evaluate on a genome wide scale using existing methods; because network inference is typically underconstrained, the presence of a difference between two regulatory interactions does not rule out an equally good network (or pair of networks) in which those regulatory interactions are similar.
+In contrast, because our method incorporates fusion penalties that favor regulatory similarity between interactions thought to be conserved, the failure to find a network in which fused regulators are similar is direct evidence that no such network is possible.
+
+Because our approach allows us to directly test -- and reject -- the hypothesis of conserved function, and because the enforcement of fusion constraints between non-conserved interactions may degrade network inference performance, we introduce an adaptive fusion method that decouples pairs of interactions for which expression data suggests a lack of conservation.
 Like the fused L2 approach, this method attempts to solve for networks which minimize a penalty function fitting the expression data and the fusion constraints.
@@
 
-One future aim is the ability to pick the $a$ parameter controlling the saturation point of the adaptive fusion penalty automatically from the expression data; this parameter is biologically interesting and indicates the difference in interaction weights which is great enough that the fusion constraints should be relaxed.
+Our approaches -- fused L2 and adaptive fusion -- represent a very general framework for simultaneous network inference and the incorporation of structured biological priors.
+These priors -- incorporated into our method as fusion constraints -- allow the use of rich sources of biological knowledge, such as orthology and operon structure, which have informed experimental design, but are typically not incorporated into genome wide network inference algorithms.
+By accomodating the simultaneous inference of multiple related networks, we can improve network inference performance by allowing the efficient reuse of data from similar, but not necessarily identical, sources.
+A method for pooling data from multiple sources holds the promise of vastly expanding the quantity of data available for analysis, particularly in less commonly used model systems.
+At the same time these methods allow us to test our assumptions on how similar biological systems relate to one another, by allowing us to rule out conservation in a principled way, and at a genome-wide scale.
+
+%One future aim is the ability to pick the $a$ parameter controlling the saturation point of the adaptive fusion penalty automatically from the expression data; this parameter is biologically interesting and indicates the difference in interaction weights which is great enough that the fusion constraints should be relaxed.
 An exhaustive search through different combinations of putative fusion constraints is prohibitive, but perhaps there is another way which avoids this.

### 2016-01-09 paper1.tex
zack committing

@@
 
-\title{Fused regression for multi-source network inference}
+\title{Fused regression for multi-source gene regulatory network inference}
 \author{Kari Y. Lam, Zachary M. Westrick, Christian L. M\"{u}ller, Lionel Christiaen, Richard Bonneau}
@@
 We present two methods for network inference based on linear estimates of gene expression dynamics, extending existing linear dynamical-systems methods for network inference \cite{bonneau_predictive_2007, arrieta-ortiz_experimentally_2015}.
-The core of both methods is the observation that biological information about the relatedness of two networks can be used to constrain which network coefficients should be similar to one another in a multi-source network inference problem (ie orthologous TFs should regulate orthologous genes), and that these constraints can be efficiently represented as penalties in a least-squares regression problem.
+The core of both methods is the observation that biological information about the relatedness of genes can be used to constrain which network coefficients should be similar to one another in a multi-source network inference problem (ie orthologous TFs should regulate orthologous genes), and that these constraints can be efficiently represented as penalties in a least-squares regression problem.
 Our first method -- fused ridge -- uses an L2 penalty on the differences between a priori similar interactions, and is useful where the relationships between networks (similarity of genes between data sets) is reliable.
@@
 However, our method allows useful pooling of data even when the overlap between genes is incomplete.
-Our second method -- adaptive fusion -- uses a non-concave saturating fusion penalty to simultaneously learn the constrained networks and to learn which constraints should be relaxed (ie which parts of the network are genuinely different).
+Our second method -- adaptive fusion -- uses a non-concave saturating fusion penalty to simultaneously infer the constrained networks and to learn which constraints should be relaxed (ie which parts of the network are genuinely different).
 With this approach, we seek to test the hypothesis of conservation of interactions across networks.
@@
 
-We test the ability of fused L2 and adaptive fusion to improve network recovery on both synthetic data and Bacillus subtilis and Bacillus anthracis, showing the applicability of our method in combining different datasets, and leveraging similarity across organisms as well as within a network in order to improve network inference.
+We test the ability of fused L2 and adaptive fusion to improve network recovery on both synthetic data and \textit{Bacillus subtilis} and \textit{Bacillus anthracis}, showing the applicability of our method in combining different datasets, and leveraging similarity across organisms as well as within a network in order to improve network inference.
 We explore the circumstances under which each approach is optimal, and evaluate the robustness of adaptive fusion to incorrect orthology, simulating the biologically relevant case of neofunctionalization.
@@
 This is important because vectorization produces a design matrix with a number of rows and columns equal to the sum of the number of rows and columns of the design matrices associated with each of the response variables (genes) being solved for.
-If the networks for a large number of Genes are solved simultaneously, the system of equations can quickly become intractible to solve.
+If the networks for a large number of genes are solved simultaneously, the system of equations can quickly become intractible to solve.
 
@@
 \section{Results}
-We tested the ability of fused regression to improve network inference performance, as well as the ability of adaptive-fusion to learn the conservation of gene interactions.
+We tested the ability of fused regression to improve network inference performance, as well as the ability of adaptive fusion to learn the conservation of gene interactions, using both synthetic networks and real data.
+For the synthetic data, we generated random pairs of networks in which orthologous genes have similar regulatory interactions, then sampled gene expression from these networks, which we used to derive networks for comparison with the true network.
+For real data, we computed recovery of a known gold standard in \textit{Bacillus subtilis}.
 We measured network inference performance with the area under the precision recall curve (AUPR).
@@
 \subsection{Using fused regression to learn related networks}
+It is known that network recovery improves with additional data \cite{bar-joseph_computational_2003}.
+Additional data is not always readily available; but often there is an abundance of related data, and we reason that data describing similar processes can be used to improve our ability to recover true networks.
 Using related data for network inference allows us to borrow statistical power from other sources, effectively increasing the sample size and boosting the sensitivity and specificity of learned interactions.
-We used synthetic data to approximate two related biological processes and show that simultaneously learning the networks describing them using fused L2 regression to share information about related interactions results in an improvement in our ability to recover the true networks.
-Adding conditions from a related dataset has an effect on network inference which is akin to adding experimental data, without doing additional experiments.
+We created synthetic networks that approximate two related biological processes, then evaluated performance of our fused L2 regression, which learns the networks simultaneously, sharing information about related interactions (Figure 1b).
+We compared recovery of the 10 TF by 200 gene networks, using fused L2 versus learning networks separately, using different amounts of simulated expression data from the two species.
+When the amount of data from species two is held constant, increasing the amount of data available for learning the network for species one results in a more accurate network, as expected.
+Using the assumption that similar genes have similar regulatory interactions, we share information about putatively conserved interactions.
+When we increase the amount of data from species two, we get performance gains on network one using fused L2 regression.
 
@@
 
-As expected, when the conserved subgraph is very similar, increasing the weight of the fusion constraint $\lambda_S$ improved network recovery (Figure 2).
-Networks with higher orthology coverage saw a greater benefit from fusion.
+As expected, when the conserved subgraph is very similar - when the interactions in the conserved subnetwork are nearly identical, simulating the case of closely related organisms or similar processes (Figure 2A) - increasing the weight of the fusion constraint $\lambda_S$ improved network recovery.
+As the size of the conserved subgraph increases, using a greater weight on the fusion penalty was beneficial.
 
-When the conservation is weak, and even the portions of the network which are conserved have deviated from each other significantly, performance may not continue to improve by increasing the fusion weight; we simulate this by adding noise to the conserved subgraph (Figure 3).
+To simulate the case of distantly related organisms, we created networks where conserved interactions are drawn from high variance distribution (Figure 2B).
+When the conservation is weak - when the interactions in the conserved subnetwork are nearly as different across species as randomly chosen interactions - performance may not continue to improve by increasing the fusion weight.
 We show that even for networks where the conserved subgraph is weakly conserved, there exists a 'sweet spot' where fusion regression improves network recovery.
@@
 \subsection{Fused regression improves performance on both the constrained and non-constrained parts of the network}
-Our method learns networks simultaneously, using predictions about conservation of TF-gene interactions to share information between sources.
-Our approach is useful for learning networks from similar sources such as different cell-lines from the same species, where there exists a one-to-one mapping of genes, as well as datasets where the networks are expected to be more different than similar, and there exists limited orthology mapping.
-For the latter case, we are interested in knowing if performance gains from fused regression are limited to those interactions which have fusion constraints.
-To test this, we used sets of 20 TF by 200 gene synthetic networks, with varying proportions of orthology information about TFs and genes.
-We varied lamS and computed AUPR on the portion of the network that had fusion constraints, the portion where TF-gene interactions from one network could be mapped to putative analogous interactions in the second network based off of orthology; the unconstrained subnetwork; and the whole network (Figure 5).
-As expected, performance improved as the fusion penalty weight increased.
+Our method learns related networks simultaneously, using predictions about conservation of TF-gene interactions to share information between sources.
+Our approach is useful for learning networks from similar sources such as different cell-lines from the same species, where there exists a one-to-one mapping of genes, as well as datasets where the orthology mapping does not span all the genes.
+This can occur when using different technology, eg microarray and RNAseq, where one method does not assay all of the genes that another one does, but the networks are expected to be very similar; or when inferring networks describing distantly related species, in which case the networks are expected to be more different than similar.
+When orthology is not complete, we are interested in knowing if performance gains from fused regression are limited to those interactions which have fusion constraints.
+To test this, we used sets of 20 TF by 200 gene synthetic networks, with varying proportions of orthologous TFs and genes.
+We divided networks into those interactions with fusion constraints (interactions where there exist putative analogous interactions in the second network based off of the orthology mapping between TF, gene pairs), which we call the constrained subnetwork, and interactions without fusion constraints, which we refer to as the non-constrained subnetwork.
+We varied the weight on the fusion penalty, lamS, and evaluated performance by computing AUPR on the constrained subnetwork, the non-constrained subnetwork, and the whole network (Figure 1C).
+Since the conserved subgraphs are similar to each other (more like Figure 2A than 2B), we expected performance to improve as the fusion penalty weight increased.
+We observed this, particularly for the constrained subnetwork.
+As lamS increased, interactions with fusion constraints were encouraged to be more similar; growth in performance slowed as the differences in interaction weights approach zero, and the small amount of added noise to the expression data and slight differences between the conserved networks limit the recovery from reaching completeness.
 Interestingly, performance gains were seen even in the portion of the network that is unconstrained by fusion.
-By constraining part of an underdetermined system, we obtain gains in even the unconstrained part.
+By constraining part of an undetermined system, we obtained gains in even the unconstrained part.
 
 \subsection{Adaptive fusion successfully identifies and unfuses 'neofunctionalized' genes}
-We recognize that orthology prediction is not a perfect proxy for functional conservation.
-We implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (Figures 5-6).
+We recognize that orthology prediction is not a perfect proxy for functional conservation \cite{gabaldon_functional_2013, studer_how_2009, nehrt_testing_2011}.
+We implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (Figure 3).
 Pairs of interactions which are very dissimilar even after fusion, which sit in the flat portion of this penalty function, are effectively ``unfused,'' and no further penalty is incurred as differences in interaction weights grow.
-Because our network inference problem is underdetermined, meaning that multiple networks would fit our data equally well, the ``unfusing'' of constraints is convincing evidence for the rejection of a network which uses orthology to predict similarity in interaction weights.
+Because our network inference problem is underdetermined, meaning that multiple networks would fit our data equally well, the ``unfusing'' or relaxation of the fusion penalty on constraints is convincing evidence for the rejection of a network which uses orthology to predict similarity in interaction weights.
 ''Unfusing'' suggests neofunctionalization of one or more of the genes involved in this pair of interactions.
 
-We performed a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved when orthology information is faulty, to mimic orthologs which are not functionally analogous  (Figure 7-8).
+We performed a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved when orthology information is faulty, to mimic orthologs which are not functionally analogous  (Figure 4).
 We generated synthetic fused networks and introduced error in the fusion constraints by adding false positives and negatives to the orthology information given to the solver.
-Because we knew which entries in the orthology mapping were ``fake'' (not reflected in the generation of the networks), we could correctly label fusion constraints that involved one or more ``fake'' mappings.
-We verified that adaptive-fusion unfused mostly ``fake'' interactions, while leaving truly analogous interactions fused.
-We compared the recovery of interaction weights which were accurately fused, and recovery of interaction weights which were inaccurately fused due to incorrect orthology information.
-Because fused L2 heavily penalizes large differences between weights which are predicted to be similar (Figures 8-9), the error for inaccurately fused interactions is higher than solving without fusion.
-When we use our adaptive fusion approach, we preserve the gain in performance on accurately fused interactions, while reducing the error for weights of interactions involving "fake" orthologies.
+Because we knew which entries in the orthology mapping were ``incorrect'' (not reflected in the generation of the networks), we could correctly label fusion constraints that involved one or more ``incorrect'' mappings.
+We verified that adaptive-fusion unfused mostly ``incorrect'' interactions, while leaving truly analogous interactions fused (Figure 4a.
+Note that while pairs of interactions between true orthologs are sampled randomly so that the difference in interactions tends to be small, and pairs of interactions between incorrect orthologs are sampled randomly so that the difference in interactions tends to be large, there is an element of randomness in the sampling.
+This results in a small fraction of interactions between incorrect othologs being similar, and similarly, a small fraction of interactions between correct orthologs being different.
+Consequently, some interactions between incorrect orthologs remain fused and some interactions between correct orthologs are unfused.
+
+We compared the recovery of interaction weights which were accurately fused, and recovery of interaction weights which were inaccurately fused due to incorrect orthology information (Figure 4b).
+Because fused L2 heavily penalizes large differences between weights which are predicted to be similar, it is able to retrieve a more accurate network for those interactions with true fusion constraints than learning networks separately.
+However, the gains accomplished by sharing information through fused regression do not extend to those interactions lacking true fusion constraints, and the error remains similar to learning networks separately.
+Because the penalty function used in adaptive fusion resembles that of fused L2 where differences in interactions are small (Figure 3), adaptive fusion's performance on accurately fused interactions is similar to that of fused L2.
+Using adaptive fusion, we are able to identify which interactions should not be similar, and therefore, which orthology mappings are not predictive of functional similarity.
+However, because these interactions remain underdetermined, and so we are not necessarily much closer to learning their values.
 
 \subsection{Cross-species network inference using bacterial data}
-We used gene-expression data from B.
-subtilis and B.
-anthracis in order to assess performance gains of fused regression on real data.
+We used gene-expression data from \textit{\textit{Bacillus subtilis}} and \textit{\textit{B.
+anthracis}} in order to assess performance gains of fused regression on real data.
 Our B subtilis data set consists of 360 time-series and steady-state observations of 4891 genes during development.
-Our B.
-anthracis dataset consists of 72 time-series and steady-state observations of 5536 genes comprising data from developmental and iron-starvation conditions.
-There were 247 known Transcription Factors (TFs) in the B.
-subtilis dataset, and 248 TFs in the B.
-anthracis dataset.
+Our \textit{\textit{B.
+anthracis}} dataset consists of 72 time-series and steady-state observations of 5536 genes comprising data from developmental and iron-starvation conditions.
+There were 247 known Transcription factors (TFs) in the \textit{\textit{B.
+subtilis}} dataset, and 248 TFs in the \textit{B.
+anthracis} dataset.
 
-We obtained 1870 one-to-one orthologs from Inparanoid \cite{ostlund_inparanoid_2010}, which produced 177650 fusion-constraints between gene interactions within the two species.
-Although this number seems very large, it represents only $14.7\%$ of the regulatory interaction matrix in B.
-subtilis and $12.9\%$ in B.
-anthracis.
+We obtained 1870 one-to-one orthologs from Inparanoid \cite{ostlund_inparanoid_2010}, which produced 177,650 fusion-constraints between gene interactions within the two species.
+This number represents only $14.7\%$ of the regulatory interaction matrix in \textit{B.
+subtilis} and $12.9\%$ in \textit{B.
+anthracis}.
 The use of such dissimilar species is a test of whether fusion can improve network-inference even when the overlap in interactions is very small.
 
-In order to evaluate network inference performance, and for use as priors, we used a gold standard of 2896 known B.
-subtilis interactions with corresponding activation and repression sign.
-Of these 2896 priors, 968 had corresponding interactions in B.
-anthracis.
+In order to evaluate network inference performance, and for use as priors, we used a gold standard of 2896 known \textit{B.
+subtilis} interactions with corresponding activation and repression sign.
+Of these 2896 priors, 968 had corresponding interactions in \textit{B.
+anthracis}.
 Based on our simulation results, we can expect the greatest gains in network-inference performance from fusion when the species of interest has a small number of available conditions, but data is abundant in a related species.
 However, in order to evaluate performance objectively a gold-standard of known interactions is necessary.
-As a result, we can only evaluate network recovery for B.
-subtilis, and B.
-subtilis also has the majority of our conditions.
-In order to simulate the data-poor regime, we subsampled our B.
-subtilis data.
-Specifically, we divided our B.
-subtilis data into $k$ folds, and then for each fold fit a network to the B.
-subtilis data from that fold alone fused to the entire 72 B.
-anthracis conditions.
+As a result, we can only evaluate network recovery for \textit{B.
+subtilis}, and \textit{B.
+subtilis} also has the majority of our conditions.
+In order to simulate the data-poor regime, we subsampled our \textit{B.
+subtilis} data.
+Specifically, we divided our \textit{B.
+subtilis} data into $k$ folds, and then for each fold fit a network to the \textit{B.
+subtilis} data from that fold alone fused to the entire 72 \textit{B.
+anthracis} conditions.
 We can then compute performance metrics (ie AUPR) for each fold, and use their variability across folds to test for significance of any improvements relating to the use of fusion constraints.
 
-When we applied fused L2 to our subsampled B.
-subtilis and B.
-anthracis datasets, we did not see a significant improvement over solving the networks separately.
-Inspection of precision recall curves shows that this improvement is a result of increased precision at low values of recall.
+When we applied fused L2 to our subsampled \textit{B.
+subtilis} and \textit{B.
+anthracis} datasets using optimized parameters, we did not see a significant improvement over solving the networks separately.
+Inspection of precision recall curves shows that there is increased precision at low values of recall.
 Essentially, precision increases for the high confidence part of the network but not for the low confidence part of the network.
-As a result, the improvements in performance may be more useful than would be suggested by AUPR alone.
-Interactions suggested by genome wide analysis need to be experimentally verified, and the ranking of interactions produced by a network inference algorithms is a useful guide to what order these experiments should be carried out in.
+As a result, the slight improvements in performance may be more useful than would be suggested by AUPR alone.
+Interactions suggested by genome wide analysis are useful for target prediction, and the ranking of interactions produced by a network inference algorithms is a useful guide to what order these experiments should be carried out in.
 The high confidence part of the network represents interactions that are likely candidates for experimental validation, so improvement in recovery of these interactions is particularly useful in guiding research.
@@
 In bacteria, genes within the same operon are under the control of the same promoter.
-We predict, therefore, that genes within the same operon will be regulated similarly by the same transcription factors.
-We apply fusion regression by creating fusion constraints between a given transcription factor and genes within the same operon, and show a boost in B.
-subtilis network recovery using within-species fusion.
+We predicted, therefore, that genes within the same operon will be regulated similarly by the same transcription factors.
+We applied fusion regression by creating fusion constraints between a given transcription factor and genes within the same operon, and showed a boost in \textit{B.
+subtilis} network recovery using within-species fusion (Figure 6b).
 
 \subsection{Integrating datasets using fusion regression}
-Although there are many large-scale collaborations which attempt to make protocols as uniform as possible for comparability between datasets generated by different labs, there still exists technical and biologial variability between many experiments attempting to capture the same or similar experimental condition.
-With the advent of new technology, such as RNAseq and single-cell RNAseq, microarray is no longer the dominant assay for genome-wide expression, but it is important to be able to combine datasets generated using different technologies, lest we lose valuable information.
+Although there are many large-scale collaborations which attempt to make protocols as uniform as possible for comparability between datasets generated by different labs \cite{kundaje_integrative_2015}, there still exists technical and biologial variability between many experiments attempting to capture the same or similar experimental condition.
+With the advent of new technology, such as RNAseq and single-cell RNAseq, microarray is no longer the dominant assay for genome-wide expression, but it is important to be able to combine datasets generated using different technologies.
 Currently, the most widely used approach to combining datasets for network inference is to learn networks from disparate datasets separately, then rank combine the networks as in Marbach et al \cite{marbach_revealing_2010}.
-We compare this approach to using fusion regression.
-We include, along with our B.
-subtilis dataset, a previously published dataset containing 269 samples covering 104 conditions, obtained using a different microarray and different strain of B.
-subtilis \cite{nicolas2012condition}.
-We compare performance when learning the networks separately, learning the networks separately followed by rank combining, and learning the networks simultaneously using fusion regression, and show improvement on network inference using our fused L2 approach.
+We compared this approach to using fusion regression.
+We included, along with our \textit{B.
+subtilis} dataset, a previously published dataset containing 269 samples covering 104 conditions, obtained using a different microarray and different strain of \textit{B.
+subtilis} \cite{nicolas2012condition}.
+We compared performance when learning the networks separately, rank combining, and learning the networks simultaneously using fusion regression, and showed improvement on network inference using our fused L2 approach (Figure 6a).
 Importantly, our approach simultaneously learns multiple networks, rather than learning networks separately than combining into one.
@@
 \subsection{Transcription factor activity estimation integrates into fusion regression approach}
-We test a combination of our fused regression approach with a method for estimating transcription factor activities (TFA).
-Rather than modeling gene expression using transcription factor mRNA abundance, we attempt to fit gene expression as a function of transcription factor activity.
-We estimate TFA based on known regulatory interactions using network component analysis \cite{liao2003network}.
-We randomly divide the prior known interactions in half, and use half to learn TFA.
-The rest we reserve as a gold standard for validation.
-When we incorporate TFA and repeat our experiment combining B.
-subtilis datasets, we see a marked improvement in performance from using TFA.
-Our gains from using fused regression are preserved and even increased with the use of TFA.
+We tested a combination of our fused regression approach with a method for estimating transcription factor activities (TFA).
+Rather than modeling gene expression using transcription factor mRNA abundance, we attempted to fit gene expression as a function of transcription factor activity, which have been shown to be better predictors of gene expression \cite{arrieta-ortiz_experimentally_2015}.
+We estimated TFA based on known regulatory interactions using network component analysis \cite{liao2003network}.
+To test the integration of this approach with our fused regression, we asssesed the integration of \texti{B.
+subtilis} datasets, as in figure 6a, with the incorporation of TFA estimation.
+We randomly divided the prior known interactions in half, and used half to learn TFA.
+The rest we reserved as a gold standard for validation.
+As in previous studies, we observed a marked improvement in network inference when using transcription factor activity, which captures activation of transcription factor through mechanisms such as dimerization and interaction with required factors, over transcription factor expression level alone \cite{fu_reconstructing_2011}.
+We also obtained AUPR improvement when using fused regression on TFA, and showed that our gains from sharing information across datasets using fused regression were preserved and even enhanced by using TFA.
 
@@
 These constraints improve the performance of the underdetermined network inference problem, by allowing information to be shared across sources.
-We apply this algorithm to the problem of network inference in two distantly related biological organisms -- B.
-Subtilis and B.
-Anthracis -- and show that network recovery is improved through the introduction of fusion constraints between pairs of orthologous genes.
+We apply this algorithm to the problem of network inference in two distantly related biological organisms -- \textit{B.
+subtilis} and \textit{B.
+anthracis} -- and show that network recovery is improved through the introduction of fusion constraints between pairs of orthologous genes.
 We further demonstrate the viability of fused L2 as a method for combining data from multiple experimental platforms, where fusion is between identical pairs of genes, and as a method for the incorporation of information about operon structure, where fusion reflects the assumption that genes in the same operons have similar regulators.
@@
 Our saturating penalty relaxes the fusion penalty, giving evidence that the fusion constraint is not predictive of gene interaction.
-We may use the analogy of a rubber band, which can become stretched when penalty functions within our objective function oppose each other; the adaptive fusion approach allows the snapping of this once the opposition is great enough.
 For the multiple species case, this represents orthologs which do not share similar interactions.

### 2016-01-11 paper1.tex
zack discussion eddits

@@
 \section{Discussion}
+Gene expression data, such as microarray or RNA seq, provide information about the relationship between genes by allowing an experimenter to measure correlations in expression value over time or across conditions.
+Many sources of information - such as the knowledge that two genes are related through orthology or belong to the same operon - provide information about the relationships between these gene-gene relationships, without necessarily providing information about the gene-gene relationships themselves.
+For example, the meta information that two genes belong to the same operon suggests that they are likely to have a similar set of regulators (CITE), but does little to inform the identity of those regulators.
+Meta-information about the structure of gene regulatory networks - specifically which pairs of interactions are \texti{a priori} likely to be similar to one another, can provide a powerful set of constraints to improve network inference performance.
+We present a general framework for gene regulatory network inference that incorporates this meta-information - termed fusion constraints - and apply the technique to the problem of simultaneous inference of regulatory networks in multiple species (\textit{B.
+subtilis} and \textit{B.
+anthracis}).
+We develop two algorithms for solving fused network inference problems, each of which has as its core the solving of a least-squares optimization problem with novel regularization constraints.
 
-We present a general framework for the simultaneous inference of gene regulatory networks in multiple sources based on the introduction of constraints that penalize differences in regulatory weights thought to be similar based on prior biological knowledge.
-In order to solve the fused network inference problem, we develop two methods for the joint learning of gene regulatory networks, based on the use of different fusion penalties.
  The first method, fused L2, penalizes the squared difference of the regulatory weights of fused interactions.
@@
 anthracis} -- and show that network recovery is improved through the introduction of fusion constraints between pairs of orthologous genes.
-We further demonstrate the viability of fused L2 as a method for combining data from multiple experimental platforms, where fusion is between identical pairs of genes, and as a method for the incorporation of information about operon structure, where fusion reflects the assumption that genes in the same operons have similar regulators.
+Existing methods for cross-species network inference operate on the conserved subset of orthologous genes [CITE].
+This may be appropriate with very closely related species, but could not be applied in this domain, where a large fraction (X\% and Y\%) of each species genomes do not have orthologs.
+Our method, in contrast, can obtain improvements in network inference performance even when the conserved subset of genes is small.
+We further demonstrate the viability of fused L2 as a method for combining data from multiple experimental platforms, where fusion is between each identical regulator-gene pair.
 Because the algorithm we developed can accomodate constraints between arbitrary pairs of regulatory interactions, any biological prior representing information about expected regulatory similarity can be represented, even if the prior provides no information about the magnitude or direction of regulation.
+We demonstrate this flexibility through the novel incorporation of operon structure into the gene-regulatory network inference problem [CITE NOVELTY].
+In this application, fusion reflects the assumption that genes in the same operons have similar regulators [CITE].
 The ability to incorporate multiple data sets describing related processes, as well as multiple data types, in a principled manner, helps us take advantage of the breadth of experimentation in biology to better learn the structure of gene regulation.
 This approach is particularly interesting in light of the diversity of model organisms used in modern biology.
-Different model systems provide different advantages and disadvantages for experimental design, but without a principled mechanism for combining data from multiple sources, it is difficult to fully leverage data obtained from even a slightly different model system.
+Different model systems provide different advantages and disadvantages for experimental design [CITE], but without a principled mechanism for combining data from multiple sources, it is difficult to fully leverage data obtained from even a slightly different model system.
 
@@
 Our cross-species network inference method is premised on the assumption that orthologous genes have similar regulators.
-This is a common assumption when comparing model systems, which is difficult to evaluate on a genome wide scale using existing methods; because network inference is typically underconstrained, the presence of a difference between two regulatory interactions does not rule out an equally good network (or pair of networks) in which those regulatory interactions are similar.
-In contrast, because our method incorporates fusion penalties that favor regulatory similarity between interactions thought to be conserved, the failure to find a network in which fused regulators are similar is direct evidence that no such network is possible.
+This is a common assumption when comparing model systems, which is difficult to evaluate on a genome wide scale using existing methods; because network inference is typically underconstrained, fitting a network that describes a particular set of experimental observations involves sampling a single network from a large set of networks that fit the data equally (or almost equally) as well.
+As a result, the existence of a difference between corresponding regulatory interactions in a pair of experimentally derived networks is weak evidence that a difference truly does exist.
+It may be that some other pair of networks exist that fit the data nearly as well, but do not exhibit this difference.
+As a result, global network inference algorithms are a very weak tool for uncovering evolutionary divergence.
+Our method, on the other hand, explicitly favors recovering networks for which evolutionarily corresponding interactions are similar.
+As a result, the failure to obtain networks that confirm evolutionary conservation is strong evidence that conservation does not exist; the next best network that does exhibit conservation must fit the data much worse to have overcome the bias built into the fusion constraints.
 
 Because our approach allows us to directly test -- and reject -- the hypothesis of conserved function, and because the enforcement of fusion constraints between non-conserved interactions may degrade network inference performance, we introduce an adaptive fusion method that decouples pairs of interactions for which expression data suggests a lack of conservation.
-Like the fused L2 approach, this method attempts to solve for networks which minimize a penalty function fitting the expression data and the fusion constraints.
-In the case of cross-species network inference, evolutionary links may be useful to guide the sharing of data.
-However, orthology may not always accurately predict functional similarity, and a fusion constraint encouraging orthologous gene, TF pairs to share similar interactions across species may not be supported by expression data.
-Then, when the fusion penalty is at odds with the penalty fitting expression data, there exists a tension between satisfying evolutionary constraints while accurately fitting experimental data.
-Our saturating penalty relaxes the fusion penalty, giving evidence that the fusion constraint is not predictive of gene interaction.
-For the multiple species case, this represents orthologs which do not share similar interactions.
+Like the fused L2 approach, this method attempts to solve for networks which minimize a function that minimizes error in fitting the expression data plus error associated with the fusion penalty.
+In the case of cross-species network inference, orthology may provide a useful set of constraints on network structure.
+However, orthology does not always accurately predict functional similarity [CITE].When a fusion constraint encourages orthologous gene, TF pairs to share similar interactions across species, but the expression data suggests that these interaction weights should not be similar, the algorithm will arrive at some compromise between these two forces.
+When the disagreement arose because the genes involved have taken on new function through evolutionary divergence [CITE], network recovery would be improved had the fusion constraint not been enforced.
+We introduce a method -- adaptive fusion -- that attempts to learn which fusion constraints should be relaxed while the network is being learned.
+This method is based on minimizing a saturating penalty function on fusion constraints, similar to a class of penalties that have been developed to minimise bias in regularized regression [CITE].
+The result of adaptive fusion is both a network and a new set of fusion constraints, describing the learned conservation weights (including which fusion constraints have been relaxed).
+For the multiple species case, relaxation of fusion constraints represents orthologs which do not share similar interactions.
 When jointly learning networks describing processes in different cell lines, this may identify interesting context-specific behavior.

### 2016-01-12 paper1.tex
hi

@@
 We tested the ability of fused regression to improve network inference performance, as well as the ability of adaptive fusion to learn the conservation of gene interactions, using both synthetic networks and real data.
-For the synthetic data, we generated random pairs of networks in which orthologous genes have similar regulatory interactions, then sampled gene expression from these networks, which we used to derive networks for comparison with the true network.
+For the synthetic data, we generated random pairs of networks in which orthologous genes have similar regulatory interactions, then sampled gene expression from these networks, which we used to derive networks for comparison with the true networks.
 For real data, we computed recovery of a known gold standard in \textit{Bacillus subtilis}.
@@
 It is known that network recovery improves with additional data \cite{bar-joseph_computational_2003}.
-Additional data is not always readily available; but often there is an abundance of related data, and we reason that data describing similar processes can be used to improve our ability to recover true networks.
+Additional data is not always readily available; but often there is an abundance of related data, and we reason that we should be able to use data describing similar processes.
 Using related data for network inference allows us to borrow statistical power from other sources, effectively increasing the sample size and boosting the sensitivity and specificity of learned interactions.
-We created synthetic networks that approximate two related biological processes, then evaluated performance of our fused L2 regression, which learns the networks simultaneously, sharing information about related interactions (Figure 1b).
-We compared recovery of the 10 TF by 200 gene networks, using fused L2 versus learning networks separately, using different amounts of simulated expression data from the two species.
-When the amount of data from species two is held constant, increasing the amount of data available for learning the network for species one results in a more accurate network, as expected.
-Using the assumption that similar genes have similar regulatory interactions, we share information about putatively conserved interactions.
-When we increase the amount of data from species two, we get performance gains on network one using fused L2 regression.
+We use the assumption that similar genes have similar regulatory interactions, and share information about putatively conserved interactions.
+We created synthetic networks to approximate two related biological processes, then evaluated performance of our fused L2 regression, which learns the networks simultaneously, sharing information about related interactions (Figure 1b).
+We compared recovery of the 10 TF by 200 gene networks, using fused L2 versus learning networks separately.
+We varied the amounts of simulated expression data made available to the solver.
+When the amount of data from species two was held constant, increasing the amount of data available for learning the network for species one resulted in a more accurate network, as expected.
+When we increased the amount of data from species two, we obtained performance gains on network one using fused L2 regression, showing our ability to improve network inference on one dataset by sharing data from a related dataset.
 
@@
 We simulated this spectrum by creating pairs of networks and varying the similarity between networks.
-The main factors governing similarity between our generated networks is the extent (and accuracy) of the orthology mapping, and the variability between analogous, or conserved, interactions.
+The main factors governing similarity between our generated networks was the extent (and accuracy) of the orthology mapping, and the variability between analogous, or conserved, interactions (in other words, how well conserved the conserved subnetwork was).
 We conducted a series of simulations to assess the effect of increasing orthology coverage on networks.
@@
 
-As expected, when the conserved subgraph is very similar - when the interactions in the conserved subnetwork are nearly identical, simulating the case of closely related organisms or similar processes (Figure 2A) - increasing the weight of the fusion constraint $\lambda_S$ improved network recovery.
-As the size of the conserved subgraph increases, using a greater weight on the fusion penalty was beneficial.
+As expected, when the conserved subgraph was very similar - when the interactions in the conserved subnetwork were nearly identical, simulating the case of closely related organisms or similar processes (Figure 2A) - increasing the weight of the fusion penalty $\lambda_S$ improved network recovery.
+As the size of the conserved subgraph increased, using a greater weight on the fusion penalty was beneficial.
 
-To simulate the case of distantly related organisms, we created networks where conserved interactions are drawn from high variance distribution (Figure 2B).
-When the conservation is weak - when the interactions in the conserved subnetwork are nearly as different across species as randomly chosen interactions - performance may not continue to improve by increasing the fusion weight.
-We show that even for networks where the conserved subgraph is weakly conserved, there exists a 'sweet spot' where fusion regression improves network recovery.
+To simulate the case of distantly related organisms, we created networks where conserved interactions were drawn from high variance distribution (Figure 2B).
+When the conservation iss weak - when the interactions in the conserved subnetwork are nearly as different across species as randomly chosen interactions - performance may not continue to improve by increasing the fusion weight.
+We showed that even for networks where the conserved subgraph was weakly conserved, there exists a 'sweet spot' where fusion regression improves network recovery.
 
@@
 We varied the weight on the fusion penalty, lamS, and evaluated performance by computing AUPR on the constrained subnetwork, the non-constrained subnetwork, and the whole network (Figure 1C).
-Since the conserved subgraphs are similar to each other (more like Figure 2A than 2B), we expected performance to improve as the fusion penalty weight increased.
+Since the conserved subgraphs were similar to each other (more like Figure 2A than 2B), we expected performance to improve as the fusion penalty weight increased.
 We observed this, particularly for the constrained subnetwork.
-As lamS increased, interactions with fusion constraints were encouraged to be more similar; growth in performance slowed as the differences in interaction weights approach zero, and the small amount of added noise to the expression data and slight differences between the conserved networks limit the recovery from reaching completeness.
-Interestingly, performance gains were seen even in the portion of the network that is unconstrained by fusion.
+As $\lambda_S$ increased, interactions with fusion constraints were encouraged to be more similar; growth in performance slowed as the differences in interaction weights approached zero, and the small amount of added noise to the expression data and slight differences between the conserved networks limited the recovery from reaching completeness.
+Interestingly, performance gains were seen even in the portion of the network that was unconstrained by fusion.
 By constraining part of an undetermined system, we obtained gains in even the unconstrained part.
@@
 We compared the recovery of interaction weights which were accurately fused, and recovery of interaction weights which were inaccurately fused due to incorrect orthology information (Figure 4b).
-Because fused L2 heavily penalizes large differences between weights which are predicted to be similar, it is able to retrieve a more accurate network for those interactions with true fusion constraints than learning networks separately.
+Because fused L2 heavily penalizes large differences between weights which are predicted to be similar, it is able to retrieve a more accurate network for those interactions with true fusion constraints than by learning networks separately.
 However, the gains accomplished by sharing information through fused regression do not extend to those interactions lacking true fusion constraints, and the error remains similar to learning networks separately.
 Because the penalty function used in adaptive fusion resembles that of fused L2 where differences in interactions are small (Figure 3), adaptive fusion's performance on accurately fused interactions is similar to that of fused L2.
-Using adaptive fusion, we are able to identify which interactions should not be similar, and therefore, which orthology mappings are not predictive of functional similarity.
-However, because these interactions remain underdetermined, and so we are not necessarily much closer to learning their values.
+Because interactions which have been unfused remain underdetermined, we are not closer to learning their values than if we learned networks separately.
+However, when using adaptive fusion, we are able to identify which interactions should not be similar, and therefore, which orthology mappings are not predictive of functional similarity.
 

### 2016-01-15 paper1.tex
hi

@@
 
+\subsection{\textit{B. subtilis} and \textit{B. anthracis} microarray data}
+We used a dataset collected for PY79, a derivative of strain 168, available on GEO with accession number GSE67023, and a dataset using BSB1, another derivative of strain 168, available at GEO with accession number GSE27219.
+We used two datasets for \texti{B.
+anthracis}, transcription profiling during iron starvation (E-MEXP-2272 on ArrayExpress), and time series over the life cycle (E-MEXP-788 on ArayExpress).
+
+\subsection{Orthology mapping using Inparanoid}
+We ran Inparanoid to obtain orthology mapping for \textit{B. subtilis} and \textit{B. anthracis} \cite{ostlund_inparanoid_2010}
+
 \subsection{Gene regulatory network}
@@
 This is similar to the formulation used in the Inferelator algorithm, which we extend to the case of simultaneously inferring multiple networks. 
+\subsection{Learning transcription factor activity and using TFA as regressors}
 
@@
 
+\subsection{Statistical approach}
+Most network inference problems have more predictors than experimental samples ($p >> N$).
+Many methods have been proposed for penalized regression, such as ridge regression \cite{hoerl_ridge_1970}, and applied to genomic data \cite{waldron_optimized_2011, li_network-constrained_2008}.
+In our application, we consider the case where related outputs (genes) are likely to be affected by common inputs (transcription factors).
+Approaches using fusion combine statistical strength across the multiple regression tasks by introducing regularization functions reflecting prior knowledge on output structure representing how outputs are related \cite{kim_tree-guided_2012,Land1997,Chen2010,Petry2011,Hebiri2011}, and approaches which attempt to simultaneously learn the regression coefficients and the output structure \cite{sohn_joint_2012}.
+Fusion has been proposed as a method for utilizing correlation information from data by penalizing pairwise differences of coefficients \cite{Daye2009,Price2014}.
+Price et al.
+and Bilgrau et al.
+use a fused ridge estimator for jointly estimating multiple inverse covariance matrices \cite{Price2014a,Bilgrau2015}; we take a similar approach, adding an fused L2 penalty to our ridge regression to incorporate prior knowledge about relationships between input-output pairs.
+We also apply a penalty function bounded by a constant to produce unbiased estimators for large coefficients, combined with an L2 penalty, similar to SCAD-L2 \cite{Zeng2012}.
+This approach allows us to simultaneously learn the regression coefficients and evaluate the validity of this prior information.
+
 \subsection{Solving fused L2 problems using augmented matrices}
@@
 This is because the structure of orthology tends to be sparse, so only a small number of genes must be solved at once.
-However, in the general case of $c_i$ response variables from each of $i \in \text{species}$ data sources the design matrix to be solved is of size $\displaystyle \sum_{c_i} rows(X^{S_i}) \times \displaystyle \sum_{c_i} cols(X^{S_i})$.
-Although the sparse structure of the augmented design matrix makes this somewhat more tractable, it is still impractical for large constraint chains.
+However, in the general case of $c_i$ response variables from each of $i \in \text{species}$ data sources the design matrix to be solved is of size $\displaystyle \sum_{c_i} rows(X^{S_i}) \times \displaystyle \sum_{c_i} cols(X^{S_i})$ plus a relatively small number of rows representing fusion constraints.
+Because the scaling of this algorithm has a complicated depence on the constraint structure used, a general description of its runtime is difficult.
+However, in the case of multi-species network inference with one-to-one orthology, the network associated with each pair of orthologous genes requires solving a linear system with approximately twice as many observations and unknowns as the single species case.
+Linear systems of this size can be solved quickly using standard techniques.
+When the number of genes linked by fusion constraints becomes large, however, the augmented design matrix approach becomes impractical.
 
@@
 
-We performed a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved when orthology information is faulty, to mimic orthologs which are not functionally analogous  (Figure 4).
+We performed a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved when orthologs are not functionally analogous  (Figure 4).
 We generated synthetic fused networks and introduced error in the fusion constraints by adding false positives and negatives to the orthology information given to the solver.
 Because we knew which entries in the orthology mapping were ``incorrect'' (not reflected in the generation of the networks), we could correctly label fusion constraints that involved one or more ``incorrect'' mappings.
-We verified that adaptive-fusion unfused mostly ``incorrect'' interactions, while leaving truly analogous interactions fused (Figure 4a.
-Note that while pairs of interactions between true orthologs are sampled randomly so that the difference in interactions tends to be small, and pairs of interactions between incorrect orthologs are sampled randomly so that the difference in interactions tends to be large, there is an element of randomness in the sampling.
-This results in a small fraction of interactions between incorrect othologs being similar, and similarly, a small fraction of interactions between correct orthologs being different.
-Consequently, some interactions between incorrect orthologs remain fused and some interactions between correct orthologs are unfused.
+We verified that adaptive-fusion unfused mostly ``incorrectly fused'' interactions, while leaving truly analogous interactions fused (Figure 4a.
+Note that while pairs of interactions between true orthologs were sampled randomly so that the difference in interactions tended to be small, and pairs of interactions between incorrect orthologs were sampled randomly so that the difference in interactions tended to be large, there was an element of randomness in the sampling.
+This resulted in a small fraction of interactions between incorrect othologs being similar, and similarly, a small fraction of interactions between correct orthologs being different.
+Consequently, some interactions between incorrect orthologs remained fused and some interactions between correct orthologs were unfused.
 
@@
 Because the penalty function used in adaptive fusion resembles that of fused L2 where differences in interactions are small (Figure 3), adaptive fusion's performance on accurately fused interactions is similar to that of fused L2.
-Because interactions which have been unfused remain underdetermined, we are not closer to learning their values than if we learned networks separately.
+Because interactions which have been unfused remain underdetermined, we are not much closer to learning their values than if we learned networks separately.
 However, when using adaptive fusion, we are able to identify which interactions should not be similar, and therefore, which orthology mappings are not predictive of functional similarity.
@@
 
-We obtained 1870 one-to-one orthologs from Inparanoid \cite{ostlund_inparanoid_2010}, which produced 177,650 fusion-constraints between gene interactions within the two species.
+We obtained 1870 one-to-one orthologs from Inparanoid \cite{ostlund_inparanoid_2010}, 95 of which are transcription factors, which produced 177,650 fusion-constraints between gene interactions within the two species.
 This number represents only $14.7\%$ of the regulatory interaction matrix in \textit{B.
@@
 anthracis} datasets using optimized parameters, we did not see a significant improvement over solving the networks separately.
-Inspection of precision recall curves shows that there is increased precision at low values of recall.
+Inspection of precision recall curves showed that there is increased precision at low values of recall.
 Essentially, precision increases for the high confidence part of the network but not for the low confidence part of the network.
-As a result, the slight improvements in performance may be more useful than would be suggested by AUPR alone.
+As a result, the improvements in performance may be more useful than would be suggested by AUPR alone.
 Interactions suggested by genome wide analysis are useful for target prediction, and the ranking of interactions produced by a network inference algorithms is a useful guide to what order these experiments should be carried out in.
@@
 Information about the similarity of TF-gene interactions can also come from knowledge about the promoter region; this may be a better predictor of target similarity than orthology.
-In bacteria, genes within the same operon are under the control of the same promoter.
+In bacteria, genes within the same operon are under the control of the same promoter \cite{lawrence_shared_2002}.
 We predicted, therefore, that genes within the same operon will be regulated similarly by the same transcription factors.
@@
 \subsection{Integrating datasets using fusion regression}
-Although there are many large-scale collaborations which attempt to make protocols as uniform as possible for comparability between datasets generated by different labs \cite{kundaje_integrative_2015}, there still exists technical and biologial variability between many experiments attempting to capture the same or similar experimental condition.
+Although there are many large-scale collaborations which attempt to make protocols as uniform as possible for comparability between datasets generated by different labs \cite{paten_nih_2015,kundaje_integrative_2015}, there still exists technical and biologial variability between many experiments attempting to capture the same or similar experimental condition.
 With the advent of new technology, such as RNAseq and single-cell RNAseq, microarray is no longer the dominant assay for genome-wide expression, but it is important to be able to combine datasets generated using different technologies.
@@
 We included, along with our \textit{B.
-subtilis} dataset, a previously published dataset containing 269 samples covering 104 conditions, obtained using a different microarray and different strain of \textit{B.
+subtilis} dataset, a previously published dataset containing 269 samples covering 104 conditions, obtained using a different tiling microarray (vs custom microarray) and different strain of \textit{B.
 subtilis} \cite{nicolas2012condition}.
@@
 We tested a combination of our fused regression approach with a method for estimating transcription factor activities (TFA).
-Rather than modeling gene expression using transcription factor mRNA abundance, we attempted to fit gene expression as a function of transcription factor activity, which have been shown to be better predictors of gene expression \cite{arrieta-ortiz_experimentally_2015}.
+Rather than modeling gene expression using transcription factor mRNA abundance, we fit gene expression as a function of transcription factor activity, as shown in \textit{B.
+subtilis} by Arrieta-Ortiz et al \cite{arrieta-ortiz_experimentally_2015}.
 We estimated TFA based on known regulatory interactions using network component analysis \cite{liao2003network}.
-To test the integration of this approach with our fused regression, we asssesed the integration of \texti{B.
+To test the integration of this approach with our fused regression, we asssesed the combination of \texti{B.
 subtilis} datasets, as in figure 6a, with the incorporation of TFA estimation.
@@
 Many sources of information - such as the knowledge that two genes are related through orthology or belong to the same operon - provide information about the relationships between these gene-gene relationships, without necessarily providing information about the gene-gene relationships themselves.
-For example, the meta information that two genes belong to the same operon suggests that they are likely to have a similar set of regulators (CITE), but does little to inform the identity of those regulators.
+For example, the meta information that two genes belong to the same operon suggests that they are likely to have a similar set of regulators \cite{lawrence_shared_2002}, but does little to inform the identity of those regulators.
 Meta-information about the structure of gene regulatory networks - specifically which pairs of interactions are \texti{a priori} likely to be similar to one another, can provide a powerful set of constraints to improve network inference performance.
@@
 anthracis} -- and show that network recovery is improved through the introduction of fusion constraints between pairs of orthologous genes.
-Existing methods for cross-species network inference operate on the conserved subset of orthologous genes [CITE].
+Existing methods for cross-species network inference operate on the conserved subset of orthologous genes \cite{dillman_comparative_2015}.
 This may be appropriate with very closely related species, but could not be applied in this domain, where a large fraction (X\% and Y\%) of each species genomes do not have orthologs.
@@
 Because the algorithm we developed can accomodate constraints between arbitrary pairs of regulatory interactions, any biological prior representing information about expected regulatory similarity can be represented, even if the prior provides no information about the magnitude or direction of regulation.
-We demonstrate this flexibility through the novel incorporation of operon structure into the gene-regulatory network inference problem [CITE NOVELTY].
-In this application, fusion reflects the assumption that genes in the same operons have similar regulators [CITE].
+We demonstrate this flexibility through the novel incorporation of operon structure into the gene-regulatory network inference problem.
+In this application, fusion reflects the assumption that genes in the same operons have similar regulators \cite{lawrence_shared_2002}.
 The ability to incorporate multiple data sets describing related processes, as well as multiple data types, in a principled manner, helps us take advantage of the breadth of experimentation in biology to better learn the structure of gene regulation.
+We illustrate this by combining two different \textit{B.
+subtilis} datasets and show that fused L2 is an improvement over current approaches to combining data \cite{marbach_revealing_2010} because of our ability to exploit the statistical power that our expanded datasets afford us.
 This approach is particularly interesting in light of the diversity of model organisms used in modern biology.
-Different model systems provide different advantages and disadvantages for experimental design [CITE], but without a principled mechanism for combining data from multiple sources, it is difficult to fully leverage data obtained from even a slightly different model system.
+Different model systems provide different advantages and disadvantages for experimental design \cite{stolfi_genetic_2012}, but without a principled mechanism for combining data from multiple sources, it is difficult to fully leverage data obtained from even a slightly different model system.
 
@@
 Our cross-species network inference method is premised on the assumption that orthologous genes have similar regulators.
-This is a common assumption when comparing model systems, which is difficult to evaluate on a genome wide scale using existing methods; because network inference is typically underconstrained, fitting a network that describes a particular set of experimental observations involves sampling a single network from a large set of networks that fit the data equally (or almost equally) as well.
+Existing approaches to the genome-wide testing of this assumption learn regulatory networks separately, then compare to identify conservation \cite{aytes_cross-species_2014, Wang2014a}.
+But because network inference is typically underconstrained, fitting a network that describes a particular set of experimental observations involves sampling a single network from a large set of networks that fit the data equally (or almost equally) as well.
 As a result, the existence of a difference between corresponding regulatory interactions in a pair of experimentally derived networks is weak evidence that a difference truly does exist.
@@
 In the case of cross-species network inference, orthology may provide a useful set of constraints on network structure.
-However, orthology does not always accurately predict functional similarity [CITE].When a fusion constraint encourages orthologous gene, TF pairs to share similar interactions across species, but the expression data suggests that these interaction weights should not be similar, the algorithm will arrive at some compromise between these two forces.
-When the disagreement arose because the genes involved have taken on new function through evolutionary divergence [CITE], network recovery would be improved had the fusion constraint not been enforced.
+However, orthology does not always accurately predict functional similarity \cite{gabaldon_functional_2013}.When a fusion constraint encourages orthologous gene, TF pairs to share similar interactions across species, but the expression data suggests that these interaction weights should not be similar, the algorithm will arrive at some compromise between these two forces.
+If the disagreement arose because the genes involved have taken on new function through evolutionary divergence \cite{kellis_proof_2004}, network recovery would be improved had the fusion constraint not been enforced.
 We introduce a method -- adaptive fusion -- that attempts to learn which fusion constraints should be relaxed while the network is being learned.
-This method is based on minimizing a saturating penalty function on fusion constraints, similar to a class of penalties that have been developed to minimise bias in regularized regression [CITE].
+This method is based on minimizing a saturating penalty function on fusion constraints, similar to a class of penalties that have been developed to minimise bias in regularized regression \cite{fan2001variable,zhang2010nearly}.
 The result of adaptive fusion is both a network and a new set of fusion constraints, describing the learned conservation weights (including which fusion constraints have been relaxed).
@@
 Adaptive fusion, therefore, is a tool for network inference as well as a method for testing network conservation.
+
+Because our model shares its basic assumptions about the role of transcription factors in gene expression dynamics with models developed for single-species network inference, we are able to leverage techniques developed for the single-species estimation of transctiption factor activity \cite{fu_reconstructing_2011}.
+The performance gains of this additional step in the cross-species case are significant.
 

### 2016-01-16 paper1.tex
hi

@@
 \section{Introduction}
-As the volume and variety of genome scale data have exploded, the goal of accurately modeling gene regulatory networks has become more attainable \cite{bonneau_predictive_2007, ciofani_validated_2012, carro_transcriptional_2010}.
-Large-scale data collection efforts have contributed to the development of high quality networks, but there remain many biological processes and organisms for which there does not exist adequate data.
+As the volume and variety of genome scale data have exploded, the goal of accurately modeling gene regulatory networks has become attainable \cite{bonneau_predictive_2007, ciofani_validated_2012, carro_transcriptional_2010}.
+Large-scale data collection efforts have contributed to the development of high quality networks which accurately recapitulate biological processes, but there remain many processes and organisms for which there does not exist adequate data.
 Furthermore, as new technologies are developed and some old ones are replaced, such as RNAseq and microarray, it becomes important to be able to combine data from multiple platforms, lest we lose valuable information from existing studies.
@@
 
-Moreover, as it becomes increasingly possible to learn genome-wide regulatory networks, we can begin to compare and to test whether there is conservation of networks across species and biological processes.
+Moreover, as it becomes possible to learn genome-wide regulatory networks, we can begin to compare and to test whether there is conservation of networks across species and biological processes.
 Our use of model organisms to study biological processes and diseases relevant to humans relies on the assumption of conservation; yet this has not been tested at the genome scale.
@@
 The core of both methods is the observation that biological information about the relatedness of genes can be used to constrain which network coefficients should be similar to one another in a multi-source network inference problem (ie orthologous TFs should regulate orthologous genes), and that these constraints can be efficiently represented as penalties in a least-squares regression problem.
+Our approach uses predictions of similarity of putative interactions, to improve our ability to accurately describe TF-gene relationships on a genome-wide scale.
+
 Our first method -- fused ridge -- uses an L2 penalty on the differences between a priori similar interactions, and is useful where the relationships between networks (similarity of genes between data sets) is reliable.
@@
 In the case of multiple species, there has been a great deal of work showing that functional conservation exists in gene regulatory networks even across large evolutionary distance \cite{satou2006gene, hinman2009evolution,tanay2005conservation,erwin2009evolution}.
-In our fused L2 approach, we make the assumption that if a pair of TFs is orthologous to one another, then their regulatory effects on a pair of orthologous genes should be similar.
+In our fused L2 approach, we make the assumption that for closely related species, if a pair of TFs is orthologous to one another, then their regulatory effects on a pair of orthologous genes should be similar.
 These orthology relationships form the basis of a set of constraints which favor -- but do not require -- networks in which orthologous transctiption factors regulate orthologous genes.
@@
 This evidence is very weak, however; because network inference is typically underconstrainted \cite{marbach_revealing_2010-1}, there could be a different set of networks for which conservation does hold, and which fit the data almost as well.
-Our adaptive fusion method allows us to learn differences in networks predicted to be similar based on orthology or gene matching: because the constraints we impose favor weight configurations in which fused interaction weights are similar to one another, a solution in which a priori similar regulatory interactions are found to be different is strong evidence that they really are different.
+We propose, with our adaptive fusion, an approach where network inference occurs simultaneously with the evaluation of conservation, to gather stronger evidence of divergence.
 
@@
 The main difference between the L2 fusion approach and this adaptive fusion approach occurs when the difference between presumed analogous interactions is large despite the fusion penalty: we assume that large differences represent evidence that gene functions have diverged and represent this, in adaptive fusion, with a relaxation of the fusion penalty.
-The resulting cost function is non-convex and difficult to optimize; however, we can approximate its solution and obtain deeper insights into functional similarity than is available through orthology or the comparison of separately learned networks.
+The resulting cost function is non-convex and difficult to optimize; however, we can approximate its solution and obtain deeper insight into functional similarity than is available through orthology or the comparison of separately learned networks.
 
@@
 
-\subsection{\textit{B. subtilis} and \textit{B. anthracis} microarray data}
+\subsection{\textit{B. subtilis} and \textit{B. anthracis} data and orthology}
 We used a dataset collected for PY79, a derivative of strain 168, available on GEO with accession number GSE67023, and a dataset using BSB1, another derivative of strain 168, available at GEO with accession number GSE27219.
@@
 anthracis}, transcription profiling during iron starvation (E-MEXP-2272 on ArrayExpress), and time series over the life cycle (E-MEXP-788 on ArayExpress).
-
-\subsection{Orthology mapping using Inparanoid}
-We ran Inparanoid to obtain orthology mapping for \textit{B. subtilis} and \textit{B. anthracis} \cite{ostlund_inparanoid_2010}
+We ran Inparanoid to obtain orthology mapping for \textit{B.
+subtilis} and \textit{B.
+anthracis} \cite{ostlund_inparanoid_2010}
 
@@
 This is similar to the formulation used in the Inferelator algorithm, which we extend to the case of simultaneously inferring multiple networks. 
-\subsection{Learning transcription factor activity and using TFA as regressors}
+\subsection{Estimating transcription factor activity and using TFA as regressors}
+When there exists a set of prior known interactions, we are able to estimate transcription factor activity (TFA) using network component analysis \cite{liao2003network}, as in \cite{arrieta-ortiz_experimentally_2015, fu_reconstructing_2011}, and use TFA as explanatory variables instead of transcription factor expression.
 
@@
 
-\subsection{Statistical approach}
+\subsection{Statistical approach and background}
 Most network inference problems have more predictors than experimental samples ($p >> N$).
@@
 In our application, we consider the case where related outputs (genes) are likely to be affected by common inputs (transcription factors).
-Approaches using fusion combine statistical strength across the multiple regression tasks by introducing regularization functions reflecting prior knowledge on output structure representing how outputs are related \cite{kim_tree-guided_2012,Land1997,Chen2010,Petry2011,Hebiri2011}, and approaches which attempt to simultaneously learn the regression coefficients and the output structure \cite{sohn_joint_2012}.
 Fusion has been proposed as a method for utilizing correlation information from data by penalizing pairwise differences of coefficients \cite{Daye2009,Price2014}.
+Approaches using fusion combine statistical strength across the multiple regression tasks by introducing regularization functions reflecting prior knowledge on output structure representing how outputs are related \cite{kim_tree-guided_2012,Land1997,Chen2010,Petry2011,Hebiri2011}.
 Price et al.
@@
 use a fused ridge estimator for jointly estimating multiple inverse covariance matrices \cite{Price2014a,Bilgrau2015}; we take a similar approach, adding an fused L2 penalty to our ridge regression to incorporate prior knowledge about relationships between input-output pairs.
-We also apply a penalty function bounded by a constant to produce unbiased estimators for large coefficients, combined with an L2 penalty, similar to SCAD-L2 \cite{Zeng2012}.
-This approach allows us to simultaneously learn the regression coefficients and evaluate the validity of this prior information.
+Sohn et al.
+attempt to simultaneously learn the regression coefficients and the output structure \cite{sohn_joint_2012}.
+We also take a simultaneous approach, where we apply a penalty function bounded by a constant to produce unbiased estimators for large coefficients, combined with an L2 penalty, similar to SCAD-L2 \cite{Zeng2012}.
+This approach allows us to simultaneously learn the regression coefficients and evaluate the validity of our prior information.
 

### 2016-01-17 paper1.tex
zackedits

@@
 As the volume and variety of genome scale data have exploded, the goal of accurately modeling gene regulatory networks has become attainable \cite{bonneau_predictive_2007, ciofani_validated_2012, carro_transcriptional_2010}.
-Large-scale data collection efforts have contributed to the development of high quality networks which accurately recapitulate biological processes, but there remain many processes and organisms for which there does not exist adequate data.
+Large-scale data collection efforts have contributed to the development of high quality networks which accurately recapitulate biological processes, but most processes and organisms remain uncharacterized at the network level.
 Furthermore, as new technologies are developed and some old ones are replaced, such as RNAseq and microarray, it becomes important to be able to combine data from multiple platforms, lest we lose valuable information from existing studies.
@@
 Moreover, as it becomes possible to learn genome-wide regulatory networks, we can begin to compare and to test whether there is conservation of networks across species and biological processes.
-Our use of model organisms to study biological processes and diseases relevant to humans relies on the assumption of conservation; yet this has not been tested at the genome scale.
+Our use of model organisms to study biological processes and diseases relevant to humans relies on the assumption of conservation; yet this has not been effectively tested at the genome scale.
 
@@
 The core of both methods is the observation that biological information about the relatedness of genes can be used to constrain which network coefficients should be similar to one another in a multi-source network inference problem (ie orthologous TFs should regulate orthologous genes), and that these constraints can be efficiently represented as penalties in a least-squares regression problem.
-Our approach uses predictions of similarity of putative interactions, to improve our ability to accurately describe TF-gene relationships on a genome-wide scale.
+Taking into account the similarity of putatively conserved interactions improves our ability to accurately describe TF-gene relationships on a genome-wide scale.
 
-Our first method -- fused ridge -- uses an L2 penalty on the differences between a priori similar interactions, and is useful where the relationships between networks (similarity of genes between data sets) is reliable.
+Our first method -- fused ridge -- uses an L2 penalty on the differences between a priori similar interactions (termed fusion penalty), and is useful where the relationships between networks (similarity of genes between data sets) is reliable.
 In the case where both networks contain an identical set of genes and TFs, this approach can be thought of as parametrically interpolating between treating the data sources separately and combining them together.
@@
 Existing multi species approaches often use orthology as a proxy for functional conservation \cite{penfold_inferring_2015, joshi_multi-species_2015, kashima_simultaneous_2009, zhang2010nearly}, or attempt to learn functional similarity via expression data \cite{gholami_cross-species_2010}.
-Orthology -- the measure of gene similarity we use -- can be approximated using readily identifiable sequence similarity.
-Sequence similarity is often a useful predictor of functional similarity \cite{wilson_assessing_2000}.
+Orthology -- the measure of gene similarity we use -- can be approximated using readily identifiable sequence similarity, which is often a useful predictor of functional similarity \cite{wilson_assessing_2000}.
 In multi-species network inference, our fused L2 approach minimizes a cost function that strives to simultaneously fit expression data in each species and produce networks that are consistent with evolutionary constraints created using orthology.
 However, we recognize that some genes may have taken on different functions and therefore may have new regulatory interactions.
-Gene duplications and neofunctionalization in multi-species network inference \cite{eisen_phylogenomics:_1998}, and changes in chromatin configuration in multi-cell line network inference \cite{li_role_2007}, are cases where assumptions about functional conservation may not hold.
+For example, gene duplications may lead to neofunctionalization \cite{eisen_phylogenomics:_1998} of the duplicated genes.
+Even in the case of inferring networks from multiple cell-lines, changes in chromatin configuration may affect our assumptions about the similarity of interactions \cite{li_role_2007}.
 
@@
 The main difference between the L2 fusion approach and this adaptive fusion approach occurs when the difference between presumed analogous interactions is large despite the fusion penalty: we assume that large differences represent evidence that gene functions have diverged and represent this, in adaptive fusion, with a relaxation of the fusion penalty.
-The resulting cost function is non-convex and difficult to optimize; however, we can approximate its solution and obtain deeper insight into functional similarity than is available through orthology or the comparison of separately learned networks.
+The resulting cost function is non-convex and difficult to optimize \cite{fan2001variable}; however, we can approximate its solution and obtain deeper insight into functional similarity than is available through orthology or the comparison of separately learned networks.
+
+Although the fusion constraints we employ can be described as arising from orthology - which links genes - it is important to note that the constraints themselves link individual regulatory interactions.
+This finer level of representational granularity is critical to the functioning of adaptive fusion, and means the method can accomodate any form of prior on expected similarity between regulatory interactions, even priors that are not decomposible into a gene to gene mapping.
+In order to solve the resulting optimization problem, we develop two algorithms for solving efficiently multi-output least-squares regression problems with pairwise L2 fusion penalties on entries of the coefficient matrix.
+We also introduce - in the form of adaptive fusion - the idea of a saturating penalty function on fusion constraints, and estimate the solution to the resulting optimization problem through iterative application of the fused L2 algorithm.
 
@@
 If TF $g^A$ in organism $A$ and TF $h^B$ in organism $B$ are orthologs, and gene $k^A$ and $l^B$ are orthologs, then we expect that the $g^A \rightarrow k^A$ interaction weight should be similar to the $h^B \rightarrow l^B$ interaction weight, and we introduce a fusion constraint between these analogous interactions.
-In terms of the above regression formulation, we expect that $\beta^A_{g,k} \approx \beta^B_{h,l}$, and include a penalty term $p_{\lambda_S}(\beta^A_{g,k} - \beta^B_{h,l})$ in the quantity being minimized in order to encourage similarity.
-The function $p_{\lambda_S}(x)$ controls the shape of the relationship between weight dissimilarity and penalty, while parameter $\lambda_S$ controls the overall scaling of the penalization of differences between fused coefficients.
+In terms of the above regression formulation, we expect that $\beta^A_{g,k} \approx \beta^B_{h,l}$, and include a penalty term $\lambda_Sp(\beta^A_{g,k} - \beta^B_{h,l})$ in the quantity being minimized in order to encourage similarity.
+The function $p(x)$ controls the shape of the relationship between weight dissimilarity and penalty, while parameter $\lambda_S$ controls the overall scaling of the penalization of differences between fused coefficients.
 $\lambda_S$ controls the tradeoff between fitting the expression data alone and producing a set of networks that conform to evolutionary prior knowledge.
 This gives us the final equation to be minimized
+
 \begin{equation}
 %\argmin_{(\beta^A, \beta^B)} \displaystyle\sum_{S \in (A, B)} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\beta^{S_1}_{g,k} \approx \beta^{S_2}_{h,l}} p_{\lambda_S}(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
-\argmin_{\beta^S} \displaystyle\sum_{S \in \text{species}} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\substack{(g,h) \in orth,\\
- (h,l) \in orth}}x p_{\lambda_S}(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
+\argmin_{\beta^S} \displaystyle\sum_{S \in \{1, 2\}} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\substack{(g,h) \in orth,\\
+ (h,l) \in orth}}p_{\lambda_S}(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
 \end{equation}
@@
 In the fused L2 algorithm we develop, the penalty function is equal to the L2 norm of the difference in regulatory weight of fused coefficients.
-That is, $p_{\lambda_S}(x)=x^2$.
+That is, $p(x)=x^2$.
 The overall objective function for network inference is then:
@@
 
-Because every component of the objective function is an L2 norm, the entire problem is convex, and can be solved through linear regression with a suitable design matrix.
+Because every component of the objective function is an L2 norm, the entire problem is convex, and can in fact be solved through linear regression with an augmented design matrix.
 %The regulators of a pair of genes can be solved independently as long as the pair is not linked by fusion constraints, or a chain of fusion constraints; thus the problem of cross-species GRN inference decomposes into a large number of manageable regression problems as long as the orthology mappings are sparse.
@@
 The choice of $\lambda_S$ allows one to interpolate between fitting each network independently ($\lambda_S=0$) and pooling data together as if it came from one source ($\lambda_S=\inf$).
-Part of the appeal of the approach, however, is that it allows pooling of data even when there is incomplete orthology.
-By introducing constraints on the similarity of individual interactions, rather than on the networks as a whole \cite{liu2011temporal}, we can pool some information across species even when an arbitrarily small fraction of genes have orthologs.
+In addition to allowing interpolation between these extremes, our method allows pooling of data even when there is incomplete orthology.
+By introducing constraints on the similarity of individual interactions, rather than on the networks as a whole \cite{liu2011temporal}, we can pool some information across species even when a small fraction of genes have orthologs.
 This is particularly useful when dealing with a large number of species; pairwise orthologies may be nearly complete even when the number of genes present in every organism is small.
@@
 Fusion has been proposed as a method for utilizing correlation information from data by penalizing pairwise differences of coefficients \cite{Daye2009,Price2014}.
-Approaches using fusion combine statistical strength across the multiple regression tasks by introducing regularization functions reflecting prior knowledge on output structure representing how outputs are related \cite{kim_tree-guided_2012,Land1997,Chen2010,Petry2011,Hebiri2011}.
+Existing approaches using fusion have combined statistical strength across the multiple regression tasks by introducing regularization functions reflecting prior knowledge on output structure representing how outputs are related \cite{kim_tree-guided_2012,Land1997,Chen2010,Petry2011,Hebiri2011}.
 Price et al.
@@
 use a fused ridge estimator for jointly estimating multiple inverse covariance matrices \cite{Price2014a,Bilgrau2015}; we take a similar approach, adding an fused L2 penalty to our ridge regression to incorporate prior knowledge about relationships between input-output pairs.
+In many cases, however, there is some uncertainty about the relationships that should be enforced.
 Sohn et al.
@@
 
-We use the fact that ridge constraints can be incorporated into a least-squares regression problem by appending a scaled identity matrix to the design matrix, and a corresponding number of zeros to the response vector.
-Similarly, a fusion constraint $\lambda_S (\beta_{i} - \beta{j})^2$ can be incorporated into a least-squares regression problem by appending a row containing $\sqrt{lambda_S}$ in the $i$th position, $-\sqrt{lambda_S}$ in the $j$th position, and $0$s elsewhere to the design matrix, and zero to the response vector.
+We begin with the observation that ridge constraints can be incorporated into a least-squares regression problem by appending a scaled identity matrix to the design matrix, and a corresponding number of zeros to the response vector.
+Similarly, a fusion constraint $\lambda_S (\beta_{i} - \beta{j})^2$ can be incorporated into a least-squares regression problem by appending a row containing $\sqrt{\lambda_S}$ in the $i$th position, $-\sqrt{\lambda_S}$ in the $j$th position, and $0$s elsewhere to the design matrix, and zero to the response vector.
 
 In order to convert an optimization over multiple response variables (indexed by $i$) and multiple sources (indexed by $j$) with design matrices $X^{S_j}$ and response variables $y_i^{S_j}$ into an optimization with a single source and response variable, we vectorize as follows.
-Let $G$ be a set containing $(i,j)$ for each of the data sources and response variables currently being optimized.
-We construct the matrix $A$ by diagonally concatenating each of $X^{S_j}$ (with the remaining entries 0) and vertically concatenating each of $y_j^{S_j}$.
+Given a set containing $(i,j)$ for each of the data sources and response variables currently being optimized, we construct the matrix $A$ by diagonally concatenating each of $X^{S_j}$ (with the remaining entries 0) and vertically concatenating each of $y_j^{S_j}$.
 We then solve
@@
 This is equivalent to the original problem due to the block structure of matrix multiplication.
-In an ordinary regression problem each response variable can be solved independently as a large number of small regression problems.
+In an ordinary regression problem each response variable can be solved independently, and vectorization is unnecessary.
 However, in fused regression, we append additional rows to the design matrix $A$ that link entries of $W$ associated with different response variables.
@@
 We can then incorporate fusion constraints as in the case of single-source single response-variable fused regression.
-In most cases, we have found the direct solve using augmented matrices to be adequate.
+In most cases, we have found the direct solution using augmented matrices to be adequate.
 This is because the structure of orthology tends to be sparse, so only a small number of genes must be solved at once.
@@
 Linear systems of this size can be solved quickly using standard techniques.
-When the number of genes linked by fusion constraints becomes large, however, the augmented design matrix approach becomes impractical.
+When the size of the groups of genes linked by fusion constraints becomes large, however, the augmented design matrix approach becomes impractical.
 
@@
 Note that this is almost identical to equation 5, but now the network $\beta$ is a function of the iteration number $t$.
-On each step, the we compute $\beta$s that minimize a penalized cost function where the fusion penalty encourages similarity of a pair of parameters, using the previous iteration's solution.
+On each step, the we compute $\beta$s that minimize a penalized cost function where the fusion penalties encourage similarity between a parameter and its fused-to parameter from the previous iteration's solution.
 This process is iterated until the estimated $\beta$s converge.
 Because each iteration reduces the error between $\beta(t)$ and $\beta(t-1)$, and because $\beta(t) = \beta(t-1)$ is the globally optimal solution, this process must eventually converge to the same network as (cite 5).
-Although we have not produced bounds on the convergence rate, in practice a small number of iterations are necessary.
+Although we have not produced bounds on the convergence rate, which also depends on the structure of constraints, in practice a small number of iterations are necessary.
 
@@
 Some network inference approaches use L1 penalties to produce sparse networks, on the basis that biological networks are thought to be sparse.
-However, since we are penalizing differences in interaction weights, rather than the weights themselves, there's no reason to assume that most differences will be exactly zero.
+However, since we are penalizing differences in interaction weights, rather than the weights themselves, there's no reason to assume that most differences will be exactly zero, and an L2 penalty - equivalent to an assumption that the differences between fused coefficients are gaussian distributed - may be more appropriate.
 
@@
 $\beta^S(t+1)$ is obtained by fitting the ridge-fused model with fusion constraints given by the above $\lambda_S$.
-This is nice because all our penalties can be treated as L2 and therefore retain the properties of ridge regression.
+This is useful because all our penalties can be treated as L2 and therefore retain the properties of ridge regression, and can be solved using the fused L2 algorithm we develop.
 
@@
 Moreover, we are primarily interested in using this saturating penalty as a way of testing the hypothesis that conservation in GRNs can be predicted based off of known similarities between genes.
-Therefore, we implement a way of selecting $a$ to be equal to a certain percentile of the distribution of differences in fused interaction weights from the networks fit without fusion.
+Therefore, we propose a method for selecting $a$ to be equal to a certain percentile of the distribution of differences in fused interaction weights from the networks fit without fusion.
 This choice represents the user's hypothesis for the fraction of both networks that have unrelated function, and should therefore be unfused.
@@
 Because the purpose of adaptive fusion is to learn the structure of conservation while also learning networks, using simulated data allows us to test the method relative to a known gold standard for network conservation.
-Our simulations are based on the model $Y=X\beta+\sigma \epsilon$ where $\epsilion \sim N(0,1)$. Each data set consists of a training set and test set.
 
@@
 $X_{T1}$ is the TF expression sub-matrix of $Y_{T1}$, and consists of columns of $Y_{T1}$ that correspond to TFs.
-Treating the decay rate as 0, the gene expression matrix at timepoint two, $Y_{T2}$ is sampled as $Y_{T2} = Y_{T1} + BX_{T1}$.
+Treating the decay rate as 0, the gene expression matrix at timepoint two, $Y_{T2}$ is sampled as $Y_{T2} = Y_{T1} + BX_{T1} + \epsilon$, where $\epsilon$ is a Gaussian noise term.
 This process is carried out separately for each network.
@@
 Following generation of simulated data, we may introduce error into the orthology mapping.
-This can take the form of discarding a specified fraction of true orthologies (governed by a false-negative rate), by introducing random false orthologies (governed by a false-positive rate), or by adding Gaussian noise so that fused interactions are not identical.
+This can take the form of discarding a specified fraction of true orthologies (governed by a false-negative rate), by introducing random false orthologies (governed by a false-positive rate), or by adding Gaussian noise so that fused interactions are not identical (described above).
 For convenience, the false-positive rate is specified in units of the number of true orthologs, and not the number of possible orthologs.
@@
 \end{equation}
-Computing residuals with respect to the data alone would disregard information gained through fusion. Instead, we used an approximation
+Computing residuals with respect to the data alone would disregard information gained through fusion, because certain iteractions may be large due to fusion, rather than their individual explanatory power.
+Instead, we used an approximation
 \begin{equation}
@@
 It is known that network recovery improves with additional data \cite{bar-joseph_computational_2003}.
-Additional data is not always readily available; but often there is an abundance of related data, and we reason that we should be able to use data describing similar processes.
+Additional data is not always readily available; but often there is an abundance of related data.
+Our method allows this data to be used in network inference.
 Using related data for network inference allows us to borrow statistical power from other sources, effectively increasing the sample size and boosting the sensitivity and specificity of learned interactions.
 We use the assumption that similar genes have similar regulatory interactions, and share information about putatively conserved interactions.
-We created synthetic networks to approximate two related biological processes, then evaluated performance of our fused L2 regression, which learns the networks simultaneously, sharing information about related interactions (Figure 1b).
-We compared recovery of the 10 TF by 200 gene networks, using fused L2 versus learning networks separately.
-We varied the amounts of simulated expression data made available to the solver.
-When the amount of data from species two was held constant, increasing the amount of data available for learning the network for species one resulted in a more accurate network, as expected.
-When we increased the amount of data from species two, we obtained performance gains on network one using fused L2 regression, showing our ability to improve network inference on one dataset by sharing data from a related dataset.
+We created synthetic networks to approximate two related biological processes, then evaluated performance of our fused L2 regression, which learns the networks simultaneously given a prior on the relatedness of interactions (Figure 1b).
+We compared recovery of the 10 TF by 200 gene networks, using fused L2 versus learning networks separately, and varied the amounts of simulated expression data made available to the solver.
+When the amount of data from the second species was held constant, increasing the amount of data available for learning the network for the first species resulted in a more accurate network, as expected.
+When we increased the amount of data from the second species, we obtained performance gains on network one using fused L2 regression, demonstrating our ability to improve network inference on one dataset through incorporation of a related dataset.
 
-The decision of what other sources to draw from is an important one; at one extreme is combining unrelated data, and at the other is resampling from the same experiments.
-We simulated this spectrum by creating pairs of networks and varying the similarity between networks.
+At one extreme, when data is generated from identical processes, the weight assigned to fusion constraints should be very large.
+On the other extreme, when the networks have diverged considerably, a large fusion weight may impair network recovery.
+We performed an experiment on simulated data in a multi-species network-inference problem to assess the effect of the similarity of networks and of the weight of fusion on network recovery performance.
+We creating pairs of networks and varying the similarity between networks.
 The main factors governing similarity between our generated networks was the extent (and accuracy) of the orthology mapping, and the variability between analogous, or conserved, interactions (in other words, how well conserved the conserved subnetwork was).
@@
 As expected, when the conserved subgraph was very similar - when the interactions in the conserved subnetwork were nearly identical, simulating the case of closely related organisms or similar processes (Figure 2A) - increasing the weight of the fusion penalty $\lambda_S$ improved network recovery.
-As the size of the conserved subgraph increased, using a greater weight on the fusion penalty was beneficial.
+As the size of the conserved subgraph increased, this effect was enhanced.
 
-To simulate the case of distantly related organisms, we created networks where conserved interactions were drawn from high variance distribution (Figure 2B).
-When the conservation iss weak - when the interactions in the conserved subnetwork are nearly as different across species as randomly chosen interactions - performance may not continue to improve by increasing the fusion weight.
+To simulate the case of distantly related organisms, we created networks where the differences between conserved interaction weights were drawn from high variance distribution, mimicking weak conservation (Figure 2B).
+When the conservation is weak - when the interactions in the conserved subnetwork are nearly as different across species as randomly chosen interactions - performance may not continue to improve by increasing the fusion weight.
 We showed that even for networks where the conserved subgraph was weakly conserved, there exists a 'sweet spot' where fusion regression improves network recovery.
@@
 This can occur when using different technology, eg microarray and RNAseq, where one method does not assay all of the genes that another one does, but the networks are expected to be very similar; or when inferring networks describing distantly related species, in which case the networks are expected to be more different than similar.
-When orthology is not complete, we are interested in knowing if performance gains from fused regression are limited to those interactions which have fusion constraints.
+When orthology is not complete, we are interested in knowing if performance gains from fused regression are limited to those interactions which have fusion constraints, or if they extend to the entire network.
 To test this, we used sets of 20 TF by 200 gene synthetic networks, with varying proportions of orthologous TFs and genes.
@@
 We observed this, particularly for the constrained subnetwork.
-As $\lambda_S$ increased, interactions with fusion constraints were encouraged to be more similar; growth in performance slowed as the differences in interaction weights approached zero, and the small amount of added noise to the expression data and slight differences between the conserved networks limited the recovery from reaching completeness.
+As $\lambda_S$ increased, interactions with fusion constraints were encouraged to be more similar.
 Interestingly, performance gains were seen even in the portion of the network that was unconstrained by fusion.
@@
 Pairs of interactions which are very dissimilar even after fusion, which sit in the flat portion of this penalty function, are effectively ``unfused,'' and no further penalty is incurred as differences in interaction weights grow.
-Because our network inference problem is underdetermined, meaning that multiple networks would fit our data equally well, the ``unfusing'' or relaxation of the fusion penalty on constraints is convincing evidence for the rejection of a network which uses orthology to predict similarity in interaction weights.
+Because our network procedure strongly favors similarity of fused interactions, and only ``unfuses'' interactions when their similarity cannot be reconciled with expression data, the ``unfusing'' or relaxation of the fusion penalty on constraints is convincing evidence for the rejection of a network in which these interactions are conserved.
 ''Unfusing'' suggests neofunctionalization of one or more of the genes involved in this pair of interactions.
@@
 Because we knew which entries in the orthology mapping were ``incorrect'' (not reflected in the generation of the networks), we could correctly label fusion constraints that involved one or more ``incorrect'' mappings.
-We verified that adaptive-fusion unfused mostly ``incorrectly fused'' interactions, while leaving truly analogous interactions fused (Figure 4a.
+We verified that adaptive-fusion unfused mostly ``incorrectly fused'' interactions, while leaving truly analogous interactions fused (Figure 4a).
 Note that while pairs of interactions between true orthologs were sampled randomly so that the difference in interactions tended to be small, and pairs of interactions between incorrect orthologs were sampled randomly so that the difference in interactions tended to be large, there was an element of randomness in the sampling.
@@
 We compared the recovery of interaction weights which were accurately fused, and recovery of interaction weights which were inaccurately fused due to incorrect orthology information (Figure 4b).
-Because fused L2 heavily penalizes large differences between weights which are predicted to be similar, it is able to retrieve a more accurate network for those interactions with true fusion constraints than by learning networks separately.
-However, the gains accomplished by sharing information through fused regression do not extend to those interactions lacking true fusion constraints, and the error remains similar to learning networks separately.
-Because the penalty function used in adaptive fusion resembles that of fused L2 where differences in interactions are small (Figure 3), adaptive fusion's performance on accurately fused interactions is similar to that of fused L2.
-Because interactions which have been unfused remain underdetermined, we are not much closer to learning their values than if we learned networks separately.
-However, when using adaptive fusion, we are able to identify which interactions should not be similar, and therefore, which orthology mappings are not predictive of functional similarity.
+Because fused L2 heavily penalizes large differences between weights which are predicted to be similar, it is able to retrieve a more accurate network for those interactions with true fusion constraints than by learning networks separately (measured by MSE between the true and inferred interaction weights) .
+In this simulation, however, the gains accomplished through fused regression do not extend to those interactions lacking true fusion constraints, and the error remains similar to learning networks separately.
+When we applied adaptive fusion, we did not observe an improvement in network recovery (relative to fused L2).
+However, we were able to identify fusion constraints reflecting incorrect orthology information that had been provided to the algorithm.
+Even in situations where adaptive fusion does not substantially improve recovery - which may occur when unfused interactions remain underconstrained - the approach shows promise in identifying non-conserved interactions, which is of direct interest to systems biology.
 
@@
 
-When we applied fused L2 to our subsampled \textit{B.
+In order to evaluate network recovery performance, we opted to evaluate AUPR on the sub-network of genes with fusion constraints.
+Although simulation results demonstrate that the non-constrained sub-network exhibit benefits from fusion, these benefits are expected to be very small given the small fraction of interactions with constraints in our cross-species application (Figure 1c).
+Moreover, the evolutionarily conserved sub-network is often of primary interest in cross-species studies.
+Application of fused L2 to our subsampled \textit{B.
 subtilis} and \textit{B.
-anthracis} datasets using optimized parameters, we did not see a significant improvement over solving the networks separately.
-Inspection of precision recall curves showed that there is increased precision at low values of recall.
+anthracis} datasets produced performance gains relative to the $\lambda_S=0$ case, but these performance gains, measured across the whole network, were very small (Figure 5b).
+However, inspection of precision recall curves showed that there is substantially increased precision at low values of recall.
 Essentially, precision increases for the high confidence part of the network but not for the low confidence part of the network.
@@
 Although there are many large-scale collaborations which attempt to make protocols as uniform as possible for comparability between datasets generated by different labs \cite{paten_nih_2015,kundaje_integrative_2015}, there still exists technical and biologial variability between many experiments attempting to capture the same or similar experimental condition.
-With the advent of new technology, such as RNAseq and single-cell RNAseq, microarray is no longer the dominant assay for genome-wide expression, but it is important to be able to combine datasets generated using different technologies.
+With the advent of new technology, such as RNAseq and single-cell RNAseq, microarray is no longer the dominant assay for genome-wide expression, but a large body of accumulated legacy data remains useful, if it can be integrated with more modern techniques.
 Currently, the most widely used approach to combining datasets for network inference is to learn networks from disparate datasets separately, then rank combine the networks as in Marbach et al \cite{marbach_revealing_2010}.
@@
 subtilis} \cite{nicolas2012condition}.
-We compared performance when learning the networks separately, rank combining, and learning the networks simultaneously using fusion regression, and showed improvement on network inference using our fused L2 approach (Figure 6a).
+We compared performance when learning the networks separately, rank combining, and learning the networks simultaneously using fusion regression, and showed large improvement on network inference using our fused L2 approach (Figure 6a).
 Importantly, our approach simultaneously learns multiple networks, rather than learning networks separately than combining into one.
@@
 We tested a combination of our fused regression approach with a method for estimating transcription factor activities (TFA).
-Rather than modeling gene expression using transcription factor mRNA abundance, we fit gene expression as a function of transcription factor activity, as shown in \textit{B.
+Rather than modeling gene expression using transcription factor mRNA abundance, we fit gene expression as a function of transcription factor activity, as applied to \textit{B.
 subtilis} by Arrieta-Ortiz et al \cite{arrieta-ortiz_experimentally_2015}.
+TFA captures activation of transcription factor through mechanisms such as dimerization and interaction with required factors, over transcription factor expression level alone \cite{fu_reconstructing_2011}.
 We estimated TFA based on known regulatory interactions using network component analysis \cite{liao2003network}.
@@
 The rest we reserved as a gold standard for validation.
-As in previous studies, we observed a marked improvement in network inference when using transcription factor activity, which captures activation of transcription factor through mechanisms such as dimerization and interaction with required factors, over transcription factor expression level alone \cite{fu_reconstructing_2011}.
+As in previous studies, we observed a marked improvement in network inference when using transcription factor activity.
 We also obtained AUPR improvement when using fused regression on TFA, and showed that our gains from sharing information across datasets using fused regression were preserved and even enhanced by using TFA.

### 2016-03-02 paper1.tex
hi

@@
 \section{Methods}
+
+\subsection{Statistical approach and background}
+We consider prediction and coefficient estimation problems with $N$ observations of $M$ dependent variables $y_{1,1}, y_{2,1}, ...y_{N,1}, y_{N,2},..., y_{N, M}$ and features $x_{i,j}$, $i=1,2,...,N, j=1,2,...,p$.
+We begin with a standard linear regression model:
+
+\begin{equation}
+y_{i,k} = \displaystyle\sum_{j}x_{ij}\beta_{j,k} + \epsilon_i
+\end{equation}
+
+with errors $\epsilon_i$ having mean 0 and constant variance, and predictors $x_{ij}$ having mean 0 and unit variance.
+We are interested in the case where $p >> N$.
+Many methods have been proposed to deal with the underconstrained case, and have been applied to genomic data \cite{waldron_optimized_2011, li_network-constrained_2008}.
+For example, ridge regression penalizes the L2 norm of the coefficients $\beta_{i,j}$ in order to avoid overfitting \cite{hoerl_ridge_1970}, and can be thought of as a mean-zero Gaussian prior on the coefficients.
+
+More complicated penalties have been developed to represent certain expected or desireable structure in a regression model's coefficients.
+For example, Land and Friedman \cite{citeulike3780356} proposed a fusion penalty:
+
+\begin{equation}
+\text{argmin}_{\beta} \displaystyle\sum_{ij} || y_i - x_{ij}\beta{j} ||^2 + \sum\limits_{j=2}^p ||\beta_j - \beta_{j-1}]|^\alpha
+
+\end{equation}
+
+Which encourages smoothness of the estimated parameter vector.
+
+%This penalty utilizes correlation information from data by penalizing pairwise differences of coefficients \cite{Daye2009,Price2014}. 
+Existing approaches have used similar fusion penalties to combine statistical strength across multiple regression tasks by introducing regularization functions reflecting prior knowledge on output structure representing how outputs are related \cite{kim_tree-guided_2012,Land1997,Chen2010,Petry2011,Hebiri2011}.
+Price et al.
+and Bilgrau et al.
+use a fused ridge estimator for jointly estimating multiple inverse covariance matrices \cite{Price2014a,Bilgrau2015}; we take a similar approach, adding an L2 penalty on the differences between coefficients to the existing ridge penalty in order to incorporate prior knowledge about relationships between input-output pairs:
+
+\begin{equation}
+
+\argmin_{\beta} \displaystyle\sum \vert \vert X\beta - Y \vert \vert ^2 + \lambda_R \vert \vert \beta \vert \vert ^2 +  \lambda_S  \displaystyle \sum_{\beta_{g,k} \approx \beta_{h,l}} || \beta_{g,k} - \beta_{h,l} ||^2
+
+\end{equation}
+
+where $X$, $Y$, and $\beta$ are matrices, and $\beta_{g,k} \approx \beta_{h,l}$ denotes fusion between arbitrary entries of $\beta$.
+Note that, like ridge regression, this penalty can be thought of as representing a Gaussian prior on the coefficients $\beta$.
+In the case where $\beta$ is a column vector, introducing this penalty is equivalent to assuming that $\beta$ is sampled from a multivariate Gaussian with inverse covariance matrix $\Sigma^{-1} = \lambda_R + \displaystyle \sum_{\beta_g \approx \beta_g} \lambda_S (1_{g,g} + 1_{h,h} - 1_{g,h} - 1_{h,g})$, where $I$ denotes the identity matrix and $1_{i,j}$ a matrix of zeros with 1 in its $i, jth$ entry.
+In the case of a two-coefficient model with fusion between the coefficients, for example, fused L2 is equivalent to assuming a prior with variance $(\lambda_R + \lambda_S)/(\lambda_R^2+2\lambda_R\lambda_S)$ and covariance $\lambda_S/(\lambda_R^2+2\lambda_R\lambda_S)$.
+
+In many cases, however, there is some uncertainty about the relationships that should be enforced.
+Sohn et al.
+attempt to simultaneously learn the regression coefficients and the output structure \cite{sohn_joint_2012}.
+We take a simultaneous approach, by applying a penalty function bounded by a constant to produce unbiased estimators for large coefficients, combined with an L2 penalty, similar to SCAD-L2 \cite{Zeng2012}.
+
+\begin{equation}
+
+\argmin_{\beta} \displaystyle\sum \vert \vert X\beta - Y \vert \vert ^2 + \lambda_R \vert \vert \beta \vert \vert ^2 + \displaystyle \lambda_S  \displaystyle \sum_{\beta_{g,k} \approx \beta_{h,l}} p_{\lambda, a} (\beta_{g,k} - \beta_{h,l})
+
+\end{equation}
+
+where the penalty $p_{\lambda, a}$ has derivative
+
+\begin{equation}
+p'_{\lambda,a}(\theta) = \left\{
+    \begin{array}{lr}
+    2\lambda\theta & \text{if } \theta \leq \lambda\\
+    \text{max}(\lambda_S(a-\theta),0) & \text{if } \theta > {a \over 2}
+    \end{array}
+    \right.
+\end{equation}
+
+ This approach allows us to simultaneously learn the regression coefficients and evaluate the validity of our prior information.
+
+\subsection{Application}
 Although our approach is generalizable to a wide variety of multi-source network inference problems, we begin with the concrete example of network inference in two related species.
@@
 We then go on to introduce a method to test the assumption of conserved network structure, and to relax the associated constraints on pairs of interactions for which data does not support conservation.
-
-\subsection{\textit{B. subtilis} and \textit{B. anthracis} data and orthology}
-We used a dataset collected for PY79, a derivative of strain 168, available on GEO with accession number GSE67023, and a dataset using BSB1, another derivative of strain 168, available at GEO with accession number GSE27219.
-We used two datasets for \texti{B.
-anthracis}, transcription profiling during iron starvation (E-MEXP-2272 on ArrayExpress), and time series over the life cycle (E-MEXP-788 on ArayExpress).
-We ran Inparanoid to obtain orthology mapping for \textit{B.
-subtilis} and \textit{B.
-anthracis} \cite{ostlund_inparanoid_2010}
 
@@
 
-\subsection{Statistical approach and background}
-Most network inference problems have more predictors than experimental samples ($p >> N$).
-Many methods have been proposed for penalized regression, such as ridge regression \cite{hoerl_ridge_1970}, and applied to genomic data \cite{waldron_optimized_2011, li_network-constrained_2008}.
-In our application, we consider the case where related outputs (genes) are likely to be affected by common inputs (transcription factors).
-Fusion has been proposed as a method for utilizing correlation information from data by penalizing pairwise differences of coefficients \cite{Daye2009,Price2014}.
-Existing approaches using fusion have combined statistical strength across the multiple regression tasks by introducing regularization functions reflecting prior knowledge on output structure representing how outputs are related \cite{kim_tree-guided_2012,Land1997,Chen2010,Petry2011,Hebiri2011}.
-Price et al.
-and Bilgrau et al.
-use a fused ridge estimator for jointly estimating multiple inverse covariance matrices \cite{Price2014a,Bilgrau2015}; we take a similar approach, adding an fused L2 penalty to our ridge regression to incorporate prior knowledge about relationships between input-output pairs.
-In many cases, however, there is some uncertainty about the relationships that should be enforced.
-Sohn et al.
-attempt to simultaneously learn the regression coefficients and the output structure \cite{sohn_joint_2012}.
-We also take a simultaneous approach, where we apply a penalty function bounded by a constant to produce unbiased estimators for large coefficients, combined with an L2 penalty, similar to SCAD-L2 \cite{Zeng2012}.
-This approach allows us to simultaneously learn the regression coefficients and evaluate the validity of our prior information.
-
 \subsection{Solving fused L2 problems using augmented matrices}
@@
 Because a small change in $\lambda_S$ will tend to produce a small change in the network, solving for the $\lambda_S+\epsilon$ network given $\beta$s initialized to the $\lambda_S$ network requires a small number of iterations.
-As a consequence, the entire solution path corresponding to a large number of values of $\lambda_S= \lambda_S_1, \lambda_S_2, ..., \lambda_S_{max}$ can be solved efficiently by using the solution of each network as the initial guess for the next.
+As a consequence, the entire solution path corresponding to a large number of values of $\lambda_S= \lambda_{S_1}, \lambda_{S_2}, ..., \lambda_S_{max}$ can be solved efficiently by using the solution of each network as the initial guess for the next.
 
@@
 As in \cite{fan2001variable}, we use solve using iterative local quadratic approximation.
-
 Specifically, $\beta^S(t)$ is the network on iteration $t$.
-
 For each fused $B^{S_1}_{g,k} \approx B^{S_2}_{h,l}$ we define:
@@
 Only in the latter case is a large weight suggestive of a true regulatory interaction.
+
+\subsection{\textit{B. subtilis} and \textit{B. anthracis} data and orthology}
+We used a dataset collected for PY79, a derivative of strain 168, available on GEO with accession number GSE67023, and a dataset using BSB1, another derivative of strain 168, available at GEO with accession number GSE27219.
+We used two datasets for \texti{B.
+anthracis}, transcription profiling during iron starvation (E-MEXP-2272 on ArrayExpress), and time series over the life cycle (E-MEXP-788 on ArayExpress).
+We ran Inparanoid to obtain orthology mapping for \textit{B.
+subtilis} and \textit{B.
+anthracis} \cite{ostlund_inparanoid_2010}
 
@@
 
-In order to evaluate network inference performance, and for use as priors, we used a gold standard of 2896 known \textit{B.
+To assess network inference performance, and for use as priors, we used a gold standard of 3040 known \textit{B.
 subtilis} interactions with corresponding activation and repression sign.
-Of these 2896 priors, 968 had corresponding interactions in \textit{B.
+Of these 3040 priors, 968 had corresponding interactions in \textit{B.
 anthracis}.

### 2016-03-03 paper1.tex
cleaned up .tex and .bib

@@
 \title{Fused regression for multi-source gene regulatory network inference}
-\author{Kari Y. Lam, Zachary M. Westrick, Christian L. M\"{u}ller, Lionel Christiaen, Richard Bonneau}
+\author{Kari Y. Lam \and Zachary M. Westrick \and Christian L. M\"{u}ller \and Lionel Christiaen \and Richard Bonneau}
 \maketitle
@@
 \text{argmin}_{\beta} \displaystyle\sum_{ij} || y_i - x_{ij}\beta{j} ||^2 + \sum\limits_{j=2}^p ||\beta_j - \beta_{j-1}]|^\alpha
-
 \end{equation}
@@
 \begin{equation}
-
 \argmin_{\beta} \displaystyle\sum \vert \vert X\beta - Y \vert \vert ^2 + \lambda_R \vert \vert \beta \vert \vert ^2 +  \lambda_S  \displaystyle \sum_{\beta_{g,k} \approx \beta_{h,l}} || \beta_{g,k} - \beta_{h,l} ||^2
-
+\label{eqn:main}
 \end{equation}
@@
 \begin{equation}
-
 \argmin_{\beta} \displaystyle\sum \vert \vert X\beta - Y \vert \vert ^2 + \lambda_R \vert \vert \beta \vert \vert ^2 + \displaystyle \lambda_S  \displaystyle \sum_{\beta_{g,k} \approx \beta_{h,l}} p_{\lambda, a} (\beta_{g,k} - \beta_{h,l})
-
 \end{equation}
@@
     2\lambda\theta & \text{if } \theta \leq \lambda\\
-    \text{max}(\lambda_S(a-\theta),0) & \text{if } \theta > {a \over 2}
+    \text{max}(\lambda_S(a-\theta),0) & \text{if } \theta > \frac{a}{2}
     \end{array}
@@
 This process is iterated until the estimated $\beta$s converge.
-Because each iteration reduces the error between $\beta(t)$ and $\beta(t-1)$, and because $\beta(t) = \beta(t-1)$ is the globally optimal solution, this process must eventually converge to the same network as (cite 5).
+Because each iteration reduces the error between $\beta(t)$ and $\beta(t-1)$, and because $\beta(t) = \beta(t-1)$ is the globally optimal solution, this process must eventually converge to the same network as equation \ref{eqn:main}.
 Although we have not produced bounds on the convergence rate, which also depends on the structure of constraints, in practice a small number of iterations are necessary.
@@
 Because a small change in $\lambda_S$ will tend to produce a small change in the network, solving for the $\lambda_S+\epsilon$ network given $\beta$s initialized to the $\lambda_S$ network requires a small number of iterations.
-As a consequence, the entire solution path corresponding to a large number of values of $\lambda_S= \lambda_{S_1}, \lambda_{S_2}, ..., \lambda_S_{max}$ can be solved efficiently by using the solution of each network as the initial guess for the next.
+As a consequence, the entire solution path corresponding to a large number of values of $\lambda_S= \lambda_{S_1}, \lambda_{S_2}, ..., \lambda_{S_{max}}$ can be solved efficiently by using the solution of each network as the initial guess for the next.
 
@@
 We used a dataset collected for PY79, a derivative of strain 168, available on GEO with accession number GSE67023, and a dataset using BSB1, another derivative of strain 168, available at GEO with accession number GSE27219.
-We used two datasets for \texti{B.
+We used two datasets for \textit{B.
 anthracis}, transcription profiling during iron starvation (E-MEXP-2272 on ArrayExpress), and time series over the life cycle (E-MEXP-788 on ArayExpress).
@@
 We estimated TFA based on known regulatory interactions using network component analysis \cite{liao2003network}.
-To test the integration of this approach with our fused regression, we asssesed the combination of \texti{B.
+To test the integration of this approach with our fused regression, we asssesed the combination of \textit{B.
 subtilis} datasets, as in figure 6a, with the incorporation of TFA estimation.
@@
 For example, the meta information that two genes belong to the same operon suggests that they are likely to have a similar set of regulators \cite{lawrence_shared_2002}, but does little to inform the identity of those regulators.
-Meta-information about the structure of gene regulatory networks - specifically which pairs of interactions are \texti{a priori} likely to be similar to one another, can provide a powerful set of constraints to improve network inference performance.
+Meta-information about the structure of gene regulatory networks - specifically which pairs of interactions are \textit{a priori} likely to be similar to one another, can provide a powerful set of constraints to improve network inference performance.
 We present a general framework for gene regulatory network inference that incorporates this meta-information - termed fusion constraints - and apply the technique to the problem of simultaneous inference of regulatory networks in multiple species (\textit{B.

### 2016-04-06 paper1.tex
hi

@@
 To this end, methods for global network inference have been developed and applied to a variety of species.
-However, most approaches consider the problem of network inference independently in each species, despite evidence that gene regulation is conserved even in distantly related species.
-We introduce a method for multi-source network inference that allows simultaneous estimation of gene regulatory networks in multiple species or biological processes through the incorporation of priors based on known gene relationships such as orthology.
-This approach improves network inference performance even when the orthology mapping in incomplete.
+However, most approaches consider the problem of network inference independently in each species, despite evidence that gene regulation can be conserved even in distantly related species.
+We introduce a method for multi-source network inference that allows simultaneous estimation of gene regulatory networks in multiple species or biological processes through the introduction of priors based on known gene relationships such as orthology.
+These priors are incorporated using fused regression.
+This approach improves network inference performance even when orthology mapping and conservation are incomplete.
 We further refine this method with an algorithm that attempts to learn the true conserved subnetwork from a larger set of potentially conserved interactions.
-We demonstrate the utility of our method not only in cross species network inference, but also in merging data collected on different experimental platform and in incorporating binding similarity predictions.
+We demonstrate the utility of our method in cross species network inference and in merging data collected on different experimental platform and in incorporating binding similarity predictions.
 \end{abstract}
@@
 \section{Introduction}
-As the volume and variety of genome scale data have exploded, the goal of accurately modeling gene regulatory networks has become attainable \cite{bonneau_predictive_2007, ciofani_validated_2012, carro_transcriptional_2010}.
+As the volume and variety of genome scale data continues to increase in quantity and quality, the goal of accurately modeling gene regulatory networks has become attainable \cite{bonneau_predictive_2007, ciofani_validated_2012, carro_transcriptional_2010}.
 Large-scale data collection efforts have contributed to the development of high quality networks which accurately recapitulate biological processes, but most processes and organisms remain uncharacterized at the network level.
@@
 The problem of inferring related -- but not necessarily identical -- structure from related -- but not identical -- data is ubiquitous in biology.
-Multi-source network inference has applications for learning multiple networks in multiple related species, for learning networks associated with distinct processes within the same species, and for learning networks based on heterogenous data sources.
-
+Multi-source network inference has applications for learning multiple networks in related species, for learning networks associated with distinct processes within the same species, and for learning networks based on heterogenous data sources.
 Moreover, as it becomes possible to learn genome-wide regulatory networks, we can begin to compare and to test whether there is conservation of networks across species and biological processes.
@@
 
-We present two methods for network inference based on linear estimates of gene expression dynamics, extending existing linear dynamical-systems methods for network inference \cite{bonneau_predictive_2007, arrieta-ortiz_experimentally_2015}.
-The core of both methods is the observation that biological information about the relatedness of genes can be used to constrain which network coefficients should be similar to one another in a multi-source network inference problem (ie orthologous TFs should regulate orthologous genes), and that these constraints can be efficiently represented as penalties in a least-squares regression problem.
+We present two methods for network inference based on linear estimates of gene expression dynamics, extending existing dynamical-systems methods for network inference \cite{bonneau_predictive_2007, arrieta-ortiz_experimentally_2015}.
+The core of both methods is the observation that biological information about the relatedness of genes can be used to select which network coefficients should be similar to one another in a multi-source network inference problem (ie orthologous TFs should regulate orthologous genes), and that these constraints can be efficiently represented as penalties in a least-squares regression problem.
 Taking into account the similarity of putatively conserved interactions improves our ability to accurately describe TF-gene relationships on a genome-wide scale.
 
-Our first method -- fused ridge -- uses an L2 penalty on the differences between a priori similar interactions (termed fusion penalty), and is useful where the relationships between networks (similarity of genes between data sets) is reliable.
+Our first method -- fused ridge -- uses an L2 penalty on the differences between \textit{a priori} similar interactions (termed fusion penalty), and is useful where the relationships between networks (similarity of genes between data sets) is reliable.
 In the case where both networks contain an identical set of genes and TFs, this approach can be thought of as parametrically interpolating between treating the data sources separately and combining them together.
-However, our method allows useful pooling of data even when the overlap between genes is incomplete.
+In the case of multi-species, simply combining two datasets is both unwise -- because the networks may differ substantially -- but also potentially impossible, because the set of common of genes may be small.
+Our method allows useful pooling of data even when the overlap between genes is incomplete, or when orthology assignments depart from a strict one-to-one mapping.
 Our second method -- adaptive fusion -- uses a non-concave saturating fusion penalty to simultaneously infer the constrained networks and to learn which constraints should be relaxed (ie which parts of the network are genuinely different).
-With this approach, we seek to test the hypothesis of conservation of interactions across networks.
+With this approach, we seek to identify both conserved and divergent interactions between related networks.
 
-In the case of multiple species, there has been a great deal of work showing that functional conservation exists in gene regulatory networks even across large evolutionary distance \cite{satou2006gene, hinman2009evolution,tanay2005conservation,erwin2009evolution}.
-In our fused L2 approach, we make the assumption that for closely related species, if a pair of TFs is orthologous to one another, then their regulatory effects on a pair of orthologous genes should be similar.
-These orthology relationships form the basis of a set of constraints which favor -- but do not require -- networks in which orthologous transctiption factors regulate orthologous genes.
+In the case of multiple species, numerous studies showed that functional conservation exists in gene regulatory networks even across large evolutionary distance \cite{satou2006gene, hinman2009evolution,tanay2005conservation,erwin2009evolution}.
+In our fused L2 approach, we assume that for closely related species, orthologous TFs could exert similar regulatory effects on orthologous target genes.
+These orthology relationships form the basis of a set of constraints which favor -- but do not require -- networks in which orthologous transcription factors regulate orthologous genes.
 As a result, data in one species can improve network inference performance in another species  (and vice versa).
@@
 This is an advantage over existing approaches to multi-species network inference, which infer only a sub-network for which orthologs exist in every species \cite{joshi_multi-species_2015}.
-This approach of introducting constraints on the similarity between specific regulatory interactions can be extended beyond the case of multi-species network-inference from orthology; any biological prior on similarity can be used in place of orthology.
-For example, we can introduce constraints that favor genes in the same operon or having similar binding sites, having similar regulators.
+
+This approach of introducing constraints on the similarity between specific regulatory interactions can be extended beyond the case of multi-species network-inference from orthology; any biological prior on similarity can be used in place of orthology.
+For example, we can use fused regression to combine datasets obtained using different platforms or experimental techniques.
+Moreover, we can introduce constraints that favor genes in the same operon or having similar binding sites, towards having similar regulators.
 
@@
 In multi-species network inference, our fused L2 approach minimizes a cost function that strives to simultaneously fit expression data in each species and produce networks that are consistent with evolutionary constraints created using orthology.
-However, we recognize that some genes may have taken on different functions and therefore may have new regulatory interactions.
+However, some genes may have taken on different functions and therefore may have new regulatory interactions.
 For example, gene duplications may lead to neofunctionalization \cite{eisen_phylogenomics:_1998} of the duplicated genes.
-Even in the case of inferring networks from multiple cell-lines, changes in chromatin configuration may affect our assumptions about the similarity of interactions \cite{li_role_2007}.
+Even in the case of comparing networks from related cell lines from the same species, changes in chromatin configuration may affect our hypotheses about the similarity of interactions between pleiotropic TFs and target genes \cite{li_role_2007}.
 
-Understanding which interactions are functionally dissimilar is of direct biological interest, but existing approaches to network inference are unable to effectively rule out conservation.
+Understanding which interactions are functionally dissimilar is of direct biological interest, but existing approaches to network inference are unable to effectively rule out the existence of a conserved subnetworks.
 Observing a large difference in the weights of regulatory interactions obtained though independent inference of multiple networks is the best evidence against conservation available with existing network inference techniques.
-This evidence is very weak, however; because network inference is typically underconstrainted \cite{marbach_revealing_2010-1}, there could be a different set of networks for which conservation does hold, and which fit the data almost as well.
+This evidence is very weak, as network inference is typically underconstrained \cite{marbach_revealing_2010-1}, there could be a different set of networks for which conservation does hold, and which fit the data almost as well.
 We propose, with our adaptive fusion, an approach where network inference occurs simultaneously with the evaluation of conservation, to gather stronger evidence of divergence.
 
-We approach this problem by introducing a saturating penalty function based on work in statistics to develop unbiased regularization penalties for regression \cite{zhang2010nearly, fan2001variable}.
-The main difference between the L2 fusion approach and this adaptive fusion approach occurs when the difference between presumed analogous interactions is large despite the fusion penalty: we assume that large differences represent evidence that gene functions have diverged and represent this, in adaptive fusion, with a relaxation of the fusion penalty.
-The resulting cost function is non-convex and difficult to optimize \cite{fan2001variable}; however, we can approximate its solution and obtain deeper insight into functional similarity than is available through orthology or the comparison of separately learned networks.
+We approach this problem by introducing a saturating penalty function based on statistical efforts to develop unbiased regularization penalties for regression \cite{zhang2010nearly, fan2001variable}.
+The main difference between the L2 fusion approach and our new adaptive fusion approach occurs when the difference between presumed analogous interactions is large despite the fusion penalty: we assume that large differences represent evidence that TF-to-target-gene interactions have diverged.
+We represent this, in adaptive fusion, with a relaxation of the fusion penalty.
+The resulting cost function is non-convex and difficult to optimize \cite{fan2001variable}; however, we can approximate its solution and obtain deeper insight into functional similarity than is available through strict orthology enforcement or the comparison of separately learned networks.
 
 Although the fusion constraints we employ can be described as arising from orthology - which links genes - it is important to note that the constraints themselves link individual regulatory interactions.
-This finer level of representational granularity is critical to the functioning of adaptive fusion, and means the method can accomodate any form of prior on expected similarity between regulatory interactions, even priors that are not decomposible into a gene to gene mapping.
-In order to solve the resulting optimization problem, we develop two algorithms for solving efficiently multi-output least-squares regression problems with pairwise L2 fusion penalties on entries of the coefficient matrix.
+This finer level of representational granularity is critical to the functioning of adaptive fusion, and means the method can accommodate any form of prior on expected similarity between regulatory interactions, even priors that cannot be decomposed into gene to gene mappings.
+We develop two algorithms for solving efficiently multi-output least-squares regression problems with pairwise L2 fusion penalties on entries of the coefficient matrix.
 We also introduce - in the form of adaptive fusion - the idea of a saturating penalty function on fusion constraints, and estimate the solution to the resulting optimization problem through iterative application of the fused L2 algorithm.
 
-We test the ability of fused L2 and adaptive fusion to improve network recovery on both synthetic data and \textit{Bacillus subtilis} and \textit{Bacillus anthracis}, showing the applicability of our method in combining different datasets, and leveraging similarity across organisms as well as within a network in order to improve network inference.
-We explore the circumstances under which each approach is optimal, and evaluate the robustness of adaptive fusion to incorrect orthology, simulating the biologically relevant case of neofunctionalization.
+We test the ability of fused L2 and adaptive fusion to improve network recovery on both synthetic data and by comparing related networks in the bacteria species \textit{Bacillus subtilis} and \textit{Bacillus anthracis}.
+This shows the applicability of our method in combining different datasets and leveraging similarity across organisms as well as within a network in order to improve network inference.
+We explore the circumstances under which each approach is optimal, and evaluate the robustness of adaptive fusion to incorrect orthology, simulating the biologically relevant cases of neo- and subfunctionalizations.
 
@@
 with errors $\epsilon_i$ having mean 0 and constant variance, and predictors $x_{ij}$ having mean 0 and unit variance.
-We are interested in the case where $p >> N$.
+We are interested in the case where $p > N$.
 Many methods have been proposed to deal with the underconstrained case, and have been applied to genomic data \cite{waldron_optimized_2011, li_network-constrained_2008}.
@@
 Our approach to multi-species network inference is based on the hypothesis that gene regulation in related species is governed by similar but not necessarily identical gene regulatory networks, due to conservation of function through evolution.
-We represent conservation of network function by introducing constraints into the objective function for network inference that penalize differences between the weights of regulatory interactions believed to be a priori similar.
+We represent conservation of network function by introducing constraints into the objective function for network inference that penalize differences between the weights of regulatory interactions believed to be conserved.
 These constraints favor the generation of similar networks for related species, and in the generally under-constrained regime of network inference can improve the accuracy of network recovery.
@@
 \subsection{Gene regulatory network}
-We model the transcription rate of each gene as a weighted sum of transcription factor (TF) expression, and seek to identify the identities and regulatory weights of these TFs.
+
+We are interested in identifying genes which affect other genes' expression, and understanding how this gives rise to a given biological process.
+We know that transcription factors (TFs) can bind to DNA and activate or repress expression of other genes.
+By restricting ourselves only to TFs and gene interactions, we vastly reduce the space of possible models while considering biologically relevant cases.
+
+We model the transcription rate of each gene as a weighted sum of transcription factor expression, and seek to identify the identities and regulatory weights of these TFs.
 This formulation matches that of the existing Inferelator algorithm, which models gene expression with linear differential equations \cite{bonneau_inferelator:_2006-1}.
@@
 \end{equation}
-for time series
+where $j \neq i$ 
+for time series and 
 \begin{equation}
-0 = \sum \beta_{i,j}x_{j}(t_k)
+\alpha_{i}x_{i}(t_k) = \sum \beta_{i,j}x_{j}(t_k)
 \end{equation}
-for steady state
+for steady state.
+
 \noindent We can summarize these equations in matrix form as
@@
 \end{equation}
-
+where $Y$ is the gene expression matrix, $X$ is the TF expression matrix, and $\beta$ is the regulatory weights we are interested in learning.
 We are interested in learning $\beta$, the matrix representation of the gene regulatory network, where the weight in a given position represents the regulatory weight of a TF on a gene.
 Positive weights represent activation, negative weights represent repression, and 0 weights represent the absence of an interaction.
-
-\noindent The matrix $\beta$ can be solved using linear regression.
+The matrix $\beta$ can be solved using linear regression.
 Because there are typically far fewer conditions than possible regressors (TFs), we introduce a ridge regularization constraint with weight $\lambda_R$ and solve
@@
 This is similar to the formulation used in the Inferelator algorithm, which we extend to the case of simultaneously inferring multiple networks.
-\subsection{Estimating transcription factor activity and using TFA as regressors}
+
+Transcription factor expression is not always the best predictor of its gene targets' expression.
 When there exists a set of prior known interactions, we are able to estimate transcription factor activity (TFA) using network component analysis \cite{liao2003network}, as in \cite{arrieta-ortiz_experimentally_2015, fu_reconstructing_2011}, and use TFA as explanatory variables instead of transcription factor expression.
@@
 These constraints penalize differences between interaction weights in the networks of multiple species that are expected to be similar based on prior biological knowledge.
-We can then solve the penalized regression problems simultaneously, in order to obtain a GRN for each species.
+We can then solve the penalized regression problems simultaneously, in order to obtain a gene regulatory network (GRN) for each species.
 Consider the case of organisms $A$ and $B$, governed by GRNs $\beta^A$ and $\beta^B$ (the following approach applies equally well to more than two species but for simplicity we continue with the case of two species).
@@
 In terms of the above regression formulation, we expect that $\beta^A_{g,k} \approx \beta^B_{h,l}$, and include a penalty term $\lambda_Sp(\beta^A_{g,k} - \beta^B_{h,l})$ in the quantity being minimized in order to encourage similarity.
-The function $p(x)$ controls the shape of the relationship between weight dissimilarity and penalty, while parameter $\lambda_S$ controls the overall scaling of the penalization of differences between fused coefficients.
+The function $p(x)$ controls the shape of the relationship between weight dissimilarity and penalty, while scalar $\lambda_S$ controls the overall scaling of the penalization of differences between fused coefficients.
 $\lambda_S$ controls the tradeoff between fitting the expression data alone and producing a set of networks that conform to evolutionary prior knowledge.
@@
 %\argmin_{(\beta^A, \beta^B)} \displaystyle\sum_{S \in (A, B)} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\beta^{S_1}_{g,k} \approx \beta^{S_2}_{h,l}} p_{\lambda_S}(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
-\argmin_{\beta^S} \displaystyle\sum_{S \in \{1, 2\}} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\substack{(g,h) \in orth,\\
- (h,l) \in orth}}p_{\lambda_S}(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
+\argmin_{\beta^S} \displaystyle\sum_{S \in \{1, 2\}} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \lambda_S \displaystyle \sum_{\substack{(g,h) \in orth,\\
+ (h,l) \in orth}}(p(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l}))
 \end{equation}
@@
 
-\subsection{Solving fused L2 problems using augmented matrices}
-We begin with the problem of constructing a design matrix to solve a fused L2 regression problem with a single response variable, because problems with multiple response variables and with multiple data sources can be converted into this form through vectorization.
-We then go on to show that, although the vectorized solution involves solving an impractically large system of equations, under typical biological conditions the structure of constraints allow the problem to be broken up into many smaller subproblems.
-
-We begin with the observation that ridge constraints can be incorporated into a least-squares regression problem by appending a scaled identity matrix to the design matrix, and a corresponding number of zeros to the response vector.
-Similarly, a fusion constraint $\lambda_S (\beta_{i} - \beta{j})^2$ can be incorporated into a least-squares regression problem by appending a row containing $\sqrt{\lambda_S}$ in the $i$th position, $-\sqrt{\lambda_S}$ in the $j$th position, and $0$s elsewhere to the design matrix, and zero to the response vector.
-
-In order to convert an optimization over multiple response variables (indexed by $i$) and multiple sources (indexed by $j$) with design matrices $X^{S_j}$ and response variables $y_i^{S_j}$ into an optimization with a single source and response variable, we vectorize as follows.
-Given a set containing $(i,j)$ for each of the data sources and response variables currently being optimized, we construct the matrix $A$ by diagonally concatenating each of $X^{S_j}$ (with the remaining entries 0) and vertically concatenating each of $y_j^{S_j}$.
-We then solve
-%Specifically, if $Y \in R^{N \times M}$ then we form $A =$ diagonally concatenate $X$, $M$ times (with the remaining entries 0),
-%$y =$ vertically concatenate the columns of $Y$. We solve
-
-$$\underset{W}{min} ||AW - y||_2^{~2}$$
-
-%We can then define $W_{i,j} = \beta_{(j-1) \times N + i, (j-1) \times M + j}$ which is the desired form.
-This is equivalent to the original problem due to the block structure of matrix multiplication.
-In an ordinary regression problem each response variable can be solved independently, and vectorization is unnecessary.
-However, in fused regression, we append additional rows to the design matrix $A$ that link entries of $W$ associated with different response variables.
-As a result, these linked response variables must be solved simultaneously through vectorization.
-Two response variables are linked by a fusion constraint if any of the regulatory weights affecting those genes are linked by a fusion constraint.
-Two response variables must be solved simultaneously if there is any chain of linked response variables connecting them.
-However, every other response variable can be solved separately.
-In biological terms, the regulators of two genes (whether in the same species, or different species) must be solved together if there is a fusion constraint linking those genes' regulators, or if there is a chain of such constraints.
-This is important because vectorization produces a design matrix with a number of rows and columns equal to the sum of the number of rows and columns of the design matrices associated with each of the response variables (genes) being solved for.
-If the networks for a large number of genes are solved simultaneously, the system of equations can quickly become intractible to solve.
-
-In order to avoid this difficulty, we use depth-first search to identify linked columns of each TF expression matrix $Y^{S_j}$, then forming design and response matrices through vectorization.
-We can then incorporate fusion constraints as in the case of single-source single response-variable fused regression.
-In most cases, we have found the direct solution using augmented matrices to be adequate.
-This is because the structure of orthology tends to be sparse, so only a small number of genes must be solved at once.
-However, in the general case of $c_i$ response variables from each of $i \in \text{species}$ data sources the design matrix to be solved is of size $\displaystyle \sum_{c_i} rows(X^{S_i}) \times \displaystyle \sum_{c_i} cols(X^{S_i})$ plus a relatively small number of rows representing fusion constraints.
-Because the scaling of this algorithm has a complicated depence on the constraint structure used, a general description of its runtime is difficult.
-However, in the case of multi-species network inference with one-to-one orthology, the network associated with each pair of orthologous genes requires solving a linear system with approximately twice as many observations and unknowns as the single species case.
-Linear systems of this size can be solved quickly using standard techniques.
-When the size of the groups of genes linked by fusion constraints becomes large, however, the augmented design matrix approach becomes impractical.
-
-%where the $j$s refer to transcription factors.
-In subproblem 1, $j_1$ regulates gene $g$; $j_1$ is orthologous to $j_2$, a transcription factor of subproblem 2, which regulates $h$, a gene which is orthologous to gene $g$.
-There exists a chain because ($j_1$, $g$) is a TF, gene pair which is fused with ($j_2$, $h$), which is fused to ($j_3$, $k$).
-%We use depth-first search to identify linked columns of $Y$, and then solve the fused regression problem directly by vectorizing those columns.
-%In order to incorporate the fusion penalty, we append to the resulting design matrix one row \emph{per connected constraint}, which contains $\lambda_S$ in the column associated with $Y_g,i,j,$ and $-\lambda_S$ to the column associated with $Y_h, i',j'$ for constraint $(Y_g,i,j,Y_h,i',j')$.
-Call this resulting matrix $X_P$.
-%The vector $Y$ is augmented by appending an appropriate number of zeros to produce $Y_P$.
-As a result, each entry in the output of $X_P b$ in the columns associated with each constraint contain $\lambda_S(b_{i,j} - b_{i',j'})$.
-%These entries each contribute  $\lambda_S||b_{i,j} - b_{i',j'}||^{2}$ to the squared error $||X_Pb - Y_P||_2^{~2}$
-
-%In the context of regression for GRN inference with multiple cell sources, each gene $j$ from source $1$ has constraints $\forall_{i \in \{TFs\}} (1, i, j, 2, i’, j’)$ where $i',j'$ are the same gene in a different cell line.
-%As a result, each gene has only one constraint and the chain has length $L=2$.
-%In general, however, solving the fused regression problem directly, even after factoring constraints, involves at least one inversion of an $L\sum_{i=0}^S K_i \times L \sum_{i=0}^S K_i$ matrix, taking $O((L\sum_{i=0}^S K_i)^3)$.
-%In general, however, solving the fused regression problem directly, even after factoring constraints, involves at least one inversion of an $p\sum_{i=0}^S K_i \times p \sum_{i=0}^S K_i$ matrix where $p$ is the size of a column subset such that there is a constraint chain $L$ linking an element of each column to an element in every other column in that subset, taking $O((p\sum_{i=0}^S K_i)^3)$.
-%Although the sparse structure of $X_P$ makes this somewhat more tractable, it is still impractical for large $L$.
-
-\subsection{Iterative solver}
-
-When the number of genes in each fusion constraints is relatively small, the augmented design matrix approach we have described has the advantage of being efficiently solvable with a closed form solution.
-However, in some cases the augmented system of equations may be too large to solve efficiently with standard methods.
-To address this limitation, we developed an iterative solver that uses coordinate-wise descent to solve for solutions corresponding to a sequence of values of fusion penalty weights; since our fused L2 method uses a convex and differentiable penalty function, this approach converges to a global minimizer.
-Although less efficient than the augmented design matrix approach we developed, the iterative solver has the advantage of computing a solution path for $\lambda_S$.
-
-On each iteration $t$ the iterative solver computes
-
-\begin{equation}
-\text{argmin}_{\beta^{S_i(t)}, \beta^{S_j(t)}} \displaystyle\sum_{S_i, S_j \in \text{species}} ||X^{S_i}\beta^{S_i}(t) - Y^{S_i}||^2 + \lambda_R||\beta^{S_i}(t)||^2 + \displaystyle \sum_{\beta_{g,k}^{S_i} \approx \beta_{h,l}^{S_j}} \lambda_S(\beta^{S_i}_{g,k}(t) - \beta_{h,l}^{S_j}(t-1)^2)
-\end{equation}
-Note that this is almost identical to equation 5, but now the network $\beta$ is a function of the iteration number $t$.
-On each step, the we compute $\beta$s that minimize a penalized cost function where the fusion penalties encourage similarity between a parameter and its fused-to parameter from the previous iteration's solution.
-This process is iterated until the estimated $\beta$s converge.
-Because each iteration reduces the error between $\beta(t)$ and $\beta(t-1)$, and because $\beta(t) = \beta(t-1)$ is the globally optimal solution, this process must eventually converge to the same network as equation \ref{eqn:main}.
-Although we have not produced bounds on the convergence rate, which also depends on the structure of constraints, in practice a small number of iterations are necessary.
-
-The ierative solver is used for multi-source regression problems with complex orthology/similarity mappings, and also for solving the regularization path in order to pick regularization and fusion penalty weights.
-
-\subsection{Fusion and regularization path}
-The iterative solver requires an initial guess $\beta^{S_i}(0)$ for each species $S_i$.
-When the initial guess is close to the final solution, a small number of iterations are necessary.
-Because a small change in $\lambda_S$ will tend to produce a small change in the network, solving for the $\lambda_S+\epsilon$ network given $\beta$s initialized to the $\lambda_S$ network requires a small number of iterations.
-As a consequence, the entire solution path corresponding to a large number of values of $\lambda_S= \lambda_{S_1}, \lambda_{S_2}, ..., \lambda_{S_{max}}$ can be solved efficiently by using the solution of each network as the initial guess for the next.
-
-We use Tibshirani's cyclical coordinate descent algorithms from the 'glmnet' package \cite{friedman_regularization_2010} to compute a ridge regularization path.
-As jointly optimizing $\lambda_R$ and $\lambda_S$ is computationally prohibitive, we opted to optimize the parameters separately.
-First, we use glmnet to compute the regularization path for the ridge penalty, and given a range of $\lambda_R$ values, we fit models for each $\lambda_R$.
-We use cross validation to select the optimal parameter, the $\lambda_R$ value which minimizes the error of prediction on the leave out set.
-Following selection of $\lambda_R$, we search for optimal $\lambda_S$ again using cross validation.
-Note that both parameters are chosen without reference to the gold standard, which is used in a separate evaluation of network quality.
-
-%\subsection{Equivalent prior}
-%Our L2 fusion penalty is equivalent to assuming a Gaussian prior with variance proportional to $\frac{1}{\lambda_S}$ on differences in parameters with fusion constraints.
-Combined with the regularization constraints, this forms a multivariate Gaussian prior on weights in both networks.
-When using our method on systems with incomplete orthology mapping, this results in different prior probability distributions for those parameters with fusion constraints and those without.
-We adjust for this by solving for a constant to multiply fused interactions by to equalize the volumes of the prior distributions.
-The intent of this adjustment is to ensure that constrained interactions are not on average more highly penalized, which may tend to drive their weight towards zero, causing them to be excluded from the network.
-
 \subsection{Adaptive fusion}
-Fusion constraints are intended to penalize dissimilarity between interactions thought to be analogous based on some a priori knowledge.
+Fusion constraints are intended to penalize dissimilarity between interactions thought to be analogous based on \textit{a priori} knowledge.
 For example, orthology can be used to predict which interactions will be similar across species.
-Because fusion constraints are L2, interaction weights which differ from each other by a large amount are excessively penalized, which effectively ensures that fused interactions are assigned similar weights.
-This may be inappropriate for interactions which are identified based on orthology as being analogous, but which are no longer similar due to evolutionary changes.
-We propose that a saturating penalty may be useful for dealing with uncertainty about which interactions are conserved.
+When fusion constraints are L2, interaction weights which differ from each other by a large amount are excessively penalized, which effectively ensures that fused interactions are assigned similar weights.
+This will be inappropriate for interactions which are identified based on orthology as being analogous, but which are no longer similar due to evolutionary changes.
+We propose that a saturating penalty -- a penalty that is relaxed once differences in weights grow beyond a certain point -- may be useful for dealing with uncertainty about which interactions are conserved.
 
@@
 Some network inference approaches use L1 penalties to produce sparse networks, on the basis that biological networks are thought to be sparse.
-However, since we are penalizing differences in interaction weights, rather than the weights themselves, there's no reason to assume that most differences will be exactly zero, and an L2 penalty - equivalent to an assumption that the differences between fused coefficients are gaussian distributed - may be more appropriate.
+However, since we are penalizing differences in interaction weights, rather than the weights themselves, there's no reason to assume that most differences will be exactly zero, and an L2 penalty - equivalent to an assumption that the differences between fused coefficients are Gaussian distributed - may be more appropriate.
 
@@
 
-As in \cite{fan2001variable}, we use solve using iterative local quadratic approximation.
+As in \cite{fan2001variable}, we solve using iterative local quadratic approximation.
 Specifically, $\beta^S(t)$ is the network on iteration $t$.
@@
 Our adaptive penalty function introduces, in addition to regularization and fusion penalty weights $\lambda_R$ and $\lambda_S$, an unknown parameter $a$.
-We could employ grid search using cross-validation to searach for the best parameters, but for many data sets, this can be computationally expensive.
+We could employ grid search using cross-validation to search for the best parameters, but for many data sets, this can be computationally expensive.
 Moreover, we are primarily interested in using this saturating penalty as a way of testing the hypothesis that conservation in GRNs can be predicted based off of known similarities between genes.
@@
 This choice represents the user's hypothesis for the fraction of both networks that have unrelated function, and should therefore be unfused.
+
+\subsection{Solving fused L2 problems using augmented matrices}
+We begin with the problem of constructing a design matrix to solve a fused L2 regression problem with a single response variable, because problems with multiple response variables and with multiple data sources can be converted into this form through vectorization.
+We then go on to show that, although the vectorized solution involves solving an impractically large system of equations, under typical biological conditions the structure of constraints allow the problem to be broken up into many smaller subproblems.
+
+We begin with the observation that ridge constraints can be incorporated into a least-squares regression problem by appending a scaled identity matrix to the design matrix, and a corresponding number of zeros to the response vector.
+Similarly, a fusion constraint $\lambda_S (\beta_{i} - \beta{j})^2$ can be incorporated into a least-squares regression problem by appending a row containing $\sqrt{\lambda_S}$ in the $i$th position, $-\sqrt{\lambda_S}$ in the $j$th position, and $0$s elsewhere to the design matrix, and zero to the response vector.
+In order to convert an optimization over multiple response variables and multiple sources into an optimization with a single source and response variable, we vectorize as follows: we construct a new design matrix by diagonally concatenating design matrices from relevant regression problems, and create a new response vector by vertically concatenating corresponding response vectors.
+
+%$$\underset{W}{min} ||AW - y||_2^{~2}$$
+
+%We can then define $W_{i,j} = \beta_{(j-1) \times N + i, (j-1) \times M + j}$ which is the desired form.
+This is equivalent to the original problem due to the block structure of matrix multiplication.
+In an ordinary regression problem each response variable can be solved independently, and vectorization is unnecessary.
+However, in fused regression, we append additional rows to the design matrix that link entries of the interaction weight matrix associated with different response variables.
+As a result, these linked response variables must be solved simultaneously through vectorization.
+Two response variables are linked by a fusion constraint if any of the regulatory weights affecting those genes are linked by a fusion constraint.
+Two response variables must be solved simultaneously if there is any chain of linked response variables connecting them.
+However, every other response variable can be solved separately.
+In biological terms, the regulators of two genes (whether in the same species, or different species) must be solved together if there is a fusion constraint linking those genes' regulators, or if there is a chain of such constraints.
+This is important because vectorization produces a design matrix with a number of rows and columns equal to the sum of the number of rows and columns of the design matrices associated with each of the response variables (genes) being solved for.
+If the networks for a large number of genes are solved simultaneously, the system of equations can quickly become intractible to solve.
+
+In order to avoid this difficulty, we use depth-first search to identify linked columns of each TF expression matrix, then form design and response matrices through vectorization.
+We can then incorporate fusion constraints as in the case of single-source single response-variable fused regression.
+In most cases, we have found the direct solution using augmented matrices to be adequate.
+This is because the structure of orthology tends to be sparse, so only a small number of genes must be solved at once.
+In the general case, the size of the design matrix is proportional to the number of response variables that must be solved simultaneously.
+Because the scaling of this algorithm has a complicated dependence on the constraint structure used, a general description of its runtime is difficult.
+However, in the case of multi-species network inference with one-to-one orthology, the network associated with each pair of orthologous genes requires solving a linear system with approximately twice as many observations and unknowns as the single species case.
+Linear systems of this size can be solved quickly using standard techniques.
+When the size of the groups of genes linked by fusion constraints becomes large, however, the augmented design matrix approach becomes impractical.
+
+%where the $j$s refer to transcription factors.
+In subproblem 1, $j_1$ regulates gene $g$; $j_1$ is orthologous to $j_2$, a transcription factor of subproblem 2, which regulates $h$, a gene which is orthologous to gene $g$.
+There exists a chain because ($j_1$, $g$) is a TF, gene pair which is fused with ($j_2$, $h$), which is fused to ($j_3$, $k$).
+%We use depth-first search to identify linked columns of $Y$, and then solve the fused regression problem directly by vectorizing those columns.
+%In order to incorporate the fusion penalty, we append to the resulting design matrix one row \emph{per connected constraint}, which contains $\lambda_S$ in the column associated with $Y_g,i,j,$ and $-\lambda_S$ to the column associated with $Y_h, i',j'$ for constraint $(Y_g,i,j,Y_h,i',j')$.
+Call this resulting matrix $X_P$.
+%The vector $Y$ is augmented by appending an appropriate number of zeros to produce $Y_P$.
+As a result, each entry in the output of $X_P b$ in the columns associated with each constraint contain $\lambda_S(b_{i,j} - b_{i',j'})$.
+%These entries each contribute  $\lambda_S||b_{i,j} - b_{i',j'}||^{2}$ to the squared error $||X_Pb - Y_P||_2^{~2}$
+
+%In the context of regression for GRN inference with multiple cell sources, each gene $j$ from source $1$ has constraints $\forall_{i \in \{TFs\}} (1, i, j, 2, i’, j’)$ where $i',j'$ are the same gene in a different cell line.
+%As a result, each gene has only one constraint and the chain has length $L=2$.
+%In general, however, solving the fused regression problem directly, even after factoring constraints, involves at least one inversion of an $L\sum_{i=0}^S K_i \times L \sum_{i=0}^S K_i$ matrix, taking $O((L\sum_{i=0}^S K_i)^3)$.
+%In general, however, solving the fused regression problem directly, even after factoring constraints, involves at least one inversion of an $p\sum_{i=0}^S K_i \times p \sum_{i=0}^S K_i$ matrix where $p$ is the size of a column subset such that there is a constraint chain $L$ linking an element of each column to an element in every other column in that subset, taking $O((p\sum_{i=0}^S K_i)^3)$.
+%Although the sparse structure of $X_P$ makes this somewhat more tractable, it is still impractical for large $L$.
+
+\subsection{Solving fused L2 problems using iterative solver}
+
+When the number of genes in each fusion constraints is relatively small, the augmented design matrix approach we have described has the advantage of being efficiently solvable with a closed form solution.
+However, in some cases the augmented system of equations may be too large to solve efficiently with standard methods.
+To address this limitation, we developed an iterative solver that uses coordinate-wise descent to solve for solutions corresponding to a sequence of values of fusion penalty weights; since our fused L2 method uses a convex and differentiable penalty function, this approach converges to a global minimizer.
+Although less efficient than the augmented design matrix approach we developed, the iterative solver has the advantage of computing a solution path for $\lambda_S$.
+
+On each iteration $t$ the iterative solver computes
+
+\begin{equation}
+\text{argmin}_{\beta^{S_i(t)}, \beta^{S_j(t)}} \displaystyle\sum_{S_i, S_j \in \text{species}} ||X^{S_i}\beta^{S_i}(t) - Y^{S_i}||^2 + \lambda_R||\beta^{S_i}(t)||^2 + \lambda_S\displaystyle \sum_{\beta_{g,k}^{S_i} \approx \beta_{h,l}^{S_j}} ||\beta^{S_i}_{g,k}(t) - \beta_{h,l}^{S_j}(t-1)||^2
+\end{equation}
+Note that this is almost identical to equation 5, but now the network $\beta$ is a function of the iteration number $t$.
+On each step, we compute $\beta$s that minimize a penalized cost function where the fusion penalties encourage similarity between a parameter and its fused-to parameter from the previous iteration's solution.
+This process is iterated until the estimated $\beta$s converge.
+Because each iteration reduces the error between $\beta(t)$ and $\beta(t-1)$, and because $\beta(t) = \beta(t-1)$ is the globally optimal solution, this process must eventually converge to the same network as equation \ref{eqn:main}.
+Although we have not produced bounds on the convergence rate, which also depends on the structure of constraints, in practice a small number of iterations are necessary.
+
+The ierative solver is used for multi-source regression problems with complex orthology/similarity mappings, and also for solving the regularization path in order to pick regularization and fusion penalty weights.
+
+\subsection{Fusion and regularization path}
+Because optimizing over both parameters, $\lambda_R$ and $\lambda_S$, is computationally prohibitive, we opted to optimize the two parameters separately.
+Our procedure first optimized $\lambda_R$ with $\lambda_S=0$, then optimized $\lambda_S$ using this value of $\lambda_R$.
+Although this procedure is sub-optimal, it is guaranteed to achieve the best unfused solution in the case when $\lambda_S$ is constrained at 0.
+As a result, any performance gains of fused regression are a lower bound on the highest achievable performance gains.
+
+To optimize $\lambda_R$ we use cyclical coordinate descent algorithms from the 'glmnet' package \cite{friedman_regularization_2010} to compute a ridge regularization path.
+We use cross validation to select the optimal $\lambda_R$ parameter from this path, selecting the $\lambda_R$ which minimizes the average error of prediction on a leave out set across cross validation folds.
+Following selection of $\lambda_R$, we search for optimal $\lambda_S$ by computing the solution path from the iterative solver (using the sequence of successive model weights) again using cross validation to select the optimal parameter.
+Note that both parameters are chosen without reference to the gold standard, which is used in a separate evaluation of network quality.
+
+%\subsection{Equivalent prior}
+%Our L2 fusion penalty is equivalent to assuming a Gaussian prior with variance proportional to $\frac{1}{\lambda_S}$ on differences in parameters with fusion constraints.
+Combined with the regularization constraints, this forms a multivariate Gaussian prior on weights in both networks.
+When using our method on systems with incomplete orthology mapping, this results in different prior probability distributions for those parameters with fusion constraints and those without.
+We adjust for this by solving for a constant to multiply fused interactions by to equalize the volumes of the prior distributions.
+The intent of this adjustment is to ensure that constrained interactions are not on average more highly penalized, which may tend to drive their weight towards zero, causing them to be excluded from the network.
 
@@
 
-\subsection{Beta scaling}
+\subsection{Ranking regulatory hypotheses}
 In previous work, betas were rescaled as to form a matrix of confidence scores $S$ as follows
@@
 \end{equation}
-Computing residuals with respect to the data alone would disregard information gained through fusion, because certain iteractions may be large due to fusion, rather than their individual explanatory power.
+Computing residuals with respect to the data alone would disregard information gained through fusion, because certain interactions may be large due to fusion, rather than their individual explanatory power.
 Instead, we used an approximation
 \begin{equation}
-S_{i,j} = \frac{\sigma^2_{\text{full model for }y_j}}{\sigma^2_{\text{full model for }y_j} + B_{i,j}^2 \times var(TF_j)}
+S_{i,j} = \frac{\sigma^2_{\text{full model for }y_j}}{\sigma^2_{\text{full model for }y_j} + \beta_{i,j}^2 \times var(TF_j)}
 \end{equation}
-When the residual is uncorrelated with the prediction, then removing regressor $i$ with weight $B_{i,j}^2$ from the model of $j$ will increase the residual by $B_{i,j}^2$ times the variance of regressor $i$.
-In the case of regularized regression, the regressor and residual need not be uncorrelated, so the approximation will not hold exactly.
+When the residual is zero, then removing regressor $i$ with weight $\beta_{i,j}$ from the model of $j$ will increase the residual by $\beta_{i,j}^2$ times the variance of regressor $i$.
 This scaling is similar to rescaling according to variance explained relative to an augmented design matrix that includes fusion constraints.
-The intuition for scaling that a large $B_{i,j}$ may be inferred because of overfitting with a TF $i$ varies little across the available data, or because the regulatory weight is strong.
+The intuition for scaling is that a large $\beta_{i,j}$ may be inferred because of overfitting with a TF that varies little across the available data, or because the regulatory weight is strong.
 Only in the latter case is a large weight suggestive of a true regulatory interaction.
@@
 \section{Results}
-We tested the ability of fused regression to improve network inference performance, as well as the ability of adaptive fusion to learn the conservation of gene interactions, using both synthetic networks and real data.
-For the synthetic data, we generated random pairs of networks in which orthologous genes have similar regulatory interactions, then sampled gene expression from these networks, which we used to derive networks for comparison with the true networks.
+We used both synthetic networks and real data to test the ability of fused regression to improve the performance of network inference, and the ability of adaptive fusion to identify conserved interactions between orthologous genes.
+For the synthetic data, we generated random pairs of networks in which orthologous genes have similar regulatory interactions, then sampled gene expression from these networks, which we used to derive learned networks for comparison with the input (true) networks.
 For real data, we computed recovery of a known gold standard in \textit{Bacillus subtilis}.
@@
 \subsection{Using fused regression to learn related networks}
-It is known that network recovery improves with additional data \cite{bar-joseph_computational_2003}.
-Additional data is not always readily available; but often there is an abundance of related data.
+It is known that the accuracy of network inference improves with additional data \cite{bar-joseph_computational_2003}.
+Additional data is not always readily available and does not always make identification of unique network weights possible; but often there is an abundance of related data.
 Our method allows this data to be used in network inference.
@@
 We use the assumption that similar genes have similar regulatory interactions, and share information about putatively conserved interactions.
-We created synthetic networks to approximate two related biological processes, then evaluated performance of our fused L2 regression, which learns the networks simultaneously given a prior on the relatedness of interactions (Figure 1b).
-We compared recovery of the 10 TF by 200 gene networks, using fused L2 versus learning networks separately, and varied the amounts of simulated expression data made available to the solver.
-When the amount of data from the second species was held constant, increasing the amount of data available for learning the network for the first species resulted in a more accurate network, as expected.
+We created synthetic networks to approximate two related biological processes, then evaluated performance of our fused L2 regression, which learns the networks simultaneously given a prior on the relatedness of interactions.
+We compared recovery of the two artificial 10 TFs by 200 genes networks, using fused L2 versus learning networks separately, and varied the amounts of simulated expression data samples made available to the solver.
+We measured network inference accuracy with the mean squared error between generated (true) network weights and inferred weights.
+When the amount of data from the second species was held constant, increasing the amount of data available for learning the network for the first species resulted in a more accurate network prediction, as expected  (figure \ref{fusion}b).
 When we increased the amount of data from the second species, we obtained performance gains on network one using fused L2 regression, demonstrating our ability to improve network inference on one dataset through incorporation of a related dataset.
 
-At one extreme, when data is generated from identical processes, the weight assigned to fusion constraints should be very large.
+When disparate data is generated from identical processes, the weight assigned to fusion constraints should be very large.
 On the other extreme, when the networks have diverged considerably, a large fusion weight may impair network recovery.
-We performed an experiment on simulated data in a multi-species network-inference problem to assess the effect of the similarity of networks and of the weight of fusion on network recovery performance.
-We creating pairs of networks and varying the similarity between networks.
-The main factors governing similarity between our generated networks was the extent (and accuracy) of the orthology mapping, and the variability between analogous, or conserved, interactions (in other words, how well conserved the conserved subnetwork was).
-We conducted a series of simulations to assess the effect of increasing orthology coverage on networks.
-In order to determine the effect of the size of the orthology mapping on network inference performance, we simulated a series of 20 TFs by 200 genes networks at several values of percent orthology coverage.
+We performed an experiment on simulated data in a multi-species network-inference problem to assess the effect of the similarity of networks, and the effect of the fusion weight on network recovery (figure \ref{syntheticgrid}).
+We created pairs of synthetic networks with varying similarity.
+The main factors governing similarity between our generated networks was the extent (and accuracy) of the orthology mapping, and the variability between conserved, interactions (in other words, how well conserved the conserved subnetwork was).
+We conducted a series of simulations to assess the effect of increasing orthology coverage on network recovery.
 
-As expected, when the conserved subgraph was very similar - when the interactions in the conserved subnetwork were nearly identical, simulating the case of closely related organisms or similar processes (Figure 2A) - increasing the weight of the fusion penalty $\lambda_S$ improved network recovery.
+As expected, when the conserved subgraphs were very similar - when the interactions in the conserved subnetwork were nearly identical, simulating the case of closely related organisms or similar processes (figure \ref{syntheticgrid}a) - increasing the weight of the fusion penalty $\lambda_S$ improved network recovery.
 As the size of the conserved subgraph increased, this effect was enhanced.
 
-To simulate the case of distantly related organisms, we created networks where the differences between conserved interaction weights were drawn from high variance distribution, mimicking weak conservation (Figure 2B).
+To simulate the case of distantly related organisms, we created networks where the differences between conserved interaction weights were drawn from high variance distribution, mimicking weak conservation (figure \ref{syntheticgrid}b).
 When the conservation is weak - when the interactions in the conserved subnetwork are nearly as different across species as randomly chosen interactions - performance may not continue to improve by increasing the fusion weight.
@@
 Our method learns related networks simultaneously, using predictions about conservation of TF-gene interactions to share information between sources.
-Our approach is useful for learning networks from similar sources such as different cell-lines from the same species, where there exists a one-to-one mapping of genes, as well as datasets where the orthology mapping does not span all the genes.
-This can occur when using different technology, eg microarray and RNAseq, where one method does not assay all of the genes that another one does, but the networks are expected to be very similar; or when inferring networks describing distantly related species, in which case the networks are expected to be more different than similar.
+Our approach is useful for learning networks from similar sources such as related cell types from the same species, where there exists a one-to-one mapping of genes, as well as datasets where the orthology mapping does not span all the genes.
+This can occur when using different technology, eg microarray and RNAseq, where one method does not assay all of the genes that another one does, but the networks are expected to be very similar; an example of the latter case occurs when inferring networks describing distantly related species, in which case the networks are expected to be more different than similar.
 When orthology is not complete, we are interested in knowing if performance gains from fused regression are limited to those interactions which have fusion constraints, or if they extend to the entire network.
 To test this, we used sets of 20 TF by 200 gene synthetic networks, with varying proportions of orthologous TFs and genes.
-We divided networks into those interactions with fusion constraints (interactions where there exist putative analogous interactions in the second network based off of the orthology mapping between TF, gene pairs), which we call the constrained subnetwork, and interactions without fusion constraints, which we refer to as the non-constrained subnetwork.
-We varied the weight on the fusion penalty, lamS, and evaluated performance by computing AUPR on the constrained subnetwork, the non-constrained subnetwork, and the whole network (Figure 1C).
-Since the conserved subgraphs were similar to each other (more like Figure 2A than 2B), we expected performance to improve as the fusion penalty weight increased.
+We divided networks into those interactions with fusion constraints (interactions where there exist putative analogous interactions in the second network based on the orthology mapping between TF, gene pairs), which we call the constrained subnetwork, and interactions without fusion constraints, which we refer to as the non-constrained subnetwork.
+We varied the weight on the fusion penalty, lamS, and evaluated performance by computing AUPR on the constrained subnetwork, the non-constrained subnetwork, and the whole network (figure \ref{fusion}c).
+Since the conserved subgraphs were similar to each other (more like figure \ref{syntheticgrid}a than \ref{syntheticgrid}b), we expected performance to improve as the fusion penalty weight increased.
 We observed this, particularly for the constrained subnetwork.
@@
 We recognize that orthology prediction is not a perfect proxy for functional conservation \cite{gabaldon_functional_2013, studer_how_2009, nehrt_testing_2011}.
-We implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (Figure 3).
+We implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (figure \ref{adapt-schematic}).
 Pairs of interactions which are very dissimilar even after fusion, which sit in the flat portion of this penalty function, are effectively ``unfused,'' and no further penalty is incurred as differences in interaction weights grow.
-Because our network procedure strongly favors similarity of fused interactions, and only ``unfuses'' interactions when their similarity cannot be reconciled with expression data, the ``unfusing'' or relaxation of the fusion penalty on constraints is convincing evidence for the rejection of a network in which these interactions are conserved.
-''Unfusing'' suggests neofunctionalization of one or more of the genes involved in this pair of interactions.
+Our network procedure strongly favors similarity of fused interactions, and only ``unfuses'' interactions when their similarity cannot be reconciled with expression data.
+Interactions are only ``unfused'' if there is no network configuration consistent with data from both organisms.
+As a result, the ``unfusing'' or relaxation of the fusion penalty on certain constraints is much more direct evidence for neofunctionalization than comparing separately fit networks could provide.
 
-We performed a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved when orthologs are not functionally analogous  (Figure 4).
+We performed a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved (figure \ref{adaptivefusion}).
 We generated synthetic fused networks and introduced error in the fusion constraints by adding false positives and negatives to the orthology information given to the solver.
 Because we knew which entries in the orthology mapping were ``incorrect'' (not reflected in the generation of the networks), we could correctly label fusion constraints that involved one or more ``incorrect'' mappings.
-We verified that adaptive-fusion unfused mostly ``incorrectly fused'' interactions, while leaving truly analogous interactions fused (Figure 4a).
+We verified that adaptive-fusion unfused mostly ``incorrectly fused'' interactions (figure \ref{adaptivefusion}a red dots), while leaving truly analogous interactions fused (figure \ref{adaptivefusion}a green dots).
 Note that while pairs of interactions between true orthologs were sampled randomly so that the difference in interactions tended to be small, and pairs of interactions between incorrect orthologs were sampled randomly so that the difference in interactions tended to be large, there was an element of randomness in the sampling.
-This resulted in a small fraction of interactions between incorrect othologs being similar, and similarly, a small fraction of interactions between correct orthologs being different.
+This resulted in a small fraction of interactions between incorrect orthologs being similar, and similarly, a small fraction of interactions between correct orthologs being different.
 Consequently, some interactions between incorrect orthologs remained fused and some interactions between correct orthologs were unfused.
 
-We compared the recovery of interaction weights which were accurately fused, and recovery of interaction weights which were inaccurately fused due to incorrect orthology information (Figure 4b).
+We compared the recovery of interaction weights which were accurately fused, and recovery of interaction weights which were inaccurately fused due to incorrect orthology information (figure \ref{adaptivefusion}b).
 Because fused L2 heavily penalizes large differences between weights which are predicted to be similar, it is able to retrieve a more accurate network for those interactions with true fusion constraints than by learning networks separately (measured by MSE between the true and inferred interaction weights) .
@@
 When we applied adaptive fusion, we did not observe an improvement in network recovery (relative to fused L2).
-However, we were able to identify fusion constraints reflecting incorrect orthology information that had been provided to the algorithm.
+However, we were able to identify fusion constraints reflecting incorrect orthology information that had been provided to the algorithm (figure \ref{adaptivefusion}a).
 Even in situations where adaptive fusion does not substantially improve recovery - which may occur when unfused interactions remain underconstrained - the approach shows promise in identifying non-conserved interactions, which is of direct interest to systems biology.
@@
 \subsection{Cross-species network inference using bacterial data}
-We used gene-expression data from \textit{\textit{Bacillus subtilis}} and \textit{\textit{B.
-anthracis}} in order to assess performance gains of fused regression on real data.
-Our B subtilis data set consists of 360 time-series and steady-state observations of 4891 genes during development.
+We used gene-expression data from \textit{Bacillus subtilis} and \textit{B.
+anthracis} in order to assess performance gains of fused regression on real data.
+Our \textit{B subtilis} data set consists of 360 time-series and steady-state observations of 4891 genes, 4100 of which are protein coding \cite{kunst_complete_1997}, during the life cycle.
 Our \textit{\textit{B.
-anthracis}} dataset consists of 72 time-series and steady-state observations of 5536 genes comprising data from developmental and iron-starvation conditions.
-There were 247 known Transcription factors (TFs) in the \textit{\textit{B.
+anthracis}} dataset consists of 72 time-series and steady-state observations of 5536 genes comprising data from distinct points in the life cycle and iron-starvation conditions.
+There were 247 known transcription factors (TFs) in the \textit{\textit{B.
 subtilis}} dataset, and 248 TFs in the \textit{B.
@@
 
-We obtained 1870 one-to-one orthologs from Inparanoid \cite{ostlund_inparanoid_2010}, 95 of which are transcription factors, which produced 177,650 fusion-constraints between gene interactions within the two species.
+We obtained 1,870 one-to-one orthologs from Inparanoid \cite{ostlund_inparanoid_2010}, 95 of which are transcription factors, which produced 177,650 fusion-constraints between gene interactions within the two species.
 This number represents only $14.7\%$ of the regulatory interaction matrix in \textit{B.
@@
 
-To assess network inference performance, and for use as priors, we used a gold standard of 3040 known \textit{B.
+To assess network inference performance, and for use as priors, we used a gold standard of 3,040 known \textit{B.
 subtilis} interactions with corresponding activation and repression sign.
-Of these 3040 priors, 968 had corresponding interactions in \textit{B.
+Of these 3,040 priors, 968 had corresponding interactions in \textit{B.
 anthracis}.
@@
 anthracis} conditions.
-We can then compute performance metrics (ie AUPR) for each fold, and use their variability across folds to test for significance of any improvements relating to the use of fusion constraints.
+We can then compute performance metrics (ie AUPR) for each fold, and use their variability across folds to test for significance of any improvements relating to the use of fusion constraints (figure \ref{xspecies-real}a).
 
-In order to evaluate network recovery performance, we opted to evaluate AUPR on the sub-network of genes with fusion constraints.
-Although simulation results demonstrate that the non-constrained sub-network exhibit benefits from fusion, these benefits are expected to be very small given the small fraction of interactions with constraints in our cross-species application (Figure 1c).
-Moreover, the evolutionarily conserved sub-network is often of primary interest in cross-species studies.
-Application of fused L2 to our subsampled \textit{B.
-subtilis} and \textit{B.
-anthracis} datasets produced performance gains relative to the $\lambda_S=0$ case, but these performance gains, measured across the whole network, were very small (Figure 5b).
-However, inspection of precision recall curves showed that there is substantially increased precision at low values of recall.
-Essentially, precision increases for the high confidence part of the network but not for the low confidence part of the network.
+Though overall performance is hindered by our subsampling -- a necessary procedure to allow evaluation of networks, we demonstrate marked improvement in learning the \textit{B.
+subtilis} network when using fused regression (figure \ref{xspecies-real}a).
+Notably, these performance gains occur mostly at low values of recall.
+Essentially, precision increases for the high confidence part of the network but less so for the low confidence part of the network.
 As a result, the improvements in performance may be more useful than would be suggested by AUPR alone.
 Interactions suggested by genome wide analysis are useful for target prediction, and the ranking of interactions produced by a network inference algorithms is a useful guide to what order these experiments should be carried out in.
-The high confidence part of the network represents interactions that are likely candidates for experimental validation, so improvement in recovery of these interactions is particularly useful in guiding research.
+The high confidence part of the network represents interactions that are likely candidates for experimental validation, so improvement in recovery of these interactions is particularly useful in guiding future experiments.
+
+\subsection{Adaptive fusion using bacterial data}
+We anticipated that the fusion constraints we introduced based on orthology reflected a mixture of conserved and non-conserved interactions.
+The goal of adaptive fusion is to unfuse constraints between non-conserved interactions, while leaving intact all other constraints.
+However, because we lacked a gold standard of known non-conserved interactions between \textit{B.
+subtilis} and \textit{B.
+anthracis}, we were unable to directly evaluate how accurately adaptive fusion identified these interactions.
+We opted instead to introduce a large number of random fusion constraints between genes not known to be orthologous.
+These fake constraints, which are unlikely to reflect any conserved network structure, served as a proxy for the unknown fraction of non-conserved interactions between orthologous genes.
+We ran adaptive fusion to learn networks for \textit{B.
+subtilis} and \texti{B.
+anthracis}, using these constraints, along with those generated by known orthology.
+We confirmed that fake fusion constraints were unfused at a higher rate than those generated by known orthologs (figure \ref{xspecies-real}b).
+Although it may seem odd that a large fraction of fake constraints were left intact, we note that biological networks tend to be sparse, so that many of the random fusion constraints are between coefficients with near zero weight (and therefore near zero difference in weight) (figure \ref{xspecies-real}c).
+
+\subsection{Integrating datasets using fused regression}
+Although there are many large-scale collaborations which attempt to make protocols as uniform as possible for comparability between datasets generated by different labs \cite{paten_nih_2015,kundaje_integrative_2015}, there still exists technical and biologial variability between many experiments attempting to capture the same or similar experimental conditions.
+With the advent of new technology, such as RNAseq and single-cell RNAseq, microarray is no longer the dominant assay for genome-wide expression, but a large body of accumulated legacy data remains useful, if it can be integrated with more modern techniques.
+Currently, the most widely used approach to combining datasets for network inference is to learn networks from disparate datasets separately, then rank combine the networks as in Marbach et al \cite{marbach_revealing_2010}.
+We compared this approach to using fusion regression.
+We included, along with our \textit{B.
+subtilis} dataset, a previously published dataset containing 269 samples covering 104 conditions, obtained using a different tiling microarray (vs custom microarray) and different strain of \textit{B.
+subtilis} \cite{nicolas2012condition}.
+We compared performance when learning the networks separately, rank combining, and learning the networks simultaneously using fusion regression, and showed large improvement on network inference using our fused L2 approach (figure \ref{fusedl2-real}a).
 
@@
 We applied fusion regression by creating fusion constraints between a given transcription factor and genes within the same operon, and showed a boost in \textit{B.
-subtilis} network recovery using within-species fusion (Figure 6b).
-
-\subsection{Integrating datasets using fusion regression}
-Although there are many large-scale collaborations which attempt to make protocols as uniform as possible for comparability between datasets generated by different labs \cite{paten_nih_2015,kundaje_integrative_2015}, there still exists technical and biologial variability between many experiments attempting to capture the same or similar experimental condition.
-With the advent of new technology, such as RNAseq and single-cell RNAseq, microarray is no longer the dominant assay for genome-wide expression, but a large body of accumulated legacy data remains useful, if it can be integrated with more modern techniques.
-Currently, the most widely used approach to combining datasets for network inference is to learn networks from disparate datasets separately, then rank combine the networks as in Marbach et al \cite{marbach_revealing_2010}.
-We compared this approach to using fusion regression.
-We included, along with our \textit{B.
-subtilis} dataset, a previously published dataset containing 269 samples covering 104 conditions, obtained using a different tiling microarray (vs custom microarray) and different strain of \textit{B.
-subtilis} \cite{nicolas2012condition}.
-We compared performance when learning the networks separately, rank combining, and learning the networks simultaneously using fusion regression, and showed large improvement on network inference using our fused L2 approach (Figure 6a).
-Importantly, our approach simultaneously learns multiple networks, rather than learning networks separately than combining into one.
+subtilis} network recovery using within-species fusion (figure \ref{fusedl2-real}b).
 
@@
 subtilis} by Arrieta-Ortiz et al \cite{arrieta-ortiz_experimentally_2015}.
-TFA captures activation of transcription factor through mechanisms such as dimerization and interaction with required factors, over transcription factor expression level alone \cite{fu_reconstructing_2011}.
+TFA captures transcription factor activities through mechanisms such as dimerization and interaction with required factors, and is a better predictor of TF function than expression level alone \cite{fu_reconstructing_2011}.
 We estimated TFA based on known regulatory interactions using network component analysis \cite{liao2003network}.
-To test the integration of this approach with our fused regression, we asssesed the combination of \textit{B.
-subtilis} datasets, as in figure 6a, with the incorporation of TFA estimation.
+To test the integration of this approach with our fused regression, we assessed the combination of \textit{B.
+subtilis} datasets, as in figure \ref{fusedl2-real}a, with the incorporation of TFA estimation.
 We randomly divided the prior known interactions in half, and used half to learn TFA.
 The rest we reserved as a gold standard for validation.
-As in previous studies, we observed a marked improvement in network inference when using transcription factor activity.
+As in previous studies, we observed a marked improvement in network inference when using transcription factor activity (figure \ref{tfa}).
 We also obtained AUPR improvement when using fused regression on TFA, and showed that our gains from sharing information across datasets using fused regression were preserved and even enhanced by using TFA.
@@
 Meta-information about the structure of gene regulatory networks - specifically which pairs of interactions are \textit{a priori} likely to be similar to one another, can provide a powerful set of constraints to improve network inference performance.
+
 We present a general framework for gene regulatory network inference that incorporates this meta-information - termed fusion constraints - and apply the technique to the problem of simultaneous inference of regulatory networks in multiple species (\textit{B.
@@
 We further demonstrate the viability of fused L2 as a method for combining data from multiple experimental platforms, where fusion is between each identical regulator-gene pair.
+
 Because the algorithm we developed can accomodate constraints between arbitrary pairs of regulatory interactions, any biological prior representing information about expected regulatory similarity can be represented, even if the prior provides no information about the magnitude or direction of regulation.
@@
 A method for pooling data from multiple sources holds the promise of vastly expanding the quantity of data available for analysis, particularly in less commonly used model systems.
-At the same time these methods allow us to test our assumptions on how similar biological systems relate to one another, by allowing us to rule out conservation in a principled way, and at a genome-wide scale.
+At the same time these methods allow us to test our assumptions on how similar biological systems relate to one another, by allowing us to rule out conservation in a principled way, and at the genome-wide scale.
 
@@
 
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{fig1.pdf}
+  \caption{\label{fusion} a.
+schematic representation of gene relationships b.
+We generate two networks with 10 TFs by 200 genes, with 75\% sparsity, 0.1 measurement noise, and 0.3 fusion noise.
+The x-axis represents the number of conditions drawn from species 1, and the y-axis tracks the number of conditions from species 2 used.
+Performance is mean square error of network 1; as the number of conditions from the first species increases, our ability to recover network 1 also improves.
+But if we increase the number of conditions from the related species 2, we are better able to learn the network describing species 1.
+c.
+We generate a series of networks with 20 TFs by 200 genes, with 50\% sparsity, 0.1 measurement noise, and 0.1 fusion noise.
+We increase amount of genes fused from 0 to 0.75, with 0.25 increments and evaluate performance on the whole network, interactions which have fusion constraints (constrained part of the network), and interactions which do not have fusion constraints (non-constrained), and the effect of increasing the fusion weight.}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{fig2.pdf}
+  \caption{\label{syntheticgrid} a.
+We generate a series of networks with 10 TFs by 200 genes, with 75\% sparsity, 0.1 measurement noise, and 0.1 fusion noise.
+We hold lamR constant at 1.0 and 2.0 respectively.
+The x-axis represents lamS and the y-axis represents the proportion of nodes which are orthologous.
+As the networks increase in similarity, performance improves and benefits from a large fusion penalty weight.
+We measure performance by showing ratio of mean squared error relative to MSE without fusion.
+b.
+Here is another set of networks of 10 TFs by 200 genes, with 75\% sparsity 0.1 measurement noise, and 0.75 fusion noise.
+We hold and lamR constant at 1.0 and 2.0 respectively and measure performance in the same way as previous figure.}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{fig3.pdf}
+  \caption{\label{adapt-schematic} a.
+Adaptive fusion is a quadratic around the origin, and tapers off and plateaus, allowing differences in fused interaction weights to grow if the expression data supports this, without incurring further penalty from the fusion.
+b.
+The adaptive fusion penalty is modified from SCAD (smoothly clipped absolute deviation) and MCP (minimax concave penalty) functions.
+It shares with these penalties the zero derivative further from the origin.
+}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{fig4_2.png}
+  \caption{\label{adaptivefusion} a.
+Here we use two network of 35 TFs by 200 genes, with 0.4 of the genes and TFs fused, 0.1 measurement noise, and 0.1 fusion noise.
+We hold lamP and lamR constant at 1, and solve without fusion, using L2 fusion, and adaptive fusion, with lamS = 10 for both L2 and adaptive fusion.
+We plot the interaction weights; each point's coordinates represents the weight for network 1 and the weight for network 2.
+The solver was given a list of orthologs; the green points represent interactions between TFs and genes whose orthology information is correct, and the red points represent interactions between TFs and genes, where the orthology of one or both is incorrectly reported to the solver.
+We see that the L2 forces the points onto a diagonal, and that fused interactions adopt similar weights, regardless of the accuracy of the fusion.
+Our adaptive fusion solver is able to correctly identify and 'unfuse' incorrectly fused interactions.
+b.
+Here we plot the mean squared error of correctly fused and incorrectly fused (green and red dots, respectively) interactions for each solver.}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{fig5_2.pdf}
+  \caption{\label{xspecies-real} a.
+Using optimal lamR and lamP values, we use 10-fold cross validation to learn B.
+subtilis network.
+We compare performance when solving without fusion, and using L2 fusion with lamS = 1.0.
+Mean AUPR for lamS = 0 (unfused): 0.0619.
+Mean AUPR for lamS = 1: 0.0839.
+b.
+We test adaptive fusion using the same setup, with the addition of false orthology information.
+We set a 30\% false positive rate, including 561 additional orthologs.
+We run adaptive fusion, setting the $a$ term equal to the value above which 40\% of constraints would unfuse, reflecting our belief that in addition to some of the known orthologs result in fusion constraints which should be relaxed.
+Here we show the percentage of constraints relaxed, from constraints created from the false orthologs and the known orthologs.
+We also show the distributions of differences in weights corresponding to fused constraints created from the false orthologs and the known orthologs, when networks are solved separately.
+}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{fig6.pdf}
+  \caption{\label{fusedl2-real} a.
+We use optimal lamR and lamP values, and solve for B.
+subtilis-1 and B.
+subtilis-2 networks using 10-fold cross validation.
+We fuse our original B.
+subtilis-1 dataset to another B.
+subtilis dataset, which we call B.
+subtilis-2, using orthology information, and evaluate performance of B.
+subtilis-1 using AUPR on gold standard.
+We compare L2 fusion with solving the networks separately without fusion, then rank combining as in Marbach et al., as well as solving B.
+subtilis on its own.
+b.
+We again use optimal lamR and lamP values, and solve for the B.
+subtilis network using 10-fold cross validation, and evaluate using AUPR on B.
+subtilis with gold standard.
+Here, we fuse together TF, gene pairs, fixing the same TF with genes in the same operon group.
+The assumption is that genes in the same operon should be regulated similarly by the same TFs.
+We compare L2 fusion performance using operons with unfused.}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=0.45]{fig7.pdf}
+  \caption{\label{tfa} We integrate transcription factor activity (TFA) into our network inference, and show that fusion improves performance.Here we learn the B.
+subtilis-1 network with and without fusion, and with and without TFA}
+\end{center}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[scale=.5,trim=0mm 0mm 0mm 0mm,clip]{party3.eps}
+    \caption{\label{cartoon} Schematic representation of design matrix construction.
+Here, the circles and hexagons correspond to different species.
+Bidirectional arrows represent orthology information and dotted arrows represent putative interactions between TFs and genes.
+Rectangles represent matrices; because weights can be solved independently unless there exists fusion constraints between them, we identify related weights and construct matrix for solving.
+}
+\end{figure}
+
 \end{document}

### 2016-04-06 paper1.tex
hi

@@
 %\setcounter{secnumdepth}{1}
+\usepackage{amsmath}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+
+\makeatletter
+\def\BState{\State\hskip-\ALG@thistlm}
+\makeatother
+
 \begin{document}
@@
 For example, gene duplications may lead to neofunctionalization \cite{eisen_phylogenomics:_1998} of the duplicated genes.
-Even in the case of comparing networks from related cell lines from the same species, changes in chromatin configuration may affect our hypotheses about the similarity of interactions between pleiotropic TFs and target genes \cite{li_role_2007}.
+In the case of comparing networks from related cell lines from the same species, changes in chromatin configuration may affect our hypotheses about the similarity of interactions between pleiotropic TFs and target genes \cite{li_role_2007}.
 
@@
 \subsection{Statistical approach and background}
-We consider prediction and coefficient estimation problems with $N$ observations of $M$ dependent variables $y_{1,1}, y_{2,1}, ...y_{N,1}, y_{N,2},..., y_{N, M}$ and features $x_{i,j}$, $i=1,2,...,N, j=1,2,...,p$.
+We consider prediction and coefficient estimation problems with $N$ observations of $M$ dependent variables $y_{1,1}, y_{2,1}, ...y_{N,1}, y_{N,2},..., y_{N, M}$ and $p$ features $x_{i,j}$, $i=1,2,...,N, j=1,2,...,p$.
 We begin with a standard linear regression model:
@@
 \argmin_{\beta^S} \displaystyle\sum_{S \in \{1, 2\}} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \lambda_S \displaystyle \sum_{\substack{(g,h) \in orth,\\
- (h,l) \in orth}}(p(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l}))
+ (h,l) \in orth}}p(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
 \end{equation}
@@
 That is, $p(x)=x^2$.
-The overall objective function for network inference is then:
-
-\begin{equation}
-\argmin_{\beta^S} \displaystyle\sum_{S \in \text{species}} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \lambda_S \sum_{\substack{(g,h) \in orth,\\
- (h,l) \in orth}} \vert \vert (\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l}) \vert \vert ^2
-\end{equation}
-
 Because every component of the objective function is an L2 norm, the entire problem is convex, and can in fact be solved through linear regression with an augmented design matrix.
@@
 
+\subsection{Approach overview}
+\begin{algorithm}
+\caption{Network inference using fused regression}\label{euclid}
+\begin{algorithmic}[1]
+%\Procedure{MyProcedure}{}
+\State load expression data
+\State load orthology
+\State create priors and fusion constraints
+\State partition gold standard into training and leave-out
+\State generate TFA matrices using gold standard training set
+\For{$k$ in folds}
+	\State partition expression data into training and leave-out set
+	\State $\lambda_R$ parameter selection using training set
+	\State $\lambda_S$ parameter selection using training set
+	\State run fused regression
+	\State return PRC and ROC curves using leaveout gold standard
+	\EndFor
+\State Average PRC / ROC curves over folds
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
 \subsection{Solving fused L2 problems using augmented matrices}
@@
 In an ordinary regression problem each response variable can be solved independently, and vectorization is unnecessary.
-However, in fused regression, we append additional rows to the design matrix that link entries of the interaction weight matrix associated with different response variables.
+However, in fused regression, we append additional rows to the design matrix that link entries of the interaction weight matrix associated with different response variables (figure \ref{cartoon}).
 As a result, these linked response variables must be solved simultaneously through vectorization.
@@
 \begin{equation}
-\text{argmin}_{\beta^{S_i(t)}, \beta^{S_j(t)}} \displaystyle\sum_{S_i, S_j \in \text{species}} ||X^{S_i}\beta^{S_i}(t) - Y^{S_i}||^2 + \lambda_R||\beta^{S_i}(t)||^2 + \lambda_S\displaystyle \sum_{\beta_{g,k}^{S_i} \approx \beta_{h,l}^{S_j}} ||\beta^{S_i}_{g,k}(t) - \beta_{h,l}^{S_j}(t-1)||^2
+\text{argmin}_{\beta^{S}} \displaystyle\sum_{S \in \{1, 2\}} ||X^{S}\beta^{S}(t) - Y^{S}||^2 + \lambda_R||\beta^{S}(t)||^2 + \lambda_S\displaystyle \sum_{\substack{(g,h) \in orth,\\
+ (h,l) \in orth}} ||\beta^{S}_{g,k}(t) - \beta_{h,l}^{S}(t-1)||^2
 \end{equation}
@@
 This process is carried out separately for TFs and non-TF genes, so that TFs and non-TF genes are never assigned to be orthologous.
-We then produce a pair of random networks ($B^1$ and $B^2$) as follows.
-For each unfilled entry in $B^1$ or $B^2$, we enumerate the set $C$ consisting of the entry along with every entry in either matrix to which it is fused.
+We then produce a pair of random networks ($\beta^1$ and $\beta^2$) as follows.
+For each unfilled entry in $\beta^1$ or $\beta^2$, we enumerate the set $C$ consisting of the entry along with every entry in either matrix to which it is fused.
 With probability equal to the sparsity rate we assign every entry in $C$ to be 0, otherwise we sample a value $v \sim \mathcal{N}(0,1)$ and independently assign each entry in $C$ to $v + \mathcal{N}(0, \sigma_f^2)$.
-$\sigma_f$ is a parameter that controls the distribution of differences in the values of fused coefficients, so that the nonzero coefficients of $B^1, B^2$ are distributed as $\mathcal{N}(0, 1 + \sigma_f^2)$.
+$\sigma_f$ is a parameter that controls the distribution of differences in the values of fused coefficients, so that the nonzero coefficients of $\beta^1, \beta^2$ are distributed as $\mathcal{N}(0, 1 + \sigma_f^2)$.
 
-Given a network $B$, we generate $N$ samples of gene expressions at each of two timepoints.
+Given a network $\beta$, we generate $N$ samples of gene expressions at each of two timepoints.
 The condition by gene expression matrix for timepoint one, $Y_{T1}$, is sampled randomly from a multivariate Gaussian distribution with identity covariance matrix.
 $X_{T1}$ is the TF expression sub-matrix of $Y_{T1}$, and consists of columns of $Y_{T1}$ that correspond to TFs.
-Treating the decay rate as 0, the gene expression matrix at timepoint two, $Y_{T2}$ is sampled as $Y_{T2} = Y_{T1} + BX_{T1} + \epsilon$, where $\epsilon$ is a Gaussian noise term.
+Treating the decay rate as 0, the gene expression matrix at timepoint two, $Y_{T2}$ is sampled as $Y_{T2} = Y_{T1} + \betaX_{T1} + \epsilon$, where $\epsilon$ is a Gaussian noise term.
 This process is carried out separately for each network.

### 2016-04-07 paper1.tex
hi
figure legeeends

@@
 \begin{equation}
-\text{argmin}_{\beta^{S}} \displaystyle\sum_{S \in \{1, 2\}} ||X^{S}\beta^{S}(t) - Y^{S}||^2 + \lambda_R||\beta^{S}(t)||^2 + \lambda_S\displaystyle \sum_{\substack{(g,h) \in orth,\\
+\argmin_{\beta^S} \displaystyle\sum_{S \in \{1, 2\}} ||X^{S}\beta^{S}(t) - Y^{S}||^2 + \lambda_R||\beta^{S}(t)||^2 + \lambda_S\displaystyle \sum_{\substack{(g,h) \in orth,\\
  (h,l) \in orth}} ||\beta^{S}_{g,k}(t) - \beta_{h,l}^{S}(t-1)||^2
@@
   \includegraphics[scale=0.45]{fig1.pdf}
-  \caption{\label{fusion} a.
-schematic representation of gene relationships b.
-We generate two networks with 10 TFs by 200 genes, with 75\% sparsity, 0.1 measurement noise, and 0.3 fusion noise.
-The x-axis represents the number of conditions drawn from species 1, and the y-axis tracks the number of conditions from species 2 used.
-Performance is mean square error of network 1; as the number of conditions from the first species increases, our ability to recover network 1 also improves.
-But if we increase the number of conditions from the related species 2, we are better able to learn the network describing species 1.
-c.
-We generate a series of networks with 20 TFs by 200 genes, with 50\% sparsity, 0.1 measurement noise, and 0.1 fusion noise.
-We increase amount of genes fused from 0 to 0.75, with 0.25 increments and evaluate performance on the whole network, interactions which have fusion constraints (constrained part of the network), and interactions which do not have fusion constraints (non-constrained), and the effect of increasing the fusion weight.}
+  \caption{\label{fusion} \textbf{A.} Schematic representation of the the generation of fusion constraints from orthology mappings.
+Dashed arrows indicate potential regulatory interactions, while solid arrows denote orthology.
+We introduce fusion constraints for pairs of interactions for which both the regulator and regulated gene are orthologs of one another.
+In this example, we would introduce a constraint between the $(A, B)$ and $(A', B')$ interactions and the $(A, C)$ and $(A', B')$ interactions.
+\textbf{B.} In order to demonstrate the utility of fused network inference in combining data, we generate two networks with 10 TFs and 200 genes (75\% sparsity).
+Mean squared error of the inferred vs.
+true coefficient matrices for network 1 are plotted as a function of the number of conditions generated for species 1 (x-axis) and the number of conditions generated for species 2 (y-axis).
+As expected, increasing the number of samples available for the species of interest improves network inference performance.
+However, because we are fusing to data from a related species, similar gains are observed when increasing the amount of data available in this second species.
+\textbf{C-F} Show the varying effects of fusion on simulated networks with different levels of conservation.
+We generate a series of networks with 20 TFs by 200 genes in two species, (50\% sparsity) while varying the fraction of gene orthologies in the simulated networks.
+For each network, we evaluated AUPR on one of the species for: all interactions (blue line), interactions with fusion constraints (green line), and interactions without fusion constraints (orange line).
+At every level of conservation, constrained interactions show the largest benefits of fusion, with the magnitude of the benefit growing with fraction of orthologous genes.
+When the networks are highly conserved, however, even interactions that are not directly constrained through fusion are recovered more accurately as $\lambda_S$ increases.
 \end{center}
@@
   \includegraphics[scale=0.45]{fig2.pdf}
-  \caption{\label{syntheticgrid} a.
-We generate a series of networks with 10 TFs by 200 genes, with 75\% sparsity, 0.1 measurement noise, and 0.1 fusion noise.
-We hold lamR constant at 1.0 and 2.0 respectively.
-The x-axis represents lamS and the y-axis represents the proportion of nodes which are orthologous.
-As the networks increase in similarity, performance improves and benefits from a large fusion penalty weight.
-We measure performance by showing ratio of mean squared error relative to MSE without fusion.
-b.
-Here is another set of networks of 10 TFs by 200 genes, with 75\% sparsity 0.1 measurement noise, and 0.75 fusion noise.
-We hold and lamR constant at 1.0 and 2.0 respectively and measure performance in the same way as previous figure.}
+  \caption{\label{syntheticgrid} This figure demonstrates the interaction between fusion weight ($\lambda_S$), the degree of orthology coverage of networks, and the degree of similarity between fused interactions.
+\textbf{A.} We generated a series of pairs of 10 TF by 200 gene networks (75\% sparsity) in two species.
+These networks had minimal fusion noise, so that pairs of interactions linking orthologs had nearly identical weight.
+We varied the fraction of genes with orthologs (y-axis) and the weight of fusion in network inference (x-axis), and measured performance as the mean-squared error of the true vs inferred network weights.
+In order to more clearly visualize the varying effect of fusion, performance is plotted in units of the of the unfused MSE (left column) for each level of orthology.
+This was necessary because each row represents a different pair of networks, generated with a different level of orthology, for which baseline performance varies.
+Performance gains from fusion are, as expected, largest when the degree of orthology is large.
+However, even when the fraction of genes with orthologs is relatively small (top row), we observe gains from fusion.
+Importantly, when fused interactions are nearly identical, larger fusion weights always outperform smaller fusion weights.
+\textbf{B.} Simulates the case where genes which are orthologs may have different regulatory weights.
+We generated a series of networks as in \textbf{A.}, but with larger (7.5\times) gaussian noise added the weights of fused interactions.
+As in \textbf{A.}, benefits from fusion were observed at every level of gene orthology.
+However, unlike \textbf{A.}, there was an optimal intermediate value of $\lambda_S$ that traded off between the benefits of fusion and the cost of combining heterogenous data.}
 \end{center}
@@
   \includegraphics[scale=0.45]{fig3.pdf}
-  \caption{\label{adapt-schematic} a.
-Adaptive fusion is a quadratic around the origin, and tapers off and plateaus, allowing differences in fused interaction weights to grow if the expression data supports this, without incurring further penalty from the fusion.
-b.
+  \caption{\label{adapt-schematic} Adaptive fusion loss function (\textbf{A}) and derivative of loss function (\textbf{B}).
+\textbf{A.} Adaptive fusion is a quadratic around the origin, begins to taper at $a/2$, and plateus at $a$.
+After the plateu, increasing the difference in interaction weight of fused interactions does not further affect the penalty incurred through fusion.
+As a result, interaction weights in this zone are effectively unfused from one another (the fusion penalty behaves like a constant).
+\textbf{B.} Shows the derivative of the adaptive fusion penalty, which is used to implement adaptive fusion through local quadratic approximation.
 The adaptive fusion penalty is modified from SCAD (smoothly clipped absolute deviation) and MCP (minimax concave penalty) functions.
-It shares with these penalties the zero derivative further from the origin.
+It shares with these penalties the zero derivative far from the origin.
 }
@@
   \includegraphics[scale=0.45]{fig4_2.png}
-  \caption{\label{adaptivefusion} a.
-Here we use two network of 35 TFs by 200 genes, with 0.4 of the genes and TFs fused, 0.1 measurement noise, and 0.1 fusion noise.
-We hold lamP and lamR constant at 1, and solve without fusion, using L2 fusion, and adaptive fusion, with lamS = 10 for both L2 and adaptive fusion.
-We plot the interaction weights; each point's coordinates represents the weight for network 1 and the weight for network 2.
-The solver was given a list of orthologs; the green points represent interactions between TFs and genes whose orthology information is correct, and the red points represent interactions between TFs and genes, where the orthology of one or both is incorrectly reported to the solver.
-We see that the L2 forces the points onto a diagonal, and that fused interactions adopt similar weights, regardless of the accuracy of the fusion.
-Our adaptive fusion solver is able to correctly identify and 'unfuse' incorrectly fused interactions.
-b.
-Here we plot the mean squared error of correctly fused and incorrectly fused (green and red dots, respectively) interactions for each solver.}
+  \caption{\label{adaptivefusion} In order to evaluate the performance of adaptive fusion in unfusing interactions that have diverged through evolution, we performed a series of simulations inferring networks given a partially false list of orthology mappings.
+These false orthologs were orthologs that were generated from random pairs of non-orthologous genes, and therefore did not reflect true similarity of regulatory interactions.
+Networks were generated with 35 TFs by 200 genes, 60\% orthology coverage and 40\% false orthology coverage.
+\textbf{A.} We plot the interaction weights between pairs of fused interactions in network 1 (x-axis) and network 2 (y-axis) following network inference without fusion ($\lambda_S=0$).
+Interactions that are generated from false orthologs, which do not reflect true similarity of regulatory weight, are marked as red, while interactions generated from true orthologs are shown in green.
+Because fused interactions in the simulated networks were, in reality, almost identical in regulatory weight, interactions with true fusion constraints should fall on the diagonal.
+However, because a very small number of samples (20 for each species) was provided to the network inference procedure, the inferred weights are only weakly correlated (green dots).
+As expected, false fusion constraints give rise to uncorrelated weights (red dots).
+\textbf{B.} When fit with fused-L2, fusion constraints give rise to very similar weights in the two species.
+This is desireable for the true fusion constraints, but introduces error for the false constraints.
+\textbf{C.} Adaptive fusion run on the same network unfuses constraints for which the inferred weights are dissimilar beyond a certain point.
+In this example, unfused interactions are almost entirely interactions between false orthologs.
+\textbf{D-F} Show the distribution of the absolute value of the difference in inferred weight for interactions with true fusion constraints (green) and false fusion constraints (orange)}
 \end{center}

### 2016-04-07 paper1.tex
hi

--- 2016-04-06 paper1.tex
@@
 \usepackage{algorithm}
-\usepackage[noend]{algpseudocode}
+\usepackage{algpseudocode}
 
@@
 	\caption{Network inference using fused regression}\label{euclid}
-\begin{algorithmic}[1]
+	\begin{algorithmic}
 %\Procedure{MyProcedure}{}
@@
 Existing methods for cross-species network inference operate on the conserved subset of orthologous genes \cite{dillman_comparative_2015}.
-This may be appropriate with very closely related species, but could not be applied in this domain, where a large fraction (X\% and Y\%) of each species genomes do not have orthologs.
+This may be appropriate with very closely related species, but could not be applied in this domain, where a large fraction (62\% and 67\%) of the \textit{B.
+subtilis} and \textit{B.
+anthracis} genomes do not have orthologs.
 Our method, in contrast, can obtain improvements in network inference performance even when the conserved subset of genes is small.
@@
 Because our approach allows us to directly test -- and reject -- the hypothesis of conserved function, and because the enforcement of fusion constraints between non-conserved interactions may degrade network inference performance, we introduce an adaptive fusion method that decouples pairs of interactions for which expression data suggests a lack of conservation.
-Like the fused L2 approach, this method attempts to solve for networks which minimize a function that minimizes error in fitting the expression data plus error associated with the fusion penalty.
-In the case of cross-species network inference, orthology may provide a useful set of constraints on network structure.
-However, orthology does not always accurately predict functional similarity \cite{gabaldon_functional_2013}.When a fusion constraint encourages orthologous gene, TF pairs to share similar interactions across species, but the expression data suggests that these interaction weights should not be similar, the algorithm will arrive at some compromise between these two forces.
+Like the fused L2 approach, this method attempts to solve for networks which minimize error in fitting the expression data plus error associated with the fusion penalty.
+When a fusion constraint encourages orthologous gene, TF pairs to share similar interactions across species, but the expression data suggests that these interaction weights should not be similar, the algorithm will arrive at some compromise between these two forces.
 If the disagreement arose because the genes involved have taken on new function through evolutionary divergence \cite{kellis_proof_2004}, network recovery would be improved had the fusion constraint not been enforced.
@@
 Genes may be fused together on the basis of similar binding sites or chromatin features, and the relaxing of the fusion penalty indicates divergence of gene function.
-Adaptive fusion, therefore, is a tool for network inference as well as a method for testing network conservation.
+Adaptive fusion, therefore, is a tool for testing network conservation as well as a tool for network inference.
 
@@
 \begin{center}
-  \includegraphics[scale=0.45]{fig4_2.png}
+  \includegraphics[scale=0.45]{fig4_2.pdf}
   \caption{\label{adaptivefusion} a.

### 2016-04-07 paper1.tex

--- 2016-04-07 paper1.tex
@@
 \begin{center}
-  \includegraphics[scale=0.45]{fig4_2.png}
+
+  \includegraphics[scale=0.45]{fig4_2.pdf}
   \caption{\label{adaptivefusion} In order to evaluate the performance of adaptive fusion in unfusing interactions that have diverged through evolution, we performed a series of simulations inferring networks given a partially false list of orthology mappings.

### 2016-07-24 paper1.tex
hi

@@
 \title{Fused regression for multi-source gene regulatory network inference}
-\author{Kari Y. Lam \and Zachary M. Westrick \and Christian L. M\"{u}ller \and Lionel Christiaen \and Richard Bonneau}
+\author{Kari Y.
+Lam\textsuperscript{1} \and Zachary M.
+Westrick\textsuperscript{1} \and Christian L.
+M\"{u}ller\textsuperscript{2}\and Lionel Christiaen\textsuperscript{1} \and Richard Bonneau\textsuperscript{1,2}}
 \maketitle
 
+  1. New York University, New York, NY 10003
+  2. Simons Foundation, New York, NY 10010
+
 \begin{abstract}
-Understanding gene regulatory networks is critical to learning how cells regulate their behavior and respond to external stimuli.
-To this end, methods for global network inference have been developed and applied to a variety of species.
-However, most approaches consider the problem of network inference independently in each species, despite evidence that gene regulation can be conserved even in distantly related species.
-We introduce a method for multi-source network inference that allows simultaneous estimation of gene regulatory networks in multiple species or biological processes through the introduction of priors based on known gene relationships such as orthology.
-These priors are incorporated using fused regression.
+Understanding gene regulatory networks is critical to understanding cellular differentiation and response to external stimuli.
+Methods for global network inference have been developed and applied to a variety of species.
+Most approaches consider the problem of network inference independently in each species, despite evidence that gene regulation can be conserved even in distantly related species. 
+Further, network inference is often confined to single data-types (single platforms) and single cell types. 
+We introduce a method for multi-source network inference that allows simultaneous estimation of gene regulatory networks in multiple species or biological processes through the introduction of priors based on known gene relationships such as orthology incorporated using fused regression.
 This approach improves network inference performance even when orthology mapping and conservation are incomplete. 
-We further refine this method with an algorithm that attempts to learn the true conserved subnetwork from a larger set of potentially conserved interactions.
-We demonstrate the utility of our method in cross species network inference and in merging data collected on different experimental platform and in incorporating binding similarity predictions.
+We refine this method by presenting an algorithm that extracts the true conserved subnetwork from a larger set of potentially conserved interactions and demonstrate the utility of our method in cross species network inference.
+Last, we demonstrate our method's utility in learning from data collected on different experimental platforms.
 \end{abstract}
+
+\section{Author Summary}
+
+Gene regulatory networks describing related biological processes are thought to share conserved interaction structure.
+This assumption motivates a great deal of work in model systems – where discovery of gene regulation may be more experimentally tractable – but is difficult to directly evaluate using existing methods.
+The presence of shared structure in a well studied model system or process should make the problem of network inference in a related process easier, but this information is not often applied to the discovery of global gene regulatory networks.
+Further, to be able to successfully translate findings between different organisms, it is important to be able to identify what is systematically different.
+We provide a method for gene regulatory network inference using data from related sources.
+We introduce approaches for the simultaneous inference of related networks, and the identification of conserved subnetworks.
+The method is benchmarked on synthetic networks, and evaluated on the problem of inferring networks in related bacterial datasets.
 
@@
 
-We present two methods for network inference based on linear estimates of gene expression dynamics, extending existing dynamical-systems methods for network inference \cite{bonneau_predictive_2007, arrieta-ortiz_experimentally_2015}.
+We present two methods for network inference based on linear estimates of gene expression dynamics, extending existing dynamical-systems methods for network inference \cite{bonneau_predictive_2007, arrieta-ortiz_experimentally_2015, yeung_reverse_2002}.
 The core of both methods is the observation that biological information about the relatedness of genes can be used to select which network coefficients should be similar to one another in a multi-source network inference problem (ie orthologous TFs should regulate orthologous genes), and that these constraints can be efficiently represented as penalties in a least-squares regression problem.
@@
 
-In the case of multiple species, numerous studies showed that functional conservation exists in gene regulatory networks even across large evolutionary distance \cite{satou2006gene, hinman2009evolution,tanay2005conservation,erwin2009evolution}.
+In the case of multiple species, numerous studies have shown that functional conservation exists in gene regulatory networks even across large evolutionary distance \cite{satou2006gene, hinman2009evolution,tanay2005conservation,erwin2009evolution}.
 In our fused L2 approach, we assume that for closely related species, orthologous TFs could exert similar regulatory effects on orthologous target genes. 
@@
 
-This approach of introducing constraints on the similarity between specific regulatory interactions can be extended beyond the case of multi-species network-inference from orthology; any biological prior on similarity can be used in place of orthology.
-For example, we can use fused regression to combine datasets obtained using different platforms or experimental techniques.
-Moreover, we can introduce constraints that favor genes in the same operon or having similar binding sites, towards having similar regulators.
+This approach of introducing constraints on the similarity between specific regulatory interactions can be extended beyond the case of multi-species network-inference from orthology; any biological prior on similarity of regulation can be used in place of orthology.
+For example, we can use fused regression to combine datasets obtained using different platforms or experimental techniques and we can introduce constraints that favor genes in the same operon (or having similar promoters) towards having similar regulators.
 
-Existing multi species approaches often use orthology as a proxy for functional conservation \cite{penfold_inferring_2015, joshi_multi-species_2015, kashima_simultaneous_2009, zhang2010nearly}, or attempt to learn functional similarity via expression data \cite{gholami_cross-species_2010}.
-Orthology -- the measure of gene similarity we use -- can be approximated using readily identifiable sequence similarity, which is often a useful predictor of functional similarity \cite{wilson_assessing_2000}.
+Existing multi species approaches often use orthology as a proxy for functional conservation \cite{roy_arboretum:_2013, penfold_inferring_2015, joshi_multi-species_2015, kashima_simultaneous_2009, zhang2010nearly}, or attempt to learn functional similarity via expression data \cite{gholami_cross-species_2010}.
+Orthology can be approximated using readily identifiable sequence similarity, which is often a useful predictor of functional similarity \cite{wilson_assessing_2000, jensen_eggnog:_2008}. 
 In multi-species network inference, our fused L2 approach minimizes a cost function that strives to simultaneously fit expression data in each species and produce networks that are consistent with evolutionary constraints created using orthology.
-However, some genes may have taken on different functions and therefore may have new regulatory interactions.
+However, many genes will have evolved different functions and therefore may have new regulatory interactions. 
 For example, gene duplications may lead to neofunctionalization \cite{eisen_phylogenomics:_1998} of the duplicated genes. 
-In the case of comparing networks from related cell lines from the same species, changes in chromatin configuration may affect our hypotheses about the similarity of interactions between pleiotropic TFs and target genes \cite{li_role_2007}.
+In the case of comparing networks from related cell lines from the same species, changes in chromatin configuration may affect our hypotheses about the similarity of interactions between pleiotropic TFs and target genes across cell types (a within-species analog to neo-functionalization) \cite{li_role_2007}.
+Identifying interactions that are present in one species but not another is of direct biological interest, but existing approaches to network inference are unable to effectively test the hypothesis of conserved subnetworks.
+Observing a large difference in the weights of regulatory interactions obtained though independent inference of multiple networks is perhaps the best (least biased) evidence against conservation of orthologous regulatory interactions (cases where target and regulator have orthologs across species).
+However, this is sometimes weak evidence, as network inference is typically underconstrained \cite{marbach_revealing_2010-1}, meaning there could be a different set of networks for which conservation does hold, and which fit the data almost as well.
+We propose using our adaptive fusion approach to simultaneously perform network inference and evaluate edge-conservation (or lack thereof).
 
-Understanding which interactions are functionally dissimilar is of direct biological interest, but existing approaches to network inference are unable to effectively rule out the existence of a conserved subnetworks.
-Observing a large difference in the weights of regulatory interactions obtained though independent inference of multiple networks is the best evidence against conservation available with existing network inference techniques.
-This evidence is very weak, as network inference is typically underconstrained \cite{marbach_revealing_2010-1}, there could be a different set of networks for which conservation does hold, and which fit the data almost as well.
-We propose, with our adaptive fusion, an approach where network inference occurs simultaneously with the evaluation of conservation, to gather stronger evidence of divergence.
-
-We approach this problem by introducing a saturating penalty function based on statistical efforts to develop unbiased regularization penalties for regression \cite{zhang2010nearly, fan2001variable}.
-The main difference between the L2 fusion approach and our new adaptive fusion approach occurs when the difference between presumed analogous interactions is large despite the fusion penalty: we assume that large differences represent evidence that TF-to-target-gene interactions have diverged.
-We represent this, in adaptive fusion, with a relaxation of the fusion penalty.
+We approach this problem by introducing a saturating penalty function based on statistical efforts to develop unbiased regularization penalties for fused regression \cite{zhang2010nearly, fan2001variable}.
+The main difference between the L2 fusion approach and our new adaptive fusion approach occurs when the difference between presumed analogous interactions is large despite the fusion penalty. 
+We assume that cases where the method is unable to reconcile the likelihood and the fusion constraint derived from orthology (in the case of multi-species fusion) or identity (in the case of multi-platform fusion) correspond to cases with evidence of divergent TF-to-target-gene interactions.
+Practically, this inability to reconcile likelihood and conservation hypothesis manifests as large differences in model weights across species or platforms. 
+We account for this possibility in our adaptive fusion model with a relaxation of the fusion penalty in cases with extreme model weight divergence. 
 The resulting cost function is non-convex and difficult to optimize \cite{fan2001variable}; however, we can approximate its solution and obtain deeper insight into functional similarity than is available through strict orthology enforcement or the comparison of separately learned networks.
@@
 This shows the applicability of our method in combining different datasets and leveraging similarity across organisms as well as within a network in order to improve network inference.
-We explore the circumstances under which each approach is optimal, and evaluate the robustness of adaptive fusion to incorrect orthology, simulating the biologically relevant cases of neo- and subfunctionalizations.
+We explore the circumstances under which each approach is optimal, and evaluate the robustness of adaptive fusion to incorrect orthology, simulating the biologically relevant cases of neo- and sub-functionalizations.
 
@@
 For example, ridge regression penalizes the L2 norm of the coefficients $\beta_{i,j}$ in order to avoid overfitting \cite{hoerl_ridge_1970}, and can be thought of as a mean-zero Gaussian prior on the coefficients.
-
-More complicated penalties have been developed to represent certain expected or desireable structure in a regression model's coefficients.
-For example, Land and Friedman \cite{citeulike3780356} proposed a fusion penalty:
-
-\begin{equation}
-\text{argmin}_{\beta} \displaystyle\sum_{ij} || y_i - x_{ij}\beta{j} ||^2 + \sum\limits_{j=2}^p ||\beta_j - \beta_{j-1}]|^\alpha
-\end{equation}
-
-Which encourages smoothness of the estimated parameter vector.
-
-%This penalty utilizes correlation information from data by penalizing pairwise differences of coefficients \cite{Daye2009,Price2014}. 
-Existing approaches have used similar fusion penalties to combine statistical strength across multiple regression tasks by introducing regularization functions reflecting prior knowledge on output structure representing how outputs are related \cite{kim_tree-guided_2012,Land1997,Chen2010,Petry2011,Hebiri2011}.
-Price et al.
-and Bilgrau et al.
-use a fused ridge estimator for jointly estimating multiple inverse covariance matrices \cite{Price2014a,Bilgrau2015}; we take a similar approach, adding an L2 penalty on the differences between coefficients to the existing ridge penalty in order to incorporate prior knowledge about relationships between input-output pairs:
+More complicated penalties have been developed to represent specific expected or desireable structure in a regression model's coefficients. 
+For example, Land and Friedman \cite{citeulike3780356} proposed a fusion penalty which encourages smoothness of the estimated parameter vector.
+Previous approaches have used  fusion penalties to draw statistical strength across multiple regression tasks \cite{kim_tree-guided_2012,Land1997,Chen2010,Petry2011,Hebiri2011}.
+Price et al. and Bilgrau et al. use a fused ridge estimator for jointly estimating multiple inverse covariance matrices \cite{Price2014a,Bilgrau2015}.
+We take a related approach to these prior works, adding an L2 penalty on the differences between coefficients to the existing ridge penalty in order to incorporate prior knowledge about relationships between input-output pairs:
 
@@
 
-where $X$, $Y$, and $\beta$ are matrices, and $\beta_{g,k} \approx \beta_{h,l}$ denotes fusion between arbitrary entries of $\beta$.
+where $X$, $Y$, and $\beta$ are matrices, and $\beta_{g,k} \approx \beta_{h,l}$ denotes fusion between entries of $\beta$ (enforcing similarity between model weights across separate data-sets). 
 Note that, like ridge regression, this penalty can be thought of as representing a Gaussian prior on the coefficients $\beta$. 
@@
 In many cases, however, there is some uncertainty about the relationships that should be enforced. 
-Sohn et al.
-attempt to simultaneously learn the regression coefficients and the output structure \cite{sohn_joint_2012}.
-We take a simultaneous approach, by applying a penalty function bounded by a constant to produce unbiased estimators for large coefficients, combined with an L2 penalty, similar to SCAD-L2 \cite{Zeng2012}.
+Sohn et al. attempt to simultaneously learn the regression coefficients and the output structure \cite{sohn_joint_2012}. 
+We develop a similar approach, by applying a penalty function bounded by a constant to produce unbiased estimators for large coefficients, combined with an L2 penalty, similar to SCAD-L2 \cite{Zeng2012}.
 
@@
     \begin{array}{lr}
-    2\lambda\theta & \text{if } \theta \leq \lambda\\
-    \text{max}(\lambda_S(a-\theta),0) & \text{if } \theta > \frac{a}{2}
+    \lambda\theta & \text{if } \theta \leq \lambda\\
+    \text{max}(\lambda(2a-\theta),0) & \text{if } \theta > a}
     \end{array}
@@
 
- This approach allows us to simultaneously learn the regression coefficients and evaluate the validity of our prior information.
+ This approach allows us to simultaneously learn the regression coefficients and evaluate the validity of our prior information (this model relaxes the fusion penalty when model components are irreconcilably different).
 
@@
 These constraints favor the generation of similar networks for related species, and in the generally under-constrained regime of network inference can improve the accuracy of network recovery. 
-We then go on to introduce a method to test the assumption of conserved network structure, and to relax the associated constraints on pairs of interactions for which data does not support conservation.
+We then go on to introduce a method to test the assumption of conserved network structure, and to relax the associated constraints on pairs of interactions for which the data does not support conservation (where conservation of a regulatory edge is implied by like model weights across data-sets).
+
+\subsection{Approach overview}
+\begin{algorithm}
+	\caption{Network inference using fused regression}\label{euclid}
+	\begin{algorithmic}
+%\Procedure{MyProcedure}{}
+\State load expression data
+\State load orthology
+\State create priors and fusion constraints
+\State partition gold standard into training and leave-out
+\State generate TFA matrices using gold standard training set
+\State set $a$ if using adaptive fusion
+\For{$k$ in folds}
+	\State partition expression data into training and leave-out set
+	\State $\lambda_R$ parameter selection using training set
+	\State $\lambda_S$ parameter selection using training set
+	\State run fused regression
+	\State return PRC and ROC curves using leaveout gold standard
+	\EndFor
+\State average PRC / ROC curves over folds
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
 
@@
 
-We are interested in identifying genes which affect other genes' expression, and understanding how this gives rise to a given biological process.
-We know that transcription factors (TFs) can bind to DNA and activate or repress expression of other genes.
-By restricting ourselves only to TFs and gene interactions, we vastly reduce the space of possible models while considering biologically relevant cases.
-
 We model the transcription rate of each gene as a weighted sum of transcription factor expression, and seek to identify the identities and regulatory weights of these TFs. 
-This formulation matches that of the existing Inferelator algorithm, which models gene expression with linear differential equations \cite{bonneau_inferelator:_2006-1}.
+This formulation matches that of the existing \textit{Inferelator} algorithm, which models gene expression with linear differential equations \cite{bonneau_inferelator:_2006-1}.
 Our primary data for learning gene regulatory networks is expression data, consisting of time-series and steady state experiments.
-We use time-series data to approximate the rate of change of expression, and treat steady-state experiments as having reached equilibrium (constant expression).
 The rate at which $x_{i}$, the observed mRNA expression of gene $i$, changes, is governed by degradation of existing transcripts with rate $\alpha$ plus a linear combination of transcription factor (TF) expressions.
@@
 We are interested in learning $\beta$, the matrix representation of the gene regulatory network, where the weight in a given position represents the regulatory weight of a TF on a gene. 
-Positive weights represent activation, negative weights represent repression, and 0 weights represent the absence of an interaction.
-The matrix $\beta$ can be solved using linear regression.
+Positive weights represent activation, negative weights represent repression, and 0 weights represent the absence of an interaction. The matrix $\beta$ can be solved using linear regression. 
 Because there are typically far fewer conditions than possible regressors (TFs), we introduce a ridge regularization constraint with weight $\lambda_R$ and solve
@@
 \end{equation}
-This is similar to the formulation used in the Inferelator algorithm, which we extend to the case of simultaneously inferring multiple networks.
+This is similar to the formulation used in the \textit{Inferelator} algorithm, which we extend to the case of simultaneously inferring multiple networks.
 
-Transcription factor expression is not always the best predictor of its gene targets' expression.
+Transcription factor expression is not always the best predictor of its gene targets' expression, so  previous network inference methods attempt to estimate transcription factor activities prior to network inference.
 When there exists a set of prior known interactions, we are able to estimate transcription factor activity (TFA) using network component analysis \cite{liao2003network}, as in \cite{arrieta-ortiz_experimentally_2015, fu_reconstructing_2011}, and use TFA as explanatory variables instead of transcription factor expression.
@@
 
-Information about the conserved structure of gene regulation is introduced through the incorporation of constraints into the above regression formulation.
+Information about the partially conserved structure of gene regulation is introduced through the incorporation of constraints into the above regression formulation.
 These constraints penalize differences between interaction weights in the networks of multiple species that are expected to be similar based on prior biological knowledge.
@@
 Consider the case of organisms $A$ and $B$, governed by GRNs $\beta^A$ and $\beta^B$ (the following approach applies equally well to more than two species but for simplicity we continue with the case of two species).
-If TF $g^A$ in organism $A$ and TF $h^B$ in organism $B$ are orthologs, and gene $k^A$ and $l^B$ are orthologs, then we expect that the $g^A \rightarrow k^A$ interaction weight should be similar to the $h^B \rightarrow l^B$ interaction weight, and we introduce a fusion constraint between these analogous interactions.
+When TF $g^A$ in organism $A$ and TF $h^B$ in organism $B$ are orthologs, and gene $k^A$ and $l^B$ are orthologs, then we expect that the $g^A \rightarrow k^A$ interaction weight should be similar to the $h^B \rightarrow l^B$ interaction weight, and we introduce a fusion constraint between these analogous interactions.
 In terms of the above regression formulation, we expect that $\beta^A_{g,k} \approx \beta^B_{h,l}$, and include a penalty term $\lambda_Sp(\beta^A_{g,k} - \beta^B_{h,l})$ in the quantity being minimized in order to encourage similarity.
 The function $p(x)$ controls the shape of the relationship between weight dissimilarity and penalty, while scalar $\lambda_S$ controls the overall scaling of the penalization of differences between fused coefficients.
-$\lambda_S$ controls the tradeoff between fitting the expression data alone and producing a set of networks that conform to evolutionary prior knowledge.
-This gives us the final equation to be minimized
+$\lambda_S$ controls the tradeoff between fitting the expression data-sets individually and producing a set of networks that conform to evolutionary prior knowledge.
+This gives us the final equation to be minimized:
 
 \begin{equation}
-%\argmin_{(\beta^A, \beta^B)} \displaystyle\sum_{S \in (A, B)} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \displaystyle \sum_{\beta^{S_1}_{g,k} \approx \beta^{S_2}_{h,l}} p_{\lambda_S}(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
 \argmin_{\beta^S} \displaystyle\sum_{S \in \{1, 2\}} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \lambda_S \displaystyle \sum_{\substack{(g,h) \in orth,\\
@@
 where the second sum is over pairs of interactions with fusion constraints. 
-In the fused L2 algorithm we develop, the penalty function is equal to the L2 norm of the difference in regulatory weight of fused coefficients.
-That is, $p(x)=x^2$.
-Because every component of the objective function is an L2 norm, the entire problem is convex, and can in fact be solved through linear regression with an augmented design matrix.
-%The regulators of a pair of genes can be solved independently as long as the pair is not linked by fusion constraints, or a chain of fusion constraints; thus the problem of cross-species GRN inference decomposes into a large number of manageable regression problems as long as the orthology mappings are sparse.
+In the fused L2 algorithm presented here, the penalty function is equal to the L2 norm of the difference in regulatory weight of fused coefficients, $p(x)=x^2$. 
+Every component of the objective function is an L2 norm and thus the problem is convex and can in fact be solved through linear regression with an augmented design matrix.
 
-In order to better understand the behavior of this objective function, consider the case where there is a one-to-one orthology between the species being considered (ie different cell-lines of the same organism).
+As an example, consider the case where there is a one-to-one orthology between the species being considered (ie different cell-lines of the same organism). 
 The choice of $\lambda_S$ allows one to interpolate between fitting each network independently ($\lambda_S=0$) and pooling data together as if it came from one source ($\lambda_S=\inf$). 
-In addition to allowing interpolation between these extremes, our method allows pooling of data even when there is incomplete orthology.
+In addition to performing well between these extremes, our method allows pooling of data even when there is incomplete orthology. 
 By introducing constraints on the similarity of individual interactions, rather than on the networks as a whole \cite{liu2011temporal}, we can pool some information across species even when a small fraction of genes have orthologs.
-This is particularly useful when dealing with a large number of species; pairwise orthologies may be nearly complete even when the number of genes present in every organism is small.
 
 \subsection{Adaptive fusion}
-Fusion constraints are intended to penalize dissimilarity between interactions thought to be analogous based on \textit{a priori} knowledge.
+Fusion constraints penalize dissimilarity between interactions thought to be analogous based on \textit{a priori} knowledge. 
 For example, orthology can be used to predict which interactions will be similar across species. 
-When fusion constraints are L2, interaction weights which differ from each other by a large amount are excessively penalized, which effectively ensures that fused interactions are assigned similar weights.
+With an L2 fusion penalty, interaction weights which differ from each other by a large amount are excessively penalized, which effectively ensures that fused interactions are assigned similar weights.
 This will be inappropriate for interactions which are identified based on orthology as being analogous, but which are no longer similar due to evolutionary changes. 
-We propose that a saturating penalty -- a penalty that is relaxed once differences in weights grow beyond a certain point -- may be useful for dealing with uncertainty about which interactions are conserved.
-
-With a saturating fusion penalty, fused interactions which appear to be very different based on expression data are allowed to unfuse from one another.
-A closely related problem has been studied in the context of LASSO regularization, where it was shown by Fan and Li that using a saturating penalty retains many of LASSO's desireable properties while removing its bias towards 0 \cite{fan2001variable}.
+We propose that a saturating penalty  that is relaxed once differences in weights grow beyond a certain point (interactions which appear to be very different based on the data are effectively unfused).
+A related problem has been studied in the context of LASSO regularization, where it was shown by Fan and Li that using a saturating penalty retains many of LASSO's desireable properties while removing its bias towards model weights of 0 \cite{fan2001variable}.
 They further showed that, although the resulting loss-function is nonconvex, good results can be obtained with a local quadratic approximation of gradient descent.
@@
 We introduce a modified form of MCP to the problem of penalizing differences between fused coefficients. 
-The principle difference between the penalty we adopt and SCAD/MCP is that both of these penalties are L1 like at the origin, producing sparse solutions.
+The principal difference between the penalty we adopt and SCAD/MCP is that both of these penalties are L1 like at the origin, producing sparse solutions. 
 Some network inference approaches use L1 penalties to produce sparse networks, on the basis that biological networks are thought to be sparse. 
-However, since we are penalizing differences in interaction weights, rather than the weights themselves, there's no reason to assume that most differences will be exactly zero, and an L2 penalty - equivalent to an assumption that the differences between fused coefficients are Gaussian distributed - may be more appropriate.
+However, as we are penalizing differences in interaction weights, rather than the weights themselves, there's no reason to assume that most differences will be exactly zero, and an L2 penalty - equivalent to an assumption that the differences between fused coefficients are Gaussian distributed - may be more appropriate.
 
-We use a penalty on the difference between fused coefficients $\theta$ which is L2 like at the origin, begins to level out at $\theta = \frac{a}{2}$, and saturates at $\theta = a$.
-Written in terms of its derivative, the penalty $p'_{\lambda, a}$
+We use a penalty on the difference between fused coefficients $\theta$ which is L2 like at the origin and saturates at $\theta = a$. Written in terms of its derivative, the penalty $p'_{\lambda, a}$
 
@@
 \end{equation}
-
-%We also implement MCP:
-
-%\begin{equation}
-%p_{\lambda,\gamma}(\theta) = \left\{
-%    \begin{array}{lr}
-%    \lambda\theta^2-{ \theta^2 \over 2\gamma} & \text{if } \theta \leq \gamma\lambda\\
-%    {1 \over 2}\gamma\lambda^2 & \text{if } \theta > \gamma\lambda
-%    \end{array}
-%    \right. 
-%\end{equation}
-
-%With derivative
-
-%\begin{equation}
-%p'_{\lambda,\gamma}(\theta) = \left\{
-%    \begin{array}{lr}
-%    2\lambda\theta - {\theta \over \gamma} & \text{if } \theta \leq \gamma\lambda\\
-%    0 & \text{if } \theta > \gamma\lambda
-%    \end{array}
-%    \right.
-%\end{equation}
 
@@
 Moreover, we are primarily interested in using this saturating penalty as a way of testing the hypothesis that conservation in GRNs can be predicted based off of known similarities between genes. 
-Therefore, we propose a method for selecting $a$ to be equal to a certain percentile of the distribution of differences in fused interaction weights from the networks fit without fusion.
-This choice represents the user's hypothesis for the fraction of both networks that have unrelated function, and should therefore be unfused.
-
-\subsection{Approach overview}
-\begin{algorithm}
-	\caption{Network inference using fused regression}\label{euclid}
-	\begin{algorithmic}
-%\Procedure{MyProcedure}{}
-\State load expression data
-\State load orthology
-\State create priors and fusion constraints
-\State partition gold standard into training and leave-out
-\State generate TFA matrices using gold standard training set
-\For{$k$ in folds}
-	\State partition expression data into training and leave-out set
-	\State $\lambda_R$ parameter selection using training set
-	\State $\lambda_S$ parameter selection using training set
-	\State run fused regression
-	\State return PRC and ROC curves using leaveout gold standard
-	\EndFor
-\State Average PRC / ROC curves over folds
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
+Therefore, we propose a user-defined $a$, where this parameter is set using the distribution of differences between fused weights from independently fit networks.
+The choice of which value in this distribution to use for $a$ represents the working hypothesis for the fraction of fused interactions which should be unfused.
 
 \subsection{Solving fused L2 problems using augmented matrices}
-We begin with the problem of constructing a design matrix to solve a fused L2 regression problem with a single response variable, because problems with multiple response variables and with multiple data sources can be converted into this form through vectorization.
+We begin with the problem of constructing a design matrix to map our problem to that of solving a fused L2 regression problem with a single response variable. 
 We then go on to show that, although the vectorized solution involves solving an impractically large system of equations, under typical biological conditions the structure of constraints allow the problem to be broken up into many smaller subproblems.
+Key to this approach is the observation that ridge constraints can be incorporated into a least-squares regression problem by appending a scaled identity matrix to the design matrix, and a corresponding number of zeros to the response vector.
+Similarly, a fusion constraint $\lambda_S (\beta_{i} - \beta{j})^2$ can be incorporated into a least-squares regression problem by appending a row containing $\sqrt{\lambda_S}$ in the $i$th position, $-\sqrt{\lambda_S}$ in the $j$th position, and $0$s elsewhere to the design matrix, and zero to the response vector.
+In order to convert an optimization over multiple response variables and multiple sources into an optimization with a single source and response variable, we vectorize as follows: we construct a new design matrix by diagonally concatenating design matrices from relevant regression problems, and create a new response vector by concatenation of corresponding response vectors.
 
-We begin with the observation that ridge constraints can be incorporated into a least-squares regression problem by appending a scaled identity matrix to the design matrix, and a corresponding number of zeros to the response vector.
-Similarly, a fusion constraint $\lambda_S (\beta_{i} - \beta{j})^2$ can be incorporated into a least-squares regression problem by appending a row containing $\sqrt{\lambda_S}$ in the $i$th position, $-\sqrt{\lambda_S}$ in the $j$th position, and $0$s elsewhere to the design matrix, and zero to the response vector.
-In order to convert an optimization over multiple response variables and multiple sources into an optimization with a single source and response variable, we vectorize as follows: we construct a new design matrix by diagonally concatenating design matrices from relevant regression problems, and create a new response vector by vertically concatenating corresponding response vectors.
-
-%$$\underset{W}{min} ||AW - y||_2^{~2}$$
-
-%We can then define $W_{i,j} = \beta_{(j-1) \times N + i, (j-1) \times M + j}$ which is the desired form.
 This is equivalent to the original problem due to the block structure of matrix multiplication. 
@@
 In biological terms, the regulators of two genes (whether in the same species, or different species) must be solved together if there is a fusion constraint linking those genes' regulators, or if there is a chain of such constraints.
-This is important because vectorization produces a design matrix with a number of rows and columns equal to the sum of the number of rows and columns of the design matrices associated with each of the response variables (genes) being solved for.
-If the networks for a large number of genes are solved simultaneously, the system of equations can quickly become intractible to solve.
+If the networks for a large number of genes are solved simultaneously, the system of equations can quickly become intractable.
 
@@
 We can then incorporate fusion constraints as in the case of single-source single response-variable fused regression. 
-In most cases, we have found the direct solution using augmented matrices to be adequate.
-This is because the structure of orthology tends to be sparse, so only a small number of genes must be solved at once.
+In most cases, we have found the direct solution using augmented matrices to be adequate (possible due to the sparse structure of orthology links; only a small number of genes must be solved at once).
 In the general case, the size of the design matrix is proportional to the number of response variables that must be solved simultaneously. 
@@
 However, in the case of multi-species network inference with one-to-one orthology, the network associated with each pair of orthologous genes requires solving a linear system with approximately twice as many observations and unknowns as the single species case.
-Linear systems of this size can be solved quickly using standard techniques.
-When the size of the groups of genes linked by fusion constraints becomes large, however, the augmented design matrix approach becomes impractical.
-
-%where the $j$s refer to transcription factors.
-In subproblem 1, $j_1$ regulates gene $g$; $j_1$ is orthologous to $j_2$, a transcription factor of subproblem 2, which regulates $h$, a gene which is orthologous to gene $g$.
-There exists a chain because ($j_1$, $g$) is a TF, gene pair which is fused with ($j_2$, $h$), which is fused to ($j_3$, $k$).
-%We use depth-first search to identify linked columns of $Y$, and then solve the fused regression problem directly by vectorizing those columns.
-%In order to incorporate the fusion penalty, we append to the resulting design matrix one row \emph{per connected constraint}, which contains $\lambda_S$ in the column associated with $Y_g,i,j,$ and $-\lambda_S$ to the column associated with $Y_h, i',j'$ for constraint $(Y_g,i,j,Y_h,i',j')$.
-Call this resulting matrix $X_P$.
-%The vector $Y$ is augmented by appending an appropriate number of zeros to produce $Y_P$.
-As a result, each entry in the output of $X_P b$ in the columns associated with each constraint contain $\lambda_S(b_{i,j} - b_{i',j'})$.
-%These entries each contribute  $\lambda_S||b_{i,j} - b_{i',j'}||^{2}$ to the squared error $||X_Pb - Y_P||_2^{~2}$
-
-%In the context of regression for GRN inference with multiple cell sources, each gene $j$ from source $1$ has constraints $\forall_{i \in \{TFs\}} (1, i, j, 2, i’, j’)$ where $i',j'$ are the same gene in a different cell line.
-%As a result, each gene has only one constraint and the chain has length $L=2$.
-%In general, however, solving the fused regression problem directly, even after factoring constraints, involves at least one inversion of an $L\sum_{i=0}^S K_i \times L \sum_{i=0}^S K_i$ matrix, taking $O((L\sum_{i=0}^S K_i)^3)$.
-%In general, however, solving the fused regression problem directly, even after factoring constraints, involves at least one inversion of an $p\sum_{i=0}^S K_i \times p \sum_{i=0}^S K_i$ matrix where $p$ is the size of a column subset such that there is a constraint chain $L$ linking an element of each column to an element in every other column in that subset, taking $O((p\sum_{i=0}^S K_i)^3)$.
-%Although the sparse structure of $X_P$ makes this somewhat more tractable, it is still impractical for large $L$.
+Linear systems of this size can be solved quickly using standard techniques, and runtime using our bacterial datasets clocks in around thirty minutes. 
+When the size of the groups of genes linked by fusion constraints becomes large (when organisms have a number of many-to-many orthologous blocks), however, the augmented design matrix approach becomes slower and we discuss further optimisations to this scheme below to enable scaling to these regimes.
 
@@
 
-When the number of genes in each fusion constraints is relatively small, the augmented design matrix approach we have described has the advantage of being efficiently solvable with a closed form solution.
-However, in some cases the augmented system of equations may be too large to solve efficiently with standard methods.
-To address this limitation, we developed an iterative solver that uses coordinate-wise descent to solve for solutions corresponding to a sequence of values of fusion penalty weights; since our fused L2 method uses a convex and differentiable penalty function, this approach converges to a global minimizer.
-Although less efficient than the augmented design matrix approach we developed, the iterative solver has the advantage of computing a solution path for $\lambda_S$.
+To address scaling limitations when many-to-many fusion constraint blocks occur, we developed an iterative solver that uses coordinate-wise descent to solve for solutions corresponding to a sequence of values of fusion penalty weights.
+As our fused L2 method uses a convex and differentiable penalty function, this approach converges to a global minimizer. 
+Although less efficient than the augmented design matrix approach we developed for cases where fusion constraints are primarily one-to-one or few-to-few, the iterative solver has the advantage of computing a solution path for $\lambda_S$ and scaling well across a wider range of biological applications.
 
@@
 Because each iteration reduces the error between $\beta(t)$ and $\beta(t-1)$, and because $\beta(t) = \beta(t-1)$ is the globally optimal solution, this process must eventually converge to the same network as equation \ref{eqn:main}.
-Although we have not produced bounds on the convergence rate, which also depends on the structure of constraints, in practice a small number of iterations are necessary.
-
-The ierative solver is used for multi-source regression problems with complex orthology/similarity mappings, and also for solving the regularization path in order to pick regularization and fusion penalty weights.
+Although we have not produced bounds on the convergence rate, which also depends on the structure of constraints, in practice a small number of iterations ($\sim$10) are necessary.
 
 \subsection{Fusion and regularization path}
-Because optimizing over both parameters, $\lambda_R$ and $\lambda_S$, is computationally prohibitive, we opted to optimize the two parameters separately.
+Optimizing over both parameters, $\lambda_R$ and $\lambda_S$, is computationally prohibitive and we opted to test a heuristic where we optimize the two parameters separately. 
 Our procedure first optimized $\lambda_R$ with $\lambda_S=0$, then optimized $\lambda_S$ using this value of $\lambda_R$. 
-Although this procedure is sub-optimal, it is guaranteed to achieve the best unfused solution in the case when $\lambda_S$ is constrained at 0.
+This procedure is guaranteed to achieve the best unfused solution in the case when $\lambda_S$ is constrained at 0. 
 As a result, any performance gains of fused regression are a lower bound on the highest achievable performance gains. 
-
 To optimize $\lambda_R$ we use cyclical coordinate descent algorithms from the 'glmnet' package \cite{friedman_regularization_2010} to compute a ridge regularization path. 
@@
 
-%\subsection{Equivalent prior}
-%Our L2 fusion penalty is equivalent to assuming a Gaussian prior with variance proportional to $\frac{1}{\lambda_S}$ on differences in parameters with fusion constraints.
-Combined with the regularization constraints, this forms a multivariate Gaussian prior on weights in both networks.
-When using our method on systems with incomplete orthology mapping, this results in different prior probability distributions for those parameters with fusion constraints and those without.
-We adjust for this by solving for a constant to multiply fused interactions by to equalize the volumes of the prior distributions.
-The intent of this adjustment is to ensure that constrained interactions are not on average more highly penalized, which may tend to drive their weight towards zero, causing them to be excluded from the network.
-
-%\subsection{Model selection}
-%We can define a leave-out set of priors and choose parameters based on cross-validated AUPR on the leave-out set.
-
 \subsection{Simulated data}
 We generate simulated data to evaluate the ability of our fused L2 approach to learn the true network and to show that sharing information between similar but not identical data sources results in more accurate network recovery.
-Because the purpose of adaptive fusion is to learn the structure of conservation while also learning networks, using simulated data allows us to test the method relative to a known gold standard for network conservation.
-
-Generation of simulated data begins with the production of random orthology mappings.
+Generation of simulated data begins with the production of random orthology mappings with sparcity simlar to that found in real data-sets. 
 We produce a one-to-one orthology by pairing random genes until a specified fraction have been assigned orthologs. 
@@
 
-Given a network $\beta$, we generate $N$ samples of gene expressions at each of two timepoints.
+Given a network $\beta$, we generate $N$ samples of gene expressions at two timepoints. 
 The condition by gene expression matrix for timepoint one, $Y_{T1}$, is sampled randomly from a multivariate Gaussian distribution with identity covariance matrix.
@@
 This process is carried out separately for each network. 
-
 Following generation of simulated data, we may introduce error into the orthology mapping. 
@@
 For convenience, the false-positive rate is specified in units of the number of true orthologs, and not the number of possible orthologs. 
-
-For the purposes of evaluating simulated network recovery, we define a gold standard network as the support of the beta matrices.
-Priors used in network inference are interactions present in the gold standard.
-The list of priors can be be manipulated to include false positives and false negatives as with ortholog mapping.
+The list of priors can in a similar fashion be manipulated to include false positives and false negatives.
 
@@
 \end{equation}
-When the residual is zero, then removing regressor $i$ with weight $\beta_{i,j}$ from the model of $j$ will increase the residual by $\beta_{i,j}^2$ times the variance of regressor $i$.
-This scaling is similar to rescaling according to variance explained relative to an augmented design matrix that includes fusion constraints.
-The intuition for scaling is that a large $\beta_{i,j}$ may be inferred because of overfitting with a TF that varies little across the available data, or because the regulatory weight is strong.
-Only in the latter case is a large weight suggestive of a true regulatory interaction.
 
@@
 \section{Results}
-We used both synthetic networks and real data to test the ability of fused regression to improve the performance of network inference, and the ability of adaptive fusion to identify conserved interactions between orthologous genes.
-For the synthetic data, we generated random pairs of networks in which orthologous genes have similar regulatory interactions, then sampled gene expression from these networks, which we used to derive learned networks for comparison with the input (true) networks.
-For real data, we computed recovery of a known gold standard in \textit{Bacillus subtilis}.
-We measured network inference performance with the area under the precision recall curve (AUPR).
+We used both synthetic networks and real data to test the ability of fused regression to improve the performance of network inference, and the ability of our adaptive fusion procedure to identify conserved interactions between orthologous genes.
+For the synthetic data, we generated random pairs of networks in which orthologous genes have similar regulatory interactions, and then sampled gene expression from these networks, which we used to derive learned networks for comparison with the input (true) networks.
+For real data, we computed recovery of a known gold standard in \textit{Bacillus subtilis} \cite{michna_subtiwikidatabase_2014}.
 
@@
 Additional data is not always readily available and does not always make identification of unique network weights possible; but often there is an abundance of related data. 
-Our method allows this data to be used in network inference.
-Using related data for network inference allows us to borrow statistical power from other sources, effectively increasing the sample size and boosting the sensitivity and specificity of learned interactions.
-We use the assumption that similar genes have similar regulatory interactions, and share information about putatively conserved interactions.
+Using related data for network inference allows us to leverage statistical power from disparate sources, effectively increasing the sample size and boosting the sensitivity and specificity of learned interactions.
 We created synthetic networks to approximate two related biological processes, then evaluated performance of our fused L2 regression, which learns the networks simultaneously given a prior on the relatedness of interactions.
 We compared recovery of the two artificial 10 TFs by 200 genes networks, using fused L2 versus learning networks separately, and varied the amounts of simulated expression data samples made available to the solver.
-We measured network inference accuracy with the mean squared error between generated (true) network weights and inferred weights.
 When the amount of data from the second species was held constant, increasing the amount of data available for learning the network for the first species resulted in a more accurate network prediction, as expected  (figure \ref{fusion}b).
@@
 On the other extreme, when the networks have diverged considerably, a large fusion weight may impair network recovery. 
-We performed an experiment on simulated data in a multi-species network-inference problem to assess the effect of the similarity of networks, and the effect of the fusion weight on network recovery (figure \ref{syntheticgrid}).
-We created pairs of synthetic networks with varying similarity.
-The main factors governing similarity between our generated networks was the extent (and accuracy) of the orthology mapping, and the variability between conserved, interactions (in other words, how well conserved the conserved subnetwork was).
-We conducted a series of simulations to assess the effect of increasing orthology coverage on network recovery.
-
-As expected, when the conserved subgraphs were very similar - when the interactions in the conserved subnetwork were nearly identical, simulating the case of closely related organisms or similar processes (figure \ref{syntheticgrid}a) - increasing the weight of the fusion penalty $\lambda_S$ improved network recovery.
+We performed an experiment on simulated data in a multi-species network-inference problem to assess the effect of the similarity of networks, and the effect of the fusion weight on network recovery (figure \ref{syntheticgrid}); pairs of synthetic networks with varying similarity were then built.
+The main factors governing similarity between our generated networks was the extent (and accuracy) of the orthology mapping, and the variability between conserved, interactions (degree of subnetwork conservation).
+We then conducted a series of simulations to assess the effect of increasing orthology coverage on network recovery.
+When the conserved subgraphs were very similar - when the interactions in the conserved subnetwork were nearly identical, emulating the case of closely related organisms or similar processes (figure \ref{syntheticgrid}a) - increasing the weight of the fusion penalty $\lambda_S$ improved network recovery.
 As the size of the conserved subgraph increased, this effect was enhanced.
-
 To simulate the case of distantly related organisms, we created networks where the differences between conserved interaction weights were drawn from high variance distribution, mimicking weak conservation (figure \ref{syntheticgrid}b).
-When the conservation is weak - when the interactions in the conserved subnetwork are nearly as different across species as randomly chosen interactions - performance may not continue to improve by increasing the fusion weight.
-We showed that even for networks where the conserved subgraph was weakly conserved, there exists a 'sweet spot' where fusion regression improves network recovery.
+We showed that even for networks where the conserved subgraph was weakly conserved, there exists a model parameterization where fusion regression improves network recovery.
 
 \subsection{Fused regression improves performance on both the constrained and non-constrained parts of the network}
-Our method learns related networks simultaneously, using predictions about conservation of TF-gene interactions to share information between sources.
-Our approach is useful for learning networks from similar sources such as related cell types from the same species, where there exists a one-to-one mapping of genes, as well as datasets where the orthology mapping does not span all the genes.
-This can occur when using different technology, eg microarray and RNAseq, where one method does not assay all of the genes that another one does, but the networks are expected to be very similar; an example of the latter case occurs when inferring networks describing distantly related species, in which case the networks are expected to be more different than similar.
-When orthology is not complete, we are interested in knowing if performance gains from fused regression are limited to those interactions which have fusion constraints, or if they extend to the entire network.
-To test this, we used sets of 20 TF by 200 gene synthetic networks, with varying proportions of orthologous TFs and genes.
-We divided networks into those interactions with fusion constraints (interactions where there exist putative analogous interactions in the second network based on the orthology mapping between TF, gene pairs), which we call the constrained subnetwork, and interactions without fusion constraints, which we refer to as the non-constrained subnetwork.
+Our approach is useful for learning networks from similar sources such as related cell types from the same species, where there exists a one-to-one mapping of genes, as well as datasets where the orthology mapping does not span all genes .
+This can occur when using different technology, eg microarray and RNAseq, where there is incomplete overlap in the genes that each method assays as well as incomplete overlap in the genes expressed in different experimental designs.
+When orthology is incomplete we are interested in knowing if performance gains from fused regression are limited to those interactions which have fusion constraints, or if they extend to the entire network.
+To test this we used multiple 20 TF by 200 gene synthetic networks with varying proportions of orthologous TFs and genes. 
+We divided networks into those interactions with fusion constraints (the constrained subnetwork) and interactions without fusion constraints (the non-constrained subnetwork). 
 We varied the weight on the fusion penalty, lamS, and evaluated performance by computing AUPR on the constrained subnetwork, the non-constrained subnetwork, and the whole network (figure \ref{fusion}c).
@@
 Interestingly, performance gains were seen even in the portion of the network that was unconstrained by fusion.
-By constraining part of an undetermined system, we obtained gains in even the unconstrained part.
 
 \subsection{Adaptive fusion successfully identifies and unfuses 'neofunctionalized' genes}
-We recognize that orthology prediction is not a perfect proxy for functional conservation \cite{gabaldon_functional_2013, studer_how_2009, nehrt_testing_2011}.
-We implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (figure \ref{adapt-schematic}).
-Pairs of interactions which are very dissimilar even after fusion, which sit in the flat portion of this penalty function, are effectively ``unfused,'' and no further penalty is incurred as differences in interaction weights grow.
+Orthology prediction is not a proxy for functional conservation \cite{gabaldon_functional_2013, studer_how_2009, nehrt_testing_2011}. 
+To allow for orthologous genes to be unfused we implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (figure \ref{adapt-schematic}).
+Pairs of interactions that are  dissimilar even after fusion, which sit in the flat portion of this penalty function, are effectively ``unfused,'' and no further penalty is incurred as differences in interaction weights grow.
 Our network procedure strongly favors similarity of fused interactions, and only ``unfuses'' interactions when their similarity cannot be reconciled with expression data. 
-Interactions are only ``unfused'' if there is no network configuration consistent with data from both organisms.
 As a result, the ``unfusing'' or relaxation of the fusion penalty on certain constraints is much more direct evidence for neofunctionalization than comparing separately fit networks could provide.
 
-We performed a simulation to assess the ability of our adaptive fusion algorithm to learn which parts of two input networks are conserved (figure \ref{adaptivefusion}).
+We performed a simulation to assess the ability of our adaptive fusion algorithm to distinguish which parts of two input networks are conserved vs.
+neofunctionalized (figure \ref{adaptivefusion}).
 We generated synthetic fused networks and introduced error in the fusion constraints by adding false positives and negatives to the orthology information given to the solver.
@@
 We verified that adaptive-fusion unfused mostly ``incorrectly fused'' interactions (figure \ref{adaptivefusion}a red dots), while leaving truly analogous interactions fused (figure \ref{adaptivefusion}a green dots).
-Note that while pairs of interactions between true orthologs were sampled randomly so that the difference in interactions tended to be small, and pairs of interactions between incorrect orthologs were sampled randomly so that the difference in interactions tended to be large, there was an element of randomness in the sampling.
-This resulted in a small fraction of interactions between incorrect orthologs being similar, and similarly, a small fraction of interactions between correct orthologs being different.
-Consequently, some interactions between incorrect orthologs remained fused and some interactions between correct orthologs were unfused.
 
-We compared the recovery of interaction weights which were accurately fused, and recovery of interaction weights which were inaccurately fused due to incorrect orthology information (figure \ref{adaptivefusion}b).
+We then compared the recovery of interaction weights which were accurately fused, and recovery of interaction weights which were inaccurately fused due to incorrect orthology information (figure \ref{adaptivefusion}b).
 Because fused L2 heavily penalizes large differences between weights which are predicted to be similar, it is able to retrieve a more accurate network for those interactions with true fusion constraints than by learning networks separately (measured by MSE between the true and inferred interaction weights) .
@@
 However, we were able to identify fusion constraints reflecting incorrect orthology information that had been provided to the algorithm (figure \ref{adaptivefusion}a).
-Even in situations where adaptive fusion does not substantially improve recovery - which may occur when unfused interactions remain underconstrained - the approach shows promise in identifying non-conserved interactions, which is of direct interest to systems biology.
 
 \subsection{Cross-species network inference using bacterial data}
-We used gene-expression data from \textit{Bacillus subtilis} and \textit{B.
-anthracis} in order to assess performance gains of fused regression on real data.
+We used gene-expression data from \textit{Bacillus subtilis} and \textit{B. anthracis} in order to assess performance gains of fused regression on real data. 
 Our \textit{B subtilis} data set consists of 360 time-series and steady-state observations of 4891 genes, 4100 of which are protein coding \cite{kunst_complete_1997}, during the life cycle. 
@@
 anthracis}} dataset consists of 72 time-series and steady-state observations of 5536 genes comprising data from distinct points in the life cycle and iron-starvation conditions.
-There were 247 known transcription factors (TFs) in the \textit{\textit{B.
-subtilis}} dataset, and 248 TFs in the \textit{B.
-anthracis} dataset.
+There were 247 known transcription factors (TFs) in the \textit{\textit{B. subtilis}} dataset, and 248 TFs in the \textit{B. anthracis} dataset.
+We obtained 1,870 one-to-one orthologs from Inparanoid \cite{ostlund_inparanoid_2010}, 95 of which are transcription factors, which produced 177,650 fusion-constraints between gene interactions within the two species.
+This number represents only $14.7\%$ of the regulatory interaction matrix in \textit{B. subtilis} and $12.9\%$ in \textit{B. anthracis}.
 
-We obtained 1,870 one-to-one orthologs from Inparanoid \cite{ostlund_inparanoid_2010}, 95 of which are transcription factors, which produced 177,650 fusion-constraints between gene interactions within the two species.
-This number represents only $14.7\%$ of the regulatory interaction matrix in \textit{B.
-subtilis} and $12.9\%$ in \textit{B.
-anthracis}.
-The use of such dissimilar species is a test of whether fusion can improve network-inference even when the overlap in interactions is very small.
-
-To assess network inference performance, and for use as priors, we used a gold standard of 3,040 known \textit{B.
-subtilis} interactions with corresponding activation and repression sign.
-Of these 3,040 priors, 968 had corresponding interactions in \textit{B.
-anthracis}.
+To assess network inference performance, and for use as priors, we used a gold standard of 3,040 known \textit{B. subtilis} interactions with corresponding activation and repression sign. 
+Of these 3,040 priors, 968 had corresponding interactions in \textit{B. anthracis}. 
 Based on our simulation results, we can expect the greatest gains in network-inference performance from fusion when the species of interest has a small number of available conditions, but data is abundant in a related species.
 However, in order to evaluate performance objectively a gold-standard of known interactions is necessary. 
-As a result, we can only evaluate network recovery for \textit{B.
-subtilis}, and \textit{B.
-subtilis} also has the majority of our conditions.
-In order to simulate the data-poor regime, we subsampled our \textit{B.
-subtilis} data.
-Specifically, we divided our \textit{B.
+As a result, we can only evaluate network recovery for \textit{B. subtilis}, and \textit{B. subtilis} also has the majority of our conditions. 
+In order to simulate the data-poor regime, we subsampled our \textit{B. subtilis} data. 
+We divided our \textit{B.
 subtilis} data into $k$ folds, and then for each fold fit a network to the \textit{B.
 subtilis} data from that fold alone fused to the entire 72 \textit{B.
-anthracis} conditions.
-We can then compute performance metrics (ie AUPR) for each fold, and use their variability across folds to test for significance of any improvements relating to the use of fusion constraints (figure \ref{xspecies-real}a).
+anthracis} conditions (figure \ref{xspecies-real}a).
+Though overall performance is hindered by our subsampling of \textit{B.
+subtilis} data (a necessary procedure to allow evaluation of networks) we demonstrate marked improvement in learning the \textit{B.
+subtilis} network when using fused regression (figure \ref{xspecies-real}a).
+Notably, these performance gains occur mostly at low values of recall (near the top of our prediction ranks, where biologists would presumably focus validation and followup experiments).
 
-Though overall performance is hindered by our subsampling -- a necessary procedure to allow evaluation of networks, we demonstrate marked improvement in learning the \textit{B.
-subtilis} network when using fused regression (figure \ref{xspecies-real}a).
-Notably, these performance gains occur mostly at low values of recall.
-Essentially, precision increases for the high confidence part of the network but less so for the low confidence part of the network.
-As a result, the improvements in performance may be more useful than would be suggested by AUPR alone.
-Interactions suggested by genome wide analysis are useful for target prediction, and the ranking of interactions produced by a network inference algorithms is a useful guide to what order these experiments should be carried out in.
-The high confidence part of the network represents interactions that are likely candidates for experimental validation, so improvement in recovery of these interactions is particularly useful in guiding future experiments.
-
-\subsection{Adaptive fusion using bacterial data}
-We anticipated that the fusion constraints we introduced based on orthology reflected a mixture of conserved and non-conserved interactions.
+\subsection{Testing adaptive fusion using bacterial data}
  The goal of adaptive fusion is to unfuse constraints between non-conserved interactions, while leaving intact all other constraints. 
-However, because we lacked a gold standard of known non-conserved interactions between \textit{B.
+ However, because we lacked a comprehensive gold standard of known non-conserved interactions between \textit{B.
 subtilis} and \textit{B.
@@
  These fake constraints, which are unlikely to reflect any conserved network structure, served as a proxy for the unknown fraction of non-conserved interactions between orthologous genes. 
-We ran adaptive fusion to learn networks for \textit{B.
-subtilis} and \texti{B.
-anthracis}, using these constraints, along with those generated by known orthology.
-We confirmed that fake fusion constraints were unfused at a higher rate than those generated by known orthologs (figure \ref{xspecies-real}b).
+ We ran adaptive fusion to learn networks for \textit{B. subtilis} and \textit{B. anthracis}, using these constraints, along with those generated by known orthology. 
+ We confirmed that our injected spurious fusion constraints were unfused at a higher rate than those generated by known orthologs (see figure \ref{xspecies-real}b & c). 
  Although it may seem odd that a large fraction of fake constraints were left intact, we note that biological networks tend to be sparse, so that many of the random fusion constraints are between coefficients with near zero weight (and therefore near zero difference in weight) (figure \ref{xspecies-real}c).
 
-\subsection{Integrating datasets using fused regression}
-Although there are many large-scale collaborations which attempt to make protocols as uniform as possible for comparability between datasets generated by different labs \cite{paten_nih_2015,kundaje_integrative_2015}, there still exists technical and biologial variability between many experiments attempting to capture the same or similar experimental conditions.
-With the advent of new technology, such as RNAseq and single-cell RNAseq, microarray is no longer the dominant assay for genome-wide expression, but a large body of accumulated legacy data remains useful, if it can be integrated with more modern techniques.
+\subsection{Integrating datasets from different platforms using fused regression}
+Although there are many large-scale collaborations which attempt to make protocols as uniform as possible for comparability between datasets generated by different labs \cite{paten_nih_2015,kundaje_integrative_2015} and several methods for removing batch effects \cite{irizarry_exploration_2003,johnson_adjusting_2007}, there still exists technical and biologial variability between many experiments attempting to capture the same or similar experimental conditions esspecially when experiments employ different experimental platforms.
+With the advent of RNAseq, for example, microarray based technologies are no longer the dominant assay for genome-wide expression, but a large body of accumulated legacy data remains useful if it can be integrated with more modern techniques.
 Currently, the most widely used approach to combining datasets for network inference is to learn networks from disparate datasets separately, then rank combine the networks as in Marbach et al \cite{marbach_revealing_2010}.
-We compared this approach to using fusion regression.
 We included, along with our \textit{B.
@@
 subtilis} \cite{nicolas2012condition}.
-We compared performance when learning the networks separately, rank combining, and learning the networks simultaneously using fusion regression, and showed large improvement on network inference using our fused L2 approach (figure \ref{fusedl2-real}a).
+We compared performance when learning the networks separately and then rank combining (as in cite Ciofani) to learning the networks simultaneously using fusion regression and show large improvement in performance using our fused L2 approach (figure \ref{fusedl2-real}a).
 
-\subsection{Within-species fusion using similarity of promoter region}
-Information about the similarity of TF-gene interactions can also come from knowledge about the promoter region; this may be a better predictor of target similarity than orthology.
-In bacteria, genes within the same operon are under the control of the same promoter \cite{lawrence_shared_2002}.
-We predicted, therefore, that genes within the same operon will be regulated similarly by the same transcription factors.
+Information about the similarity of TF-gene interactions can also come from knowledge about the promoter region or the structure, for bacteria, of polycistronic transcripts. 
+In bacteria, genes within the same operon are typically under the control of the same promoter \cite{lawrence_shared_2002}. 
+We posited, therefore, that genes within the same operon will be regulated similarly by the same transcription factors. 
 We applied fusion regression by creating fusion constraints between a given transcription factor and genes within the same operon, and showed a boost in \textit{B.
@@
 subtilis} by Arrieta-Ortiz et al \cite{arrieta-ortiz_experimentally_2015}.
-TFA captures transcription factor activities through mechanisms such as dimerization and interaction with required factors, and is a better predictor of TF function than expression level alone \cite{fu_reconstructing_2011}.
-We estimated TFA based on known regulatory interactions using network component analysis \cite{liao2003network}.
+TFA activity estimates  transcription factor activities that are modulated through mechanisms such as dimerization and interaction with required factors.
+TFA activity estimates have been shown prior to be better predictors of TF function than expression level alone in several contexts including similar network inference tasks\cite{fu_reconstructing_2011} \cite{arrieta-ortiz_experimentally_2015}.
+We estimate TFA based on known regulatory interactions using network component analysis \cite{liao2003network}.
 To test the integration of this approach with our fused regression, we assessed the combination of \textit{B.
 subtilis} datasets, as in figure \ref{fusedl2-real}a, with the incorporation of TFA estimation.
-We randomly divided the prior known interactions in half, and used half to learn TFA.
-The rest we reserved as a gold standard for validation.
+We randomly divided the prior known interactions in half, and used half to learn TFA and to generate priors on network structure. 
+The remaining interactions were reserved as a gold standard for validation. 
 As in previous studies, we observed a marked improvement in network inference when using transcription factor activity (figure \ref{tfa}). 
@@
 Gene expression data, such as microarray or RNA seq, provide information about the relationship between genes by allowing an experimenter to measure correlations in expression value over time or across conditions.
-Many sources of information - such as the knowledge that two genes are related through orthology or belong to the same operon - provide information about the relationships between these gene-gene relationships, without necessarily providing information about the gene-gene relationships themselves.
-For example, the meta information that two genes belong to the same operon suggests that they are likely to have a similar set of regulators \cite{lawrence_shared_2002}, but does little to inform the identity of those regulators.
-Meta-information about the structure of gene regulatory networks - specifically which pairs of interactions are \textit{a priori} likely to be similar to one another, can provide a powerful set of constraints to improve network inference performance.
-
+Many sources of information - such as the knowledge that two genes are related through orthology or belong to the same operon - provide additional information about the relationships between these gene-gene relationships.
+For example,  two genes that belong to the same operon are likely to have a similar set of regulators \cite{lawrence_shared_2002}, but knowing that two genes are members of a polycistronic transcript does little to inform the identity (strength, sign) of those regulators.
+Meta-information about the structure of gene regulatory networks, specifically which pairs of interactions are \textit{a priori} likely to be similar to one another, can provide a powerful set of constraints to improve network inference performance \cite{roy_arboretum:_2013, pierson_sharing_2015}.
 We present a general framework for gene regulatory network inference that incorporates this meta-information - termed fusion constraints - and apply the technique to the problem of simultaneous inference of regulatory networks in multiple species (\textit{B.
@@
 anthracis}).
-We develop two algorithms for solving fused network inference problems, each of which has as its core the solving of a least-squares optimization problem with novel regularization constraints.
 
-The first method, fused L2, penalizes the squared difference of the regulatory weights of fused interactions.
-These constraints improve the performance of the underdetermined network inference problem, by allowing information to be shared across sources.
 We apply this algorithm to the problem of network inference in two distantly related biological organisms -- \textit{B.
@@
 anthracis} -- and show that network recovery is improved through the introduction of fusion constraints between pairs of orthologous genes.
-Existing methods for cross-species network inference operate on the conserved subset of orthologous genes \cite{dillman_comparative_2015}.
-This may be appropriate with very closely related species, but could not be applied in this domain, where a large fraction (62\% and 67\%) of the \textit{B.
+Many previous methods for cross-species network inference operate on the conserved subset of orthologous genes \cite{dillman_comparative_2015}. 
+This assumption may be appropriate with very closely related species, but could not be applied in this domain, where a large fraction (62\% and 67\%) of the \textit{B.
 subtilis} and \textit{B.
-anthracis} genomes do not have orthologs.
+anthracis} genomes do not have clear orthologs (and many orthologs have ambiguous many-to-many groupings).
 Our method, in contrast, can obtain improvements in network inference performance even when the conserved subset of genes is small. 
+This approach is particularly interesting in light of the diversity of important model organisms used in modern biology. 
+Different model systems provide different advantages and disadvantages for experimental design \cite{stolfi_genetic_2012}, but without a principled mechanism for combining data from multiple sources, it is difficult to fully leverage data obtained from even a slightly different model system.
 We further demonstrate the viability of fused L2 as a method for combining data from multiple experimental platforms, where fusion is between each identical regulator-gene pair. 
-
 Because the algorithm we developed can accomodate constraints between arbitrary pairs of regulatory interactions, any biological prior representing information about expected regulatory similarity can be represented, even if the prior provides no information about the magnitude or direction of regulation.
@@
 subtilis} datasets and show that fused L2 is an improvement over current approaches to combining data \cite{marbach_revealing_2010} because of our ability to exploit the statistical power that our expanded datasets afford us.
-This approach is particularly interesting in light of the diversity of model organisms used in modern biology.
-Different model systems provide different advantages and disadvantages for experimental design \cite{stolfi_genetic_2012}, but without a principled mechanism for combining data from multiple sources, it is difficult to fully leverage data obtained from even a slightly different model system.
 
@@
 Existing approaches to the genome-wide testing of this assumption learn regulatory networks separately, then compare to identify conservation \cite{aytes_cross-species_2014, Wang2014a}. 
-But because network inference is typically underconstrained, fitting a network that describes a particular set of experimental observations involves sampling a single network from a large set of networks that fit the data equally (or almost equally) as well.
+Because network inference is typically underconstrained, fitting a network that describes a particular set of experimental observations involves sampling a single network from a large set of networks that fit the data equally (or almost equally) as well.
 As a result, the existence of a difference between corresponding regulatory interactions in a pair of experimentally derived networks is weak evidence that a difference truly does exist.
-It may be that some other pair of networks exist that fit the data nearly as well, but do not exhibit this difference.
-As a result, global network inference algorithms are a very weak tool for uncovering evolutionary divergence.
-Our method, on the other hand, explicitly favors recovering networks for which evolutionarily corresponding interactions are similar.
-As a result, the failure to obtain networks that confirm evolutionary conservation is strong evidence that conservation does not exist; the next best network that does exhibit conservation must fit the data much worse to have overcome the bias built into the fusion constraints.
+Uncoupled global network inference algorithms are a very weak tool for uncovering evolutionary divergence. 
+Our method explicitly favors recovering networks for which evolutionarily corresponding interactions are similar. 
+As a result, the failure to obtain networks that confirm evolutionary conservation is stronger evidence that conservation does not exist; the next best network that does exhibit conservation must fit the data much worse to have overcome the bias built into the fusion constraints.
 
-Because our approach allows us to directly test -- and reject -- the hypothesis of conserved function, and because the enforcement of fusion constraints between non-conserved interactions may degrade network inference performance, we introduce an adaptive fusion method that decouples pairs of interactions for which expression data suggests a lack of conservation.
-Like the fused L2 approach, this method attempts to solve for networks which minimize error in fitting the expression data plus error associated with the fusion penalty.
-When a fusion constraint encourages orthologous gene, TF pairs to share similar interactions across species, but the expression data suggests that these interaction weights should not be similar, the algorithm will arrive at some compromise between these two forces.
-If the disagreement arose because the genes involved have taken on new function through evolutionary divergence \cite{kellis_proof_2004}, network recovery would be improved had the fusion constraint not been enforced.
-We introduce a method -- adaptive fusion -- that attempts to learn which fusion constraints should be relaxed while the network is being learned.
+We have described a method -- adaptive fusion -- that attempts to learn which fusion constraints should be relaxed while the network is being learned. 
 This method is based on minimizing a saturating penalty function on fusion constraints, similar to a class of penalties that have been developed to minimise bias in regularized regression \cite{fan2001variable,zhang2010nearly}.
-The result of adaptive fusion is both a network and a new set of fusion constraints, describing the learned conservation weights (including which fusion constraints have been relaxed).
-For the multiple species case, relaxation of fusion constraints represents orthologs which do not share similar interactions.
+The result of adaptive fusion is both a network and a new set of fusion constraints, describing the learned fusion weights (including which fusion constraints have been relaxed). 
+For the multiple species case, relaxation of fusion constraints represents orthologs which do not share similar interactions presumably due to evolution of regulatory circuitry \cite{kellis_proof_2004}.
+%not sure what radid is
 When jointly learning networks describing processes in different cell lines, this may identify interesting context-specific behavior. 
 Genes may be fused together on the basis of similar binding sites or chromatin features, and the relaxing of the fusion penalty indicates divergence of gene function.
-Adaptive fusion, therefore, is a tool for testing network conservation as well as a tool for network inference.
 
@@
 The performance gains of this additional step in the cross-species case are significant.
-
 Our approaches -- fused L2 and adaptive fusion -- represent a very general framework for simultaneous network inference and the incorporation of structured biological priors. 
@@
 At the same time these methods allow us to test our assumptions on how similar biological systems relate to one another, by allowing us to rule out conservation in a principled way, and at the genome-wide scale.
-
-%One future aim is the ability to pick the $a$ parameter controlling the saturation point of the adaptive fusion penalty automatically from the expression data; this parameter is biologically interesting and indicates the difference in interaction weights which is great enough that the fusion constraints should be relaxed.
-An exhaustive search through different combinations of putative fusion constraints is prohibitive, but perhaps there is another way which avoids this.
 
@@
 We introduce fusion constraints for pairs of interactions for which both the regulator and regulated gene are orthologs of one another.
-In this example, we would introduce a constraint between the $(A, B)$ and $(A', B')$ interactions and the $(A, C)$ and $(A', B')$ interactions.
-\textbf{B.} In order to demonstrate the utility of fused network inference in combining data, we generate two networks with 10 TFs and 200 genes (75\% sparsity).
+  In this example, we would introduce a constraint between the $(A, B)$ and $(A', B')$ interactions and the $(A, C)$ and $(A', B')$ interactions. \textbf{B.} 
+  In order to demonstrate the utility of fused network inference in combining data, we generate two networks with 10 TFs and 200 genes (75\% sparsity).
 Mean squared error of the inferred vs.
@@
 At every level of conservation, constrained interactions show the largest benefits of fusion, with the magnitude of the benefit growing with fraction of orthologous genes.
-When the networks are highly conserved, however, even interactions that are not directly constrained through fusion are recovered more accurately as $\lambda_S$ increases.
+  When the networks are highly conserved, however, even interactions that are not directly constrained through fusion are recovered more accurately as $\lambda_S$ increases.}
 \end{center}
@@
 We varied the fraction of genes with orthologs (y-axis) and the weight of fusion in network inference (x-axis), and measured performance as the mean-squared error of the true vs inferred network weights.
-In order to more clearly visualize the varying effect of fusion, performance is plotted in units of the of the unfused MSE (left column) for each level of orthology.
+  In order to more clearly visualize the varying effect of fusion, performance is plotted in relative units of the of the unfused MSE (left column) for each level of orthology. 
   This was necessary because each row represents a different pair of networks, generated with a different level of orthology, for which baseline performance varies. 
@@
 However, even when the fraction of genes with orthologs is relatively small (top row), we observe gains from fusion.
-Importantly, when fused interactions are nearly identical, larger fusion weights always outperform smaller fusion weights.
+  When fused interactions are nearly identical, larger fusion weights always outperform smaller fusion weights. 
   \textbf{B.} Simulates the case where genes which are orthologs may have different regulatory weights. 
-We generated a series of networks as in \textbf{A.}, but with larger (7.5\times) gaussian noise added the weights of fused interactions.
+  We generated a series of networks as in \textbf{A.}, but with larger (7.5 $\times$) gaussian noise added the weights of fused interactions. 
   As in \textbf{A.}, benefits from fusion were observed at every level of gene orthology.
@@
   \textbf{B.} Shows the derivative of the adaptive fusion penalty, which is used to implement adaptive fusion through local quadratic approximation.
-The adaptive fusion penalty is modified from SCAD (smoothly clipped absolute deviation) and MCP (minimax concave penalty) functions.
-It shares with these penalties the zero derivative far from the origin.
+The adaptive fusion penalty is modified from SCAD (smoothly clipped absolute deviation) and MCP (minimax concave penalty) functions and like these penalties has a zero derivative far from the origin.
 }
@@
 \begin{center}
-
   \includegraphics[scale=0.45]{fig4_2.pdf}
-  \caption{\label{adaptivefusion} In order to evaluate the performance of adaptive fusion in unfusing interactions that have diverged through evolution, we performed a series of simulations inferring networks given a partially false list of orthology mappings.
-These false orthologs were orthologs that were generated from random pairs of non-orthologous genes, and therefore did not reflect true similarity of regulatory interactions.
+  \caption{\label{adaptivefusion} In order to evaluate the performance of adaptive fusion in unfusing interactions that have diverged through evolution, we performed a series of simulations inferring networks given a partially corrupted list of orthology mappings.
   Networks were generated with 35 TFs by 200 genes, 60\% orthology coverage and 40\% false orthology coverage.
 \textbf{A.} We plot the interaction weights between pairs of fused interactions in network 1 (x-axis) and network 2 (y-axis) following network inference without fusion ($\lambda_S=0$).
-Interactions that are generated from false orthologs, which do not reflect true similarity of regulatory weight, are marked as red, while interactions generated from true orthologs are shown in green.
-Because fused interactions in the simulated networks were, in reality, almost identical in regulatory weight, interactions with true fusion constraints should fall on the diagonal.
-However, because a very small number of samples (20 for each species) was provided to the network inference procedure, the inferred weights are only weakly correlated (green dots).
-As expected, false fusion constraints give rise to uncorrelated weights (red dots).
-\textbf{B.} When fit with fused-L2, fusion constraints give rise to very similar weights in the two species.
-This is desireable for the true fusion constraints, but introduces error for the false constraints.
+  Interactions that are generated from false orthologs are marked as red, while interactions generated from true orthologs are shown in green. 
+  As expected, false fusion constraints give rise to uncorrelated weights (red dots). \textbf{B.} 
+  When fit with fused-L2, fusion constraints give rise to very similar weights in the two species for 'true' and 'false' interactions. 
   \textbf{C.} Adaptive fusion run on the same network unfuses constraints for which the inferred weights are dissimilar beyond a certain point. 
-In this example, unfused interactions are almost entirely interactions between false orthologs.
+  Here, unfused interactions are almost entirely interactions between false orthologs.
 \textbf{D-F} Show the distribution of the absolute value of the difference in inferred weight for interactions with true fusion constraints (green) and false fusion constraints (orange)}
@@
   \includegraphics[scale=0.45]{fig5_2.pdf}
-  \caption{\label{xspecies-real} a.
-Using optimal lamR and lamP values, we use 10-fold cross validation to learn B.
-subtilis network.
-We compare performance when solving without fusion, and using L2 fusion with lamS = 1.0.
-Mean AUPR for lamS = 0 (unfused): 0.0619.
-Mean AUPR for lamS = 1: 0.0839.
-b.
-We test adaptive fusion using the same setup, with the addition of false orthology information.
-We set a 30\% false positive rate, including 561 additional orthologs.
+  \caption{\label{xspecies-real} a. Using optimal lamR and lamP values, we use 20-fold cross validation to learn B. subtilis network. 
+  We comparedare performance when solving without fusion, and using L2 fusion with lamS = 1.0. Mean AUPR for lamS = 0 (unfused): 0.0298. 
+  Mean AUPR for lamS = 1: 0.0388. b. We test adaptive fusion using the same setup, with the addition of false orthology information.
+  We set a 30\% false potentialsitive rate, including 561 additional orthologs.
 We run adaptive fusion, setting the $a$ term equal to the value above which 40\% of constraints would unfuse, reflecting our belief that in addition to some of the known orthologs result in fusion constraints which should be relaxed.
   Here we show the percentage of constraints relaxed, from constraints created from the false orthologs and the known orthologs. 
-We also show the distributions of differences in weights corresponding to fused constraints created from the false orthologs and the known orthologs, when networks are solved separately.
-}
+  We also show the distributions of differences in weights corresponding to fused constraints created from the false orthologs and the known orthologs, when networks are solved separately. }
 \end{center}
@@
 subtilis with gold standard.
-Here, we fuse together TF, gene pairs, fixing the same TF with genes in the same operon group.
-The assumption is that genes in the same operon should be regulated similarly by the same TFs.
-We compare L2 fusion performance using operons with unfused.}
+  Here, we fuse genes in the same operon group and compare L2 fusion performance using operons with unfused network inference.}
 \end{center}
@@
   \includegraphics[scale=0.45]{fig7.pdf}
-  \caption{\label{tfa} We integrate transcription factor activity (TFA) into our network inference, and show that fusion improves performance.Here we learn the B.
-subtilis-1 network with and without fusion, and with and without TFA}
+  \caption{\label{tfa} We integrate transcription factor activity (TFA) into our network inference, and solve for B.
+subtilis-1 and B.
+subtilis-2 networks using 10-fold cross validation, using fused L2 with and without TFA.
+As in figure 6, we evaluate performance on B.
+subtilis-1 using AUPR on gold standard.}
 \end{center}
@@
     \includegraphics[scale=.5,trim=0mm 0mm 0mm 0mm,clip]{party3.eps}
-    \caption{\label{cartoon} Schematic representation of design matrix construction.
-Here, the circles and hexagons correspond to different species.
+    \caption{\label{cartoon} Schematic representation of design matrix construction. Here, the circles and hexagons correspond to different species. 
     Bidirectional arrows represent orthology information and dotted arrows represent putative interactions between TFs and genes. 
-Rectangles represent matrices; because weights can be solved independently unless there exists fusion constraints between them, we identify related weights and construct matrix for solving.
-}
+    Rectangles represent matrices; because weights can be solved independently unless there exists fusion constraints between them, we identify related weights and construct matrix for solving. }
 \end{figure}

### 2016-07-25 paper_revisions.tex
hi

--- 2016-07-24 paper1.tex
@@
 Observing a large difference in the weights of regulatory interactions obtained though independent inference of multiple networks is perhaps the best (least biased) evidence against conservation of orthologous regulatory interactions (cases where target and regulator have orthologs across species).
-However, this is sometimes weak evidence, as network inference is typically underconstrained \cite{marbach_revealing_2010-1}, meaning there could be a different set of networks for which conservation does hold, and which fit the data almost as well.
+However, this is sometimes weak evidence, as network inference is typically underconstrained \cite{marbach_revealing_2010}, meaning there could be a different set of networks for which conservation does hold, and which fit the data almost as well.
 We propose using our adaptive fusion approach to simultaneously perform network inference and evaluate edge-conservation (or lack thereof).
@@
 \argmin_{\beta^S} \displaystyle\sum_{S \in \{1, 2\}} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \lambda_S \displaystyle \sum_{\substack{(g,h) \in orth,\\
- (h,l) \in orth}}p(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
+ (k,l) \in orth}}p(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
 \end{equation}
@@
 In addition to performing well between these extremes, our method allows pooling of data even when there is incomplete orthology. 
-By introducing constraints on the similarity of individual interactions, rather than on the networks as a whole \cite{liu2011temporal}, we can pool some information across species even when a small fraction of genes have orthologs.
+By introducing constraints on the similarity of individual interactions, rather than on the networks as a whole \cite{parikh2011treegl}, we can pool some information across species even when a small fraction of genes have orthologs.
 
@@
 It is known that the accuracy of network inference improves with additional data \cite{bar-joseph_computational_2003}. 
-Additional data is not always readily available and does not always make identification of unique network weights possible; but often there is an abundance of related data. 
 Using related data for network inference allows us to leverage statistical power from disparate sources, effectively increasing the sample size and boosting the sensitivity and specificity of learned interactions.
 We created synthetic networks to approximate two related biological processes, then evaluated performance of our fused L2 regression, which learns the networks simultaneously given a prior on the relatedness of interactions.
-We compared recovery of the two artificial 10 TFs by 200 genes networks, using fused L2 versus learning networks separately, and varied the amounts of simulated expression data samples made available to the solver.
+We compared recovery of the artificial 10 TFs by 200 genes networks, using fused L2 versus learning networks separately, and varied the amounts of simulated expression data samples made available to the solver.
 When the amount of data from the second species was held constant, increasing the amount of data available for learning the network for the first species resulted in a more accurate network prediction, as expected  (figure \ref{fusion}b).
@@
 
-When disparate data is generated from identical processes, the weight assigned to fusion constraints should be very large. 
-On the other extreme, when the networks have diverged considerably, a large fusion weight may impair network recovery. 
-We performed an experiment on simulated data in a multi-species network-inference problem to assess the effect of the similarity of networks, and the effect of the fusion weight on network recovery (figure \ref{syntheticgrid}); pairs of synthetic networks with varying similarity were then built.
-The main factors governing similarity between our generated networks was the extent (and accuracy) of the orthology mapping, and the variability between conserved, interactions (degree of subnetwork conservation).
-We then conducted a series of simulations to assess the effect of increasing orthology coverage on network recovery.
-When the conserved subgraphs were very similar - when the interactions in the conserved subnetwork were nearly identical, emulating the case of closely related organisms or similar processes (figure \ref{syntheticgrid}a) - increasing the weight of the fusion penalty $\lambda_S$ improved network recovery.
-As the size of the conserved subgraph increased, this effect was enhanced.
-To simulate the case of distantly related organisms, we created networks where the differences between conserved interaction weights were drawn from high variance distribution, mimicking weak conservation (figure \ref{syntheticgrid}b).
-We showed that even for networks where the conserved subgraph was weakly conserved, there exists a model parameterization where fusion regression improves network recovery.
-
 \subsection{Fused regression improves performance on both the constrained and non-constrained parts of the network}
-Our approach is useful for learning networks from similar sources such as related cell types from the same species, where there exists a one-to-one mapping of genes, as well as datasets where the orthology mapping does not span all genes .
+Our approach is useful for learning networks from similar sources such as related cell types from the same species, where there exists a one-to-one mapping of genes, as well as datasets where the orthology mapping does not span all genes.
 This can occur when using different technology, eg microarray and RNAseq, where there is incomplete overlap in the genes that each method assays as well as incomplete overlap in the genes expressed in different experimental designs.
@@
 We divided networks into those interactions with fusion constraints (the constrained subnetwork) and interactions without fusion constraints (the non-constrained subnetwork). 
-We varied the weight on the fusion penalty, lamS, and evaluated performance by computing AUPR on the constrained subnetwork, the non-constrained subnetwork, and the whole network (figure \ref{fusion}c).
-Since the conserved subgraphs were similar to each other (more like figure \ref{syntheticgrid}a than \ref{syntheticgrid}b), we expected performance to improve as the fusion penalty weight increased. 
+We varied the weight on the fusion penalty, $\lambda_S$, and evaluated performance by computing AUPR on the constrained subnetwork, the non-constrained subnetwork, and the whole network (figure \ref{fusion}c).
+Since the conserved subgraphs were similar to each other, we expected performance to improve as the fusion penalty weight increased. 
 We observed this, particularly for the constrained subnetwork. 
@@
 Interestingly, performance gains were seen even in the portion of the network that was unconstrained by fusion.
+This is because a gene may have some some interactions that are constrained by fusion - regulation by TFs with orthologs - and some interactions that are unconstrained - regulation by TFs without orthologs.
+Because both constrained and unconstrained components compete to explain the same pattern of gene expression, improving recovery of the constrained sub-network will tend to improve recovery for the unconstrained sub-network as well.
 
@@
 When we applied adaptive fusion, we did not observe an improvement in network recovery (relative to fused L2).
-However, we were able to identify fusion constraints reflecting incorrect orthology information that had been provided to the algorithm (figure \ref{adaptivefusion}a).
+However, we were able to identify fusion constraints reflecting incorrect orthology information that had been provided to the algorithm (figure \ref{adaptivefusion}subsection).
 
-\subsection{Cross-species network inference using bacterial data}
+\c{Cross-species network inference using bacterial data}
 We used gene-expression data from \textit{Bacillus subtilis} and \textit{B. anthracis} in order to assess performance gains of fused regression on real data. 
@@
 Our cross-species network inference method is premised on the assumption that orthologous genes have similar regulators. 
-Existing approaches to the genome-wide testing of this assumption learn regulatory networks separately, then compare to identify conservation \cite{aytes_cross-species_2014, Wang2014a}. 
+Existing approaches to the genome-wide testing of this assumption learn regulatory networks separately, then compare to identify conservation \cite{aytes_cross-species_2014}. 
 Because network inference is typically underconstrained, fitting a network that describes a particular set of experimental observations involves sampling a single network from a large set of networks that fit the data equally (or almost equally) as well.

### 2016-07-25 paper_revisions.tex
zack edits to paper revisions. minor clarifications up to methods, major reordering of methods

@@
 The presence of shared structure in a well studied model system or process should make the problem of network inference in a related process easier, but this information is not often applied to the discovery of global gene regulatory networks.
-Further, to be able to successfully translate findings between different organisms, it is important to be able to identify what is systematically different.
-We provide a method for gene regulatory network inference using data from related sources.
-We introduce approaches for the simultaneous inference of related networks, and the identification of conserved subnetworks.
-The method is benchmarked on synthetic networks, and evaluated on the problem of inferring networks in related bacterial datasets.
+Further, to be able to successfully translate findings between different organisms, it is important to be able to identify where regulatory structure is different.
+We provide a method based on penalized fused regression for inferring gene regulatory networks given prior knowledge about the similarity of interactions in each network.
+This method is demonstrated on synthetic data, and applied to the problem of inferring networks in distantly related bacterial organisms.
+We then introduce an extension of the method to deal with the condition of uncertainty over the degree of regulatory conservation by simultaneously inferring gene conservation and interaction weights.
 
@@
 As the volume and variety of genome scale data continues to increase in quantity and quality, the goal of accurately modeling gene regulatory networks has become attainable \cite{bonneau_predictive_2007, ciofani_validated_2012, carro_transcriptional_2010}.
-Large-scale data collection efforts have contributed to the development of high quality networks which accurately recapitulate biological processes, but most processes and organisms remain uncharacterized at the network level.
+Large-scale data collection efforts have contributed to the development of high quality networks which accurately represent biological processes, but most processes and organisms remain uncharacterized at the network level.
 Furthermore, as new technologies are developed and some old ones are replaced, such as RNAseq and microarray, it becomes important to be able to combine data from multiple platforms, lest we lose valuable information from existing studies.
@@
 The core of both methods is the observation that biological information about the relatedness of genes can be used to select which network coefficients should be similar to one another in a multi-source network inference problem (ie orthologous TFs should regulate orthologous genes), and that these constraints can be efficiently represented as penalties in a least-squares regression problem.
-Taking into account the similarity of putatively conserved interactions improves our ability to accurately describe TF-gene relationships on a genome-wide scale.
 
-Our first method -- fused ridge -- uses an L2 penalty on the differences between \textit{a priori} similar interactions (termed fusion penalty), and is useful where the relationships between networks (similarity of genes between data sets) is reliable.
-In the case where both networks contain an identical set of genes and TFs, this approach can be thought of as parametrically interpolating between treating the data sources separately and combining them together.
-In the case of multi-species, simply combining two datasets is both unwise -- because the networks may differ substantially -- but also potentially impossible, because the set of common of genes may be small.
-Our method allows useful pooling of data even when the overlap between genes is incomplete, or when orthology assignments depart from a strict one-to-one mapping.
-Our second method -- adaptive fusion -- uses a non-concave saturating fusion penalty to simultaneously infer the constrained networks and to learn which constraints should be relaxed (ie which parts of the network are genuinely different).
-With this approach, we seek to identify both conserved and divergent interactions between related networks.
+Numerous studies have shown that functional conservation exists in gene regulatory networks even across large evolutionary distance \cite{satou2006gene, hinman2009evolution,tanay2005conservation,erwin2009evolution}.
+Our first method -- fused ridge -- uses an L2 penalty on the differences between \textit{a priori} similar interactions (termed fusion penalty).
+These constraints favor, but do not require, network configurations in which orthologous genes have similar regulators.
+Because network inference problems are typically underconstrained, these additional constraints allow data in one species to improve network inference performance in another.
 
-In the case of multiple species, numerous studies have shown that functional conservation exists in gene regulatory networks even across large evolutionary distance \cite{satou2006gene, hinman2009evolution,tanay2005conservation,erwin2009evolution}.
-In our fused L2 approach, we assume that for closely related species, orthologous TFs could exert similar regulatory effects on orthologous target genes. 
-These orthology relationships form the basis of a set of constraints which favor -- but do not require -- networks in which orthologous transcription factors regulate orthologous genes.
-As a result, data in one species can improve network inference performance in another species  (and vice versa).
-This general framework for multi-species network inference can be extended to an arbitrary number of distinct organisms, each contributing data with only partially overlapping sets of genes. 
+By fusing arbitrary interactions within the networks rather than the networks themselves \cite{}, the method allows pooling of data even when there is not a one-to-one correspondence between genes, which is critical to the method's application to cross-species network inference.
+This general framework can be extended to an arbitrary number of distinct organisms, each contributing data with only partially overlapping sets of genes.
 This is an advantage over existing approaches to multi-species network inference, which infer only a sub-network for which orthologs exist in every species \cite{joshi_multi-species_2015}.
 
-This approach of introducing constraints on the similarity between specific regulatory interactions can be extended beyond the case of multi-species network-inference from orthology; any biological prior on similarity of regulation can be used in place of orthology.
-For example, we can use fused regression to combine datasets obtained using different platforms or experimental techniques and we can introduce constraints that favor genes in the same operon (or having similar promoters) towards having similar regulators.
+This general framework for multi-species network inference has broad applications.
+Any set of pairwise L2 penalties on the differences between regulatory interactions (equivalent to a multivariate Gaussian prior on networks; see methods) can be optimized by the algorithm we develop.
+As a result, any biological prior on similarity of regulation can be used in place of orthology.
+For example, we demonstrate improvements in network inference performance with the introduction of penalties based on the expectation that genes in the same operon will have similar regulators.
 
@@
 Orthology can be approximated using readily identifiable sequence similarity, which is often a useful predictor of functional similarity \cite{wilson_assessing_2000, jensen_eggnog:_2008}.
-In multi-species network inference, our fused L2 approach minimizes a cost function that strives to simultaneously fit expression data in each species and produce networks that are consistent with evolutionary constraints created using orthology.
 However, many genes will have evolved different functions and therefore may have new regulatory interactions.
 For example, gene duplications may lead to neofunctionalization \cite{eisen_phylogenomics:_1998} of the duplicated genes.
-In the case of comparing networks from related cell lines from the same species, changes in chromatin configuration may affect our hypotheses about the similarity of interactions between pleiotropic TFs and target genes across cell types (a within-species analog to neo-functionalization) \cite{li_role_2007}.
+Or, when comparing regulation across cell lines, changes in chromatin configuration may affect our hypotheses about the similarity of interactions between pleiotropic TFs and target genes across cell types (a within-species analog to neo-functionalization) \cite{li_role_2007}.
+
 Identifying interactions that are present in one species but not another is of direct biological interest, but existing approaches to network inference are unable to effectively test the hypothesis of conserved subnetworks.
@@
 However, this is sometimes weak evidence, as network inference is typically underconstrained \cite{marbach_revealing_2010}, meaning there could be a different set of networks for which conservation does hold, and which fit the data almost as well.
-We propose using our adaptive fusion approach to simultaneously perform network inference and evaluate edge-conservation (or lack thereof).
+Solving the networks jointly with fusion addresses this problem, but may be biased when gene function is not conserved.
 
-We approach this problem by introducing a saturating penalty function based on statistical efforts to develop unbiased regularization penalties for fused regression \cite{zhang2010nearly, fan2001variable}.
+Our second method -- adaptive fusion -- attempts to solve the problem of identifying evolutionary divergence using a non-concave saturating fusion penalty to simultaneously infer the constrained networks and to learn which constraints should be relaxed (ie which parts of the network are genuinely different).
+This penalty is based on statistical techniques intended to provide unbiased regularization penalties for fused regression \cite{zhang2010nearly, fan2001variable}.
+
 The main difference between the L2 fusion approach and our new adaptive fusion approach occurs when the difference between presumed analogous interactions is large despite the fusion penalty.
-We assume that cases where the method is unable to reconcile the likelihood and the fusion constraint derived from orthology (in the case of multi-species fusion) or identity (in the case of multi-platform fusion) correspond to cases with evidence of divergent TF-to-target-gene interactions.
-Practically, this inability to reconcile likelihood and conservation hypothesis manifests as large differences in model weights across species or platforms. 
-We account for this possibility in our adaptive fusion model with a relaxation of the fusion penalty in cases with extreme model weight divergence. 
+This condition means that no network exists which simultaneously provides an adequate explanation of expression data and which conforms to our prior on similarity.
+In this case, we assume that the constrained interactions have diverged and relax the fusion constraint between them (corresponding to the saturating regime of the penalty).
+
 The resulting cost function is non-convex and difficult to optimize \cite{fan2001variable}; however, we can approximate its solution and obtain deeper insight into functional similarity than is available through strict orthology enforcement or the comparison of separately learned networks.
 
-Although the fusion constraints we employ can be described as arising from orthology - which links genes - it is important to note that the constraints themselves link individual regulatory interactions.
-This finer level of representational granularity is critical to the functioning of adaptive fusion, and means the method can accommodate any form of prior on expected similarity between regulatory interactions, even priors that cannot be decomposed into gene to gene mappings.
-We develop two algorithms for solving efficiently multi-output least-squares regression problems with pairwise L2 fusion penalties on entries of the coefficient matrix. 
+We develop two algorithms for solving efficiently multi-output least-squares regression problems with pairwise L2 fusion penalties on entries of the coefficient matrix, and discuss conditions under which each is suitable.
 We also introduce - in the form of adaptive fusion - the idea of a saturating penalty function on fusion constraints, and estimate the solution to the resulting optimization problem through iterative application of the fused L2 algorithm.
 
-We test the ability of fused L2 and adaptive fusion to improve network recovery on both synthetic data and by comparing related networks in the bacteria species \textit{Bacillus subtilis} and \textit{Bacillus anthracis}.
-This shows the applicability of our method in combining different datasets and leveraging similarity across organisms as well as within a network in order to improve network inference.
-We explore the circumstances under which each approach is optimal, and evaluate the robustness of adaptive fusion to incorrect orthology, simulating the biologically relevant cases of neo- and sub-functionalizations.
+We test the ability of fused L2 and adaptive fusion to improve network recovery on both synthetic data and on real data, by fitting  networks from two strains of tthe bacteria species \textit{Bacillus subtilis} and along with the distantly related \textit{Bacillus anthracis}.
+We then explore the applications of adaptive fusion by demonstrating the technique's ability to identify incorrect orthology information in biological data, simulating the biologically relevant cases of neo- and sub-functionalizations.
 
@@
 \subsection{Statistical approach and background}
-We consider prediction and coefficient estimation problems with $N$ observations of $M$ dependent variables $y_{1,1}, y_{2,1}, ...y_{N,1}, y_{N,2},..., y_{N, M}$ and $p$ features $x_{i,j}$, $i=1,2,...,N, j=1,2,...,p$.
+\subsubsection{Fused L2}
+We consider prediction and coefficient estimation problems with $N$ observations of $M$ dependent variables $y_{1,1}, y_{2,1}, ...y_{N,1},..., y_{N,2},..., y_{N, M}$ and $p$ features $x_{i,j}$, $i=1,2,...,N, j=1,2,...,p$.
 We begin with a standard linear regression model:
@@
 More complicated penalties have been developed to represent specific expected or desireable structure in a regression model's coefficients. 
-For example, Land and Friedman \cite{citeulike3780356} proposed a fusion penalty which encourages smoothness of the estimated parameter vector.
+For example, Land and Friedman \cite{citeulike3780356} proposed a fusion penalty (or a penalty on the differences between certain interaction weights) which encourages smoothness of the estimated parameter vector.
 Previous approaches have used  fusion penalties to draw statistical strength across multiple regression tasks \cite{kim_tree-guided_2012,Land1997,Chen2010,Petry2011,Hebiri2011}.
@@
 
-In many cases, however, there is some uncertainty about the relationships that should be enforced. 
-Sohn et al. attempt to simultaneously learn the regression coefficients and the output structure \cite{sohn_joint_2012}. 
-We develop a similar approach, by applying a penalty function bounded by a constant to produce unbiased estimators for large coefficients, combined with an L2 penalty, similar to SCAD-L2 \cite{Zeng2012}.
+\subsubsection{Solving fused L2 problems using augmented matrices}
+We begin with the problem of constructing a design matrix to map our problem to that of solving a fused L2 regression problem with a single response variable. 
+We then go on to show that, although the vectorized solution involves solving an impractically large system of equations, under typical biological conditions the structure of constraints allow the problem to be broken up into many smaller subproblems.
+Key to this approach is the observation that ridge constraints can be incorporated into a least-squares regression problem by appending a scaled identity matrix to the design matrix, and a corresponding number of zeros to the response vector.
+Similarly, a fusion constraint $\lambda_S (\beta_{i} - \beta{j})^2$ can be incorporated into a least-squares regression problem by appending a row containing $\sqrt{\lambda_S}$ in the $i$th position, $-\sqrt{\lambda_S}$ in the $j$th position, and $0$s elsewhere to the design matrix, and zero to the response vector.
+In order to convert an optimization over multiple response variables and multiple sources into an optimization with a single source and response variable, we vectorize as follows: we construct a new design matrix by diagonally concatenating design matrices from relevant regression problems, and create a new response vector by concatenation of corresponding response vectors.
+
+This is equivalent to the original problem due to the block structure of matrix multiplication. 
+In an ordinary regression problem each response variable can be solved independently, and vectorization is unnecessary. 
+However, in fused regression, we append additional rows to the design matrix that link entries of the interaction weight matrix associated with different response variables (figure \ref{cartoon}). 
+As a result, these linked response variables must be solved simultaneously through vectorization. 
+Two response variables are linked by a fusion constraint if any of the regulatory weights affecting those genes are linked by a fusion constraint. 
+Two response variables must be solved simultaneously if there is any chain of linked response variables connecting them. 
+However, every other response variable can be solved separately. 
+In biological terms, the regulators of two genes (whether in the same species, or different species) must be solved together if there is a fusion constraint linking those genes' regulators, or if there is a chain of such constraints.
+If the networks for a large number of genes are solved simultaneously, the system of equations can quickly become intractable.
+
+In order to avoid this difficulty, we use depth-first search to identify linked columns of each TF expression matrix, then form design and response matrices through vectorization. 
+We can then incorporate fusion constraints as in the case of single-source single response-variable fused regression. 
+In most cases, we have found the direct solution using augmented matrices to be adequate (possible due to the sparse structure of orthology links; only a small number of genes must be solved at once).
+In the general case, the size of the design matrix is proportional to the number of response variables that must be solved simultaneously. 
+Because the scaling of this algorithm has a complicated dependence on the constraint structure used, a general description of its runtime is difficult. 
+However, in the case of multi-species network inference with one-to-one orthology, the network associated with each pair of orthologous genes requires solving a linear system with approximately twice as many observations and unknowns as the single species case.
+Linear systems of this size can be solved quickly using standard techniques, and runtime using our bacterial datasets clocks in around thirty minutes. 
+When the size of the groups of genes linked by fusion constraints becomes large (when organisms have a number of many-to-many orthologous blocks), however, the augmented design matrix approach becomes slower and we discuss further optimisations to this scheme below to enable scaling to these regimes.
+
+\subsubsection{Solving fused L2 problems using iterative solver}
+
+To address scaling limitations when many-to-many fusion constraint blocks occur, we developed an iterative solver that uses coordinate-wise descent to solve for solutions corresponding to a sequence of values of fusion penalty weights.
+As our fused L2 method uses a convex and differentiable penalty function, this approach converges to a global minimizer. 
+Although less efficient than the augmented design matrix approach we developed for cases where fusion constraints are primarily one-to-one or few-to-few, the iterative solver has the advantage of computing a solution path for $\lambda_S$ and scaling well across a wider range of biological applications.
+
+On each iteration $t$ the iterative solver computes
 
 \begin{equation}
-\argmin_{\beta} \displaystyle\sum \vert \vert X\beta - Y \vert \vert ^2 + \lambda_R \vert \vert \beta \vert \vert ^2 + \displaystyle \lambda_S  \displaystyle \sum_{\beta_{g,k} \approx \beta_{h,l}} p_{\lambda, a} (\beta_{g,k} - \beta_{h,l})
+\argmin_{\beta^S} \displaystyle\sum_{S \in \{1, 2\}} ||X^{S}\beta^{S}(t) - Y^{S}||^2 + \lambda_R||\beta^{S}(t)||^2 + \lambda_S\displaystyle \sum_{\substack{(g,h) \in orth,\\
+ (h,l) \in orth}} ||\beta^{S}_{g,k}(t) - \beta_{h,l}^{S}(t-1)||^2
 \end{equation}
+Note that this is almost identical to equation 5, but now the network $\beta$ is a function of the iteration number $t$.
+On each step, we compute $\beta$s that minimize a penalized cost function where the fusion penalties encourage similarity between a parameter and its fused-to parameter from the previous iteration's solution.
+This process is iterated until the estimated $\beta$s converge. 
+Because each iteration reduces the error between $\beta(t)$ and $\beta(t-1)$, and because $\beta(t) = \beta(t-1)$ is the globally optimal solution, this process must eventually converge to the same network as equation \ref{eqn:main}.
+Although we have not produced bounds on the convergence rate, which also depends on the structure of constraints, in practice a small number of iterations ($\sim$10) are necessary.
 
-where the penalty $p_{\lambda, a}$ has derivative
+\subsubsection{Fusion and regularization path}
+Optimizing over both parameters, $\lambda_R$ and $\lambda_S$, is computationally prohibitive and we opted to test a heuristic where we optimize the two parameters separately. 
+Our procedure first optimized $\lambda_R$ with $\lambda_S=0$, then optimized $\lambda_S$ using this value of $\lambda_R$. 
+This procedure is guaranteed to achieve the best unfused solution in the case when $\lambda_S$ is constrained at 0. 
+As a result, any performance gains of fused regression are a lower bound on the highest achievable performance gains. 
+To optimize $\lambda_R$ we use cyclical coordinate descent algorithms from the 'glmnet' package \cite{friedman_regularization_2010} to compute a ridge regularization path. 
+We use cross validation to select the optimal $\lambda_R$ parameter from this path, selecting the $\lambda_R$ which minimizes the average error of prediction on a leave out set across cross validation folds.
+Following selection of $\lambda_R$, we search for optimal $\lambda_S$ by computing the solution path from the iterative solver (using the sequence of successive model weights) again using cross validation to select the optimal parameter.
+Note that both parameters are chosen without reference to the gold standard, which is used in a separate evaluation of network quality.
 
-\begin{equation}
-p'_{\lambda,a}(\theta) = \left\{
-    \begin{array}{lr}
-    \lambda\theta & \text{if } \theta \leq \lambda\\
-    \text{max}(\lambda(2a-\theta),0) & \text{if } \theta > a}
-    \end{array}
-    \right.
-\end{equation}
-
- This approach allows us to simultaneously learn the regression coefficients and evaluate the validity of our prior information (this model relaxes the fusion penalty when model components are irreconcilably different).
-
-\subsection{Application}
-Although our approach is generalizable to a wide variety of multi-source network inference problems, we begin with the concrete example of network inference in two related species. 
-Our approach to multi-species network inference is based on the hypothesis that gene regulation in related species is governed by similar but not necessarily identical gene regulatory networks, due to conservation of function through evolution.
-We represent conservation of network function by introducing constraints into the objective function for network inference that penalize differences between the weights of regulatory interactions believed to be conserved.
-These constraints favor the generation of similar networks for related species, and in the generally under-constrained regime of network inference can improve the accuracy of network recovery. 
-We then go on to introduce a method to test the assumption of conserved network structure, and to relax the associated constraints on pairs of interactions for which the data does not support conservation (where conservation of a regulatory edge is implied by like model weights across data-sets).
-
-\subsection{Approach overview}
-\begin{algorithm}
-	\caption{Network inference using fused regression}\label{euclid}
-	\begin{algorithmic}
-%\Procedure{MyProcedure}{}
-\State load expression data
-\State load orthology
-\State create priors and fusion constraints
-\State partition gold standard into training and leave-out
-\State generate TFA matrices using gold standard training set
-\State set $a$ if using adaptive fusion
-\For{$k$ in folds}
-	\State partition expression data into training and leave-out set
-	\State $\lambda_R$ parameter selection using training set
-	\State $\lambda_S$ parameter selection using training set
-	\State run fused regression
-	\State return PRC and ROC curves using leaveout gold standard
-	\EndFor
-\State average PRC / ROC curves over folds
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-\subsection{Gene regulatory network}
-
-We model the transcription rate of each gene as a weighted sum of transcription factor expression, and seek to identify the identities and regulatory weights of these TFs. 
-This formulation matches that of the existing \textit{Inferelator} algorithm, which models gene expression with linear differential equations \cite{bonneau_inferelator:_2006-1}.
-Our primary data for learning gene regulatory networks is expression data, consisting of time-series and steady state experiments.
-The rate at which $x_{i}$, the observed mRNA expression of gene $i$, changes, is governed by degradation of existing transcripts with rate $\alpha$ plus a linear combination of transcription factor (TF) expressions.
-\begin{equation}
-\frac{\mathrm d}{\mathrm d t} x_i = -\alpha_{i}x_{i} + \sum \beta_{i,j}x_{j}
-\end{equation}
-where $\beta_{i,j}$ represents the weight of TF $j$ on gene $i$, and $\alpha$ is the decay rate of gene $i$.
-We fix the decay rate $\alpha$ for all genes, and set it assuming a time-constant of 10 minutes \cite{hambraeus_genome-wide_2003, selinger_global_2003}, as in \cite{greenfield_robust_2013}.
-Let $x_i(t)$ be the expression of gene $i$ at time $t$.
-Given time-series data on the expression of gene $i$ at timepoints $t_k$ and $t_{k+1}$, we can approximate the rate of change of $x_i$ as $x_i'(t_k)=\frac{x_i(t_{k+1})-x_i(t_k)}{t_{k+1}-t_k}$. 
-We treat steady-state data as having a derivative of zero. 
-This gives us, for each gene $i$ and time $t_{k}$ an equation
-
-\begin{equation}
-\frac{x_i(t_{k+1})-x_i(t_k)}{t_{k+1}-t_k} + \alpha_{i}x_{i}(t_k)= \sum \beta_{i,j}x_{j}(t_k)
-\end{equation}
-where $j \neq i$ 
-for time series and 
-\begin{equation}
-\alpha_{i}x_{i}(t_k) = \sum \beta_{i,j}x_{j}(t_k)
-\end{equation}
-for steady state.
-
-\noindent We can summarize these equations in matrix form as
-\begin{equation}
-Y = X \beta 
-\end{equation}
-where $Y$ is the gene expression matrix, $X$ is the TF expression matrix, and $\beta$ is the regulatory weights we are interested in learning.
-We are interested in learning $\beta$, the matrix representation of the gene regulatory network, where the weight in a given position represents the regulatory weight of a TF on a gene. 
-Positive weights represent activation, negative weights represent repression, and 0 weights represent the absence of an interaction. The matrix $\beta$ can be solved using linear regression. 
-Because there are typically far fewer conditions than possible regressors (TFs), we introduce a ridge regularization constraint with weight $\lambda_R$ and solve
-\begin{equation}
-\argmin_\beta\vert \vert X\beta - Y_2 \vert \vert ^2 + \lambda_R \vert \vert \beta \vert \vert ^2
-\end{equation}
-This is similar to the formulation used in the \textit{Inferelator} algorithm, which we extend to the case of simultaneously inferring multiple networks.
-
-Transcription factor expression is not always the best predictor of its gene targets' expression, so  previous network inference methods attempt to estimate transcription factor activities prior to network inference.
-When there exists a set of prior known interactions, we are able to estimate transcription factor activity (TFA) using network component analysis \cite{liao2003network}, as in \cite{arrieta-ortiz_experimentally_2015, fu_reconstructing_2011}, and use TFA as explanatory variables instead of transcription factor expression.
-
-\subsection{Fused gene regulatory networks}
-
-Information about the partially conserved structure of gene regulation is introduced through the incorporation of constraints into the above regression formulation.
-These constraints penalize differences between interaction weights in the networks of multiple species that are expected to be similar based on prior biological knowledge.
-We can then solve the penalized regression problems simultaneously, in order to obtain a gene regulatory network (GRN) for each species. 
-Consider the case of organisms $A$ and $B$, governed by GRNs $\beta^A$ and $\beta^B$ (the following approach applies equally well to more than two species but for simplicity we continue with the case of two species).
-When TF $g^A$ in organism $A$ and TF $h^B$ in organism $B$ are orthologs, and gene $k^A$ and $l^B$ are orthologs, then we expect that the $g^A \rightarrow k^A$ interaction weight should be similar to the $h^B \rightarrow l^B$ interaction weight, and we introduce a fusion constraint between these analogous interactions.
-In terms of the above regression formulation, we expect that $\beta^A_{g,k} \approx \beta^B_{h,l}$, and include a penalty term $\lambda_Sp(\beta^A_{g,k} - \beta^B_{h,l})$ in the quantity being minimized in order to encourage similarity.
-The function $p(x)$ controls the shape of the relationship between weight dissimilarity and penalty, while scalar $\lambda_S$ controls the overall scaling of the penalization of differences between fused coefficients.
-$\lambda_S$ controls the tradeoff between fitting the expression data-sets individually and producing a set of networks that conform to evolutionary prior knowledge.
-This gives us the final equation to be minimized:
-
-\begin{equation}
-\argmin_{\beta^S} \displaystyle\sum_{S \in \{1, 2\}} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \lambda_S \displaystyle \sum_{\substack{(g,h) \in orth,\\
- (k,l) \in orth}}p(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
-\end{equation}
-
-where the second sum is over pairs of interactions with fusion constraints. 
-In the fused L2 algorithm presented here, the penalty function is equal to the L2 norm of the difference in regulatory weight of fused coefficients, $p(x)=x^2$. 
-Every component of the objective function is an L2 norm and thus the problem is convex and can in fact be solved through linear regression with an augmented design matrix.
-
-As an example, consider the case where there is a one-to-one orthology between the species being considered (ie different cell-lines of the same organism). 
-The choice of $\lambda_S$ allows one to interpolate between fitting each network independently ($\lambda_S=0$) and pooling data together as if it came from one source ($\lambda_S=\inf$). 
-In addition to performing well between these extremes, our method allows pooling of data even when there is incomplete orthology. 
-By introducing constraints on the similarity of individual interactions, rather than on the networks as a whole \cite{parikh2011treegl}, we can pool some information across species even when a small fraction of genes have orthologs.
-
-\subsection{Adaptive fusion}
+\subsubsection{Adaptive fusion}
 Fusion constraints penalize dissimilarity between interactions thought to be analogous based on \textit{a priori} knowledge. 
@@
 
-\subsection{Solving fused L2 problems using augmented matrices}
-We begin with the problem of constructing a design matrix to map our problem to that of solving a fused L2 regression problem with a single response variable. 
-We then go on to show that, although the vectorized solution involves solving an impractically large system of equations, under typical biological conditions the structure of constraints allow the problem to be broken up into many smaller subproblems.
-Key to this approach is the observation that ridge constraints can be incorporated into a least-squares regression problem by appending a scaled identity matrix to the design matrix, and a corresponding number of zeros to the response vector.
-Similarly, a fusion constraint $\lambda_S (\beta_{i} - \beta{j})^2$ can be incorporated into a least-squares regression problem by appending a row containing $\sqrt{\lambda_S}$ in the $i$th position, $-\sqrt{\lambda_S}$ in the $j$th position, and $0$s elsewhere to the design matrix, and zero to the response vector.
-In order to convert an optimization over multiple response variables and multiple sources into an optimization with a single source and response variable, we vectorize as follows: we construct a new design matrix by diagonally concatenating design matrices from relevant regression problems, and create a new response vector by concatenation of corresponding response vectors.
+\subsection{Biological application}
+Although our approach is generalizable to a wide variety of multi-source network inference problems, we begin with the concrete example of network inference in two related species. 
+Our approach to multi-species network inference is based on the hypothesis that gene regulation in related species is governed by similar but not necessarily identical gene regulatory networks, due to conservation of function through evolution.
+We represent conservation of network function by introducing constraints into the objective function for network inference that penalize differences between the weights of regulatory interactions believed to be conserved.
+These constraints favor the generation of similar networks for related species, and in the generally under-constrained regime of network inference can improve the accuracy of network recovery. 
+We then go on to introduce a method to test the assumption of conserved network structure, and to relax the associated constraints on pairs of interactions for which the data does not support conservation.
+Finally, we demonstrate the flexibility of the method by using fusion constraints based on operon membership to improve network inference performance.
 
-This is equivalent to the original problem due to the block structure of matrix multiplication. 
-In an ordinary regression problem each response variable can be solved independently, and vectorization is unnecessary. 
-However, in fused regression, we append additional rows to the design matrix that link entries of the interaction weight matrix associated with different response variables (figure \ref{cartoon}). 
-As a result, these linked response variables must be solved simultaneously through vectorization. 
-Two response variables are linked by a fusion constraint if any of the regulatory weights affecting those genes are linked by a fusion constraint. 
-Two response variables must be solved simultaneously if there is any chain of linked response variables connecting them. 
-However, every other response variable can be solved separately. 
-In biological terms, the regulators of two genes (whether in the same species, or different species) must be solved together if there is a fusion constraint linking those genes' regulators, or if there is a chain of such constraints.
-If the networks for a large number of genes are solved simultaneously, the system of equations can quickly become intractable.
+\subsubsection{Approach overview}
+\begin{algorithm}
+	\caption{Network inference using fused regression}\label{euclid}
+	\begin{algorithmic}
+%\Procedure{MyProcedure}{}
+\State load expression data
+\State load orthology
+\State create priors and fusion constraints
+\State partition gold standard into training and leave-out
+\State generate TFA matrices using gold standard training set
+\State set $a$ if using adaptive fusion
+\For{$k$ in folds}
+	\State partition expression data into training and leave-out set
+	\State $\lambda_R$ parameter selection using training set
+	\State $\lambda_S$ parameter selection using training set
+	\State run fused regression
+	\State return PRC and ROC curves using leaveout gold standard
+	\EndFor
+\State average PRC / ROC curves over folds
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
 
-In order to avoid this difficulty, we use depth-first search to identify linked columns of each TF expression matrix, then form design and response matrices through vectorization. 
-We can then incorporate fusion constraints as in the case of single-source single response-variable fused regression. 
-In most cases, we have found the direct solution using augmented matrices to be adequate (possible due to the sparse structure of orthology links; only a small number of genes must be solved at once).
-In the general case, the size of the design matrix is proportional to the number of response variables that must be solved simultaneously. 
-Because the scaling of this algorithm has a complicated dependence on the constraint structure used, a general description of its runtime is difficult. 
-However, in the case of multi-species network inference with one-to-one orthology, the network associated with each pair of orthologous genes requires solving a linear system with approximately twice as many observations and unknowns as the single species case.
-Linear systems of this size can be solved quickly using standard techniques, and runtime using our bacterial datasets clocks in around thirty minutes. 
-When the size of the groups of genes linked by fusion constraints becomes large (when organisms have a number of many-to-many orthologous blocks), however, the augmented design matrix approach becomes slower and we discuss further optimisations to this scheme below to enable scaling to these regimes.
+\subsubsection{Gene regulatory network}
 
-\subsection{Solving fused L2 problems using iterative solver}
-
-To address scaling limitations when many-to-many fusion constraint blocks occur, we developed an iterative solver that uses coordinate-wise descent to solve for solutions corresponding to a sequence of values of fusion penalty weights.
-As our fused L2 method uses a convex and differentiable penalty function, this approach converges to a global minimizer. 
-Although less efficient than the augmented design matrix approach we developed for cases where fusion constraints are primarily one-to-one or few-to-few, the iterative solver has the advantage of computing a solution path for $\lambda_S$ and scaling well across a wider range of biological applications.
-
-On each iteration $t$ the iterative solver computes
+We model the transcription rate of each gene as a weighted sum of transcription factor expression, and seek to identify the identities and regulatory weights of these TFs. 
+This formulation matches that of the existing \textit{Inferelator} algorithm, which models the linear dynamics of gene expression \cite{bonneau_inferelator:_2006-1}.
+Our primary data for learning gene regulatory networks is expression data, consisting of both time-series and steady state experiments.
+The rate at which $x_{i}$, the observed mRNA expression of gene $i$, changes, is governed by degradation of existing transcripts with rate $\alpha$ plus a linear combination of transcription factor (TF) expressions.
+\begin{equation}
+\frac{\mathrm d}{\mathrm d t} x_i = -\alpha_{i}x_{i} + \sum \beta_{i,j}x_{j}
+\end{equation}
+where $\beta_{i,j}$ represents the weight of TF $j$ on gene $i$, and $\alpha$ is the decay rate of gene $i$.
+We fix the decay rate $\alpha$ for all genes, and set it assuming a time-constant of 10 minutes \cite{hambraeus_genome-wide_2003, selinger_global_2003}, as in \cite{greenfield_robust_2013}.
+Let $x_i(t)$ be the expression of gene $i$ at time $t$.
+Given time-series data on the expression of gene $i$ at timepoints $t_k$ and $t_{k+1}$, we can approximate the rate of change of $x_i$ as $x_i'(t_k)=\frac{x_i(t_{k+1})-x_i(t_k)}{t_{k+1}-t_k}$. 
+We treat steady-state data as having a derivative of zero. 
+This gives us, for each gene $i$ and time $t_{k}$ an equation
 
 \begin{equation}
-\argmin_{\beta^S} \displaystyle\sum_{S \in \{1, 2\}} ||X^{S}\beta^{S}(t) - Y^{S}||^2 + \lambda_R||\beta^{S}(t)||^2 + \lambda_S\displaystyle \sum_{\substack{(g,h) \in orth,\\
- (h,l) \in orth}} ||\beta^{S}_{g,k}(t) - \beta_{h,l}^{S}(t-1)||^2
+\frac{x_i(t_{k+1})-x_i(t_k)}{t_{k+1}-t_k} + \alpha_{i}x_{i}(t_k)= \sum \beta_{i,j}x_{j}(t_k)
 \end{equation}
-Note that this is almost identical to equation 5, but now the network $\beta$ is a function of the iteration number $t$.
-On each step, we compute $\beta$s that minimize a penalized cost function where the fusion penalties encourage similarity between a parameter and its fused-to parameter from the previous iteration's solution.
-This process is iterated until the estimated $\beta$s converge. 
-Because each iteration reduces the error between $\beta(t)$ and $\beta(t-1)$, and because $\beta(t) = \beta(t-1)$ is the globally optimal solution, this process must eventually converge to the same network as equation \ref{eqn:main}.
-Although we have not produced bounds on the convergence rate, which also depends on the structure of constraints, in practice a small number of iterations ($\sim$10) are necessary.
+where $j \neq i$ 
+for time series and 
+\begin{equation}
+\alpha_{i}x_{i}(t_k) = \sum \beta_{i,j}x_{j}(t_k)
+\end{equation}
+for steady state.
 
-\subsection{Fusion and regularization path}
-Optimizing over both parameters, $\lambda_R$ and $\lambda_S$, is computationally prohibitive and we opted to test a heuristic where we optimize the two parameters separately. 
-Our procedure first optimized $\lambda_R$ with $\lambda_S=0$, then optimized $\lambda_S$ using this value of $\lambda_R$. 
-This procedure is guaranteed to achieve the best unfused solution in the case when $\lambda_S$ is constrained at 0. 
-As a result, any performance gains of fused regression are a lower bound on the highest achievable performance gains. 
-To optimize $\lambda_R$ we use cyclical coordinate descent algorithms from the 'glmnet' package \cite{friedman_regularization_2010} to compute a ridge regularization path. 
-We use cross validation to select the optimal $\lambda_R$ parameter from this path, selecting the $\lambda_R$ which minimizes the average error of prediction on a leave out set across cross validation folds.
-Following selection of $\lambda_R$, we search for optimal $\lambda_S$ by computing the solution path from the iterative solver (using the sequence of successive model weights) again using cross validation to select the optimal parameter.
-Note that both parameters are chosen without reference to the gold standard, which is used in a separate evaluation of network quality.
+\noindent We can summarize these equations in matrix form as
+\begin{equation}
+Y = X \beta 
+\end{equation}
+where $Y$ is the gene expression matrix, $X$ is the TF expression matrix, and $\beta$ is the regulatory weights we are interested in learning.
+We are interested in learning $\beta$, the matrix representation of the gene regulatory network, where the weight in a given position represents the regulatory weight of a TF on a gene. 
+Positive weights represent activation, negative weights represent repression, and 0 weights represent the absence of an interaction. The matrix $\beta$ can be solved using linear regression. 
+Because there are typically far fewer conditions than possible regressors (TFs), we introduce a ridge regularization constraint with weight $\lambda_R$ and solve
+\begin{equation}
+\argmin_\beta\vert \vert X\beta - Y_2 \vert \vert ^2 + \lambda_R \vert \vert \beta \vert \vert ^2
+\end{equation}
+This is similar to the formulation used in the \textit{Inferelator} algorithm, except that we use an $L2$ penalty in place of the more standard $L1$ penalty.
+Although $L1$ penalties have the theoretical appeal of sparsity - which matches our assumption of sparsity in biological networks - we are primarily interested here in ranking interactions, rather than recovering a specific network.
+Moreover, $L2$ penalized regression followed by thresholding can be competitive with $L1$ penalized regression for support recovery \cite{}.
+Although our approach can be easily modified to use an $L1$ penalty, we did not observe performance gains when doing so, and the version presented here is in some ways conceptually simpler (with the combined penalty function being equivalent to a multivariate Gaussian prior).
 
-\subsection{Simulated data}
+Because a transcription factor's raw abundance is not always a good predictor of its influence on its gene targets' expression, previous network inference methods have attempted to estimate transcription factor activities prior to network inference.
+When there exists a set of prior known interactions, we are able to estimate transcription factor activity (TFA) using network component analysis \cite{liao2003network}, as in \cite{arrieta-ortiz_experimentally_2015, fu_reconstructing_2011}, and use TFA as explanatory variables instead of transcription factor expression.
+
+\subsubsection{Fused gene regulatory networks}
+
+Information about the partially conserved structure of gene regulation is introduced through the incorporation of constraints into the above regression formulation.
+These constraints penalize differences between interaction weights in the networks of multiple species that are expected to be similar based on prior biological knowledge.
+We can then solve the penalized regression problems simultaneously, in order to obtain a gene regulatory network (GRN) for each species. 
+Consider the case of organisms $A$ and $B$, governed by GRNs $\beta^A$ and $\beta^B$ (the following approach applies equally well to more than two species but for simplicity we continue with the case of two species).
+When TF $g^A$ in organism $A$ and TF $h^B$ in organism $B$ are orthologs, and gene $k^A$ and $l^B$ are orthologs, then we expect that the $g^A \rightarrow k^A$ interaction weight should be similar to the $h^B \rightarrow l^B$ interaction weight, and we introduce a fusion constraint between these analogous interactions.
+In terms of the above regression formulation, we expect that $\beta^A_{g,k} \approx \beta^B_{h,l}$, and include a penalty term $\lambda_Sp(\beta^A_{g,k} - \beta^B_{h,l})$ in the quantity being minimized in order to encourage similarity.
+The function $p(x)$ controls the shape of the relationship between weight dissimilarity and penalty, while scalar $\lambda_S$ controls the overall scaling of the penalization of differences between fused coefficients.
+$\lambda_S$ controls the tradeoff between fitting the expression data-sets individually and producing a set of networks that conform to evolutionary prior knowledge.
+This gives us the final equation to be minimized:
+
+\begin{equation}
+\argmin_{\beta^S} \displaystyle\sum_{S \in \{1, 2\}} \vert \vert X^S\beta^S - Y^S \vert \vert ^2 + \lambda_R \vert \vert \beta^S \vert \vert ^2 + \lambda_S \displaystyle \sum_{\substack{(g,h) \in orth,\\
+ (k,l) \in orth}}p(\beta^{S_1}_{g,k} - \beta^{S_2}_{h,l})
+\end{equation}
+
+where the second sum is over pairs of interactions with fusion constraints. 
+In the fused L2 algorithm presented here, the penalty function is equal to the L2 norm of the difference in regulatory weight of fused coefficients, $p(x)=x^2$. 
+Every component of the objective function is an L2 norm and thus the problem is convex and can in fact be solved through linear regression with an augmented design matrix.
+
+As an example, consider the case where there is a one-to-one orthology between the species being considered (ie different cell-lines of the same organism). 
+The choice of $\lambda_S$ allows one to interpolate between fitting each network independently ($\lambda_S=0$) and pooling data together as if it came from one source ($\lambda_S=\inf$). 
+In addition to performing well between these extremes, our method allows pooling of data even when there is incomplete orthology. 
+By introducing constraints on the similarity of individual interactions, rather than on the networks as a whole \cite{parikh2011treegl}, we can pool some information across species even when a small fraction of genes have orthologs.
+
+\subsubsection{Simulated data}
 We generate simulated data to evaluate the ability of our fused L2 approach to learn the true network and to show that sharing information between similar but not identical data sources results in more accurate network recovery.
@@
 
-\subsection{Ranking regulatory hypotheses}
+\subsubsection{Ranking regulatory hypotheses}
 In previous work, betas were rescaled as to form a matrix of confidence scores $S$ as follows
@@
 
-\subsection{\textit{B. subtilis} and \textit{B. anthracis} data and orthology}
+\subsubsection{\textit{B. subtilis} and \textit{B. anthracis} data and orthology}
 We used a dataset collected for PY79, a derivative of strain 168, available on GEO with accession number GSE67023, and a dataset using BSB1, another derivative of strain 168, available at GEO with accession number GSE27219.

### 2016-07-26 paper_revisions.tex
zack writing revisions done. TODO detailed model comparison section

@@
 Moreover, we are primarily interested in using this saturating penalty as a way of testing the hypothesis that conservation in GRNs can be predicted based off of known similarities between genes. 
-Therefore, we propose a user-defined $a$, where this parameter is set using the distribution of differences between fused weights from independently fit networks.
-The choice of which value in this distribution to use for $a$ represents the working hypothesis for the fraction of fused interactions which should be unfused.
+Therefore, we propose setting $a$ based on the distribution of differences between fused weights from independently fit networks.
+These networks can be fit without fusion, prior to the adaptive fusion procedure.
+The choice of which percentile of this distribution to use for $a$ represents the working hypothesis for the fraction of fused interactions which should be unfused.
 
@@
 We treat steady-state data as having a derivative of zero.
-This gives us, for each gene $i$ and time $t_{k}$ an equation
+
+For timeseries conditions this gives us, for each gene $i$ and time $t_{k}$ an equation
 
@@
 where $j \neq i$
-for time series and 
+
+And for steady state conditions
 \begin{equation}
@@
 \end{equation}
-for steady state.
-
-\noindent We can summarize these equations in matrix form as
+Both kinds of conditions can be included in the set of equations to be solved. We can summarize these equations in matrix form as
 \begin{equation}
@@
 We can then solve the penalized regression problems simultaneously, in order to obtain a gene regulatory network (GRN) for each species.
-Consider the case of organisms $A$ and $B$, governed by GRNs $\beta^A$ and $\beta^B$ (the following approach applies equally well to more than two species but for simplicity we continue with the case of two species).
+
+Consider the case of organisms $A$ and $B$, governed by GRNs $\beta^A$ and $\beta^B$ (the following approach applies equally well to more than two species - where the set of constraints is the enumeration of all constraints between pairs of species).
 When TF $g^A$ in organism $A$ and TF $h^B$ in organism $B$ are orthologs, and gene $k^A$ and $l^B$ are orthologs, then we expect that the $g^A \rightarrow k^A$ interaction weight should be similar to the $h^B \rightarrow l^B$ interaction weight, and we introduce a fusion constraint between these analogous interactions.
+This is shown schematically in Figure \ref{fusion}.
 In terms of the above regression formulation, we expect that $\beta^A_{g,k} \approx \beta^B_{h,l}$, and include a penalty term $\lambda_Sp(\beta^A_{g,k} - \beta^B_{h,l})$ in the quantity being minimized in order to encourage similarity.
-The function $p(x)$ controls the shape of the relationship between weight dissimilarity and penalty, while scalar $\lambda_S$ controls the overall scaling of the penalization of differences between fused coefficients.
+
+The function $p(x)$ controls the shape of the penalty function, while scalar $\lambda_S$ controls the overall scaling of the penalization of differences between fused coefficients.
 $\lambda_S$ controls the tradeoff between fitting the expression data-sets individually and producing a set of networks that conform to evolutionary prior knowledge.
@@
 where the second sum is over pairs of interactions with fusion constraints.
-In the fused L2 algorithm presented here, the penalty function is equal to the L2 norm of the difference in regulatory weight of fused coefficients, $p(x)=x^2$. 
-Every component of the objective function is an L2 norm and thus the problem is convex and can in fact be solved through linear regression with an augmented design matrix.
 
-As an example, consider the case where there is a one-to-one orthology between the species being considered (ie different cell-lines of the same organism). 
+As an example of how this formulation allows pooling of data across multiple sources, consider the case where there is a one-to-one orthology between the species being considered (ie different cell-lines of the same organism).
 The choice of $\lambda_S$ allows one to interpolate between fitting each network independently ($\lambda_S=0$) and pooling data together as if it came from one source ($\lambda_S=\inf$).
-In addition to performing well between these extremes, our method allows pooling of data even when there is incomplete orthology. 
-By introducing constraints on the similarity of individual interactions, rather than on the networks as a whole \cite{parikh2011treegl}, we can pool some information across species even when a small fraction of genes have orthologs.
+In the case of cross-species network inference, there is unlikely to be a complete one-to-one orthology.
+However, because we  constraints the similarity of individual interactions, rather than on the networks as a whole \cite{parikh2011treegl}, we can pool some information across species even when a small fraction of genes have orthologs.
 
@@
 
+[KARI - INSERT HOW GOLD STANDARD WAS OBTAINED]
+
 \section{Results}
@@
 For the synthetic data, we generated random pairs of networks in which orthologous genes have similar regulatory interactions, and then sampled gene expression from these networks, which we used to derive learned networks for comparison with the input (true) networks.
-For real data, we computed recovery of a known gold standard in \textit{Bacillus subtilis} \cite{michna_subtiwikidatabase_2014}.
+For real data, we examined recovery of a known gold standard in \textit{Bacillus subtilis} \cite{michna_subtiwikidatabase_2014}.
 
-\subsection{Using fused regression to learn related networks}
-It is known that the accuracy of network inference improves with additional data \cite{bar-joseph_computational_2003}. 
-Using related data for network inference allows us to leverage statistical power from disparate sources, effectively increasing the sample size and boosting the sensitivity and specificity of learned interactions.
+\subsection{Simulations}
+\subsubsection{Using fused regression to learn related networks}
 We created synthetic networks to approximate two related biological processes, then evaluated performance of our fused L2 regression, which learns the networks simultaneously given a prior on the relatedness of interactions.
-We compared recovery of the artificial 10 TFs by 200 genes networks, using fused L2 versus learning networks separately, and varied the amounts of simulated expression data samples made available to the solver.
-When the amount of data from the second species was held constant, increasing the amount of data available for learning the network for the first species resulted in a more accurate network prediction, as expected  (figure \ref{fusion}b).
-When we increased the amount of data from the second species, we obtained performance gains on network one using fused L2 regression, demonstrating our ability to improve network inference on one dataset through incorporation of a related dataset.
+Simulated networks consisted of pairs of related 10 TF by 200 gene networks, with varying numbers of expression data samples made available to the solver.
+In order to compare the relative contribution of data in the species of interest (species 1) and in a related organism (species 2) to network recovery, we varied the number of conditions in each species while evaluating the mean-squared-error between the recovered network for species 1 and its known true network.
+When the amount of data from the related species was held constant, increasing the amount of data available for learning the network for the species of interest resulted in a more accurate network prediction, as expected  (figure \ref{fusion}b).
+Similarly, when we increased the amount of data from the related species, we obtained performance gains on network one using fused L2 regression, demonstrating our ability to improve network inference on one dataset through incorporation of a related dataset.
 
-\subsection{Fused regression improves performance on both the constrained and non-constrained parts of the network}
+\subsubsection{Fused regression improves performance on both the constrained and non-constrained parts of the network}
 Our approach is useful for learning networks from similar sources such as related cell types from the same species, where there exists a one-to-one mapping of genes, as well as datasets where the orthology mapping does not span all genes.
-This can occur when using different technology, eg microarray and RNAseq, where there is incomplete overlap in the genes that each method assays as well as incomplete overlap in the genes expressed in different experimental designs.
+This latter case is typical not only for cross-species network inference, but can occur when using different technology, eg microarray and RNAseq, where there is incomplete overlap in the genes that each method assays as well as incomplete overlap in the genes expressed in different experimental designs.
 When orthology is incomplete we are interested in knowing if performance gains from fused regression are limited to those interactions which have fusion constraints, or if they extend to the entire network.
@@
 We divided networks into those interactions with fusion constraints (the constrained subnetwork) and interactions without fusion constraints (the non-constrained subnetwork).
-We varied the weight on the fusion penalty, $\lambda_S$, and evaluated performance by computing AUPR on the constrained subnetwork, the non-constrained subnetwork, and the whole network (figure \ref{fusion}c).
-Since the conserved subgraphs were similar to each other, we expected performance to improve as the fusion penalty weight increased. 
-We observed this, particularly for the constrained subnetwork. 
-As $\lambda_S$ increased, interactions with fusion constraints were encouraged to be more similar. 
-Interestingly, performance gains were seen even in the portion of the network that was unconstrained by fusion.
+We varied the weight on the fusion penalty, $\lambda_S$, and evaluated performance by computing AUPR on the constrained subnetwork, the non-constrained subnetwork, and the whole network (Figures \ref{fusion}c-f).
+Because the conserved subgraphs were known to be similar to each other, we expected performance to improve as the fusion penalty weight increased (Figures \ref{fusion}c-f, Constrained).
+
+Interestingly, performance gains were seen even in the portion of the network that was unconstrained by fusion (Figures \ref{fusion}c-f, Non-constrained).
 This is because a gene may have some some interactions that are constrained by fusion - regulation by TFs with orthologs - and some interactions that are unconstrained - regulation by TFs without orthologs.
@@
 
-\subsection{Adaptive fusion successfully identifies and unfuses 'neofunctionalized' genes}
+\subsubsection{Adaptive fusion successfully identifies and unfuses 'neofunctionalized' genes}
 Orthology prediction is not a proxy for functional conservation \cite{gabaldon_functional_2013, studer_how_2009, nehrt_testing_2011}. 
@@
 We then compared the recovery of interaction weights which were accurately fused, and recovery of interaction weights which were inaccurately fused due to incorrect orthology information (figure \ref{adaptivefusion}b).
-Because fused L2 heavily penalizes large differences between weights which are predicted to be similar, it is able to retrieve a more accurate network for those interactions with true fusion constraints than by learning networks separately (measured by MSE between the true and inferred interaction weights) .
+Because fused L2 heavily penalizes large differences between weights which are predicted to be similar, it is able to retrieve a more accurate network for those interactions with true fusion constraints than by learning networks separately (measured by MSE between the true and inferred interaction weights).
 In this simulation, however, the gains accomplished through fused regression do not extend to those interactions lacking true fusion constraints, and the error remains similar to learning networks separately.
 When we applied adaptive fusion, we did not observe an improvement in network recovery (relative to fused L2).
-However, we were able to identify fusion constraints reflecting incorrect orthology information that had been provided to the algorithm (figure \ref{adaptivefusion}subsection).
+However, we were able to identify fusion constraints reflecting incorrect orthology information that had been provided to the algorithm (Figure \ref{adaptivefusion} subsection).
+[ZACK - WE NEED TO WRITE MORE ABOUT THIS]
 
-\c{Cross-species network inference using bacterial data}
+\subsection{Fused network inference with bacterial data}
 We used gene-expression data from \textit{Bacillus subtilis} and \textit{B. anthracis} in order to assess performance gains of fused regression on real data. 
@@
 subtilis} network when using fused regression (figure \ref{xspecies-real}a).
-Notably, these performance gains occur mostly at low values of recall (near the top of our prediction ranks, where biologists would presumably focus validation and followup experiments).
+Notably, these performance gains occur mostly at low values of recall.
+That is: the highest confidence part of the network is inferred more accurately, with minimal gains for interactions which are more uncertain.
+Because these interactions are likely to be the focus of followup experiments and validation, gains here are more valuable for prioritizing the order in which interactions are investigated.
 
-\subsection{Testing adaptive fusion using bacterial data}
+\subsubsection{Adaptive fusion using bacterial data}
+[ZACK MORE HERE TOO]
  The goal of adaptive fusion is to unfuse constraints between non-conserved interactions, while leaving intact all other constraints. 
@@
 
-\subsection{Integrating datasets from different platforms using fused regression}
-Although there are many large-scale collaborations which attempt to make protocols as uniform as possible for comparability between datasets generated by different labs \cite{paten_nih_2015,kundaje_integrative_2015} and several methods for removing batch effects \cite{irizarry_exploration_2003,johnson_adjusting_2007}, there still exists technical and biologial variability between many experiments attempting to capture the same or similar experimental conditions esspecially when experiments employ different experimental platforms.
+ \subsubsection{Single species applications}
+
+Although the approach we describe was developed with cross-species network inference in mind, the framework of defining fusion constraints on TF gene interactions expected to be similar is quite general, and can be applied to data from a single organism.
+Despite the many large-scale collaborations which attempt to make protocols as uniform as possible for comparability between datasets generated by different labs \cite{paten_nih_2015,kundaje_integrative_2015} and several methods for removing batch effects \cite{irizarry_exploration_2003,johnson_adjusting_2007}, there still exists technical and biologial variability between many experiments attempting to capture the same or similar experimental conditions esspecially when experiments employ different experimental platforms.
 With the advent of RNAseq, for example, microarray based technologies are no longer the dominant assay for genome-wide expression, but a large body of accumulated legacy data remains useful if it can be integrated with more modern techniques.
+
 Currently, the most widely used approach to combining datasets for network inference is to learn networks from disparate datasets separately, then rank combine the networks as in Marbach et al \cite{marbach_revealing_2010}.
@@
 
-Information about the similarity of TF-gene interactions can also come from knowledge about the promoter region or the structure, for bacteria, of polycistronic transcripts. 
+Our approach can also make use of forms of biological data other than RNA expression.
+Information about the expected similarity of TF-gene interactions can come from knowledge about the promoter region or the structure, for bacteria, of polycistronic transcripts.
 In bacteria, genes within the same operon are typically under the control of the same promoter \cite{lawrence_shared_2002}.
@@
 As in previous studies, we observed a marked improvement in network inference when using transcription factor activity (figure \ref{tfa}).
-We also obtained AUPR improvement when using fused regression on TFA, and showed that our gains from sharing information across datasets using fused regression were preserved and even enhanced by using TFA.
+We also obtained AUPR improvement when using fused regression alongside TFA; gains from sharing information across datasets using fused regression were preserved and even enhanced by using TFA.
 
@@
 subtilis} and \textit{B.
-anthracis}).
+anthracis}), as well as to the problems of combining data from multiple experimental platforms and information about operon structure.
 
-We apply this algorithm to the problem of network inference in two distantly related biological organisms -- \textit{B.
+\textit{B.
 subtilis} and \textit{B.
-anthracis} -- and show that network recovery is improved through the introduction of fusion constraints between pairs of orthologous genes.
+anthracis} are distantly related bacterial species with limited gene orthology [ZACK - HOW DISTANTLY?].
+Nevertheless we show that network recovery in \textit{B.
+subtilis} can be improved through the inclusion of expression data from \textif{B.
+anthracis}.
 Many previous methods for cross-species network inference operate on the conserved subset of orthologous genes \cite{dillman_comparative_2015}. 
@@
 anthracis} genomes do not have clear orthologs (and many orthologs have ambiguous many-to-many groupings).
-Our method, in contrast, can obtain improvements in network inference performance even when the conserved subset of genes is small. 
+Our method, in contrast, can obtain improvements in network inference performance even when the conserved subset of genes is small. [DISCUSS RELATIONSHIP TO OTHER PAPERS HERE]
+
 This approach is particularly interesting in light of the diversity of important model organisms used in modern biology.
 Different model systems provide different advantages and disadvantages for experimental design \cite{stolfi_genetic_2012}, but without a principled mechanism for combining data from multiple sources, it is difficult to fully leverage data obtained from even a slightly different model system.
-We further demonstrate the viability of fused L2 as a method for combining data from multiple experimental platforms, where fusion is between each identical regulator-gene pair. 
+
 Because the algorithm we developed can accomodate constraints between arbitrary pairs of regulatory interactions, any biological prior representing information about expected regulatory similarity can be represented, even if the prior provides no information about the magnitude or direction of regulation.
@@
 In this application, fusion reflects the assumption that genes in the same operons have similar regulators \cite{lawrence_shared_2002}.
-The ability to incorporate multiple data sets describing related processes, as well as multiple data types, in a principled manner, helps us take advantage of the breadth of experimentation in biology to better learn the structure of gene regulation.
-We illustrate this by combining two different \textit{B.
-subtilis} datasets and show that fused L2 is an improvement over current approaches to combining data \cite{marbach_revealing_2010} because of our ability to exploit the statistical power that our expanded datasets afford us.
 
@@
 Existing approaches to the genome-wide testing of this assumption learn regulatory networks separately, then compare to identify conservation \cite{aytes_cross-species_2014}.
-Because network inference is typically underconstrained, fitting a network that describes a particular set of experimental observations involves sampling a single network from a large set of networks that fit the data equally (or almost equally) as well.
+Because network inference is typically underconstrained, fitting a network that describes a particular set of experimental observations amounts to sampling a single network from a large set of networks that fit the data equally (or almost equally) as well.
 As a result, the existence of a difference between corresponding regulatory interactions in a pair of experimentally derived networks is weak evidence that a difference truly does exist.
@@
 
-We have described a method -- adaptive fusion -- that attempts to learn which fusion constraints should be relaxed while the network is being learned. 
+Although fused regression allows more accurate identification of pairs of non-conserved interactions, the weights obtained for these interactions will be biased towards one another.
+We have described a method -- adaptive fusion -- that attempts to address this bias by simultaneously learning the networks and constraint weights.
 This method is based on minimizing a saturating penalty function on fusion constraints, similar to a class of penalties that have been developed to minimise bias in regularized regression \cite{fan2001variable,zhang2010nearly}.
-The result of adaptive fusion is both a network and a new set of fusion constraints, describing the learned fusion weights (including which fusion constraints have been relaxed). 
-For the multiple species case, relaxation of fusion constraints represents orthologs which do not share similar interactions presumably due to evolution of regulatory circuitry \cite{kellis_proof_2004}.
+The result of adaptive fusion is both a network and a new set of fusion constraints, the weights of which can be interpreted as describing conservation structure across the networks.For the multiple species case, relaxation of fusion constraints represents orthologs which do not share similar interactions presumably due to evolution of regulatory circuitry \cite{kellis_proof_2004}.
 %not sure what radid is
@@
 
+\pagebreak
+\begin{figure}[h!]
+    \centering
+    \includegraphics[scale=.5,trim=0mm 0mm 0mm 0mm,clip]{party3.eps}
+    \caption{\label{cartoon} Schematic representation of design matrix construction. Here, the circles and hexagons correspond to different species. 
+    Bidirectional arrows represent orthology information and dotted arrows represent putative interactions between TFs and genes. 
+    Rectangles represent matrices; because weights can be solved independently unless there exists fusion constraints between them, we identify related weights and construct matrix for solving. }
+\end{figure}
+\pagebreak
+
 \begin{figure}
@@
   \textbf{C-F} Show the varying effects of fusion on simulated networks with different levels of conservation.
-We generate a series of networks with 20 TFs by 200 genes in two species, (50\% sparsity) while varying the fraction of gene orthologies in the simulated networks.
+We generate a series of networks with 20 TFs by 200 genes in two species, (50\% sparsity) while varying the fraction of gene orthologies in the simulated networks (Amount fused, top).
 For each network, we evaluated AUPR on one of the species for: all interactions (blue line), interactions with fusion constraints (green line), and interactions without fusion constraints (orange line).
@@
 \end{figure}
+\pagebreak
 
-\begin{figure}
-\begin{center}
-  \includegraphics[scale=0.45]{fig2.pdf}
-  \caption{\label{syntheticgrid} This figure demonstrates the interaction between fusion weight ($\lambda_S$), the degree of orthology coverage of networks, and the degree of similarity between fused interactions.
+%\begin{figure}
+%\begin{center}
+%  \includegraphics[scale=0.45]{fig2.pdf}
+%  \caption{\label{syntheticgrid} This figure demonstrates the interaction between fusion weight ($\lambda_S$), the degree of orthology coverage of networks, and the degree of similarity between fused interactions.
 \textbf{A.} We generated a series of pairs of 10 TF by 200 gene networks (75\% sparsity) in two species.
@@
 We varied the fraction of genes with orthologs (y-axis) and the weight of fusion in network inference (x-axis), and measured performance as the mean-squared error of the true vs inferred network weights.
-  In order to more clearly visualize the varying effect of fusion, performance is plotted in relative units of the of the unfused MSE (left column) for each level of orthology. 
-  This was necessary because each row represents a different pair of networks, generated with a different level of orthology, for which baseline performance varies. 
-  Performance gains from fusion are, as expected, largest when the degree of orthology is large.
+%  In order to more clearly visualize the varying effect of fusion, performance is plotted in relative units of the of the unfused MSE (left column) for each level of orthology. 
+%  This was necessary because each row represents a different pair of networks, generated with a different level of orthology, for which baseline performance varies. 
+%  Performance gains from fusion are, as expected, largest when the degree of orthology is large.
 However, even when the fraction of genes with orthologs is relatively small (top row), we observe gains from fusion.
-  When fused interactions are nearly identical, larger fusion weights always outperform smaller fusion weights. 
-  \textbf{B.} Simulates the case where genes which are orthologs may have different regulatory weights. 
-  We generated a series of networks as in \textbf{A.}, but with larger (7.5 $\times$) gaussian noise added the weights of fused interactions. 
-  As in \textbf{A.}, benefits from fusion were observed at every level of gene orthology.
-However, unlike \textbf{A.}, there was an optimal intermediate value of $\lambda_S$ that traded off between the benefits of fusion and the cost of combining heterogenous data.}
-\end{center}
-\end{figure}
+%  When fused interactions are nearly identical, larger fusion weights always outperform smaller fusion weights. 
+%  \textbf{B.} Simulates the case where genes which are orthologs may have different regulatory weights. 
+%  We generated a series of networks as in \textbf{A.}, but with larger (7.5 $\times$) gaussian noise added the weights of fused interactions. 
+%  As in \textbf{A.}, benefits from fusion were observed at every level of gene orthology.
+However, unlike \textbf{A.}, there was an optimal intermediate value of $\lambda_S$ that traded off between the benefits of fusion and the cost of% combining heterogenous data.}
+%\end{center}
+%\end{figure}
 
@@
 \end{figure}
+\pagebreak
 
@@
 
+\pagebreak
+
 \begin{figure}
@@
 
+\pagebreak
 \begin{figure}
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[scale=.5,trim=0mm 0mm 0mm 0mm,clip]{party3.eps}
-    \caption{\label{cartoon} Schematic representation of design matrix construction. Here, the circles and hexagons correspond to different species. 
-    Bidirectional arrows represent orthology information and dotted arrows represent putative interactions between TFs and genes. 
-    Rectangles represent matrices; because weights can be solved independently unless there exists fusion constraints between them, we identify related weights and construct matrix for solving. }
-\end{figure}
-
 \end{document}

### 2016-07-28 paper_revisions.tex
moved stuff around

@@
 Any set of pairwise L2 penalties on the differences between regulatory interactions (equivalent to a multivariate Gaussian prior on networks; see methods) can be optimized by the algorithm we develop.
-As a result, any biological prior on similarity of regulation can be used in place of orthology.
-For example, we demonstrate improvements in network inference performance with the introduction of penalties based on the expectation that genes in the same operon will have similar regulators.
+As a result, any biological prior on similarity of regulation can be used in place of orthology, including non one-to-one mappings such the prior on genes within the same operon.
+We demonstrate improvements in network inference performance with the introduction of penalties based on the expectation that genes in the same operon will have similar regulators.
 
@@
 
-We test the ability of fused L2 and adaptive fusion to improve network recovery on both synthetic data and on real data, by fitting  networks from two strains of tthe bacteria species \textit{Bacillus subtilis} and along with the distantly related \textit{Bacillus anthracis}.
-We then explore the applications of adaptive fusion by demonstrating the technique's ability to identify incorrect orthology information in biological data, simulating the biologically relevant cases of neo- and sub-functionalizations.
+We test the ability of fused L2 and adaptive fusion to improve network recovery on both synthetic data and on real data.
+We demonstrate the applications of our method to multi-species network inference, using the bacteria species \texit{Bacillus subtilis} and the distantly related \texti{Bacillus anthracis}; multi-platform network inference using \texit{B.
+subtilis}; and intra-species fused network inference using operons in \textit{B.
+subtilis}.
+We then discuss adaptive fusion and show the technique's ability to identify incorrect orthology information which has been introduced to a biological dataset, suggesting the technique's applicability to discovering neo- and sub-functionalizations.
+
+\section{Results}
+We used both synthetic networks and real data to evaluate our methods.
+For the synthetic data, we generated random pairs of networks in which orthologous genes have similar regulatory interactions, and then sampled gene expression from these networks, which we used to derive learned networks for comparison with the input (true) networks.
+For real data, we examined recovery of a known gold standard in \textit{Bacillus subtilis} \cite{michna_subtiwikidatabase_2014}.
+
+\subsection{Simulations}
+\subsubsection{Using fused regression to learn related networks}
+We created synthetic networks to approximate two related biological processes, then evaluated performance of our fused L2 regression, which learns the networks simultaneously given a prior on the relatedness of interactions.
+Simulated networks consisted of pairs of related 10 TF by 200 gene networks, with varying numbers of expression data samples made available to the solver.
+In order to compare the relative contribution of data in the species of interest (species 1) and in a related organism (species 2) to network recovery, we varied the number of conditions in each species while evaluating the mean-squared-error between the recovered network for species 1 and its known true network.
+When the amount of data from the related species was held constant, increasing the amount of data available for learning the network for the species of interest resulted in a more accurate network prediction, as expected  (figure \ref{fusion}b).
+Similarly, when we increased the amount of data from the related species, we obtained performance gains on network one using fused L2 regression, demonstrating our ability to improve network inference on one dataset through incorporation of a related dataset.
+
+\subsubsection{Fused regression improves performance on both the constrained and non-constrained parts of the network}
+Our approach is useful for learning networks from similar sources such as related cell types from the same species, where there exists a one-to-one mapping of genes, as well as datasets where the orthology mapping does not span all genes.
+This latter case is typical not only for cross-species network inference, but can occur when using different technology, eg microarray and RNAseq, where there is incomplete overlap in the genes that each method assays as well as incomplete overlap in the genes expressed in different experimental designs.
+When orthology is incomplete we are interested in knowing if performance gains from fused regression are limited to those interactions which have fusion constraints, or if they extend to the entire network.
+To test this we used multiple 20 TF by 200 gene synthetic networks with varying proportions of orthologous TFs and genes.
+We divided networks into those interactions with fusion constraints (the constrained subnetwork) and interactions without fusion constraints (the non-constrained subnetwork).
+We varied the weight on the fusion penalty, $\lambda_S$, and evaluated performance by computing AUPR on the constrained subnetwork, the non-constrained subnetwork, and the whole network (Figures \ref{fusion}c-f).
+Because the conserved subgraphs were known to be similar to each other, we expected performance to improve as the fusion penalty weight increased (Figures \ref{fusion}c-f, Constrained).
+
+Interestingly, performance gains were seen even in the portion of the network that was unconstrained by fusion (Figures \ref{fusion}c-f, Non-constrained).
+This is because a gene may have some interactions that are constrained by fusion - regulation by TFs with orthologs - and some interactions that are unconstrained - regulation by TFs without orthologs.
+Because both constrained and unconstrained components compete to explain the same pattern of gene expression, improving recovery of the constrained sub-network will tend to improve recovery for the unconstrained sub-network as well.
+
+\subsubsection{Adaptive fusion successfully identifies and unfuses 'neofunctionalized' genes}
+Orthology prediction is not alwyas a good proxy for functional conservation \cite{gabaldon_functional_2013, studer_how_2009, nehrt_testing_2011}. 
+To allow for orthologous genes to be unfused we implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (figure \ref{adapt-schematic}).
+Pairs of interactions that are dissimilar even after fusion, which sit in the flat portion of this penalty function, are effectively ``unfused,'' and no further penalty is incurred as differences in interaction weights grow.
+Our network procedure strongly favors similarity of fused interactions, and only ``unfuses'' interactions when their similarity cannot be reconciled with expression data. 
+As a result, the ``unfusing'' or relaxation of the fusion penalty on certain constraints is much more direct evidence for neofunctionalization than comparing separately fit networks could provide.
+
+We performed a simulation to assess the ability of our adaptive fusion algorithm to distinguish which parts of two input networks are conserved vs.
+neofunctionalized (figure \ref{adaptivefusion}).
+We generated synthetic fused networks and introduced error in the fusion constraints by adding false positives and negatives to the orthology information given to the solver.
+Because we knew which entries in the orthology mapping were ``incorrect'' (not reflected in the generation of the networks), we could correctly label fusion constraints that involved one or more ``incorrect'' mappings.
+We verified that adaptive-fusion unfused mostly ``incorrectly fused'' interactions (figure \ref{adaptivefusion}a red dots), while leaving truly analogous interactions fused (figure \ref{adaptivefusion}a green dots).
+
+We then compared the recovery of interaction weights which were accurately fused, and recovery of interaction weights which were inaccurately fused due to incorrect orthology information (figure \ref{adaptivefusion}b).
+Because fused L2 heavily penalizes large differences between weights which are predicted to be similar, it is able to retrieve a more accurate network for those interactions with true fusion constraints than by learning networks separately (measured by MSE between the true and inferred interaction weights).
+In this simulation, however, the gains accomplished through fused regression do not extend to those interactions lacking true fusion constraints, and the error remains similar to learning networks separately.
+When we applied adaptive fusion, we did not observe an improvement in network recovery (relative to fused L2).
+However, we were able to identify fusion constraints reflecting incorrect orthology information that had been provided to the algorithm (Figure \ref{adaptivefusion} subsection).
+[ZACK - WE NEED TO WRITE MORE ABOUT THIS]
+
+\subsection{Fused network inference with bacterial data}
+We used gene-expression data from \textit{Bacillus subtilis} and \textit{B. anthracis} in order to assess performance gains of fused regression on real data. 
+Our \textit{B subtilis} data set consists of 360 time-series and steady-state observations of 4891 genes, 4100 of which are protein coding \cite{kunst_complete_1997}, during the life cycle.
+Our \textit{\textit{B.
+anthracis}} dataset consists of 72 time-series and steady-state observations of 5536 genes comprising data from distinct points in the life cycle and iron-starvation conditions.
+There were 247 known transcription factors (TFs) in the \textit{\textit{B. subtilis}} dataset, and 248 TFs in the \textit{B. anthracis} dataset.
+We obtained 1,870 one-to-one orthologs from Inparanoid \cite{ostlund_inparanoid_2010}, 95 of which are transcription factors, which produced 177,650 fusion-constraints between gene interactions within the two species.
+This number represents only $14.7\%$ of the regulatory interaction matrix in \textit{B. subtilis} and $12.9\%$ in \textit{B. anthracis}.
+
+To assess network inference performance, and for use as priors, we used a gold standard of 3,040 known \textit{B. subtilis} interactions with corresponding activation and repression sign. 
+Of these 3,040 priors, 968 had corresponding interactions in \textit{B. anthracis}. 
+Based on our simulation results, we can expect the greatest gains in network-inference performance from fusion when the species of interest has a small number of available conditions, but data is abundant in a related species.
+However, in order to evaluate performance objectively a gold-standard of known interactions is necessary. 
+As a result, we can only evaluate network recovery for \textit{B. subtilis}, and \textit{B. subtilis} also has the majority of our conditions. 
+In order to simulate the data-poor regime, we subsampled our \textit{B. subtilis} data. 
+We divided our \textit{B.
+subtilis} data into $k$ folds, and then for each fold fit a network to the \textit{B.
+subtilis} data from that fold alone fused to the entire 72 \textit{B.
+anthracis} conditions (figure \ref{xspecies-real}a).
+Though overall performance is hindered by our subsampling of \textit{B.
+subtilis} data (a necessary procedure to allow evaluation of networks) we demonstrate marked improvement in learning the \textit{B.
+subtilis} network when using fused regression (figure \ref{xspecies-real}a).
+Notably, these performance gains occur mostly at low values of recall.
+That is: the highest confidence part of the network is inferred more accurately, with minimal gains for interactions which are more uncertain.
+Because these interactions are likely to be the focus of followup experiments and validation, gains here are more valuable for prioritizing the order in which interactions are investigated.
+
+\subsubsection{Adaptive fusion using bacterial data}
+[ZACK MORE HERE TOO]
+ The goal of adaptive fusion is to unfuse constraints between non-conserved interactions, while leaving intact all other constraints. 
+ However, because we lacked a comprehensive gold standard of known non-conserved interactions between \textit{B.
+subtilis} and \textit{B.
+anthracis}, we were unable to directly evaluate how accurately adaptive fusion identified these interactions.
+  We opted instead to introduce a large number of random fusion constraints between genes not known to be orthologous. 
+ These fake constraints, which are unlikely to reflect any conserved network structure, served as a proxy for the unknown fraction of non-conserved interactions between orthologous genes. 
+ We ran adaptive fusion to learn networks for \textit{B. subtilis} and \textit{B. anthracis}, using these constraints, along with those generated by known orthology. 
+ We confirmed that our injected spurious fusion constraints were unfused at a higher rate than those generated by known orthologs (see figure \ref{xspecies-real}b & c). 
+ Although it may seem odd that a large fraction of fake constraints were left intact, we note that biological networks tend to be sparse, so that many of the random fusion constraints are between coefficients with near zero weight (and therefore near zero difference in weight) (figure \ref{xspecies-real}c).
+
+ \subsubsection{Single species applications}
+
+Although the approach we describe was developed with cross-species network inference in mind, the framework of defining fusion constraints on TF-gene interactions expected to be similar is quite general, and can be applied to data from a single organism.
+Despite the many large-scale collaborations which attempt to make protocols as uniform as possible for comparability between datasets generated by different labs \cite{paten_nih_2015,kundaje_integrative_2015} and several methods for removing batch effects \cite{irizarry_exploration_2003,johnson_adjusting_2007}, there still exists technical and biologial variability between many experiments attempting to capture the same or similar experimental conditions esspecially when experiments employ different experimental platforms.
+With the advent of RNAseq, for example, microarray based technologies are no longer the dominant assay for genome-wide expression, but a large body of accumulated legacy data remains useful if it can be integrated with more modern techniques.
+
+Currently, the most widely used approach to combining datasets for network inference is to learn networks from disparate datasets separately, then rank combine the networks as in Marbach et al \cite{marbach_revealing_2010}.
+We included, along with our \textit{B.
+subtilis} dataset, a previously published dataset containing 269 samples covering 104 conditions, obtained using a different tiling microarray (vs custom microarray) and different strain of \textit{B.
+subtilis} \cite{nicolas2012condition}.
+We compared performance when learning the networks separately and then rank combining (as in cite Ciofani) to learning the networks simultaneously using fusion regression and show improvement in performance using our fused L2 approach (figure \ref{fusedl2-real}a).
+
+Our approach can also make use of other forms of biological data beyond orthology.
+Information about the expected similarity of TF-gene interactions can come from knowledge about the promoter region or the structure, for bacteria, of polycistronic transcripts.
+In bacteria, genes within the same operon are typically under the control of the same promoter \cite{lawrence_shared_2002}.
+We posited, therefore, that genes within the same operon will be regulated similarly by the same transcription factors.
+We applied fusion regression by creating fusion constraints between a given transcription factor and genes within the same operon, and showed a boost in \textit{B.
+subtilis} network recovery using within-species fusion (figure \ref{fusedl2-real}b).
+
+\subsection{Transcription factor activity estimation integrates into fusion regression approach}
+We tested a combination of our fused regression approach with a method for estimating transcription factor activities (TFA). 
+Rather than modeling gene expression using transcription factor mRNA abundance, we fit gene expression as a function of transcription factor activity, as applied to \textit{B.
+subtilis} by Arrieta-Ortiz et al \cite{arrieta-ortiz_experimentally_2015}.
+TFA activity estimates  transcription factor activities that are modulated through mechanisms such as dimerization and interaction with required factors.
+TFA activity estimates have been shown prior to be better predictors of TF function than expression level alone in several contexts including similar network inference tasks\cite{fu_reconstructing_2011} \cite{arrieta-ortiz_experimentally_2015}.
+We estimate TFA based on known regulatory interactions using network component analysis \cite{liao2003network}.
+To test the integration of this approach with our fused regression, we assessed the combination of \textit{B.
+subtilis} datasets, as in figure \ref{fusedl2-real}a, with the incorporation of TFA estimation.
+We randomly divided the prior known interactions in half, and used half to learn TFA and to generate priors on network structure. 
+The remaining interactions were reserved as a gold standard for validation. 
+As in previous studies, we observed a marked improvement in network inference when using transcription factor activity (figure \ref{tfa}).
+We also obtained AUPR improvement when using fused regression alongside TFA; gains from sharing information across datasets using fused regression were preserved and even enhanced by using TFA.
 
@@
 
-where $X$, $Y$, and $\beta$ are matrices, and $\beta_{g,k} \approx \beta_{h,l}$ denotes fusion between entries of $\beta$ (enforcing similarity between model weights across separate data-sets). 
+where $X$, $Y$, and $\beta$ are matrices, and $\beta_{g,k} \approx \beta_{h,l}$ denotes fusion between entries of $\beta$ (enforcing similarity between model weights across separate datasets). 
 Note that, like ridge regression, this penalty can be thought of as representing a Gaussian prior on the coefficients $\beta$. 
@@
 As a result, these linked response variables must be solved simultaneously through vectorization. 
-Two response variables are linked by a fusion constraint if any of the regulatory weights affecting those genes are linked by a fusion constraint. 
+Two response variables are linked if any of the regulatory weights affecting those genes are linked by a fusion constraint. 
 Two response variables must be solved simultaneously if there is any chain of linked response variables connecting them. 
@@
 \begin{equation}
-\argmin_{\beta^S} \displaystyle\sum_{S \in \{1, 2\}} ||X^{S}\beta^{S}(t) - Y^{S}||^2 + \lambda_R||\beta^{S}(t)||^2 + \lambda_S\displaystyle \sum_{\substack{(g,h) \in orth,\\
+\argmin_{\beta^S} \displaystyle\sum_{S \in \{1, 2\}} ||X^{S}\beta^{S}(t) - Y^{S}||^2 + \lambda_R||\beta^{S}(t)||^2 + \lambda_S\displaystyle \sum_{\substack{(g,k) \in orth,\\
  (h,l) \in orth}} ||\beta^{S}_{g,k}(t) - \beta_{h,l}^{S}(t-1)||^2
@@
 This procedure is guaranteed to achieve the best unfused solution in the case when $\lambda_S$ is constrained at 0. 
-As a result, any performance gains of fused regression are a lower bound on the highest achievable performance gains. 
 To optimize $\lambda_R$ we use cyclical coordinate descent algorithms from the 'glmnet' package \cite{friedman_regularization_2010} to compute a ridge regularization path. 
@@
 This will be inappropriate for interactions which are identified based on orthology as being analogous, but which are no longer similar due to evolutionary changes. 
-We propose that a saturating penalty  that is relaxed once differences in weights grow beyond a certain point (interactions which appear to be very different based on the data are effectively unfused).
+We propose a saturating penalty that is relaxed once differences in weights grow beyond a certain point (interactions which appear to be very different based on the data are effectively unfused). 
 A related problem has been studied in the context of LASSO regularization, where it was shown by Fan and Li that using a saturating penalty retains many of LASSO's desireable properties while removing its bias towards model weights of 0 \cite{fan2001variable}.
@@
 \subsubsection{Approach overview}
+For a high-level view, we summarize our approach:
+
 \begin{algorithm}
@@
 The function $p(x)$ controls the shape of the penalty function, while scalar $\lambda_S$ controls the overall scaling of the penalization of differences between fused coefficients.
-$\lambda_S$ controls the tradeoff between fitting the expression data-sets individually and producing a set of networks that conform to evolutionary prior knowledge.
+$\lambda_S$ controls the tradeoff between fitting the expression datasets individually and producing a set of networks that conform to evolutionary prior knowledge.
 This gives us the final equation to be minimized:
@@
 In the case of cross-species network inference, there is unlikely to be a complete one-to-one orthology.
-However, because we  constraints the similarity of individual interactions, rather than on the networks as a whole \cite{parikh2011treegl}, we can pool some information across species even when a small fraction of genes have orthologs.
+However, because we create constraints the similarity of individual interactions, rather than on the networks as a whole \cite{parikh2011treegl}, we can pool some information across species even when a small fraction of genes have orthologs.
 
@@
 We generate simulated data to evaluate the ability of our fused L2 approach to learn the true network and to show that sharing information between similar but not identical data sources results in more accurate network recovery.
-Generation of simulated data begins with the production of random orthology mappings with sparcity simlar to that found in real data-sets. 
-We produce a one-to-one orthology by pairing random genes until a specified fraction have been assigned orthologs. 
+Generation of simulated data begins with the production of random orthology mappings. 
+We produce orthology by pairing random genes until a specified fraction have been assigned orthologs. 
 This process is carried out separately for TFs and non-TF genes, so that TFs and non-TF genes are never assigned to be orthologous. 
-We then produce a pair of random networks ($\beta^1$ and $\beta^2$) as follows.
-For each unfilled entry in $\beta^1$ or $\beta^2$, we enumerate the set $C$ consisting of the entry along with every entry in either matrix to which it is fused.
-With probability equal to the sparsity rate we assign every entry in $C$ to be 0, otherwise we sample a value $v \sim \mathcal{N}(0,1)$ and independently assign each entry in $C$ to $v + \mathcal{N}(0, \sigma_f^2)$.
+We then produce a pair of random networks ($\beta^1$ and $\beta^2$) as follows: for each unfilled entry in $\beta^1$ or $\beta^2$, we enumerate the set $C$ consisting of the entry along with every entry in either matrix to which it is fused.
+We set a sparsity rate and with probability equal to this sparsity rate, we assign every entry in $C$ to be 0, otherwise we sample a value $v \sim \mathcal{N}(0,1)$ and independently assign each entry in $C$ to $v + \mathcal{N}(0, \sigma_f^2)$.
 $\sigma_f$ is a parameter that controls the distribution of differences in the values of fused coefficients, so that the nonzero coefficients of $\beta^1, \beta^2$ are distributed as $\mathcal{N}(0, 1 + \sigma_f^2)$.
@@
 subtilis} and \textit{B.
-anthracis} \cite{ostlund_inparanoid_2010}
-
-[KARI - INSERT HOW GOLD STANDARD WAS OBTAINED]
-
-\section{Results}
-We used both synthetic networks and real data to test the ability of fused regression to improve the performance of network inference, and the ability of our adaptive fusion procedure to identify conserved interactions between orthologous genes.
-For the synthetic data, we generated random pairs of networks in which orthologous genes have similar regulatory interactions, and then sampled gene expression from these networks, which we used to derive learned networks for comparison with the input (true) networks.
-For real data, we examined recovery of a known gold standard in \textit{Bacillus subtilis} \cite{michna_subtiwikidatabase_2014}.
-
-\subsection{Simulations}
-\subsubsection{Using fused regression to learn related networks}
-We created synthetic networks to approximate two related biological processes, then evaluated performance of our fused L2 regression, which learns the networks simultaneously given a prior on the relatedness of interactions.
-Simulated networks consisted of pairs of related 10 TF by 200 gene networks, with varying numbers of expression data samples made available to the solver.
-In order to compare the relative contribution of data in the species of interest (species 1) and in a related organism (species 2) to network recovery, we varied the number of conditions in each species while evaluating the mean-squared-error between the recovered network for species 1 and its known true network.
-When the amount of data from the related species was held constant, increasing the amount of data available for learning the network for the species of interest resulted in a more accurate network prediction, as expected  (figure \ref{fusion}b).
-Similarly, when we increased the amount of data from the related species, we obtained performance gains on network one using fused L2 regression, demonstrating our ability to improve network inference on one dataset through incorporation of a related dataset.
-
-\subsubsection{Fused regression improves performance on both the constrained and non-constrained parts of the network}
-Our approach is useful for learning networks from similar sources such as related cell types from the same species, where there exists a one-to-one mapping of genes, as well as datasets where the orthology mapping does not span all genes.
-This latter case is typical not only for cross-species network inference, but can occur when using different technology, eg microarray and RNAseq, where there is incomplete overlap in the genes that each method assays as well as incomplete overlap in the genes expressed in different experimental designs.
-When orthology is incomplete we are interested in knowing if performance gains from fused regression are limited to those interactions which have fusion constraints, or if they extend to the entire network.
-To test this we used multiple 20 TF by 200 gene synthetic networks with varying proportions of orthologous TFs and genes.
-We divided networks into those interactions with fusion constraints (the constrained subnetwork) and interactions without fusion constraints (the non-constrained subnetwork).
-We varied the weight on the fusion penalty, $\lambda_S$, and evaluated performance by computing AUPR on the constrained subnetwork, the non-constrained subnetwork, and the whole network (Figures \ref{fusion}c-f).
-Because the conserved subgraphs were known to be similar to each other, we expected performance to improve as the fusion penalty weight increased (Figures \ref{fusion}c-f, Constrained).
-
-Interestingly, performance gains were seen even in the portion of the network that was unconstrained by fusion (Figures \ref{fusion}c-f, Non-constrained).
-This is because a gene may have some some interactions that are constrained by fusion - regulation by TFs with orthologs - and some interactions that are unconstrained - regulation by TFs without orthologs.
-Because both constrained and unconstrained components compete to explain the same pattern of gene expression, improving recovery of the constrained sub-network will tend to improve recovery for the unconstrained sub-network as well.
-
-\subsubsection{Adaptive fusion successfully identifies and unfuses 'neofunctionalized' genes}
-Orthology prediction is not a proxy for functional conservation \cite{gabaldon_functional_2013, studer_how_2009, nehrt_testing_2011}. 
-To allow for orthologous genes to be unfused we implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (figure \ref{adapt-schematic}).
-Pairs of interactions that are dissimilar even after fusion, which sit in the flat portion of this penalty function, are effectively ``unfused,'' and no further penalty is incurred as differences in interaction weights grow.
-Our network procedure strongly favors similarity of fused interactions, and only ``unfuses'' interactions when their similarity cannot be reconciled with expression data. 
-As a result, the ``unfusing'' or relaxation of the fusion penalty on certain constraints is much more direct evidence for neofunctionalization than comparing separately fit networks could provide.
-
-We performed a simulation to assess the ability of our adaptive fusion algorithm to distinguish which parts of two input networks are conserved vs.
-neofunctionalized (figure \ref{adaptivefusion}).
-We generated synthetic fused networks and introduced error in the fusion constraints by adding false positives and negatives to the orthology information given to the solver.
-Because we knew which entries in the orthology mapping were ``incorrect'' (not reflected in the generation of the networks), we could correctly label fusion constraints that involved one or more ``incorrect'' mappings.
-We verified that adaptive-fusion unfused mostly ``incorrectly fused'' interactions (figure \ref{adaptivefusion}a red dots), while leaving truly analogous interactions fused (figure \ref{adaptivefusion}a green dots).
-
-We then compared the recovery of interaction weights which were accurately fused, and recovery of interaction weights which were inaccurately fused due to incorrect orthology information (figure \ref{adaptivefusion}b).
-Because fused L2 heavily penalizes large differences between weights which are predicted to be similar, it is able to retrieve a more accurate network for those interactions with true fusion constraints than by learning networks separately (measured by MSE between the true and inferred interaction weights).
-In this simulation, however, the gains accomplished through fused regression do not extend to those interactions lacking true fusion constraints, and the error remains similar to learning networks separately.
-When we applied adaptive fusion, we did not observe an improvement in network recovery (relative to fused L2).
-However, we were able to identify fusion constraints reflecting incorrect orthology information that had been provided to the algorithm (Figure \ref{adaptivefusion} subsection).
-[ZACK - WE NEED TO WRITE MORE ABOUT THIS]
-
-\subsection{Fused network inference with bacterial data}
-We used gene-expression data from \textit{Bacillus subtilis} and \textit{B. anthracis} in order to assess performance gains of fused regression on real data. 
-Our \textit{B subtilis} data set consists of 360 time-series and steady-state observations of 4891 genes, 4100 of which are protein coding \cite{kunst_complete_1997}, during the life cycle.
-Our \textit{\textit{B.
-anthracis}} dataset consists of 72 time-series and steady-state observations of 5536 genes comprising data from distinct points in the life cycle and iron-starvation conditions.
-There were 247 known transcription factors (TFs) in the \textit{\textit{B. subtilis}} dataset, and 248 TFs in the \textit{B. anthracis} dataset.
-We obtained 1,870 one-to-one orthologs from Inparanoid \cite{ostlund_inparanoid_2010}, 95 of which are transcription factors, which produced 177,650 fusion-constraints between gene interactions within the two species.
-This number represents only $14.7\%$ of the regulatory interaction matrix in \textit{B. subtilis} and $12.9\%$ in \textit{B. anthracis}.
-
-To assess network inference performance, and for use as priors, we used a gold standard of 3,040 known \textit{B. subtilis} interactions with corresponding activation and repression sign. 
-Of these 3,040 priors, 968 had corresponding interactions in \textit{B. anthracis}. 
-Based on our simulation results, we can expect the greatest gains in network-inference performance from fusion when the species of interest has a small number of available conditions, but data is abundant in a related species.
-However, in order to evaluate performance objectively a gold-standard of known interactions is necessary. 
-As a result, we can only evaluate network recovery for \textit{B. subtilis}, and \textit{B. subtilis} also has the majority of our conditions. 
-In order to simulate the data-poor regime, we subsampled our \textit{B. subtilis} data. 
-We divided our \textit{B.
-subtilis} data into $k$ folds, and then for each fold fit a network to the \textit{B.
-subtilis} data from that fold alone fused to the entire 72 \textit{B.
-anthracis} conditions (figure \ref{xspecies-real}a).
-Though overall performance is hindered by our subsampling of \textit{B.
-subtilis} data (a necessary procedure to allow evaluation of networks) we demonstrate marked improvement in learning the \textit{B.
-subtilis} network when using fused regression (figure \ref{xspecies-real}a).
-Notably, these performance gains occur mostly at low values of recall.
-That is: the highest confidence part of the network is inferred more accurately, with minimal gains for interactions which are more uncertain.
-Because these interactions are likely to be the focus of followup experiments and validation, gains here are more valuable for prioritizing the order in which interactions are investigated.
-
-\subsubsection{Adaptive fusion using bacterial data}
-[ZACK MORE HERE TOO]
- The goal of adaptive fusion is to unfuse constraints between non-conserved interactions, while leaving intact all other constraints. 
- However, because we lacked a comprehensive gold standard of known non-conserved interactions between \textit{B.
-subtilis} and \textit{B.
-anthracis}, we were unable to directly evaluate how accurately adaptive fusion identified these interactions.
-  We opted instead to introduce a large number of random fusion constraints between genes not known to be orthologous. 
- These fake constraints, which are unlikely to reflect any conserved network structure, served as a proxy for the unknown fraction of non-conserved interactions between orthologous genes. 
- We ran adaptive fusion to learn networks for \textit{B. subtilis} and \textit{B. anthracis}, using these constraints, along with those generated by known orthology. 
- We confirmed that our injected spurious fusion constraints were unfused at a higher rate than those generated by known orthologs (see figure \ref{xspecies-real}b & c). 
- Although it may seem odd that a large fraction of fake constraints were left intact, we note that biological networks tend to be sparse, so that many of the random fusion constraints are between coefficients with near zero weight (and therefore near zero difference in weight) (figure \ref{xspecies-real}c).
-
- \subsubsection{Single species applications}
-
-Although the approach we describe was developed with cross-species network inference in mind, the framework of defining fusion constraints on TF gene interactions expected to be similar is quite general, and can be applied to data from a single organism.
-Despite the many large-scale collaborations which attempt to make protocols as uniform as possible for comparability between datasets generated by different labs \cite{paten_nih_2015,kundaje_integrative_2015} and several methods for removing batch effects \cite{irizarry_exploration_2003,johnson_adjusting_2007}, there still exists technical and biologial variability between many experiments attempting to capture the same or similar experimental conditions esspecially when experiments employ different experimental platforms.
-With the advent of RNAseq, for example, microarray based technologies are no longer the dominant assay for genome-wide expression, but a large body of accumulated legacy data remains useful if it can be integrated with more modern techniques.
-
-Currently, the most widely used approach to combining datasets for network inference is to learn networks from disparate datasets separately, then rank combine the networks as in Marbach et al \cite{marbach_revealing_2010}.
-We included, along with our \textit{B.
-subtilis} dataset, a previously published dataset containing 269 samples covering 104 conditions, obtained using a different tiling microarray (vs custom microarray) and different strain of \textit{B.
-subtilis} \cite{nicolas2012condition}.
-We compared performance when learning the networks separately and then rank combining (as in cite Ciofani) to learning the networks simultaneously using fusion regression and show large improvement in performance using our fused L2 approach (figure \ref{fusedl2-real}a).
-
-Our approach can also make use of forms of biological data other than RNA expression.
-Information about the expected similarity of TF-gene interactions can come from knowledge about the promoter region or the structure, for bacteria, of polycistronic transcripts.
-In bacteria, genes within the same operon are typically under the control of the same promoter \cite{lawrence_shared_2002}.
-We posited, therefore, that genes within the same operon will be regulated similarly by the same transcription factors.
-We applied fusion regression by creating fusion constraints between a given transcription factor and genes within the same operon, and showed a boost in \textit{B.
-subtilis} network recovery using within-species fusion (figure \ref{fusedl2-real}b).
-
-\subsection{Transcription factor activity estimation integrates into fusion regression approach}
-We tested a combination of our fused regression approach with a method for estimating transcription factor activities (TFA). 
-Rather than modeling gene expression using transcription factor mRNA abundance, we fit gene expression as a function of transcription factor activity, as applied to \textit{B.
-subtilis} by Arrieta-Ortiz et al \cite{arrieta-ortiz_experimentally_2015}.
-TFA activity estimates  transcription factor activities that are modulated through mechanisms such as dimerization and interaction with required factors.
-TFA activity estimates have been shown prior to be better predictors of TF function than expression level alone in several contexts including similar network inference tasks\cite{fu_reconstructing_2011} \cite{arrieta-ortiz_experimentally_2015}.
-We estimate TFA based on known regulatory interactions using network component analysis \cite{liao2003network}.
-To test the integration of this approach with our fused regression, we assessed the combination of \textit{B.
-subtilis} datasets, as in figure \ref{fusedl2-real}a, with the incorporation of TFA estimation.
-We randomly divided the prior known interactions in half, and used half to learn TFA and to generate priors on network structure. 
-The remaining interactions were reserved as a gold standard for validation. 
-As in previous studies, we observed a marked improvement in network inference when using transcription factor activity (figure \ref{tfa}).
-We also obtained AUPR improvement when using fused regression alongside TFA; gains from sharing information across datasets using fused regression were preserved and even enhanced by using TFA.
+anthracis} \cite{ostlund_inparanoid_2010} To evaluate our approach, we compare with a set of experimentally validated regulatory interactions from SubtiWiki \cite{michna_subtiwikidatabase_2014}.
 

### 2016-07-29 paper_revisions.tex
zack more edits. added paragraph in intro and discussion about other methods. general cleanup. fix to operon explanation (but not figure legends yet)

--- 2016-07-26 paper_revisions.tex
@@
 Numerous studies have shown that functional conservation exists in gene regulatory networks even across large evolutionary distance \cite{satou2006gene, hinman2009evolution,tanay2005conservation,erwin2009evolution}.
-Our first method -- fused ridge -- uses an L2 penalty on the differences between \textit{a priori} similar interactions (termed fusion penalty).
-These constraints favor, but do not require, network configurations in which orthologous genes have similar regulators.
+Our first method -- fused L2 -- takes advantage of this similarity by imposing an L2 penalty on the differences between \textit{a priori} similar interactions (termed fusion penalty).
+These constraints favor  network configurations in which orthologous genes have similar regulators.
 Because network inference problems are typically underconstrained, these additional constraints allow data in one species to improve network inference performance in another.
 
-By fusing arbitrary interactions within the networks rather than the networks themselves \cite{}, the method allows pooling of data even when there is not a one-to-one correspondence between genes, which is critical to the method's application to cross-species network inference.
-This general framework can be extended to an arbitrary number of distinct organisms, each contributing data with only partially overlapping sets of genes.
-This is an advantage over existing approaches to multi-species network inference, which infer only a sub-network for which orthologs exist in every species \cite{joshi_multi-species_2015}.
+A closely related algorithm -- fused LASSO -- has been applied gene regulatory network inference in multiple cell lines, or in multiple experimental regimes \cite{omranian2016gene, parikh2011treegl, pierson2015sharing}.
+These approaches have been shown to perform well in fusing data from multiple sources, but because the fusion is between whole networks rather than individual coefficients, they can only be applied when sources share the same set of expressed genes.
+This all but rules out cross-species applications, and can present difficulty even when data describing a single species sourced from multiple experimental techniques or platforms.
+Existing approaches to multi-species network inference have inferred only a sub-network of genes with orthologs in every species \cite{joshi_multi-species_2015}, or fit networks separately  \cite{marbach_revealing_2010}.
+Our approach allows multiple networks to be inferred with biologically motivated constraints on similarity even when the data sources have a relatively small set of overlapping genes.
 
-This general framework for multi-species network inference has broad applications.
-Any set of pairwise L2 penalties on the differences between regulatory interactions (equivalent to a multivariate Gaussian prior on networks; see methods) can be optimized by the algorithm we develop.
+This general framework for multi-species network inference has broad applications, extending beyond the case of combining multiple networks.
+Any set of pairwise L2 penalties on the differences between regulatory interactions (equivalent to a multivariate Gaussian prior on networks; see methods) can be utilized by the algorithm we develop.
 As a result, any biological prior on similarity of regulation can be used in place of orthology.
@@
 In the case of cross-species network inference, there is unlikely to be a complete one-to-one orthology.
-However, because we  constraints the similarity of individual interactions, rather than on the networks as a whole \cite{parikh2011treegl}, we can pool some information across species even when a small fraction of genes have orthologs.
+However, because we  constraints the similarity of individual interactions, rather than on the networks as a whole \cite{parikh2011treegl,omranian2016gene,}, we can pool some information across species even when a small fraction of genes have orthologs.
 
@@
 \subsubsection{Adaptive fusion successfully identifies and unfuses 'neofunctionalized' genes}
-Orthology prediction is not a proxy for functional conservation \cite{gabaldon_functional_2013, studer_how_2009, nehrt_testing_2011}. 
+Although we expect that orthology provides some evidence of regulatory similarity, we know that it is not a perfect proxy for functional conservation \cite{gabaldon_functional_2013, studer_how_2009, nehrt_testing_2011}.
+Some gene interactions may be non-conserved, and in order to minimize bias it would be desireable to identify these interactions and relax the fusion constraints on them.
 To allow for orthologous genes to be unfused we implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (figure \ref{adapt-schematic}).
@@
 We posited, therefore, that genes within the same operon will be regulated similarly by the same transcription factors.
-We applied fusion regression by creating fusion constraints between a given transcription factor and genes within the same operon, and showed a boost in \textit{B.
-subtilis} network recovery using within-species fusion (figure \ref{fusedl2-real}b).
+We applied fusion regression by creating fusion constraints between a given transcription factor and genes within the same operon.
+We also used a subset ($50\%$ [ZACK CHECK THIS]) of the gold-standard as priors by relaxing the $\lambda_R$ parameter for these interactions, as in \cite{greenfield_robust_2013}, and evaluated on the remaining set of priors.
+Incorporation of operon-derived fusion constraints in \textit{B.
+subtilis} improved network recovery (figure \ref{fusedl2-real}b).
 
@@
 
+A number of existing approaches have applied fused regression to related problems in network inference.
+TreeGL applied fused LASSO to the problem of estimating partial correlations of gene expression in a breast cancer dataset consisting of multiple cell types \cite{parikh2011treegl}.
+They imposed fusion between the coefficients of models learned for different cell types whenever there was an edge between those cell types in a geneology graph describing the cancer cells' development.
+A similar approach was recently applied to the problem of inferring TF to gene regulatory weights from data sources describing multiple environmental conditions in \textit{E.
+Coli}, \textif{Mycobacterium tuberculosis}, and \textit{Mus musculus}.
+In this formulation, networks associated with each data source were fit simultaneously, with an $L1$ constraint on the similarity of (arbitrarily ordered) adjacent networks.
+In both of these approaches, fusion was restricted to be between corresponding entries of the coefficient matrices.
+In terms of our formulation in the cross-species case, this is equivalent to the constraints that would be generated by a one-to-one orthology mapping between species, limiting to the (typical) case where orthology is partial or not one-to-one.
+By optimizing a more general objective function -- in which fusion constraints can be placed on arbitrary pairs of coefficients associated with otherwise unrelated regression tasks -- we can extend this work to the cross-species case.
+
 \textit{B.
 subtilis} and \textit{B.
-anthracis} are distantly related bacterial species with limited gene orthology [ZACK - HOW DISTANTLY?].
+anthracis} are distantly related bacterial species with limited gene orthology.
 Nevertheless we show that network recovery in \textit{B.
@@
 anthracis} genomes do not have clear orthologs (and many orthologs have ambiguous many-to-many groupings).
-Our method, in contrast, can obtain improvements in network inference performance even when the conserved subset of genes is small. [DISCUSS RELATIONSHIP TO OTHER PAPERS HERE]
+Our method, in contrast, can obtain improvements in network inference performance even when the conserved subset of genes is small.
 
 This approach is particularly interesting in light of the diversity of important model organisms used in modern biology.
-Different model systems provide different advantages and disadvantages for experimental design \cite{stolfi_genetic_2012}, but without a principled mechanism for combining data from multiple sources, it is difficult to fully leverage data obtained from even a slightly different model system.
+Different model systems provide different advantages and disadvantages for experimental design \cite{stolfi_genetic_2012}, but in many cases work in less used systems is hampered by lack of available data.
+Our intent is to provide a principled method for combining data from diverse sources, so that results in specialized systems can be integrated with data from well studied organism.
 
-Because the algorithm we developed can accomodate constraints between arbitrary pairs of regulatory interactions, any biological prior representing information about expected regulatory similarity can be represented, even if the prior provides no information about the magnitude or direction of regulation.
-We demonstrate this flexibility through the novel incorporation of operon structure into the gene-regulatory network inference problem. 
+Although we developed the approach initially with cross-species network inference in mind, the core algorithm -- an implementation of L2-fused multi-task regression with arbitrary constraint graphs -- is flexible enough to represent other useful constraint structures.
+Any biological prior representing information about expected regulatory similarity can be utilized, even if the prior provides no information about the magnitude or direction of regulation.
+We demonstrate this flexibility through the incorporation of operon structure into the gene-regulatory network inference problem, and show an improvement in performance in bacterial data when combined with a prior consisting of some known interactions (Figure \ref{fusedl2-real}).
 In this application, fusion reflects the assumption that genes in the same operons have similar regulators \cite{lawrence_shared_2002}.
 
-Although it is important to take advantage of the similarities of related organisms for generating improved models of gene regulation, it is also critically important to understand how systems differ from one another.
-Our cross-species network inference method is premised on the assumption that orthologous genes have similar regulators.
-Existing approaches to the genome-wide testing of this assumption learn regulatory networks separately, then compare to identify conservation \cite{aytes_cross-species_2014}.
+Our work is premised on the assumption that distinct organisms share many conserved regulatory interactions, so that networks obtained in related species should similar to one another in a manner predictable by the evolutionary relationships of the organisms' genes.
+However, we know that orthology is an imperfect predictor of functional similarity\cite{gabaldon_functional_2013, studer_how_2009, nehrt_testing_2011}.
+Moreover, it can be critically important to understand how systems differ from one another.
+Existing approaches to the genome-wide testing of conserved function learn regulatory networks separately, then compare to identify conservation \cite{aytes_cross-species_2014}.
 Because network inference is typically underconstrained, fitting a network that describes a particular set of experimental observations amounts to sampling a single network from a large set of networks that fit the data equally (or almost equally) as well.
-As a result, the existence of a difference between corresponding regulatory interactions in a pair of experimentally derived networks is weak evidence that a difference truly does exist.
+As a result, the existence of a difference between corresponding regulatory interactions in a pair of independently fit networks is weak evidence that a difference truly does exist; some other networks, which do exhibit conservation, may have fit the data almost as well.
 Uncoupled global network inference algorithms are a very weak tool for uncovering evolutionary divergence.
-Our method explicitly favors recovering networks for which evolutionarily corresponding interactions are similar. 
-As a result, the failure to obtain networks that confirm evolutionary conservation is stronger evidence that conservation does not exist; the next best network that does exhibit conservation must fit the data much worse to have overcome the bias built into the fusion constraints.
+Our method is biased to recover networks in which conservation does hold; a failure to find conservation is therefore much much more direct evidence that the networks have truly diverged.
 
@@
 Genes may be fused together on the basis of similar binding sites or chromatin features, and the relaxing of the fusion penalty indicates divergence of gene function.
+[ZACK - CONSIDER CUTTING, SPECULATIVE]
 

### 2016-07-29 paper_revisions.tex
slashed and burned intro

--- 2016-07-28 paper_revisions.tex
@@
 Numerous studies have shown that functional conservation exists in gene regulatory networks even across large evolutionary distance \cite{satou2006gene, hinman2009evolution,tanay2005conservation,erwin2009evolution}.
-Our first method -- fused ridge -- uses an L2 penalty on the differences between \textit{a priori} similar interactions (termed fusion penalty).
+Our first method -- fused L2 -- takes advantage of this similariy by imposing an L2 penalty on the differences between \texit{a prior} similar interactions (termed fusion penalty).
 These constraints favor, but do not require, network configurations in which orthologous genes have similar regulators.
@@
 
-By fusing arbitrary interactions within the networks rather than the networks themselves \cite{}, the method allows pooling of data even when there is not a one-to-one correspondence between genes, which is critical to the method's application to cross-species network inference.
-This general framework can be extended to an arbitrary number of distinct organisms, each contributing data with only partially overlapping sets of genes.
-This is an advantage over existing approaches to multi-species network inference, which infer only a sub-network for which orthologs exist in every species \cite{joshi_multi-species_2015}.
+%A closely related algorithm -- fused LASSO -- has been applied to gene regulatory network inference in multiple cell lines, or multiple experimental regimes \cite{omranian2016gene, parikh2011treegl, pierson2015sharing}.
+These approaches have been shown to perform well in fusing data from multiple sources, but because the fusion is between whole networks rather than individual coefficients, they can only be applied when sources share the same set of expressed genes.
+This all but rules out cross-species applications, and can present difficulty even when data describing a single species sourced from multiple experimental techniques or platforms.
+Existing approaches to multi-species network inference have inferred only a sub-network of genes with orthologs in every species \cite{joshi_multi-species_2015}, or fit networks separately  \cite{marbach_revealing_2010}.
+Our approach allows multiple networks to be inferred with biologically motivated constraints on similarity even when the data sources have a relatively small set of overlapping genes.
 
-This general framework for multi-species network inference has broad applications.
-Any set of pairwise L2 penalties on the differences between regulatory interactions (equivalent to a multivariate Gaussian prior on networks; see methods) can be optimized by the algorithm we develop.
-As a result, any biological prior on similarity of regulation can be used in place of orthology, including non one-to-one mappings such the prior on genes within the same operon.
-We demonstrate improvements in network inference performance with the introduction of penalties based on the expectation that genes in the same operon will have similar regulators.
+%This general framework for multi-species network inference has broad applications, extending beyond the case of combining multiple networks.
+Any set of pairwise L2 penalties on the differences between regulatory interactions (equivalent to a multivariate Gaussian prior on networks; see methods) can be utilized by the algorithm we develop.
+As a result, any biological prior on similarity of regulation can be used in place of orthology.
+For example, we demonstrate improvements in network inference performance with the introduction of penalties based on the expectation that genes in the same operon will have similar regulators.
 
@@
 In this case, we assume that the constrained interactions have diverged and relax the fusion constraint between them (corresponding to the saturating regime of the penalty).
-
 The resulting cost function is non-convex and difficult to optimize \cite{fan2001variable}; however, we can approximate its solution and obtain deeper insight into functional similarity than is available through strict orthology enforcement or the comparison of separately learned networks.
@@
 
+\section{Results}
 We test the ability of fused L2 and adaptive fusion to improve network recovery on both synthetic data and on real data.
@@
 We then discuss adaptive fusion and show the technique's ability to identify incorrect orthology information which has been introduced to a biological dataset, suggesting the technique's applicability to discovering neo- and sub-functionalizations.
-
-\section{Results}
-We used both synthetic networks and real data to evaluate our methods.
-For the synthetic data, we generated random pairs of networks in which orthologous genes have similar regulatory interactions, and then sampled gene expression from these networks, which we used to derive learned networks for comparison with the input (true) networks.
-For real data, we examined recovery of a known gold standard in \textit{Bacillus subtilis} \cite{michna_subtiwikidatabase_2014}.
 
@@
 \subsubsection{Adaptive fusion successfully identifies and unfuses 'neofunctionalized' genes}
-Orthology prediction is not alwyas a good proxy for functional conservation \cite{gabaldon_functional_2013, studer_how_2009, nehrt_testing_2011}. 
+Although we expect that orthology provides some evidence of regulatory similarity, we know that it is not a perfect proxy for functional conservation \cite{gabaldon_functional_2013, studer_how_2009, nehrt_testing_2011}.
+Some gene interactions may be non-conserved, and in order to minimize bias it would be desireable to identify these interactions and relax the fusion constraints on them.
 To allow for orthologous genes to be unfused we implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (figure \ref{adapt-schematic}).
@@
 
-We then compared the recovery of interaction weights which were accurately fused, and recovery of interaction weights which were inaccurately fused due to incorrect orthology information (figure \ref{adaptivefusion}b).
-Because fused L2 heavily penalizes large differences between weights which are predicted to be similar, it is able to retrieve a more accurate network for those interactions with true fusion constraints than by learning networks separately (measured by MSE between the true and inferred interaction weights).
-In this simulation, however, the gains accomplished through fused regression do not extend to those interactions lacking true fusion constraints, and the error remains similar to learning networks separately.
-When we applied adaptive fusion, we did not observe an improvement in network recovery (relative to fused L2).
-However, we were able to identify fusion constraints reflecting incorrect orthology information that had been provided to the algorithm (Figure \ref{adaptivefusion} subsection).
-[ZACK - WE NEED TO WRITE MORE ABOUT THIS]
+We then compared the interaction weights which were accurately fused, and interaction weights which were inaccurately fused due to incorrect orthology information (figure \ref{adaptivefusion}b).
+Because fused L2 heavily penalizes large differences between weights which are predicted to be similar, all learned interaction weights with fusion constraints are similar.
+Those constraints that are unfused by adaptive fusion are allowed to differ; some of the incorrectly fused constraints have large differences in interaction weights.
 
@@
 subtilis} \cite{nicolas2012condition}.
-We compared performance when learning the networks separately and then rank combining (as in cite Ciofani) to learning the networks simultaneously using fusion regression and show improvement in performance using our fused L2 approach (figure \ref{fusedl2-real}a).
+We compared performance when learning the networks separately and then rank combining (as in \cite{ciofani_validated_2012}) to learning the networks simultaneously using fusion regression, and we show improvement in performance using our fused L2 approach (figure \ref{fusedl2-real}a).
 
-Our approach can also make use of other forms of biological data beyond orthology.
+Our approach can also make use of forms of biological data other than RNA expression.
 Information about the expected similarity of TF-gene interactions can come from knowledge about the promoter region or the structure, for bacteria, of polycistronic transcripts.
@@
 We posited, therefore, that genes within the same operon will be regulated similarly by the same transcription factors.
-We applied fusion regression by creating fusion constraints between a given transcription factor and genes within the same operon, and showed a boost in \textit{B.
-subtilis} network recovery using within-species fusion (figure \ref{fusedl2-real}b).
+We applied fusion regression by creating fusion constraints between a given transcription factor and genes within the same operon.
+We also used a subset of the gold-standard as priors by relaxing the $\lambda_R$ parameter for these interactions, as in \cite{greenfield_robust_2013}, and evaluated on the remaining set of priors.
+Incorporation of operon-derived fusion constraints in \textit{B.
+subtilis} improved network recovery (figure \ref{fusedl2-real}b).
 
@@
 In the case of cross-species network inference, there is unlikely to be a complete one-to-one orthology.
-However, because we create constraints the similarity of individual interactions, rather than on the networks as a whole \cite{parikh2011treegl}, we can pool some information across species even when a small fraction of genes have orthologs.
+However, because we create constraints the similarity of individual interactions, rather than on the networks as a whole \cite{parikh2011treegl, omranian2016gene}, we can pool some information across species even when a small fraction of genes have orthologs.
 
@@
 
+A number of existing approaches have applied fused regression to related problems in network inference.
+TreeGL applied fused LASSO to the problem of estimating partial correlations of gene expression in a breast cancer dataset consisting of multiple cell types \cite{parikh2011treegl}.
+They imposed fusion between the coefficients of models learned for different cell types whenever there was an edge between those cell types in a geneology graph describing the cancer cells' development.
+A similar approach was recently applied to the problem of inferring TF to gene regulatory weights from data sources describing multiple environmental conditions in \textit{E.
+Coli}, \textit{Mycobacterium tuberculosis}, and \textit{Mus musculus} \cite{omranian2016gene}.
+In this formulation, networks associated with each data source were fit simultaneously, with an $L1$ constraint on the similarity of (arbitrarily ordered) adjacent networks.
+In both of these approaches, fusion was restricted to be between corresponding entries of the coefficient matrices.
+In terms of our formulation in the cross-species case, this is equivalent to the constraints that would be generated by a one-to-one orthology mapping between species, limiting to the (typical) case where orthology is partial or not one-to-one.
+By optimizing a more general objective function -- in which fusion constraints can be placed on arbitrary pairs of coefficients associated with otherwise unrelated regression tasks -- we can extend this work to the cross-species case.
+
 \textit{B.
 subtilis} and \textit{B.
-anthracis} are distantly related bacterial species with limited gene orthology [ZACK - HOW DISTANTLY?].
+anthracis} are distantly related bacterial species with limited gene orthology.
 Nevertheless we show that network recovery in \textit{B.
@@
 subtilis} and \textit{B.
-anthracis} genomes do not have clear orthologs (and many orthologs have ambiguous many-to-many groupings).
-Our method, in contrast, can obtain improvements in network inference performance even when the conserved subset of genes is small. [DISCUSS RELATIONSHIP TO OTHER PAPERS HERE]
+anthracis} genomes do not have clear orthologs.
+Our method, in contrast, can obtain improvements in network inference performance even when the conserved subset of genes is small.
 
 This approach is particularly interesting in light of the diversity of important model organisms used in modern biology.
-Different model systems provide different advantages and disadvantages for experimental design \cite{stolfi_genetic_2012}, but without a principled mechanism for combining data from multiple sources, it is difficult to fully leverage data obtained from even a slightly different model system.
-
-Because the algorithm we developed can accomodate constraints between arbitrary pairs of regulatory interactions, any biological prior representing information about expected regulatory similarity can be represented, even if the prior provides no information about the magnitude or direction of regulation.
-We demonstrate this flexibility through the novel incorporation of operon structure into the gene-regulatory network inference problem. 
-In this application, fusion reflects the assumption that genes in the same operons have similar regulators \cite{lawrence_shared_2002}.
+Different model systems provide different advantages and disadvantages for experimental design \cite{stolfi_genetic_2012}, but in many cases work in less used systems is hampered by lack of available data.
+Our intent is to provide a principled method for combining data from diverse sources, so that results in specialized systems can be integrated with data from well studied organism.
 
 Although it is important to take advantage of the similarities of related organisms for generating improved models of gene regulation, it is also critically important to understand how systems differ from one another.
-Our cross-species network inference method is premised on the assumption that orthologous genes have similar regulators.
-Existing approaches to the genome-wide testing of this assumption learn regulatory networks separately, then compare to identify conservation \cite{aytes_cross-species_2014}.
+Existing approaches to the genome-wide testing of the assumption that orthologous genes have similar regulators learn regulatory networks separately, then compare to identify conservation \cite{aytes_cross-species_2014}.
 Because network inference is typically underconstrained, fitting a network that describes a particular set of experimental observations amounts to sampling a single network from a large set of networks that fit the data equally (or almost equally) as well.
@@
 Our method explicitly favors recovering networks for which evolutionarily corresponding interactions are similar. 
-As a result, the failure to obtain networks that confirm evolutionary conservation is stronger evidence that conservation does not exist; the next best network that does exhibit conservation must fit the data much worse to have overcome the bias built into the fusion constraints.
+As a result, the failure to obtain networks that confirm evolutionary conservation is much more direct evidence that the networks have truly diverged.
 

### 2016-07-29 paper_revisions.tex

@@
 
-Numerous studies have shown that functional conservation exists in gene regulatory networks even across large evolutionary distance \cite{satou2006gene, hinman2009evolution,tanay2005conservation,erwin2009evolution}.
-Our first method -- fused L2 -- takes advantage of this similariy by imposing an L2 penalty on the differences between \texit{a prior} similar interactions (termed fusion penalty).
-These constraints favor, but do not require, network configurations in which orthologous genes have similar regulators.
-Because network inference problems are typically underconstrained, these additional constraints allow data in one species to improve network inference performance in another.
-
 %A closely related algorithm -- fused LASSO -- has been applied to gene regulatory network inference in multiple cell lines, or multiple experimental regimes \cite{omranian2016gene, parikh2011treegl, pierson2015sharing}.
@@
 For example, we demonstrate improvements in network inference performance with the introduction of penalties based on the expectation that genes in the same operon will have similar regulators.
+
+Numerous studies have shown that functional conservation exists in gene regulatory networks even across large evolutionary distance \cite{satou2006gene, hinman2009evolution,tanay2005conservation,erwin2009evolution}.
+Our first method -- fused L2 -- takes advantage of this similarity by imposing an L2 penalty on the differences between \textit{a priori} similar interactions (termed fusion penalty).
+These constraints favor  network configurations in which orthologous genes have similar regulators.
+Because network inference problems are typically underconstrained, these additional constraints allow data in one species to improve network inference performance in another.
 
@@
 In the case of cross-species network inference, there is unlikely to be a complete one-to-one orthology.
-However, because we create constraints the similarity of individual interactions, rather than on the networks as a whole \cite{parikh2011treegl, omranian2016gene}, we can pool some information across species even when a small fraction of genes have orthologs.
+However, because we  constraints the similarity of individual interactions, rather than on the networks as a whole \cite{parikh2011treegl,omranian2016gene}, we can pool some information across species even when a small fraction of genes have orthologs.
 

### 2016-07-29 paper_revisions.tex
restored some stuff to intro, spellcheck, fixed bib

@@
 The problem of inferring related -- but not necessarily identical -- structure from related -- but not identical -- data is ubiquitous in biology.
-Multi-source network inference has applications for learning multiple networks in related species, for learning networks associated with distinct processes within the same species, and for learning networks based on heterogenous data sources.
+Multi-source network inference has applications for learning multiple networks in related species, for learning networks associated with distinct processes within the same species, and for learning networks based on heterogeneous data sources.
 Moreover, as it becomes possible to learn genome-wide regulatory networks, we can begin to compare and to test whether there is conservation of networks across species and biological processes. 
@@
 These constraints favor  network configurations in which orthologous genes have similar regulators.
-Because network inference problems are typically underconstrained, these additional constraints allow data in one species to improve network inference performance in another.
+Because network inference problems are typically under-constrained, these additional constraints allow data in one species to improve network inference performance in another.
 
@@
 Observing a large difference in the weights of regulatory interactions obtained though independent inference of multiple networks is perhaps the best (least biased) evidence against conservation of orthologous regulatory interactions (cases where target and regulator have orthologs across species).
-However, this is sometimes weak evidence, as network inference is typically underconstrained \cite{marbach_revealing_2010}, meaning there could be a different set of networks for which conservation does hold, and which fit the data almost as well.
+However, this is sometimes weak evidence, as network inference is typically under-constrained \cite{marbach_revealing_2010}, meaning there could be a different set of networks for which conservation does hold, and which fit the data almost as well.
 Solving the networks jointly with fusion addresses this problem, but may be biased when gene function is not conserved.
@@
 Our second method -- adaptive fusion -- attempts to solve the problem of identifying evolutionary divergence using a non-concave saturating fusion penalty to simultaneously infer the constrained networks and to learn which constraints should be relaxed (ie which parts of the network are genuinely different).
-This penalty is based on statistical techniques intended to provide unbiased regularization penalties for fused regression \cite{zhang2010nearly, fan2001variable}.
+This penalty is based on statistical techniques intended to provide unbiased regularization penalties for regularized regression \cite{zhang2010nearly, fan2001variable}.
+We extend these techniques fused regression, and provide an algorithm that approximates the solution to the resulting non-convex loss function.
 
-The main difference between the L2 fusion approach and our new adaptive fusion approach occurs when the difference between presumed analogous interactions is large despite the fusion penalty.
+%The main difference between the L2 fusion approach and our new adaptive fusion approach occurs when the difference between presumed analogous interactions is large despite the fusion penalty.
 This condition means that no network exists which simultaneously provides an adequate explanation of expression data and which conforms to our prior on similarity.
@@
 We also introduce - in the form of adaptive fusion - the idea of a saturating penalty function on fusion constraints, and estimate the solution to the resulting optimization problem through iterative application of the fused L2 algorithm.
+We start with a discussion of the fused L2 and adaptive fusion algorithms, and describe their performance on synthetic datasets intended to represent related gene regulatory networks.
+We then demonstrate the applications of the fused L2 network inference algorithm on biological data, first demonstrating gains in cross-species network inference using the bacteria species \texit{Bacillus subtilis} and the distantly related \texti{Bacillus anthracis}, then moving on to multi-platform network inference using \texit{B.
+subtilis}, and finally intra-species fused network inference using priors based on operons in \textit{B.
+subtilis}.
+We then discuss adaptive fusion and show the technique's ability to identify incorrect orthology information which has been introduced to a biological dataset, suggesting the technique's applicability to discovering neo- and sub-functionalizations.
 
 \section{Results}
-We test the ability of fused L2 and adaptive fusion to improve network recovery on both synthetic data and on real data.
-We demonstrate the applications of our method to multi-species network inference, using the bacteria species \texit{Bacillus subtilis} and the distantly related \texti{Bacillus anthracis}; multi-platform network inference using \texit{B.
-subtilis}; and intra-species fused network inference using operons in \textit{B.
-subtilis}.
-We then discuss adaptive fusion and show the technique's ability to identify incorrect orthology information which has been introduced to a biological dataset, suggesting the technique's applicability to discovering neo- and sub-functionalizations.
 
@@
 Although we expect that orthology provides some evidence of regulatory similarity, we know that it is not a perfect proxy for functional conservation \cite{gabaldon_functional_2013, studer_how_2009, nehrt_testing_2011}.
-Some gene interactions may be non-conserved, and in order to minimize bias it would be desireable to identify these interactions and relax the fusion constraints on them.
+Some gene interactions may be non-conserved, and in order to minimize bias it would be desirable to identify these interactions and relax the fusion constraints on them.
 To allow for orthologous genes to be unfused we implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (figure \ref{adapt-schematic}).
@@
 Although the approach we describe was developed with cross-species network inference in mind, the framework of defining fusion constraints on TF-gene interactions expected to be similar is quite general, and can be applied to data from a single organism.
-Despite the many large-scale collaborations which attempt to make protocols as uniform as possible for comparability between datasets generated by different labs \cite{paten_nih_2015,kundaje_integrative_2015} and several methods for removing batch effects \cite{irizarry_exploration_2003,johnson_adjusting_2007}, there still exists technical and biologial variability between many experiments attempting to capture the same or similar experimental conditions esspecially when experiments employ different experimental platforms.
+Despite the many large-scale collaborations which attempt to make protocols as uniform as possible for comparability between datasets generated by different labs \cite{paten_nih_2015,kundaje_integrative_2015} and several methods for removing batch effects \cite{irizarry_exploration_2003,johnson_adjusting_2007}, there still exists technical and biological variability between many experiments attempting to capture the same or similar experimental conditions especially when experiments employ different experimental platforms.
 With the advent of RNAseq, for example, microarray based technologies are no longer the dominant assay for genome-wide expression, but a large body of accumulated legacy data remains useful if it can be integrated with more modern techniques.
@@
 We are interested in the case where $p > N$.
-Many methods have been proposed to deal with the underconstrained case, and have been applied to genomic data \cite{waldron_optimized_2011, li_network-constrained_2008}.
+Many methods have been proposed to deal with the under-constrained case, and have been applied to genomic data \cite{waldron_optimized_2011, li_network-constrained_2008}.
 For example, ridge regression penalizes the L2 norm of the coefficients $\beta_{i,j}$ in order to avoid overfitting \cite{hoerl_ridge_1970}, and can be thought of as a mean-zero Gaussian prior on the coefficients.
-More complicated penalties have been developed to represent specific expected or desireable structure in a regression model's coefficients. 
+More complicated penalties have been developed to represent specific expected or desirable structure in a regression model's coefficients. 
 For example, Land and Friedman \cite{citeulike3780356} proposed a fusion penalty (or a penalty on the differences between certain interaction weights) which encourages smoothness of the estimated parameter vector.
@@
 Linear systems of this size can be solved quickly using standard techniques, and runtime using our bacterial datasets clocks in around thirty minutes. 
-When the size of the groups of genes linked by fusion constraints becomes large (when organisms have a number of many-to-many orthologous blocks), however, the augmented design matrix approach becomes slower and we discuss further optimisations to this scheme below to enable scaling to these regimes.
+When the size of the groups of genes linked by fusion constraints becomes large (when organisms have a number of many-to-many orthologous blocks), however, the augmented design matrix approach becomes slower and we discuss further optimizations to this scheme below to enable scaling to these regimes.
 
@@
 We propose a saturating penalty that is relaxed once differences in weights grow beyond a certain point (interactions which appear to be very different based on the data are effectively unfused). 
-A related problem has been studied in the context of LASSO regularization, where it was shown by Fan and Li that using a saturating penalty retains many of LASSO's desireable properties while removing its bias towards model weights of 0 \cite{fan2001variable}.
-They further showed that, although the resulting loss-function is nonconvex, good results can be obtained with a local quadratic approximation of gradient descent.
+A related problem has been studied in the context of LASSO regularization, where it was shown by Fan and Li that using a saturating penalty retains many of LASSO's desirable properties while removing its bias towards model weights of 0 \cite{fan2001variable}.
+They further showed that, although the resulting loss-function is non-convex, good results can be obtained with a local quadratic approximation of gradient descent.
 Several saturating penalties, such as SCAD \cite{fan2001variable} and MCP \cite{zhang2010nearly}, have been discussed in the context of sparse regression.
@@
 	\State run fused regression
-	\State return PRC and ROC curves using leaveout gold standard
+	\State return PRC and ROC curves using leave-out gold standard
 	\EndFor
@@
 TreeGL applied fused LASSO to the problem of estimating partial correlations of gene expression in a breast cancer dataset consisting of multiple cell types \cite{parikh2011treegl}.
-They imposed fusion between the coefficients of models learned for different cell types whenever there was an edge between those cell types in a geneology graph describing the cancer cells' development.
+They imposed fusion between the coefficients of models learned for different cell types whenever there was an edge between those cell types in a genealogy graph describing the cancer cells' development.
 A similar approach was recently applied to the problem of inferring TF to gene regulatory weights from data sources describing multiple environmental conditions in \textit{E.
@@
 Existing approaches to the genome-wide testing of the assumption that orthologous genes have similar regulators learn regulatory networks separately, then compare to identify conservation \cite{aytes_cross-species_2014}.
-Because network inference is typically underconstrained, fitting a network that describes a particular set of experimental observations amounts to sampling a single network from a large set of networks that fit the data equally (or almost equally) as well.
+Because network inference is typically under-constrained, fitting a network that describes a particular set of experimental observations amounts to sampling a single network from a large set of networks that fit the data equally (or almost equally) as well.
 As a result, the existence of a difference between corresponding regulatory interactions in a pair of experimentally derived networks is weak evidence that a difference truly does exist.
@@
 We have described a method -- adaptive fusion -- that attempts to address this bias by simultaneously learning the networks and constraint weights.
-This method is based on minimizing a saturating penalty function on fusion constraints, similar to a class of penalties that have been developed to minimise bias in regularized regression \cite{fan2001variable,zhang2010nearly}.
+This method is based on minimizing a saturating penalty function on fusion constraints, similar to a class of penalties that have been developed to minimize bias in regularized regression \cite{fan2001variable,zhang2010nearly}.
 The result of adaptive fusion is both a network and a new set of fusion constraints, the weights of which can be interpreted as describing conservation structure across the networks.For the multiple species case, relaxation of fusion constraints represents orthologs which do not share similar interactions presumably due to evolution of regulatory circuitry \cite{kellis_proof_2004}.
@@
 
-Because our model shares its basic assumptions about the role of transcription factors in gene expression dynamics with models developed for single-species network inference, we are able to leverage techniques developed for the single-species estimation of transctiption factor activity \cite{fu_reconstructing_2011}.
+Because our model shares its basic assumptions about the role of transcription factors in gene expression dynamics with models developed for single-species network inference, we are able to leverage techniques developed for the single-species estimation of transcription factor activity \cite{fu_reconstructing_2011}.
 The performance gains of this additional step in the cross-species case are significant.
@@
 These priors -- incorporated into our method as fusion constraints -- allow the use of rich sources of biological knowledge, such as orthology and operon structure, which have informed experimental design, but are typically not incorporated into genome wide network inference algorithms.
-By accomodating the simultaneous inference of multiple related networks, we can improve network inference performance by allowing the efficient reuse of data from similar, but not necessarily identical, sources.
+By accommodating the simultaneous inference of multiple related networks, we can improve network inference performance by allowing the efficient reuse of data from similar, but not necessarily identical, sources.
 A method for pooling data from multiple sources holds the promise of vastly expanding the quantity of data available for analysis, particularly in less commonly used model systems. 
@@
   \caption{\label{adapt-schematic} Adaptive fusion loss function (\textbf{A}) and derivative of loss function (\textbf{B}). 
-  \textbf{A.} Adaptive fusion is a quadratic around the origin, begins to taper at $a/2$, and plateus at $a$.
-After the plateu, increasing the difference in interaction weight of fused interactions does not further affect the penalty incurred through fusion.
+  \textbf{A.} Adaptive fusion is a quadratic around the origin, begins to taper at $a/2$, and plateaus at $a$.
+After the plateau, increasing the difference in interaction weight of fused interactions does not further affect the penalty incurred through fusion.
 As a result, interaction weights in this zone are effectively unfused from one another (the fusion penalty behaves like a constant).

### 2016-07-30 paper_revisions.tex
zack changes. Figure legends may need to be changed more if the layout changes

@@
 In order to compare the relative contribution of data in the species of interest (species 1) and in a related organism (species 2) to network recovery, we varied the number of conditions in each species while evaluating the mean-squared-error between the recovered network for species 1 and its known true network.
-When the amount of data from the related species was held constant, increasing the amount of data available for learning the network for the species of interest resulted in a more accurate network prediction, as expected  (figure \ref{fusion}b).
+When the amount of data from the related species was held constant, increasing the amount of data available for learning the network for the species of interest resulted in a more accurate network prediction, as expected  (Figure \ref{fusion}b).
 Similarly, when we increased the amount of data from the related species, we obtained performance gains on network one using fused L2 regression, demonstrating our ability to improve network inference on one dataset through incorporation of a related dataset.
@@
 
-\subsubsection{Adaptive fusion successfully identifies and unfuses 'neofunctionalized' genes}
+\subsubsection{Adaptive fusion with simulated data}
 Although we expect that orthology provides some evidence of regulatory similarity, we know that it is not a perfect proxy for functional conservation \cite{gabaldon_functional_2013, studer_how_2009, nehrt_testing_2011}.
 Some gene interactions may be non-conserved, and in order to minimize bias it would be desirable to identify these interactions and relax the fusion constraints on them.
-To allow for orthologous genes to be unfused we implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (figure \ref{adapt-schematic}).
+To this end, we developed an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (Figure \ref{adapt-schematic}).
 Pairs of interactions that are dissimilar even after fusion, which sit in the flat portion of this penalty function, are effectively ``unfused,'' and no further penalty is incurred as differences in interaction weights grow.
 Our network procedure strongly favors similarity of fused interactions, and only ``unfuses'' interactions when their similarity cannot be reconciled with expression data.
-As a result, the ``unfusing'' or relaxation of the fusion penalty on certain constraints is much more direct evidence for neofunctionalization than comparing separately fit networks could provide.
+As a result, the ``unfusing'' or relaxation of the fusion penalty on certain constraints serves not only to reduce bias in the inferred network weights, but can be interpreted directly as evidence against conservation.
 
-We performed a simulation to assess the ability of our adaptive fusion algorithm to distinguish which parts of two input networks are conserved vs.
-neofunctionalized (figure \ref{adaptivefusion}).
-We generated synthetic fused networks and introduced error in the fusion constraints by adding false positives and negatives to the orthology information given to the solver.
-Because we knew which entries in the orthology mapping were ``incorrect'' (not reflected in the generation of the networks), we could correctly label fusion constraints that involved one or more ``incorrect'' mappings.
-We verified that adaptive-fusion unfused mostly ``incorrectly fused'' interactions (figure \ref{adaptivefusion}a red dots), while leaving truly analogous interactions fused (figure \ref{adaptivefusion}a green dots).
-
-We then compared the interaction weights which were accurately fused, and interaction weights which were inaccurately fused due to incorrect orthology information (figure \ref{adaptivefusion}b).
-Because fused L2 heavily penalizes large differences between weights which are predicted to be similar, all learned interaction weights with fusion constraints are similar.
-Those constraints that are unfused by adaptive fusion are allowed to differ; some of the incorrectly fused constraints have large differences in interaction weights.
+We performed a simulation to assess the performance of adaptive fusion in the context of partially conserved networks.
+We generated synthetic fused networks based on a partial orthology mapping, then randomly assigned the remaining genes unrelated orthologs.
+These false orthologs gave rise to false fusion constraints, which if enforced would degrade network inference performance.
+Slightlymore than half of the fusion constraints in the generated network were false.
+We were interested in the ability of adaptive fusion to identify these false fusion constraints, as well as improve the accuracy of the inferred networks by reducing the bias associated with fusion.
+We verified that solving these partially conserved networks with fused-L2 forced interactions bound by both true and false fusion constraints to be similar (Figure \ref{adaptivefusion}b).
+Fusion led to improved performance on the conserved part of the network and degraded performance on the non-conserved part of the network (Figure \ref{adaptivefusion}d, Fused-L2).
+Applying adaptive fusion with the ``a'' parameter set to a percentile that roughly matched the fraction of false fusion constraints preserved the performance gains obtained through fusion on the conserved part of the network, while substantially reducing the loss on the non-conserved part of the network.
+Many of the incorrectly fused interactions were relaxed, allowing these interactions to freely fit the data without bias from a faulty prior.
+These non-conserved interactions -- which have diverged from one another and left the diagonal in  Figure \ref{adaptivefusion}c -- can be identified through inspection of the $\lambda_S$ parameter following adaptive fusion, which falls to zero for unfused interactions.
+This application suggests that adaptive fusion has use as both a tool to reduce bias while retaining many of the benefits of fused regression, and as a method for determining conservation across networks.
 
@@
 subtilis} data from that fold alone fused to the entire 72 \textit{B.
-anthracis} conditions (figure \ref{xspecies-real}a).
+anthracis} conditions (Figure \ref{xspecies-real}a).
 Though overall performance is hindered by our subsampling of \textit{B.
 subtilis} data (a necessary procedure to allow evaluation of networks) we demonstrate marked improvement in learning the \textit{B.
-subtilis} network when using fused regression (figure \ref{xspecies-real}a).
+subtilis} network when using fused regression (Figure \ref{xspecies-real}a).
 Notably, these performance gains occur mostly at low values of recall.
@@
 \subsubsection{Adaptive fusion using bacterial data}
-[ZACK MORE HERE TOO]
- The goal of adaptive fusion is to unfuse constraints between non-conserved interactions, while leaving intact all other constraints. 
+When we applied adaptive fusion to our bacterial datasets we did not see a significant improvement in network recovery relative to fused-L2.
+Based on simulation results, we expected that non-conserved interactions weights would be fit more accurately with adaptive-fusion than with fused-L2, but that both algorithms would perform worse than independently fitting the networks on this subset of interactions.
+It may be that the quality of the independently fit network was too poor in this case for fusion constraints on non-conserved interactions to meaningfully degrade network inference performance.
+Nevertheless, we were interested in the application of adaptive fusion to identifying neo-functionalized or non-conserved interactions.
+Essentially, we wanted to identify constraints arising from orthology that does not correspond to functional conservation.
 However, because we lacked a comprehensive gold standard of known non-conserved interactions between \textit{B.
 subtilis} and \textit{B.
-anthracis}, we were unable to directly evaluate how accurately adaptive fusion identified these interactions.
-  We opted instead to introduce a large number of random fusion constraints between genes not known to be orthologous. 
- These fake constraints, which are unlikely to reflect any conserved network structure, served as a proxy for the unknown fraction of non-conserved interactions between orthologous genes. 
- We ran adaptive fusion to learn networks for \textit{B. subtilis} and \textit{B. anthracis}, using these constraints, along with those generated by known orthology. 
- We confirmed that our injected spurious fusion constraints were unfused at a higher rate than those generated by known orthologs (see figure \ref{xspecies-real}b & c). 
- Although it may seem odd that a large fraction of fake constraints were left intact, we note that biological networks tend to be sparse, so that many of the random fusion constraints are between coefficients with near zero weight (and therefore near zero difference in weight) (figure \ref{xspecies-real}c).
+anthracis}, we were forced to indirectly evaluate the performance of adaptive fusion in identifying these interactions.
+
+We obtained a known set of non-conserved interactions by randomly generating orthology mappings between genes not known to be orthologous.
+These orthology mappings generated false constraints, as in our simulation studies,  which are unlikely to reflect any conserved network structure.
+We used these constraints served as a proxy for the unknown fraction of non-conserved interactions between orthologous genes.
+ We ran adaptive fusion to learn networks for \textit{B.
+subtilis} and \textit{B.
+anthracis}, using these constraints along with those generated by known orthology.
+We confirmed that our injected spurious fusion constraints were unfused at a higher rate than those generated by known orthologs (see Figure \ref{xspecies-real}b & c).
+Although it may seem odd that a large fraction of fake constraints were left intact, we note that biological networks tend to be sparse, so that many of the random fusion constraints are between coefficients with near zero weight (and therefore near zero difference in weight) (Figure \ref{xspecies-real}c).
+It is unclear to what extent the relatively large fraction of fusion constraints between true orthologs that were unfused reflects an inadequacy of the method, or a relatively high rate of neofunctionalizaton in these organisms.
 
@@
 Currently, the most widely used approach to combining datasets for network inference is to learn networks from disparate datasets separately, then rank combine the networks as in Marbach et al \cite{marbach_revealing_2010}.
-We included, along with our \textit{B.
+We included, along with our main \textit{B.
 subtilis} dataset, a previously published dataset containing 269 samples covering 104 conditions, obtained using a different tiling microarray (vs custom microarray) and different strain of \textit{B.
 subtilis} \cite{nicolas2012condition}.
-We compared performance when learning the networks separately and then rank combining (as in \cite{ciofani_validated_2012}) to learning the networks simultaneously using fusion regression, and we show improvement in performance using our fused L2 approach (figure \ref{fusedl2-real}a).
+We compared performance when learning the networks separately and then rank combining (as in \cite{ciofani_validated_2012}) to learning the networks simultaneously using fusion regression, and we show improvement in performance using our fused L2 approach (Figure \ref{fusedl2-real}a).
 
@@
 Incorporation of operon-derived fusion constraints in \textit{B.
-subtilis} improved network recovery (figure \ref{fusedl2-real}b).
+subtilis} improved network recovery (Figure \ref{fusedl2-real}b).
 
@@
 To test the integration of this approach with our fused regression, we assessed the combination of \textit{B.
-subtilis} datasets, as in figure \ref{fusedl2-real}a, with the incorporation of TFA estimation.
+subtilis} datasets, as in Figure \ref{fusedl2-real}a, with the incorporation of TFA estimation.
 We randomly divided the prior known interactions in half, and used half to learn TFA and to generate priors on network structure. 
 The remaining interactions were reserved as a gold standard for validation. 
-As in previous studies, we observed a marked improvement in network inference when using transcription factor activity (figure \ref{tfa}).
+As in previous studies, we observed a marked improvement in network inference when using transcription factor activity (Figure \ref{tfa}).
 We also obtained AUPR improvement when using fused regression alongside TFA; gains from sharing information across datasets using fused regression were preserved and even enhanced by using TFA.
@@
 In an ordinary regression problem each response variable can be solved independently, and vectorization is unnecessary. 
-However, in fused regression, we append additional rows to the design matrix that link entries of the interaction weight matrix associated with different response variables (figure \ref{cartoon}). 
+However, in fused regression, we append additional rows to the design matrix that link entries of the interaction weight matrix associated with different response variables (Figure \ref{cartoon}). 
 As a result, these linked response variables must be solved simultaneously through vectorization. 
@@
 Although $L1$ penalties have the theoretical appeal of sparsity - which matches our assumption of sparsity in biological networks - we are primarily interested here in ranking interactions, rather than recovering a specific network.
-Moreover, $L2$ penalized regression followed by thresholding can be competitive with $L1$ penalized regression for support recovery \cite{}.
+Moreover, $L2$ penalized regression followed by thresholding can be competitive with $L1$ penalized regression for support recovery \cite{} [LOOK UP THIS REF].
 Although our approach can be easily modified to use an $L1$ penalty, we did not observe performance gains when doing so, and the version presented here is in some ways conceptually simpler (with the combined penalty function being equivalent to a multivariate Gaussian prior).
@@
 Nevertheless we show that network recovery in \textit{B.
-subtilis} can be improved through the inclusion of expression data from \textif{B.
+subtilis} can be improved through the inclusion of expression data from \textit{B.
 anthracis}.
@@
     \includegraphics[scale=.5,trim=0mm 0mm 0mm 0mm,clip]{party3.eps}
-    \caption{\label{cartoon} Schematic representation of design matrix construction. Here, the circles and hexagons correspond to different species. 
+    \caption{\label{cartoon} Schematic representation of design matrix construction.
+Here, the circles and hexagons correspond to different species.
 Bidirectional arrows represent orthology information and dotted arrows represent putative interactions between TFs and genes.
-    Rectangles represent matrices; because weights can be solved independently unless there exists fusion constraints between them, we identify related weights and construct matrix for solving. }
+Rectangles under ``Data'' represent TF $\times$ condition matrices of gene expression values in species 1 (top row) and species 2 (bottom row), colored in correspondence with the gene orthology diagram, with bidirectional arrows representing orthology between the two species.
+Rectangles under ``Weights'' represent gene regulatory interactions in each species, with lines linking coefficients that are fused due to the orthology information shown to the left.
+Networks associated with genes in each species can be solved independently unless there exist a fusion constraint constraining coefficients of each gene, or a path of such constraints.
+When a path does exist, these genes must be solved simultaneously.
+In this example, genes $B$, $C$, and $B'$ must be solved simultaneously; the lower right corner shows a representation of the design matrix necessary to solve the this fused regression problem.
+    }
 \end{figure}
@@
 \begin{center}
-  \includegraphics[scale=0.45]{fig1.pdf}
+  \includegraphics[scale=0.30]{fig1.pdf}
   \caption{\label{fusion} \textbf{A.} Schematic representation of the the generation of fusion constraints from orthology mappings.
@@
 We introduce fusion constraints for pairs of interactions for which both the regulator and regulated gene are orthologs of one another.
-  In this example, we would introduce a constraint between the $(A, B)$ and $(A', B')$ interactions and the $(A, C)$ and $(A', B')$ interactions. \textbf{B.} 
+In this example, we would introduce a constraint between the $(A, B)$ and $(A', B')$ interactions and the $(A, C)$ and $(A', B')$ interactions.
+\textbf{B.}
   In order to demonstrate the utility of fused network inference in combining data, we generate two networks with 10 TFs and 200 genes (75\% sparsity).
@@
   \textbf{C-F} Show the varying effects of fusion on simulated networks with different levels of conservation.
-We generate a series of networks with 20 TFs by 200 genes in two species, (50\% sparsity) while varying the fraction of gene orthologies in the simulated networks (Amount fused, top).
+We generate a series of networks with 20 TFs by 200 genes in two species, (50\% sparsity) while varying the fraction of gene orthologies in the simulated networks (Amount fused, shown above each figure).
 For each network, we evaluated AUPR on one of the species for: all interactions (blue line), interactions with fusion constraints (green line), and interactions without fusion constraints (orange line).
@@
 \begin{center}
-  \includegraphics[scale=0.45]{fig3.pdf}
+  \includegraphics[scale=0.30]{fig3.pdf}
   \caption{\label{adapt-schematic} Adaptive fusion loss function (\textbf{A}) and derivative of loss function (\textbf{B}). 
@@
 \begin{center}
-  \includegraphics[scale=0.45]{fig4_2.pdf}
-  \caption{\label{adaptivefusion} In order to evaluate the performance of adaptive fusion in unfusing interactions that have diverged through evolution, we performed a series of simulations inferring networks given a partially corrupted list of orthology mappings.
-  Networks were generated with 35 TFs by 200 genes, 60\% orthology coverage and 40\% false orthology coverage.
-\textbf{A.} We plot the interaction weights between pairs of fused interactions in network 1 (x-axis) and network 2 (y-axis) following network inference without fusion ($\lambda_S=0$).
-  Interactions that are generated from false orthologs are marked as red, while interactions generated from true orthologs are shown in green. 
-  As expected, false fusion constraints give rise to uncorrelated weights (red dots). \textbf{B.} 
-  When fit with fused-L2, fusion constraints give rise to very similar weights in the two species for 'true' and 'false' interactions. 
-  \textbf{C.} Adaptive fusion run on the same network unfuses constraints for which the inferred weights are dissimilar beyond a certain point. 
-  Here, unfused interactions are almost entirely interactions between false orthologs.
-\textbf{D-F} Show the distribution of the absolute value of the difference in inferred weight for interactions with true fusion constraints (green) and false fusion constraints (orange)}
+  \includegraphics[scale=0.30]{figure_4_rev1.pdf}
+  \caption{\label{adaptivefusion} In order to evaluate the performance of adaptive fusion in the presence of non-conserved interactions, we performed a series of simulations inferring networks given a partially corrupted list of orthology mappings (or, orthology mappings that do not represent functional similarity).
+    Networks were generated with 35 TFs by 200 genes, 60\% orthology coverage, 40\% false orthology coverage, and 30 samples per network.
+\textbf{A.} Top: we plot the interaction weights between pairs of fused interactions in network 1 (x-axis) and network 2 (y-axis) following network inference without fusion ($\lambda_S=0$).
+True fusion constraints are generated from pairs of true orthologs, while false fusion constraints derive from a pair of orthologs at least one of which is false.
+When $\lambda_S=0$ -- equivalent to fitting the networks independently -- interactions linked by false constraints are uncorrelated with one another, while interactions linked by true constraints are correlated, reflecting their functional similarity.
+Below we plot the distribution of differences in weights joined by true and false constraints; true constraints have differences on average closer to zero.\textbf{B.} shows the weights of interactions when networks are fit using fused-L2.
+In this condition, interactions bound by both true and false fusion constraints are forced to be very similar across the two species.
+\textbf{C.} With adaptive fusion, many of the false fusion constraints are relaxed.
+These constraints are no longer forced to be similar, and some of the original structure is restored.
+\textbf{D.} Shows the effect of fused-L2 and adaptive fusion on the accuracy of network recovery, as measured by MSE between the known simulated network weights and the inferred network weights.
+When fusion is introduced, recovery of the correctly fused part of the network improves, but the bias induced by fused-L2 degrades performance on the incorrectly fused part of the network.
+With adaptive fusion, the gains on the conserved part of the network are preserved with significantly less loss of accuracy on the non-conserved part of the network.
+Error bars are $95\%$ confidence intervals on the squared error of coefficients.}
 \end{center}
@@
 \begin{center}
-  \includegraphics[scale=0.45]{fig5_2.pdf}
-  \caption{\label{xspecies-real} a. Using optimal lamR and lamP values, we use 20-fold cross validation to learn B. subtilis network. 
-  We comparedare performance when solving without fusion, and using L2 fusion with lamS = 1.0. Mean AUPR for lamS = 0 (unfused): 0.0298. 
-  Mean AUPR for lamS = 1: 0.0388. b. We test adaptive fusion using the same setup, with the addition of false orthology information.
-  We set a 30\% false potentialsitive rate, including 561 additional orthologs.
-We run adaptive fusion, setting the $a$ term equal to the value above which 40\% of constraints would unfuse, reflecting our belief that in addition to some of the known orthologs result in fusion constraints which should be relaxed.
-  Here we show the percentage of constraints relaxed, from constraints created from the false orthologs and the known orthologs. 
-  We also show the distributions of differences in weights corresponding to fused constraints created from the false orthologs and the known orthologs, when networks are solved separately. }
+  \includegraphics[scale=0.30]{fig5_2.pdf}
+  \caption{\label{xspecies-real} \textbf{A} We first optimized $\lambda_R$, then performed a 20-fold cross-validation procedure in which subsets of the \textit{B.
+subtilis} network were fit either independently or fused to the \textit{B.
+anthracis network}.
+Each \textit{B.
+subtilis} network was then evaluated on a gold standard of known interactions, to simulate results of network inference with a small amount of available data in the species of interest (\textit{B.
+subtilis}) but more data available in a related species (\textit{B.
+anthracis}).
+Fusion with $\lambda_S=1.0$ improved performance, yielding a mean AUPR across the learned networks of 0.0388 vs.
+0.0298 fit independently of  \textit{B.
+anthracis}.
+Plotted is the resulting precision-recall curve, with significant performan gains at low values of recall.
+In \textbf{B.,C.} we tested adaptive fusion using the same procedure, adding an additional 561 randomly selected additional orthologs.
+We set the $a$ parameter of adaptive fusion equal to $60t$ percentile of the differences in fused interaction weights when networks were fit with $\lambda_S=0$, so that approximately $40\%$ of constraints were relaxed.
+This reflected our belief that many of the fusion constraints arising from known orthologies did not necessarily reflect functional similarity, and could be unfused.
+\textbf{B.} Shows the fraction of fusion constraints arising from known orthologies and false random orthologies which were relaxed by adaptive fusion; a larger fraction of constraints known to not reflect functional conservation were relaxed.
+\textbf{C.} After fitting the networks independently, we plot the distribution of the differences in weight of interactions fused by known false fusion constraints and fusion constraints arising from known orthology.}
 \end{center}
@@
 \begin{center}
-  \includegraphics[scale=0.45]{fig6.pdf}
-  \caption{\label{fusedl2-real} a.
-We use optimal lamR and lamP values, and solve for B.
-subtilis-1 and B.
-subtilis-2 networks using 10-fold cross validation.
-We fuse our original B.
-subtilis-1 dataset to another B.
-subtilis dataset, which we call B.
-subtilis-2, using orthology information, and evaluate performance of B.
-subtilis-1 using AUPR on gold standard.
-We compare L2 fusion with solving the networks separately without fusion, then rank combining as in Marbach et al., as well as solving B.
-subtilis on its own.
-b.
-We again use optimal lamR and lamP values, and solve for the B.
-subtilis network using 10-fold cross validation, and evaluate using AUPR on B.
-subtilis with gold standard.
-  Here, we fuse genes in the same operon group and compare L2 fusion performance using operons with unfused network inference.}
+  \includegraphics[scale=0.30]{fig6.pdf}
+  \caption{\label{fusedl2-real} Demonstration of the application of fused-L2 to intra-species network inference \textit{B.
+subtilis}.
+In each example, $\lambda_R$ is optimized separately without fusion and 10-fold cross validation is used when fitting networks (although, in \textbf{A} the gold-standard was not used in fitting the network and did not vary across folds) \textbf{A.} We compared the performance of independently fitting our main \textit{B.
+Subtilis} dataset with two methods for incorporating data from another strain of \textit{B.
+subtilis}.
+We evaluated performance on a gold-standard of known interactions.
+Adaptive fusion outperforms both an independently fitting the first \textit{B.
+subtilis} dataset and fitting both \textit{B.
+subtilis} datasets then rank-combining the results, as in Marbach et al.
+\textbf{B} We demonstrate the application of a prior based on operon membership.
+We generated fusion constraints between pairs of interactions for which both the TF and gene belonged to the same operon respectively.
+We then held out half of the gold-standard and used it as a prior on individual interactions, as in \cite{greenfield_robust_2013}.
+We fit the \textit{B.
+subtilis} network with and without fusion, then evaluated on the remaining gold-standard.
+In this example, using fusion constraints to enforce a prior based on co-regulation of genes in the same operon improved network inference performance.}
 \end{center}
@@
 \begin{center}
-  \includegraphics[scale=0.45]{fig7.pdf}
-  \caption{\label{tfa} We integrate transcription factor activity (TFA) into our network inference, and solve for B.
-subtilis-1 and B.
-subtilis-2 networks using 10-fold cross validation, using fused L2 with and without TFA.
-As in figure 6, we evaluate performance on B.
-subtilis-1 using AUPR on gold standard.}
+  \includegraphics[scale=0.30]{fig7.pdf}
+  \caption{\label{tfa} Demonstrates the integration of transcription factor activity (TFA) in fused network inference.
+The procedure was identical to Figure \ref{fusedl2-real}a, except for the additional pre-processing step of transforming transcription factor abundances into an estimate of their activity (see Methods).
+We then compared performance on the main \textit{B.
+Subtilis} strain with and without fusion to the second strain, and with and without TFA.
+TFA outperforms both fused-L2 and unfused inference based on transcription factor abundance, but TFA combined with fusion dramatically outperforms all three methods.}
 \end{center}

### 2016-08-04 paper_revisions.tex
hi

@@
 We start with a discussion of the fused L2 and adaptive fusion algorithms, and describe their performance on synthetic datasets intended to represent related gene regulatory networks.
-We then demonstrate the applications of the fused L2 network inference algorithm on biological data, first demonstrating gains in cross-species network inference using the bacteria species \texit{Bacillus subtilis} and the distantly related \texti{Bacillus anthracis}, then moving on to multi-platform network inference using \texit{B.
+We then demonstrate the applications of the fused L2 network inference algorithm on biological data, first demonstrating gains in cross-species network inference using the bacteria species \texit{Bacillus subtilis} and the distantly related \textit{Bacillus anthracis}, then moving on to multi-platform network inference using \texit{B.
 subtilis}, and finally intra-species fused network inference using priors based on operons in \textit{B.
@@
 Although $L1$ penalties have the theoretical appeal of sparsity - which matches our assumption of sparsity in biological networks - we are primarily interested here in ranking interactions, rather than recovering a specific network.
-Moreover, $L2$ penalized regression followed by thresholding can be competitive with $L1$ penalized regression for support recovery \cite{} [LOOK UP THIS REF].
+Moreover, $L2$ penalized regression followed by thresholding can be competitive with $L1$ penalized regression for support recovery \cite{shao_estimation_2012}.
 Although our approach can be easily modified to use an $L1$ penalty, we did not observe performance gains when doing so, and the version presented here is in some ways conceptually simpler (with the combined penalty function being equivalent to a multivariate Gaussian prior).
@@
 Genes may be fused together on the basis of similar binding sites or chromatin features, and the relaxing of the fusion penalty indicates divergence of gene function.
-[ZACK - CONSIDER CUTTING, SPECULATIVE]
 
@@
 \begin{center}
-  \includegraphics[scale=0.30]{fig1.pdf}
+  \includegraphics[scale=0.30]{fig2_revisions.pdf}
   \caption{\label{fusion} \textbf{A.} Schematic representation of the the generation of fusion constraints from orthology mappings.
@@
 \textbf{B.}
-  In order to demonstrate the utility of fused network inference in combining data, we generate two networks with 10 TFs and 200 genes (75\% sparsity).
+  In order to demonstrate the utility of fused network inference in combining data, we generate two networks with 10 TFs and 200 genes, 75\% of which are orthologous (75\% sparsity).
 Mean squared error of the inferred vs.
@@
 \begin{center}
-  \includegraphics[scale=0.30]{fig5_2.pdf}
+  \includegraphics[scale=0.30]{fig5_3.pdf}
   \caption{\label{xspecies-real} \textbf{A} We first optimized $\lambda_R$, then performed a 20-fold cross-validation procedure in which subsets of the \textit{B.
@@
 anthracis}.
-Plotted is the resulting precision-recall curve, with significant performan gains at low values of recall.
+Plotted is the resulting precision-recall curve, with significant performance gains at low values of recall.
+We show 95\% bootstrap confidence interval.
 In \textbf{B.,C.} we tested adaptive fusion using the same procedure, adding an additional 561 randomly selected additional orthologs.
